---
title: "Plant species information"
author: "Marcela Aragon"
date: "24/1/2024"
output: html_document
---

#General Settings

let's first set working directory and load packages that we might need
```{r}

#Set working directory to project directory 
getwd() #ok

#let's deactivate renv for this script
renv::deactivate()

#Now load your (global) libraries:
library(plyr)
library(knitr)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(ggpubr)
library(tibble)
library(readxl)
library(readr)
library(ape)
library(phyloseq)

```

```{r}
## Settings for plots

# Set axis looks
theme_set(theme_bw())
axis_looks <- theme(axis.text.x = element_text(colour = "black", size = 9,
                                               face = "bold", angle=90, vjust = 0.5, hjust=1))+
  theme(axis.text.y = element_text(colour = "black", size = 9, face = "bold"))+
  theme(axis.title=element_text(size=10, face = "bold"))

```

#Get data

I would like to create a df with all the right information per species, including correct name, correct tribe (Kasper's tree), Family experiment code and all of the information that was used to select these species to represent the Brassicaceae family phylogeny. 

Furthermore, I would like to use this information to arrange the df with species_growth and re-create the regressions that were used to classify plants as fast-medium-slow.

The data is scattered, so let's start by gathering all the available data 

```{r}

#1
#Has the raw data of germination percentage and selection criteria used in 2021, only selected spp for Experiment
selection_experiment_full <- read_excel("./16_Plant_species_information/data/Family_spp_MegaExcel.xlsx",
                        sheet = "Selection_criter")

#2
#Has the raw data of germination percentage and selection criteria used in 2021, all of the species 
plant_sp_candidates_full <- read_excel("./16_Plant_species_information/data/Family_spp_MegaExcel.xlsx",
                        sheet = "Rounds_info")

#3
#Has species full name, species speed, and species abbreviation name
#this is the same info as the one contained in the original ps_16S object
phylo_ps <- read_csv("16_Plant_species_information/data/Plant_species_phylogeny_metadata.csv")

#4
#Has the info from Kasper on plants used for the alignment + phylogenetic tree
#It contains the most updated names for species
df_Kasper_tree <- read_excel("./16_Plant_species_information/data/2023-11-16_Brassicaceae_crop_genetic_data_for_MarcelaAG.xlsx")

#5
#Has all the raw data of rounds of growth 
growth_speed <- read.csv("./16_Plant_species_information/data/plantgrowth.csv")

#6
#Final plant phylogeny from Kasper
Bra_tree <- read.tree("./17_Plant_phylogeny/Results/Supermatrix_Kasper_tree_renamed_rotated_rooted.tre")

#quick look
plot(Bra_tree, type="phylogram", align.tip.label=TRUE)
phylo_tips <- rev(Bra_tree$tip.label) #vector with phylogeny order of 22 final species

#7
#SuperTribe classification of species
supertribe <- read_excel("./16_Plant_species_information/data/Super_Tribe.xlsx")

#8
#Has the Final metadata df from ps object

#this will load the CSS-normalized 16S and ITS data as a list of two objects, used for analysis
#we only need the sample_data info
load("./Data/phyloseq_objects/CSS_BacFun_rhizoplane_HighFilt_ps_l.RData")

#Get only metadata df
df_full_ps <- data.frame(sample_data(CSS_BacFun_rhizoplane_HighFilt_ps_l$Bac))

#clean
rm(CSS_BacFun_rhizoplane_HighFilt_ps_l)

```

## 1. Species taxonomic info

First, make a df with the key names for species name, species abbreviation names,the current correct taxonomic names and super tribe classification

```{r}

#You'll need:
df_full_ps #sample data from the phyloseq object, this was used to analyze everything.
df_Kasper_tree #df with accurate and most up to date names of species + phylogenetic tree info
supertribe #classification of species in SuperTribes

#make small df with the key between full species name, sp abb name and MY presence: 22 species 
key_for_spp <- df_full_ps %>% 
               select(Sp_full_name, Sp_abb_name, Sp_Tribe, MY_tree_presence, Speed)%>%
               distinct(Sp_full_name, Sp_abb_name, Sp_Tribe, MY_tree_presence, Speed)%>% #keep only the unique 
               tibble::remove_rownames()%>%                                       #combination once.
               filter(MY_tree_presence=="inside_MY_tree")%>%
               droplevels()%>%
               arrange(factor(Sp_abb_name, levels=phylo_tips))

#Now make a single df with accurate phylogenetic tree information
df_sp_names_complete <- left_join(key_for_spp, df_Kasper_tree, by="Sp_abb_name")

df_sp_names_complete <- df_sp_names_complete %>% 
                        rename(Species_new_name_correct = Species_name_new)%>%
                        rename(Species_old_name_correct = Species_name_old)%>% 
                        rename(Tribe_new_correct = Tribe)%>%
                        relocate(Sp_abb_name, Sp_full_name, Species_new_name_correct, Tribe_new_correct)

df_sp_names_complete <- left_join(df_sp_names_complete, supertribe, by="Sp_abb_name")
df_sp_names_complete <- df_sp_names_complete %>% 
                        relocate(SuperTribe, .after=Tribe_new_correct)

#add "_" to blank spaces in text to avoid errors later
df_sp_names_complete$Species_new_name_correct <- gsub(" ", "_", df_sp_names_complete$Species_new_name_correct)
df_sp_names_complete$Species_old_name_correct <- gsub(" ", "_", df_sp_names_complete$Species_old_name_correct)
df_sp_names_complete$SuperTribe <- gsub(" ", "_", df_sp_names_complete$SuperTribe)

#save it
#write.csv(df_sp_names_complete, #"./16_Plant_species_information/data/df_sp_names_complete_clean.csv", row.names=FALSE)

```

## 2. Selection criteria info

Now, get extra info of selected species 
 
```{r}

#You'll need:
plant_sp_candidates_full #df with everything, not cleaned but useful to document process of selection
selection_experiment_full #df with final species selection, get info from here 

##Process of species selection

#Arrange df's
plant_sp_candidates_full <- plant_sp_candidates_full[,c(2:12)] #keep only necessary columns
colnames(plant_sp_candidates_full)[3] <- "Species"

#how many species/accessions were planned to be tested?
unique(plant_sp_candidates_full$Species) #53

#from those, how many were actually tested?
plant_sp_candidates_full <- plant_sp_candidates_full %>% drop_na(Germinated) #47 

#from those, how many did not germinate in MiCRop soil?
plant_sp_candidates_full %>% dplyr::count(Germinated) #13, leaving 34 to choose from 
plant_sp_candidates_full <- plant_sp_candidates_full %>% filter(Germinated=="yes")

#how fast were these 34 species growing?
plant_sp_candidates_full %>% dplyr::count(Speed) #more or less distributed equally between F,M and S*


# *See 2.DF with species speed to check how speed/phenology was calculated

## Germination information

#Then, the 34 "semi-finalist" species (selection df) were selected based on (1)speed of growth,
#(2)Lineage-Tribe distribution, (3)Germination success and how easy it was to Dip(4)
#we want to document this in a single plot, for this I need to first clean the df 

selection_experiment_full %>% dplyr::count(Family) #Let's see how many per Family we had in the experiment

#Add sp_abb column to make it easier to select
#arrange names
selection_experiment <- selection_experiment_full %>%
                    filter(Family=="Brassicaceae")%>% #keep only the Brass family
                    dplyr::rename(Sp_full_name=Species)%>%#to merge later
                    filter(!(Sp_full_name == "Brassica oleraceae" & `Growth speed` %in% c("Fast", "Slow")))
                           
selection_experiment$Sp_full_name <- gsub(" ", "_", selection_experiment$Sp_full_name)

#correctnames
library(plyr)
selection_experiment$Sp_full_name
selection_experiment <- transform(selection_experiment, Sp_full_name=revalue(Sp_full_name,
                                  c("Euclidium_syriacum"="Euclydium_syriacum",
                                    "Brassica_oleraceae"="Brassica_oleraceae_Rivera_Medium",
                                    "Camelina_sativa_crop"="Camelina_sativa", #we used the crop Cas
                                    "Camelina_sativa"="Camelina_sativa_wild",
                                    "Arabis_glabra"="Turritis_glabra",
                                    "Descurainia_sophia"="Descurainia_sophiades")))
#merge df's
selection_experiment <- left_join(selection_experiment, key_for_spp, by="Sp_full_name") %>%
                        relocate(Sp_full_name, Sp_abb_name, MY_tree_presence)%>%
                        dplyr::filter(MY_tree_presence=="inside_MY_tree")%>%
                        mutate(as.factor(Sp_abb_name))

#quick check
table(selection_experiment$Sp_abb_name)
table(key_for_spp$Sp_abb_name) #ok!

#make it into factors/numeric
selection_experiment <- selection_experiment %>% 
                        select(-c(Total, 'Per speed'))%>%
                        dplyr::relocate(Germination, .after=last_col())

selection_experiment[,1:15] <- lapply(selection_experiment[,1:15], factor)
selection_experiment$Germination <- as.numeric(as.character(selection_experiment$Germination))

str(selection_experiment)

selection_criteria <- selection_experiment #change name so its more informative

#clean
#rm(selection_experiment)

```

## 3. Growth rate 


```{r}

#In here there's all the data for plant growth (number of true leaves vs days after sowing)
plantgrowth <- growth_speed

#let's select only those 22 species of interest
unique(plantgrowth$species) #43 species

#filtering
#Keeping only those that were fertilized with hoagland==1 and controls (so, not treated with MeJA as a pilot)
plantgrowth_selection <- plantgrowth %>% 
                         dplyr::rename(Sp_full_name = species)%>%
                         #filter(!hoagland == 3 & !hoagland == 2 & !hoagland == 0)%>% #removes test with Hoagland
                         filter(treated == "control") %>%
                         filter(!lineage == "Cucurbitaceae" & !lineage =="Fabaceae" 
                 & !lineage == "Poaceae" & !lineage =="Solanaceae")%>%
                         select(-c(concentration, dipped, tribe, lineage, hoagland, treated))%>% #not needed
                         droplevels()

#numbers are characters, so have to change it to number
plantgrowth_selection[,c(4:39)] <-lapply(plantgrowth_selection[,c(4:39)], as.numeric)
table(plantgrowth_selection$Sp_full_name, plantgrowth_selection$round) 

#separating X_DAS to a column with number and another column with value of leaves (making it long format)
long <-plantgrowth_selection %>%
    pivot_longer(
    cols = ends_with("das"), 
    names_to = "DAS",
    values_to = "number_of_true_leaves",
    values_drop_na = TRUE) #removes NA's

long$DAS <- gsub("X", "", long$DAS) #remove X from DAS
long$DAS <- gsub("_das", "", long$DAS) #remove _das from DAS
long[,c(4:5)] <-lapply(long[,c(4:5)], as.numeric)
long[,c(1:3)]<-lapply(long[,c(1:3)], as.factor)

#correct species name
long <- transform(long, Sp_full_name=revalue(Sp_full_name,
                                  c("Euclidium_syriacum"="Euclydium_syriacum",
                                    "Brassica_oleracea_rivera"="Brassica_oleraceae_Rivera_Medium",
                                    "Camelina_sativa_crop"="Camelina_sativa", #we used the crop Cas
                                    "Camelina_sativa"="Camelina_sativa_wild",
                                    "Arabis_glabra"="Turritis_glabra")))


#now, select only those present in the MY tree
#Before that, let's fix the name in the key for spp from Descurainia sophiades to Descurainia sophia 
key_for_spp <- key_for_spp %>%
  mutate(Sp_full_name = if_else(Sp_full_name == "Descurainia_sophiades", 
                                "Descurainia_sophia", 
                                Sp_full_name))

#merge df's
long_selection <- left_join(long, key_for_spp, by="Sp_full_name") %>%
                        relocate(Sp_full_name, Sp_abb_name, MY_tree_presence)%>%
                        dplyr::filter(MY_tree_presence=="inside_MY_tree")%>%
                        mutate(as.factor(Sp_abb_name))%>%
                        select(-c(Sp_Tribe, MY_tree_presence))

#quick check
table(long_selection$Sp_abb_name)
unique(long_selection$Sp_abb_name)
table(key_for_spp$Sp_abb_name) #ok!
table(long_selection$round)

#I want to know if some species were repeated in more rounds
table(long_selection$Sp_abb_name, long_selection$round) #some are repeated, let's select 1 round

#filter more
long_selection2 <- long_selection %>% 
                   filter(!(round == "2" & Sp_abb_name == "Br"))%>%
                   filter(!(round == "8" & Sp_abb_name == "Sar"))%>%
                   filter(!(round == "8" & Sp_abb_name == "Tg"))%>%
                   filter(!(round == "2" & Sp_abb_name == "Bi"))%>%
                   filter(!(round %in% c("1", "2", "3", "5", "7", "8") & Sp_abb_name == "Bo_M"))
                  
                   
table(long_selection2$Sp_abb_name, long_selection2$round)


detach(package:plyr) #need to unload plyr, otherwise dplyr::summarise doesn't work
#Now, let's make a summary table to plot it

long_summary <- long_selection2 %>% 
                dplyr::group_by(Sp_abb_name, ID)%>% #I cluster your data per species
                filter(number_of_true_leaves %in% 2.5:5.5) %>% #Retain only the observations with 3 to 5 leaves 
                summarise(Average_days = mean(DAS),
                          SD = sd(DAS),
                          Min_days = min(DAS),
                          Max_days = max(DAS))

#Now get the mean of summary functions by species 
growth_perSp <- long_summary%>% 
                      select(-ID)%>% 
                      dplyr::group_by(Sp_abb_name)%>%
                      summarise_all(mean, na.rm = TRUE)%>%
                      as.data.frame()

str(growth_perSp)#ok! 

```

#Plot

```{r}

#Finally, make a single df and plot it
df_sp_names_complete
growth_perSp
selection_criteria

#let's remove those columns that are repeated from selection_criteria
selection_criteria2 <- selection_criteria %>% 
                      select(-c("MY_tree_presence", "Sp_Tribe", "Speed", "as.factor(Sp_abb_name)","Growth speed"))

#join df's
df_plant_sp_full <- left_join(df_sp_names_complete, growth_perSp, by="Sp_abb_name")
df_plant_sp_full <- left_join(df_plant_sp_full, selection_criteria2, by="Sp_abb_name")

str(df_plant_sp_full) #check

#move columns around
df_plant_sp_full <- df_plant_sp_full %>% 
                     relocate(Germination, .after=Species_old_name_correct)%>%
                     relocate(c(Sp_Tribe,MY_tree_presence,Species_old_name_correct),
                              .after=last_col()) %>% 
                     relocate(c(Family,Lineage, Tribe,Sample,`Sampling comment`,
                                `Taxonomic comment`),
                              .after=last_col())

df_plant_sp_full[,c(1:6)] <- lapply(df_plant_sp_full[,c(1:6)], as.factor)
df_plant_sp_full$Sp_abb_name <- factor(df_plant_sp_full$Sp_abb_name, levels=phylo_tips)

#done!
#save
#save it as Rdata and csv
saveRDS(df_plant_sp_full, "./16_Plant_species_information/data/df_full_sp_info_clean.rds")
write.csv(df_plant_sp_full, "./16_Plant_species_information/data/df_full_sp_info_clean.csv", row.names=FALSE)

```


```{r}

#1. Germination rates 
germination <- ggplot(df_plant_sp_full, aes(x=Sp_abb_name, y=Germination*100))+
        geom_bar(stat='identity', fill="black")+
        axis_looks+
        ylab("Germination %")+
        theme(aspect.ratio=1/4)+
        theme(axis.title.x = element_blank(), axis.text.x = element_blank())
        #scale_x_discrete(limits = rev(levels(df_plant_sp_full$Sp_abb_name)))+ #reversing order
        #coord_flip()

#2. Growth speed
growthspeed <- ggplot(df_plant_sp_full, aes(x=Sp_abb_name, y = Average_days, colour=Speed)) +
  geom_point(size=3) + 
  geom_linerange(aes(ymin = Average_days - SD, ymax = Average_days + SD, colour=Speed), size=1)+
  scale_color_manual(values=c("chartreuse4", "gold2", "red4"))+
  ylab("Average days to 3:5 leaves stage")+
  axis_looks+
  #scale_x_discrete(limits = rev(levels(df_plant_sp_full$Sp_abb_name)))+ #reversing order
  #coord_flip()+
  geom_hline(yintercept = 15, linetype ="dashed", size = 0.8, color = "chartreuse4")+
  geom_hline(yintercept = 20, linetype ="dashed", size = 0.8, color = "gold2")+
  geom_hline(yintercept = 23, linetype ="dashed", size = 0.8, color = "red4")+
  theme(aspect.ratio=1/3)+
  theme(axis.title.x = element_blank(), axis.text.x = element_blank())

#3. Wild vs Cultivated
temp <- df_plant_sp_full %>%  mutate(category="1")
colnames(temp)


wild_cultivated <- ggplot(temp, aes(x=Sp_abb_name, y = category, fill=`Wild/ Cultivated`)) +
  geom_point(size=5, shape=21) + 
  #geom_linerange(aes(ymin = Average_days - SD, ymax = Average_days + SD, colour=Speed), size=1)+
  scale_fill_manual(values=c("black", "white"))+
  ylab("Wild/Cultivated")+
  axis_looks+
  theme(aspect.ratio=1/6)+
  theme(axis.title.x = element_blank(), axis.text.x = element_blank())
  #scale_x_discrete(limits = rev(levels(df_plant_sp_full$Sp_abb_name)))+ #reversing order
  #coord_flip()

#---- 4. Tribe [Not used] ----
library(RColorBrewer)
display.brewer.all()

colourCount<-length(unique(temp$Tribe_new_correct))
getPalette <- colorRampPalette(brewer.pal(16, "Dark2"))

tribe <- ggplot(temp, aes(x=Sp_abb_name, y = category, fill=Tribe_new_correct)) +
  geom_point(size=5, shape=21) + 
  #geom_linerange(aes(ymin = Average_days - SD, ymax = Average_days + SD, colour=Speed), size=1)+
  scale_fill_manual(values=getPalette(colourCount))+
  ylab("Tribe")+
  axis_looks+
  theme(aspect.ratio=1/6)

#not super clear, many colors

#4.1 Lineage [Not used]
colourCount<-length(unique(temp$Lineage))
getPalette <- colorRampPalette(brewer.pal(4, "Set1"))

lineage <- ggplot(temp, aes(x=Sp_abb_name, y = category, fill=Lineage)) +
  geom_point(size=5, shape=21) + 
  #geom_linerange(aes(ymin = Average_days - SD, ymax = Average_days + SD, colour=Speed), size=1)+
  scale_fill_manual(values=getPalette(colourCount))+
  ylab("Lineage")+
  axis_looks+
  theme(aspect.ratio=1/6)

#----

#5. SuperTribe
unique(temp[,c('SuperTribe','Color_HEX')])

colors <- c("Camelinodae_(I)"="#de70a8", #same colors as BrassiTol
            "Arabodae_(IV)"="#9474e1",
            "Hesperodae_(III)"="#5ab2eb",
            "Brassicodae_(II)"="#10a072",
            "Unplaced"="#e99c00",
            "Heliophilodae_(V)"="#e05a00")

Supertribe <- ggplot(temp, aes(x=Sp_abb_name, y = category, fill=SuperTribe)) +
  geom_point(size=5, shape=21) + 
  #geom_linerange(aes(ymin = Average_days - SD, ymax = Average_days + SD, colour=Speed), size=1)+
  scale_fill_manual(values=colors)+
  ylab("Super Tribe")+
  axis_looks+
  theme(aspect.ratio=1/6)


# Combine the plots with patchwork
library(patchwork)

#only with Supetribe
wild_cultivated/growthspeed/germination/Supertribe 

ggsave("plot_full_species_info_noPhylo.pdf",
       path = "./16_Plant_species_information/results/",
        width = 2200, height = 3000 , units = "px", dpi=300, scale=0.9) #Supplementary Fig X

```

#Table

Finally, let's make a table with the right information
```{r}

colnames(df_plant_sp_full)
df_plant_sp_full 

#I need another column: "Data_from/Reference:" 
#We need a table with species abb code, species name (correct), tribe and lineage + sample code
table <- df_plant_sp_full %>% 
                    select(Sp_abb_name, Species_new_name_correct,
                           Lineage, Tribe_new_correct, Sample, `Sampling comment`)

table <- table %>% 
          mutate(Data_from=
                   case_when(
                     grepl("Nikolov", `Sampling comment`) ~ "Nikolov et al., (2019)",
                     grepl("Elena", `Sampling comment`) ~ "Castillo-Lorenzo, unpublished data",
                     TRUE ~ "Hendriks et al., (2023)"))

table$Species_new_name_correct <- gsub("_", " ", table$Species_new_name_correct)

library(plyr)
table <- transform(table, Lineage=revalue(Lineage,
                                  c("Expanded Lineage 2"="Exp. Lineage II",
                                    "Lineage 1"="Lineage I",
                                    "Lineage 2"="Lineage II",
                                    "Lineage 3"="Lineage III")))

table <- table %>%  select(-Sampling.comment)
colnames(table) <- c("Species code", "Species name", "Lineage", "Tribe", "NCBI SRA", "Data from:" )

library(rempsyc)
table2 <- nice_table(table,
              title = c("Phylogenetic information of Brassicaceae plant species"),
              note = "NCBI SRA is for now a tentative code")

flextable::save_as_docx(table2, path = "./16_Plant_species_information/results/Table_full_species_info.docx")

```


