---
title: "15_Plant_Biomass_cohen's_D"
author: "Zulema Carracedo Lorenzo"
date: "August 25, 2021"
output: html_document
editor_options: 
  chunk_output_type: console
---
# This script is based in the script 10_plant_traits from the MeJA pilot and written by Marcela. 

#During the experiment, shoot plant biomass was measured to characterize the plant's phenotypic response to the different stress treatments. As the species of the experiment are phenotypically different *per se* we compared the effects of the treatments between them using Cohen's D 

# 15.0 set up working environemnt

## 15.01 load libraries, improve memory use

```{r results="hide"}
# this will load all key libraries for the whole project
#source("./renv/0_load_libraries.R")

# packages needed for Cohen's D
library(effsize)    
library(lsr)

#Load the metadata
metadata <- read.csv("./Data/mapping_file_generation/Plant_phenotype_metadata.csv")

# increases memory limit used by R
memory.limit(size = 350000)
```


## Settings for plots
```{r}
# Setting up white background for plots

theme_set(theme_bw())

# Set colors per stress

color_stress <- c("#E69F00", "#009E73") #"MeJA", "SA"


# Set axis looks

axis_looks <- theme(axis.text.x = element_text(colour = "black", size = 9,
                                               
                                               face = "bold", angle=0, hjust=0.5))+
  
  theme(axis.text.y = element_text(colour = "black", size = 9, face = "bold"))+
  
  theme(axis.title=element_text(size=11, face = "bold"))+
  
  theme(axis.title.y = element_text(margin = ggplot2::margin(t = 0, r = 10, b = 0, l = 0)))+ #specify ggplot2 here to avoid "margin" function being overridden by a synonymous function from randomForest
  
  theme(panel.grid.major = element_blank(),
        
        panel.grid.minor = element_blank())+
  
  theme(legend.position="none")+
  
  theme(plot.background = element_blank(),
        
        panel.grid.major = element_blank(),
        
        panel.grid.minor = element_blank(),
        
        panel.border = element_blank())+
  
  theme(axis.line = element_line(colour = "black", 
                                 
                                 size = 0.8, linetype = "solid"))


#stress_names <- c("control" = "Control", "MeJA_0.1" = "0.1 mM \n MeJA",

#"MeJA_1.0" = "1 mM \n MeJA", "oral_secretion"="P. brassicae \n OS")



```
#Add species name column to data frame

```{r}
#Our metadata is missing the species names, it only contains species speed information. For our plots we need to have full names. Horrible hard code but didn't want to spend time figuring it out how to this in a more automated way...

metadata <- mutate(metadata, Species =
                     case_when(Sp_speed == "S1" ~ "Brassica_oleraceae Rivera_Slow",
                               Sp_speed == "S2" ~ "Solanum_lycopersicum",
                               Sp_speed == "S3" ~ "Nicotiana_benthamiana",
                               Sp_speed == "S4" ~ "Cucumis_sativus",
                               Sp_speed == "S5" ~ "Hirschfeldia_incana",
                               Sp_speed == "S6" ~ "Iberis_amara",
                               Sp_speed == "S7" ~ "Cochlearia_officinalis",
                               Sp_speed == "S8" ~ "Isatis_tinctoria",
                               Sp_speed == "S9" ~ "Brassica_oleraceae_Kimmeridge",
                               Sp_speed == "S10" ~ "Sinapis_arvensis",
                               Sp_speed == "S11" ~ "Diplotaxis_tenuifolia",
                               Sp_speed == "M1" ~ "Brassica_oleraceae_Rivera",
                               Sp_speed == "M2" ~ "Medicago_trunculata",
                               Sp_speed == "M3" ~ "Brachypodium",
                               Sp_speed == "M4" ~ "Arabidopsis_lyrata",
                               Sp_speed == "M5" ~ "Descurainia_sophiades",
                               Sp_speed == "M6" ~ "Camelina_sativa",
                               Sp_speed == "M7" ~ "Turritis_glabra",
                               Sp_speed == "M8" ~ "Brassica_nigra",
                               Sp_speed == "M9" ~ "Brassica_rapa",
                               Sp_speed == "M10" ~ "Sinapis_alba",
                               Sp_speed == "M11" ~ "Sysimbrium_irio",
                               Sp_speed == "F1" ~ "Brassica_oleraceae_Rivera_Fast",
                               Sp_speed == "F2" ~ "Pisum_sativum",
                               Sp_speed == "F3" ~ "Zea_mays",
                               Sp_speed == "F4" ~ "Arabidopsis_thaliana",
                               Sp_speed == "F5" ~ "Erysium_cheirantoides",
                               Sp_speed == "F6" ~ "Barbarea_vulgaris",
                               Sp_speed == "F7" ~ "Malcomia_maritima",
                               Sp_speed == "F8" ~ "Lepidium_ruderale",
                               Sp_speed == "F9" ~ "Lobularia_maritima",
                               Sp_speed == "F10" ~ "Berteroa_incana",
                               Sp_speed == "F11" ~ "Euclydium_syriacum",
                     ))

#We make a vector with the species names in the same order than in the phylogeny tree. In this way, our plots will have the species list organized like in the phylogeny. 

#this was the old order
phylogeny_vector <- c("Arabidopsis_lyrata", "Arabidopsis_thaliana", "Camelina_sativa", "Turritis_glabra","Erysium_cheirantoides","Malcomia_maritima","Descurainia_sophiades","Lepidium_ruderale", "Barbarea_vulgaris",  "Berteroa_incana", "Euclydium_syriacum", "Brassica_rapa",  "Brassica_oleraceae_Rivera","Hirschfeldia_incana","Sinapis_arvensis","Diplotaxis_tenuifolia","Brassica_nigra", "Isatis_tinctoria", "Sysimbrium_irio","Cochlearia_officinalis", "Lobularia_maritima", "Iberis_amara") #needed for the plots


#Be aware that not all the species are in the phylogeny (for instance, species from other families). We will have to removed them at some point. Let's make a new metadata table containing only the species from our phylogeny. It will be useful for our plots.

metadata_phylogeny <- subset(metadata, Species %in% phylogeny_vector)
```



## 15.1 Cohen's D

```{r}
 #making costume function using lsr package to calculate Cohen's D
  
cohensD_per_trait <- function(df, column, Sp_speed){
                                    
                                    
                                    
                     #df= data frame
                     #column = column name with trait of interest, no quotes 
                      #species = species name to be put in df, no quotes 
                                    
                       trait_name <- deparse(substitute(column)) #gets actual name (" ") of the column to be substituted
                                    
                       species_name <- deparse(substitute(Sp_speed)) #idem for species
                       
                       Control<- df %>%  #gets only values for control and the trait 
                                      
                       dplyr::filter(Stress == "Control") %>%  pull(trait_name) 
                                    
                                    
                                    
                                    MeJA <- df %>% 
                                      
                                      dplyr::filter(Stress == "MeJA") %>% pull(trait_name)
                                    
                                    
                                    
                                    SA <- df %>% 
                                      
                                      dplyr::filter(Stress == "SA") %>% pull(trait_name)
                                    
                          
                                    
                                    d_MeJA <- cohen.d(MeJA, Control) #calculates Cohen's D, Ctrl should be 2nd position
                                    
                                    d_SA <- cohen.d(SA, Control)   #so estimate matches the effect: 
                                    
                                              #(smaller or larger than ctrl)
                                    
                                    
                                    
                                    a <- as.data.frame(t(unlist(d_MeJA))) %>% #makes new df with values from Cohen's D estimation
                                      
                                      mutate(Stress="MeJA")           #and adds a new column with the treatment 
                                    
                                    
                                    
                                    b <- as.data.frame(t(unlist(d_SA))) %>% 
                                      
                                      mutate(Stress="SA")  
                                    
                                    
                                     df.cohensD <- rbind(a,b)%>%             #merge 2 df's
                                      
                                      mutate(trait=trait_name)%>% #makes a new column with the trait name 
                                      
                                      mutate(Sp_speed=species_name)%>% #idem for species 
                                      
                                      mutate(eff_size=                #Adds a new column replacing numeric "magnitude" 
                                               
                                               case_when(             #values (1-4) coming from cohen.d function 
                                                 
                                                 magnitude == 1 ~ "negligible", #with the category written in text 
                                                 
                                                 magnitude == 2 ~ "small",
                                                 
                                                 magnitude == 3 ~ "medium",
                                                 
                                                 magnitude == 4 ~ "large"))%>%
                                      
                                      relocate(trait, Sp_speed, Stress, eff_size)
                                    
                                    
                                    
                                    
                                    
                                    return(df.cohensD) #returns full df 
                                    
                                  }  
                                  
                                  
                                  
#calculate cohen's D for shoot dry weight

metadata <- drop_na(metadata) #we dont have shoot biomass for all plants, drop rows with no biomass value
                        
cohen_d_F1 <- cohensD_per_trait(filter(metadata, Sp_speed == "F1"), Shoot_.DryWeight, F1)
cohen_d_F2 <- cohensD_per_trait(filter(metadata, Sp_speed == "F2"), Shoot_.DryWeight, F2)
cohen_d_F3 <- cohensD_per_trait(filter(metadata, Sp_speed == "F3"), Shoot_.DryWeight, F3)
cohen_d_F4 <- cohensD_per_trait(filter(metadata, Sp_speed == "F4"), Shoot_.DryWeight, F4)
cohen_d_F5 <- cohensD_per_trait(filter(metadata, Sp_speed == "F5"), Shoot_.DryWeight, F5)
cohen_d_F6 <- cohensD_per_trait(filter(metadata, Sp_speed == "F6"), Shoot_.DryWeight, F6)
cohen_d_F7 <- cohensD_per_trait(filter(metadata, Sp_speed == "F7"), Shoot_.DryWeight, F7)
cohen_d_F8 <- cohensD_per_trait(filter(metadata, Sp_speed == "F8"), Shoot_.DryWeight, F8)
cohen_d_F9 <- cohensD_per_trait(filter(metadata, Sp_speed == "F9"), Shoot_.DryWeight, F9)
cohen_d_F10 <- cohensD_per_trait(filter(metadata, Sp_speed == "F10"), Shoot_.DryWeight, F10)
cohen_d_F11 <- cohensD_per_trait(filter(metadata, Sp_speed == "F11"), Shoot_.DryWeight, F11)
cohen_d_M1 <- cohensD_per_trait(filter(metadata, Sp_speed == "M1"), Shoot_.DryWeight, M1)
cohen_d_M2 <- cohensD_per_trait(filter(metadata, Sp_speed == "M2"), Shoot_.DryWeight, M2)
cohen_d_M3 <- cohensD_per_trait(filter(metadata, Sp_speed == "M3"), Shoot_.DryWeight, M3)
cohen_d_M4 <- cohensD_per_trait(filter(metadata, Sp_speed == "M4"), Shoot_.DryWeight, M4)
cohen_d_M5 <- cohensD_per_trait(filter(metadata, Sp_speed == "M5"), Shoot_.DryWeight, M5)
cohen_d_M6 <- cohensD_per_trait(filter(metadata, Sp_speed == "M6"), Shoot_.DryWeight, M6)
cohen_d_M7 <- cohensD_per_trait(filter(metadata, Sp_speed == "M7"), Shoot_.DryWeight, M7)
cohen_d_M8 <- cohensD_per_trait(filter(metadata, Sp_speed == "M8"), Shoot_.DryWeight, M8)
cohen_d_M9 <- cohensD_per_trait(filter(metadata, Sp_speed == "M9"), Shoot_.DryWeight, M9)
cohen_d_M10 <- cohensD_per_trait(filter(metadata, Sp_speed == "M10"), Shoot_.DryWeight, M10)
cohen_d_M11 <- cohensD_per_trait(filter(metadata, Sp_speed == "M11"), Shoot_.DryWeight, M11)
cohen_d_S1 <- cohensD_per_trait(filter(metadata, Sp_speed == "S1"), Shoot_.DryWeight, S1)
cohen_d_S2 <- cohensD_per_trait(filter(metadata, Sp_speed == "S2"), Shoot_.DryWeight, S2)
cohen_d_S3 <- cohensD_per_trait(filter(metadata, Sp_speed == "S3"), Shoot_.DryWeight, S3)
cohen_d_S4 <- cohensD_per_trait(filter(metadata, Sp_speed == "S4"), Shoot_.DryWeight, S4)
cohen_d_S5 <- cohensD_per_trait(filter(metadata, Sp_speed == "S5"), Shoot_.DryWeight, S5)
cohen_d_S6 <- cohensD_per_trait(filter(metadata, Sp_speed == "S6"), Shoot_.DryWeight, S6)
cohen_d_S7 <- cohensD_per_trait(filter(metadata, Sp_speed == "S7"), Shoot_.DryWeight, S7)
cohen_d_S8 <- cohensD_per_trait(filter(metadata, Sp_speed == "S8"), Shoot_.DryWeight, S8)
cohen_d_S9 <- cohensD_per_trait(filter(metadata, Sp_speed == "S9"), Shoot_.DryWeight, S9)
cohen_d_S10 <- cohensD_per_trait(filter(metadata, Sp_speed == "S10"), Shoot_.DryWeight, S10)
cohen_d_S11 <- cohensD_per_trait(filter(metadata, Sp_speed == "S11"), Shoot_.DryWeight, S11)

all_cohen_df <- rbind(cohen_d_F1, cohen_d_F2, cohen_d_F3, cohen_d_F4, cohen_d_F5, cohen_d_F6, cohen_d_F7, cohen_d_F8, cohen_d_F9, cohen_d_F10, cohen_d_F11, cohen_d_M1, cohen_d_M2, cohen_d_M3, cohen_d_M4, cohen_d_M5, cohen_d_M6, cohen_d_M7, cohen_d_M8, cohen_d_M9, cohen_d_M10, cohen_d_M11, cohen_d_S1, cohen_d_S2, cohen_d_S3, cohen_d_S4, cohen_d_S5, cohen_d_S6, cohen_d_S7, cohen_d_S8, cohen_d_S9, cohen_d_S10, cohen_d_S11)

#Add full species name
species_naming_df <- metadata %>% dplyr::select(Sp_speed, Species) %>% unique() #creates df with sp_speed and species name information

all_cohen_df <- merge(all_cohen_df, species_naming_df, by = "Sp_speed") #adds species names to cohen d results

#save Cohen's D results
save(all_cohen_df,  
     file = "./15_Plant_Biomass/Objects/shoot_biomass_CohenD.RData")


#estimate and interval need to be numeric to be use in plot

all_cohen_df <- transform(all_cohen_df, estimate = as.numeric(estimate))
all_cohen_df <- transform(all_cohen_df, conf.int.lower = as.numeric(conf.int.lower))
all_cohen_df <- transform(all_cohen_df, conf.int.upper = as.numeric(conf.int.upper))
#Select only the species from the phylogeny

phylogeny_cohen_df <-  subset(all_cohen_df, Species %in% phylogeny_vector)

phylogeny_cohen_df <- phylogeny_cohen_df %>%
    arrange(match(Species, phylogeny_vector)) %>%
    arrange(match(Stress, c("MeJA", "SA")))

#write.csv(phylogeny_cohen_df, "./15_Plant_Biomass/Results/shoot_biomass_CohenD_tree.csv")



```



## 15.2 Plot (Supplementary Figure S3)

```{r}
phylogeny_cohen_df

phylogeny_cohen_df$Species <- gsub("_", " ", phylogeny_cohen_df$Species) #fixes the species names


 phylogeny_cohen_df <- phylogeny_cohen_df %>%  mutate(Sp_full_name =
                                                     case_when(Species == "Brassica oleraceae Rivera" ~ "Brassica oleracea",
                                                     TRUE ~ Species)) #Renaming B. oleracea


phylogeny_vector <- c("Iberis amara", "Lobularia maritima", "Cochlearia officinalis", "Sysimbrium irio", "Isatis tinctoria", "Brassica nigra", "Diplotaxis tenuifolia", "Sinapis arvensis", "Hirschfeldia incana", "Brassica oleracea", "Brassica rapa", "Euclydium syriacum", "Berteroa incana", "Barbarea vulgaris", "Lepidium ruderale", "Descurainia sophiades",  "Malcomia maritima", "Erysium cheirantoides", "Turritis glabra", "Camelina sativa","Arabidopsis thaliana", "Arabidopsis lyrata")

phylogeny_cohen_df$Sp_full_name <- factor(phylogeny_cohen_df$Sp_full_name, levels = phylogeny_vector)#arrange order for ggplot

names(phylogeny_cohen_df)[names(phylogeny_cohen_df) == "Stress"] <- "Treatment" #change stress for treatment

phylogeny_cohen_df <- phylogeny_cohen_df %>%
  dplyr::mutate(
    estimate = as.numeric(estimate),
    conf.int.lower = as.numeric(conf.int.lower),
    conf.int.upper = as.numeric(conf.int.upper)
  )


cohen_plot_phylogeny <- ggplot(phylogeny_cohen_df, aes(x=Sp_full_name, y=estimate, ymin=conf.int.lower, ymax=conf.int.upper,col=Treatment,fill=Treatment)) + 
   geom_linerange(size=5, position=position_dodge(width = 0.5)) +
                                    
   geom_hline(aes(yintercept=0), lty=1) + 
                                    
   geom_hline(aes(yintercept=0.8), lty=2, colour="red") +
                                    
   geom_hline(aes(yintercept=-0.8), lty=2, colour="red") +
                                    
   geom_point(size=3, shape=20, colour="black", stroke = 0.5,
                                               
    position=position_dodge(width = 0.5)) +
  geom_point(size=3, shape=21, colour="white", stroke = 0.5,position=position_dodge(width = 0.5)) +
  scale_fill_manual(values=color_stress)+
  scale_color_manual(values=color_stress)+
 scale_x_discrete(name="Plant species") +
  scale_y_continuous(name="Cohen's D estimate with 95% confidence interval", limits = c(-4,4))+
  coord_flip() +
  #theme_minimal() +
  #theme(axis.text.y = element_text(colour = "black", size = 12,
  #face = "bold", angle=0, hjust=0.5))+
  theme() +
  theme(legend.position="top", text = element_text(size = 18), axis.text.y=element_text(hjust=1, face=c(rep("italic", 25), "plain"))) +
  facet_wrap(~ Treatment)

ggsave(cohen_plot_phylogeny,
       height = 10, width = 15, dpi=500,
       file="./15_Plant_Biomass/Results/Shoot_biomass_CohenD_phylogeny_paper_Oct24.pdf")

```




