---
title: "MU_07_consolidate_NMandRFandDA_ForGoldenTaxonomies"
author: "Melissa"
date: "2023-04-11"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---
## 5_1.0 load libraries, improve memory use
```{r results="hide"}
# this will load all key libraries for the whole project
source("./renv/0_load_libraries.R")
```

The objective of this script is to determine Golden ASVs and taxonomies as those most consistently responsive to the treatments by being selected by three different methodologies when comparing the Control microbial communities with those of each of the treatments. 

#DA: Load and prepare (22 sp)
```{r}
#Load ASVs detected by two DA tests
load("./05_Differential_abundance/Objects/MeJA_Bac_conca_two_f.RData")
load("./05_Differential_abundance/Objects/SA_Bac_conca_two_f.RData")
load("./05_Differential_abundance/Objects/MeJA_Fun_conca_two_f.RData")
load("./05_Differential_abundance/Objects/SA_Fun_conca_two_f.RData")

# define species present in the MY tree
phylo_tips <- c("At", "Al", "Cas", "Tg", "Ec", "Mm", "Bv",
              "Lr", "Ds", "Bi", "Bn", "Sar", "Hi", "Br", "Bo_M",
              "Dt", "Si", "It", "Co", "Lm", "Ia", "Es")


DA_MeJA_Bac_22sp <- unlist(MeJA_Bac_conca_two_f[names(MeJA_Bac_conca_two_f) %in%  phylo_tips], use.names = F)
DA_SA_Bac_22sp <- unlist(SA_Bac_conca_two_f[names(SA_Bac_conca_two_f) %in%  phylo_tips],use.names = F)
DA_MeJA_Fun_22sp <- unlist(MeJA_Fun_conca_two_f[names(MeJA_Fun_conca_two_f) %in%  phylo_tips], use.names = F)
DA_SA_Fun_22sp <- unlist(SA_Fun_conca_two_f[names(SA_Fun_conca_two_f) %in%  phylo_tips], use.names = F)

DA_ControlMeJA_ControlSA_ASVs_22sp <- list(DA_MeJA_Bac_22sp = DA_MeJA_Bac_22sp, DA_SA_Bac_22sp = DA_SA_Bac_22sp, DA_MeJA_Fun_22sp = DA_MeJA_Fun_22sp, DA_SA_Fun_22sp = DA_SA_Fun_22sp)
```


```{r}
save(DA_ControlMeJA_ControlSA_ASVs_22sp, 
     file = "./07_consolidate_all_analysis/Results/DA_ControlMeJA_ControlSA_ASVs_22sp.RData")
```

#RF: Load and prepare
```{r}
#Load list of all detected ASVs
load("./10_Machine_learning/r_objects/RF_ControlMeJA_ControlSA_ASVs_22sp.RData")



```

#NM: Load and prepare
AGREED WITH MELISSA THAT THIS VERSION IS BETTER
this lsit does not include ASVs that are above neutral in both treatment and control of an individual species. it also removes any asvs that is below in both treatment and control of the same individual species. essentially, it removes some NM ASVs that do not help differentaiting between controls and treatments
```{r}
load (file = "./04_Neutral_models/r_objects/differing_ASVs_above_below_l_22sp.RData")

NM_ControlMeJA_ControlSA_ASVs_22sp<- differing_ASVs_above_below_l_22sp
```


#Make sure all ASVs and taxonomies are unique inside their own lists (to prevent having one list with an ASV more than once)
```{r}
#ASVs
DA_ControlMeJA_ControlSA_ASVs_22sp_unique <- lapply(DA_ControlMeJA_ControlSA_ASVs_22sp, function(x) unique(x))
RF_ControlMeJA_ControlSA_ASVs_22sp_unique <- lapply(RF_ControlMeJA_ControlSA_ASVs_22sp, function(x) unique(x))
NM_ControlMeJA_ControlSA_ASVs_22sp_unique <- lapply(NM_ControlMeJA_ControlSA_ASVs_22sp, function(x) unique(x))


```


#Three methods overlap in ASVs: GOLDEN ASVs
```{r}
DA_RF_NM_ASVs_intersect <- mapply(function(x,y,z)
Reduce(intersect, list(x,y,z)),
x = DA_ControlMeJA_ControlSA_ASVs_22sp_unique,
y = RF_ControlMeJA_ControlSA_ASVs_22sp_unique,
z = NM_ControlMeJA_ControlSA_ASVs_22sp_unique,
  SIMPLIFY = FALSE)


# overwrrite fungi, keeping both golden and "silver" ASVs
DA_RF_NM_ASVs_intersect[3:4] <-
mapply(function(x,y,z){
      
    DA_RF<-Reduce(intersect, list(x,y))
    DA_NM<-Reduce(intersect, list(x,z))
    RF_NM<-Reduce(intersect, list(y,z))
    
    output<-c(DA_RF, DA_NM, RF_NM)%>%unique()
    
    return(output)
    },
    
x = DA_ControlMeJA_ControlSA_ASVs_22sp_unique[3:4],
y = RF_ControlMeJA_ControlSA_ASVs_22sp_unique[3:4],
z = NM_ControlMeJA_ControlSA_ASVs_22sp_unique[3:4],
  SIMPLIFY = FALSE)



names(DA_RF_NM_ASVs_intersect) <- c("MeJA_Bac_GoldenASVs", "SA_Bac_GoldenASVs", "MeJA_Fun_GoldenASVs", "SA_Fun_GoldenASVs")
```


```{r}
save(DA_RF_NM_ASVs_intersect, 
     file = "./07_consolidate_all_analysis/Results/DA_RF_NM_ASVs_intersect.RData")
```


```{r}
#VennDiagram
library("ggVennDiagram")
#Three methods overlap in taxonomies
Venn_DA_RF_NM_ASVs_intersect <- mapply(function(x,y,z)
ggVennDiagram(list(x,y,z), category.names = c("DA", "RF","NM"), edge_lty = "dashed", edge_size = 0.01, set_size = 5)
+ scale_fill_gradient(low = "cadetblue1", high = "cadetblue4"),
x = DA_ControlMeJA_ControlSA_ASVs_22sp_unique,
y = RF_ControlMeJA_ControlSA_ASVs_22sp_unique,
z = NM_ControlMeJA_ControlSA_ASVs_22sp_unique,
  SIMPLIFY = FALSE)

ggexport(plotlist = Venn_DA_RF_NM_ASVs_intersect, filename = "./07_consolidate_all_analysis/Results/Venn_DA_RF_NM_ASVs_intersect.pdf",
 ncol = 1, nrow = 1, height=10, width = 10)

```

