---
title: "Annotated phylotree for Golden ASVs"
author: "Marcela Aragon"
date: "6/11/2023"
output: html_document
---

This script leads to Figure 3 and Supplementary Figure S19. It creates the .txt files to draw the phylogenetic tree with the metadata in iTOL.

# Load libraries

```{r results="hide"}

#Set working directory to project directory 
getwd() #ok

#let's deactivate renv for this script
renv::deactivate()

#Load libraries

#microbiome data
library(phyloseq)
library(microbiome)
library(microbiomeutilities) #ternary plots
library(vegan)
library(Biostrings) #sequence data

#data management
library(tidyr)
library(dplyr)
library(tibble)
library(stringr)
library(purrr) #to handle lists
library(itol.toolkit)#handles data for iTOL use 
library(varhandle) #rm.all.but() function to clean

#plots
library(ggpubr)
library(ggplot2)
library(RColorBrewer) #nice colors for plots
library(phytools)
library(ape) #to handle phylogenetic trees 


```


# 16S

(Optional!!)
If you want to skip running the code and just have the results from line 57 until line 691 just upload this Rdata that contains the saved environment.

```{r}

load("./07_consolidate_all_analysis/Results/16S_R-environment.RData")

```

Otherwise, continue here: 

## Phyloseq

```{r}

#load Golden ASVs names
load("./07_consolidate_all_analysis/Results/DA_RF_NM_ASVs_intersect.RData")

#check
DA_RF_NM_ASVs_intersect #4 lists, one per treatment per 16S or ITS

#### 16S ####
MeJA_asvs <- DA_RF_NM_ASVs_intersect$MeJA_Bac_GoldenASVs #130
SA_asvs <- DA_RF_NM_ASVs_intersect$SA_Bac_GoldenASVs #138

#Now I need to get the unique names 
bac_sig_asvs <- c(MeJA_asvs, SA_asvs) #268
bac_sig_asvs <- unique(bac_sig_asvs) #234 unique ones, there are 34 shared ASVs 
bac_intersect <- intersect(MeJA_asvs, SA_asvs) #get the shared ones 

# this will load the CSS-normalized 16S and ITS data as a list of two objects
load("./Data/phyloseq_objects/unnormalized_BacFun_rhizoplane_HighFilt_ps_l.RData")
unnormalized_BacFun_rhizoplane_HighFilt_ps_l

#Now, I need to remove all those samples that belong to guest species + others not present in Kasper's tree
Spp_abb_name_Kasper <- c("Hi","Sal","Al","Tg","Bo_M","Dt","Ds","Br","Ia","Cas",
                         "Ec","Es","At","Mm"," Si"," Lr","Bi","Lm","Bv", "Co", "Bn","Sar", "It") 

ps_Bac <- subset_samples(unnormalized_BacFun_rhizoplane_HighFilt_ps_l$Bac,
                       Sp_abb_name %in% Spp_abb_name_Kasper) #only those plant sp present in the tree

#prune ps to keep only those that belong to the Golden ASVs
ps_golden <- prune_taxa(bac_sig_asvs, ps_Bac)

#check
ps_golden #ok, 234

#remove big ps object to clear up space, you'll load it later for fungi
rm(ps_Bac, unnormalized_BacFun_rhizoplane_HighFilt_ps_l)

```

### Make new bacterial phylogenetic tree
For now we skip this part as it has already been done, go directly to line 133.
```{r}

#get the sequences and save them as FASTA format
#ps_golden %>%
#refseq() %>%
#Biostrings::writeXStringSet("./07_consolidate_all_analysis/Results/Annotated_microbial_phylogenetic_tree/16S/16S_goldenASVs_sequences.fasta",
#             append=FALSE,compress=FALSE,
#             compression_level=NA, format="fasta")


#Go to MAFTT [...]
```

As phylogenetic tree from Qiime2 sometimes is not very accurate for the microbial phylogeny I will first export the sequences from the golden_asvs and then follow a similar approach to the `Phylogenetic_analysis_good.Rmd` (you can find it in "./Plant_phylogeny/Code/") but in the HPC Anunna. 

I did the alignment in MAFFT online (https://mafft.cbrc.jp/alignment/server/index.html) and then used that alignment in fasta format to run the following slurm script to run IQ-tree in the HPC which means that iqtree will run with a DNA sequence type, 10,000 replicates in bootstrap analysis and that finds the best substitution model automatically.

```{bash}

#!/bin/bash -login

#SBATCH --time=4:00:00
#SBATCH --ntasks=1
#SBATCH --cpus-per-task=8
#SBATCH --mem=4G
#SBATCH --job-name=16S_tree
#SBATCH --output=16S_tree.output


cd /lustre/nobackup/INDIVIDUAL/arago004/IQ_tree/16S ##place where I had the .fasta alignment

iqtree2 -s 16S_alignment_MAFFT.fasta -st DNA -m MFP -bb 10000 -nt AUTO ##

```

Once I got the tree (it took less than 1 hour) I imported back to R

```{r}
#Import tree
tree_16S <- read.tree("./07_consolidate_all_analysis/Results/Annotated_microbial_phylogenetic_tree/16S/16S_alignment_MAFFT.fasta.treefile")

#check
tree_16S
plot(tree_16S) #good 
tree_16S <- multi2di(tree_16S) #root it

#save it in .tre format to be imported in iTOL
#ape::write.tree(tree_16S, "./07_consolidate_all_analysis/Results/Annotated_microbial_phylogenetic_tree/16S_28-08-24/16S_tree.tre")

#change tree in ps object 
ps_golden@phy_tree <- tree_16S

```

## Taxonomy dataframe

```{r}
#make a new df with two columns; one for MeJA and another one for SA
df_bac_sig_asvs <- as.data.frame(bac_sig_asvs) 
colnames(df_bac_sig_asvs) <- "asv"                   
df_bac_sig_asvs <- df_bac_sig_asvs %>%
                   mutate(MeJA=case_when(asv %in% MeJA_asvs ~ "1",
                                         TRUE ~ "-1"))%>% #0 is for empty shape, -1 is no shape
                   mutate(SA=case_when(asv %in% SA_asvs ~ "1",
                                         TRUE ~ "-1"))

#let's get the taxonomy
#clean it first
tax_table(ps_golden)[, colnames(tax_table(ps_golden))] <- gsub(tax_table(ps_golden)[, colnames(tax_table(ps_golden))], pattern = "[a-z]__", replacement = "")

#put it in a df 
df_taxonomy <- as.data.frame(ps_golden@tax_table)
order_asvs <- rownames(df_taxonomy) #get order of rows to keep using it for rest of df's

#re-order sig_asvs df 
df_bac_sig_asvs <- df_bac_sig_asvs %>%
                   column_to_rownames(var="asv") #first make new rownames
#now order it 
df_bac_sig_asvs <- df_bac_sig_asvs[order(match(rownames(df_bac_sig_asvs),
                                   order_asvs)), , drop=FALSE] 

#now you can merge df's
df_complete <- cbind(df_bac_sig_asvs, df_taxonomy)
df_complete <- df_complete %>% select(-c(Species,Confidence)) #remove non-informative columns 

```

## Option 1: L2FC per Sp from Melissa [Not used for iTOL]

```{r}

#load df with log2-fold change
load("./07_consolidate_all_analysis/r_objects/MU_HeatmapsGoldenASVs/lfcres_GoldenASVs_FinalDF_l_replaceNA.RData")

#check
lfcres_GoldenASVs_FinalDF_l_replaceNA

#first, only MeJA
lfc_MeJA <- lfcres_GoldenASVs_FinalDF_l_replaceNA$lfcres_df_MeJA_Bac

#add MeJA suffix
lfc_MeJA <- lfc_MeJA %>% 
            mutate(across(everything(), ~ na_if(., 0.00)))%>% #first, change 0.0 to NA's
            rename_with(~ paste0(., "_MeJA"))

#let's make a column with the amount of times an ASV was significant
#(if 0.0 I asume it was NA and therefore not significant)
lfc_average <- lfc_MeJA %>% 
             mutate(LFC_average_MeJA = rowMeans(., na.rm=T))%>%  #get average
             select(LFC_average_MeJA)

lfc_ocurrences <- lfc_MeJA %>% 
             mutate(num_sp_sig_MeJA= rowSums(!is.na(.)))%>%
             mutate(num_sp_NA_MeJA=rowSums(is.na(.)))%>% 
             mutate(negative_lfc_MeJA=rowSums(lfc_MeJA < 0, na.rm=T))%>% 
             mutate(positive_lfc_MeJA=num_sp_sig_MeJA-negative_lfc_MeJA) 

#add the average value
lfc_average$ASV <- rownames(lfc_average)
lfc_ocurrences$ASV <- rownames(lfc_ocurrences)
lfc_MeJA <- left_join(lfc_ocurrences, lfc_average) 
lfc_MeJA <- lfc_MeJA %>% relocate(ASV)


#then, SA
lfc_SA <- lfcres_GoldenASVs_FinalDF_l_replaceNA$lfcres_df_SA_Bac

#add SA suffix
lfc_SA <- lfc_SA %>% 
            mutate(across(everything(), ~ na_if(., 0.00)))%>% #first, change 0.0 to NA's
            rename_with(~ paste0(., "_SA"))

#let's make a column with the amount of times an ASV was significant
#(if 0.0 I asume it was NA and therefore not significant)
lfc_average <- lfc_SA %>% 
             mutate(LFC_average_SA = rowMeans(., na.rm=T))%>%  #get average
             select(LFC_average_SA)

lfc_ocurrences <- lfc_SA %>% 
             mutate(num_sp_sig_SA= rowSums(!is.na(.)))%>%
             mutate(num_sp_NA_SA=rowSums(is.na(.)))%>% 
             mutate(negative_lfc_SA=rowSums(lfc_SA < 0, na.rm=T))%>% 
             mutate(positive_lfc_SA=num_sp_sig_SA-negative_lfc_SA) 

#add the average value
lfc_average$ASV <- rownames(lfc_average)
lfc_ocurrences$ASV <- rownames(lfc_ocurrences)
lfc_SA <- left_join(lfc_ocurrences, lfc_average) 
lfc_SA <- lfc_SA %>% relocate(ASV)

#Now, let's join the df's
lfc <- full_join(lfc_MeJA, lfc_SA)

#Let's make a new df first with MeJA information
df_complete #our dataframe with metadata
lfc #the df we want to add

#add column named ASV
df_complete$ASV <- rownames(df_complete) #make new column 'ASV'

#join df 
df_complete <- full_join(df_complete, lfc, by="ASV")

#order
df_complete <- df_complete %>% 
                select(-OTU) %>% 
                dplyr::relocate(ASV, MeJA, SA)

rownames(df_complete) <- df_complete$ASV


```

I think I stopped here because I didn't know how to integrate both values, perhaps an idea is to not include the lfc value if it was not significant, to include the average per species and add as metadata in how many species it was positive and negative as it becomes a problem to reduce it to a single value if we are talking about 22 different species. 

## Option 2: LFC from Ancombc [This one was used]

Let's try to use Ancom just to calculate the log-fold change per ASVs between Control and Stress-treatments. This will be an average value across species (I think this was the way Pedro plotted the heattrees but needs confirmation)

```{r}

#This code comes from the psf_microbiome project

ps_golden #this is my ps object with golden asvs for Bac

library(ANCOMBC)

#Just for Bac-GoldenASVs but using 3 groups
set.seed(123)
ancom <- ANCOMBC::ancombc(phyloseq = ps_golden, 
                      formula = "Stress", 
                      p_adj_method = "fdr", #same as "BH" Benjamini & Hochberg
                      #zero_cut = 1, #default setting
                      lib_cut = 0, 
                      prv_cut=0, #otherwise removes some ASVs
                      group = "Stress", 
                      struc_zero = TRUE, #with structural zero's
                      neg_lb = FALSE, #classifies structural zero's if it's zero and not its asymptotic lower bound 
                      tol = 1e-5,               
                      max_iter = 100, 
                      conserve = TRUE, 
                      alpha = 0.05,
                      global=TRUE)

head(ancom$res$lfc) #LFC

#Get LFC values
lfc_ancom <- ancom$res$lfc #log-fold change according to control --> not removing any sample

#### Now I have a single value of LFC per each golden ASV for both MeJA and SA vs Control in the lfc object ####
lfc_ancom

lfc_ancom <- lfc_ancom %>% 
             mutate(ASV=taxon)%>% 
             select(c(ASV, StressMeJA, StressSA))%>%
             rename_with(~ paste0(., "_ancombc"))
colnames(lfc_ancom)[1] <- "ASV"

#finally, join it with full df 
df_complete <- left_join(df_complete, lfc_ancom)

#save df
save(df_complete, file="./07_consolidate_all_analysis/Results/Annotated_microbial_phylogenetic_tree/16S/df_complete.Rdata")

```



## iTOL


### Metadata in iTOL format

To add annotations iTOL works in layers (see https://itol.embl.de/help.cgi#anno), therefore we'll need one dataframe in .txt format per type of annotation we would like to see. The package `itol.toolkit::` helps you create dataframes in iTOL format. 

First, arrange data and colors to be used 
```{r}

#load full df (this contains ANCOMBC results)
load(file="./07_consolidate_all_analysis/Results/Annotated_microbial_phylogenetic_tree/16S/df_complete.Rdata")
rownames(df_complete) <- df_complete$ASV

#First gather what you need 
df_complete #our dataframe with taxonomy and lfc data for both MeJA and SA
tree_16S #our tree

#Set-up colors

#### colors for Treatment ####
color_MeJA <- "#E69F00"
color_SA <- "#009E73"
color_treatment <- c("#E69F00","#009E73")

```


#### Treatment

MeJA (yellow dots)
```{r}
Treatment_MeJA <- df_complete %>%
                  dplyr::select(ASV,MeJA)%>%
                  mutate(id=ASV)%>% #make new column named "ID" for iTOL
                  relocate(id)%>% #put it first
                  select(-ASV) #remove repeated column
  
DF1 <- itol.toolkit::create_unit(data = Treatment_MeJA, 
                         key = "DF1_Treatment-MeJA", 
                         type = "DATASET_BINARY", #type of dataset
                         tree = tree_16S)

DF1@field$colors <- color_MeJA #change colors
DF1@field$shapes #circle is 2, so ok

#changed shape by hand in the .txt file, from circle(2) to triangle(5)

#save it

#itol.toolkit::write_unit(DF1, #"./07_consolidate_all_analysis/Results/Annotated_microbial_phylogenetic_tree/16S/")

```

SA (green dots)
```{r}
Treatment_SA <- df_complete %>%
                  dplyr::select(ASV,SA)%>%
                  mutate(id=ASV)%>% #make new column named "ID" for iTOL
                  relocate(id)%>% #put it first
                  select(-ASV) #remove repeated column
  
DF2 <- itol.toolkit::create_unit(data = Treatment_SA, 
                         key = "DF2_Treatment-SA", 
                         type = "DATASET_BINARY", #type of dataset
                         tree = tree_16S)

DF2@field$colors <- color_SA #change colors
DF2@field$shapes #check

#save it
#itol.toolkit::write_unit(DF2, #"./07_consolidate_all_analysis/Results/Annotated_microbial_phylogenetic_tree/16S/")

```

#### Genus

Written genus name with number of ASV (Genus-Number)
```{r}

df_Genus <- df_complete %>%
             dplyr::select(Genus)%>%
             mutate(id=rownames(.))%>% #make new column named "ID" for iTOL
             relocate(id) #put it first

#replace Genus names with names of ASV and add number to Genus so its clear that it's different ASVs (Marcel's suggestion)
#it also replaces NA's by the name of the ASV
df_Genus <- df_Genus %>% mutate(number=str_split_i(df_Genus$id, "_", 2))%>% #get number of ASV
                                mutate(Genus=paste(.$Genus, .$number, sep= "-"))%>% #make new with number and genus
                                mutate(Genus=gsub("NA", "bASV", .$Genus))%>%
                                dplyr::select(-number) #ok

#Fix very long name of Allorhizobium-Neorhizobium-Pararhizobium-Rhizobium
#change names
df_Genus$Genus <- gsub(df_Genus$Genus, pattern = "Allorhizobium-Neorhizobium-Pararhizobium-Rhizobium", replacement = "Rhizobium_group")

#make df 
DF3 <- create_unit(data = df_Genus,
                       key = "DF3_label_Genus",
                       type = "LABELS",
                       tree = tree_16S)

#itol.toolkit::write_unit(DF3, #"./07_consolidate_all_analysis/Results/Annotated_microbial_phylogenetic_tree/16S/")

```

#### Phylum
How to extend color palettes: #https://www.datanovia.com/en/blog/easy-way-to-expand-color-palettes-in-r/
```{r}

df_Phylum <- df_complete %>%
             dplyr::select(Phylum)%>%
             mutate(id=rownames(.))%>% #make new column named "ID" for iTOL
             relocate(id) #put it first
          
df_Phylum[is.na(df_Phylum)] <- "Not_assigned" #doesn't work if there's an NA
        
#add color to Phylum
Phylum <- unique(df_Phylum$Phylum) 
colorCount <- 15
color_phylum <- colorRampPalette(brewer.pal(colorCount, name = 'Set3'))(colorCount) #gives you a palette
color_phylum #copy-paste output so the colors can be seen
color_phylum <- c("#8DD3C7", "#E6F5B7", "#D9D7C9", "#D3A5B4", "#E9877F", "#88ADCC", "#D9B382", "#D8C965", "#000001", "#F9CDE4", "#DED7DA", "#C69FC7", "#C2ADC5", "#D6EBB2","#FFED6F")

#make it a df
df_Phylum_color <- data.frame(Phylum, color_phylum)

#merge with df_Phylum
df_Phylum <- df_Phylum%>%
             dplyr::inner_join(df_Phylum_color)

#option of just lines
#DF4 <- itol.toolkit::create_unit(data = df_Phylum, 
#                      key = "DF4_Phylum_colors", 
#                      type = "TREE_COLORS", 
#                      subtype = "clade",
#                      size_factor=10,
#                      tree = tree_16S)

#option with circles at the end
DF4 <- itol.toolkit::create_unit(data = df_Phylum, 
                      key = "DF4_Phylum_colors", 
                      type = "DATASET_SYMBOL",
                      position=1,
                      tree = tree_16S,
                      fill=1)

DF4@specific_themes$basic_plot$size_max <- 4

#change it for legend
DF4@common_themes$legend$shapes <- c("2", "2", "2", "2","2", "2","2", "2","2", "2","2", "2","2", "2","2")
DF4@common_themes$legend$shape_scales <- c(1,1,1,1,1,1,1,1,1,1,1,1,1,1)
DF4@common_themes$legend$horizontal <- "1"

#save it
#itol.toolkit::write_unit(DF4, #"./07_consolidate_all_analysis/Results/Annotated_microbial_phylogenetic_tree/16S/")

```
This works to color the tips, but not to have a legend I don't know why, Info has to be added by hand.

#### Class

```{r}

df_Class<- df_complete %>%
             dplyr::select(Class, Phylum)%>%
             mutate(id=rownames(.))%>% #make new column named "ID" for iTOL
             relocate(id) #put it first
          
df_Class[is.na(df_Class)] <- "Not_assigned" #doesn't work if there's an NA
 
#merge with df_Phylum
df_Class2 <- df_Class%>%
             dplyr::inner_join(df_Phylum_color)%>%
             dplyr::rename(color_class=color_phylum)

#get unique combinations
df_Class3 <- df_Class2 %>% distinct(Phylum, Class, color_class)%>% arrange(Phylum, Class)

#Now, replace colors in those that have more than one class per Phylum

"#E9877F"
Acidobacteriota <-  c("#E89586","#E97C76", "#E9807A", "#E98B84")
"#D9D7C9"
Actinobacteriota <- c("#D7D5C7","#DBD9CC")
"#8DD3C7"
Proteobacteria <- c("#8ED6CA","#A0E2D6")

#replace values 
df_Class4 <- df_Class3 %>% mutate(color_class = case_when(
  Class == "Acidobacteriae" ~ Acidobacteriota[1],
  Class == "Blastocatellia" ~ Acidobacteriota[2],
  Class == "Holophagae" ~ Acidobacteriota[3],
  Class == "Vicinamibacteria" ~ Acidobacteriota[4],
  Class == "Acidimicrobiia" ~ Actinobacteriota[1],
  Class == "Alphaproteobacteria" ~ Proteobacteria[1],
  Class == "Gammaproteobacteria" ~ Proteobacteria[2],
  TRUE ~ color_class
)) %>% select(-Phylum)

#Now, make it a single df 
df_Class5 <- df_Class %>% 
             inner_join(df_Class4)%>%
             select(-Phylum)

#prepare df for iTOL
DF5 <- itol.toolkit::create_unit(data = df_Class5, 
                      key = "DF5_Class_colors", 
                      type = "TREE_COLORS", 
                      subtype = "range",
                      tree = tree_16S)

#save it
#itol.toolkit::write_unit(DF5, #"./07_consolidate_all_analysis/Results/Annotated_microbial_phylogenetic_tree/16S/")

```

####LFC-1: Values from Meli

```{r}

#For this, first you'll take the summary values of the L2FC value per ASV per stress and this will be
#the input for iTOL (same as Pedro did)

df_lfc <- df_complete %>% 
          select(ASV,LFC_average_MeJA, LFC_average_SA)%>%
          dplyr::rename(id=ASV)

#make df 
DF6 <- create_unit(data = df_lfc,
                       key = "DF6_LFC_MeJA-SA_Heatmap",
                       type = "DATASET_HEATMAP",
                       tree = tree_16S)

DF6@data$tip[DF6@data$tip == 0] <- NA #replace 0's for NAs

#select colors 
DF6@specific_themes$heatmap$color$min <- "#2986cc"
DF6@specific_themes$heatmap$color$max <- "#ff0000"
DF6@specific_themes$heatmap$color$mid <- "#ebeceb"
DF6@specific_themes$heatmap$value$mid <- 0
DF6@specific_themes$heatmap$color$nan <- "#FFFFFF" #if NA, it will be blank

DF6@specific_themes$heatmap$tree$tree_display <- 0 #no display of tree in heatmap

#final check
DF6

#save it
#itol.toolkit::write_unit(DF6, #"./07_consolidate_all_analysis/Results/Annotated_microbial_phylogenetic_tree/16S/")

```

####LFC-2: Values from ANCOMBC

The difference is that it calculates the overall LFC of stress vs control per ASV, without taking species into account (in order to have an average value)

```{r}

colnames(df_complete)

df_lfc_ancom <- df_complete %>% 
                select(ASV,StressMeJA_ancombc, StressSA_ancombc)%>%
                dplyr::rename(id=ASV)

#make df 
DF7 <- create_unit(data = df_lfc_ancom,
                       key = "DF7_LFC_MeJA-SA_Heatmap_ancombc",
                       type = "DATASET_HEATMAP",
                       tree = tree_16S)

#select colors 
DF7@specific_themes$heatmap$color$min <- "#2986cc"
DF7@specific_themes$heatmap$color$max <- "#ff0000"
DF7@specific_themes$heatmap$color$mid <- "#ebeceb"
DF7@specific_themes$heatmap$value$mid <- 0
DF7@specific_themes$heatmap$color$nan <- "#FFFFFF" #if NA, it will be blank
DF7@specific_themes$heatmap$tree$tree_display <- 0 #no display of tree in heatmap

#final check
DF7

#save it
#itol.toolkit::write_unit(DF7, #"./07_consolidate_all_analysis/Results/Annotated_microbial_phylogenetic_tree/16S/")

```

####LFC: Direction 

To know how many species had a negative or positive LFC value on stress with respect to control

MeJA
```{r}

colnames(df_complete)

#MeJA
df_lfc_direction <- df_complete %>% 
                    select(ASV, negative_lfc_MeJA, positive_lfc_MeJA)%>%
                    dplyr::rename(id=ASV)
                     
#make df 
DF8 <- create_unit(data = df_lfc_direction,
                       key = "DF8_LFC_MeJA-Direction",
                       type = "DATASET_MULTIBAR",
                       tree = tree_16S)

DF8@field$labels <- c("#Sp with negative LFC", "#Sp with positive LFC")

#choose colors
display.brewer.pal(n=8, name='RdBu') #see palette
color_heatmap <- brewer.pal(n = 8, name = "RdBu") #get colors
negative <- color_heatmap[7]
positive <- color_heatmap[2]

DF8@field$colors <- c(negative, positive)


#final check
DF8

#save it
#itol.toolkit::write_unit(DF8, #"./07_consolidate_all_analysis/Results/Annotated_microbial_phylogenetic_tree/16S/")

```
SA
```{r}

#SA
df_lfc_direction <- df_complete %>% 
                    select(ASV, negative_lfc_SA, positive_lfc_SA)%>%
                    dplyr::rename(id=ASV)
                     
#make df 
DF9 <- create_unit(data = df_lfc_direction,
                       key = "DF9_LFC_SA-Direction",
                       type = "DATASET_MULTIBAR",
                       tree = tree_16S)

DF9@field$labels <- c("#Sp with negative LFC", "#Sp with positive LFC")

#choose colors
display.brewer.pal(n=8, name='RdBu') #see palette
color_heatmap <- brewer.pal(n = 8, name = "RdBu") #get colors
negative <- color_heatmap[7]
positive <- color_heatmap[2]

DF9@field$colors <- c(negative, positive)


#final check
DF9

#save it
#itol.toolkit::write_unit(DF9, #"./07_consolidate_all_analysis/Results/Annotated_microbial_phylogenetic_tree/16S/")

```

Now it's finished, and we'll do the same with the fungal tree. For now let's save this environment so we don't have to run it again.

```{r}

save.image(file="./07_consolidate_all_analysis/Results/Annotated_microbial_phylogenetic_tree/16S/16S_R-environment.Rdata")

#clean everything
rm.all.but(keep=c("lfcres_GoldenASVs_FinalDF_l_replaceNA", "DA_RF_NM_ASVs_intersect", "Spp_abb_name_Kasper"),
                keep_functions = F)
```


# ITS

## Phyloseq

```{r}

#load Golden ASVs names
DA_RF_NM_ASVs_intersect

#### ITS ####
MeJA_asvs <- DA_RF_NM_ASVs_intersect$MeJA_Fun_GoldenASVs #82
SA_asvs <- DA_RF_NM_ASVs_intersect$SA_Fun_GoldenASVs #97

#Now I need to get the unique names 
fun_sig_asvs <- c(MeJA_asvs, SA_asvs) #179
fun_sig_asvs <- unique(fun_sig_asvs) #119 unique ones, there are 60 shared ASVs 

# this will load the 16S and ITS data as a list of two objects
load("./Data/phyloseq_objects/unnormalized_BacFun_rhizoplane_HighFilt_ps_l.RData")
unnormalized_BacFun_rhizoplane_HighFilt_ps_l

#Now, I need to remove all those samples that belong to guest species + others not present in the MY tree
ps_Fun <- subset_samples(unnormalized_BacFun_rhizoplane_HighFilt_ps_l$Fun,
                       Sp_abb_name %in% Spp_abb_name_Kasper) #only those plant sp present in the tree

#prune ps to keep only those that belong to the Golden ASVs
ps_golden <- prune_taxa(fun_sig_asvs, ps_Fun)

#check
ps_golden #ok

#remove big ps object to clear up space
rm(ps_Fun, unnormalized_BacFun_rhizoplane_HighFilt_ps_l)

```

### Make new phylogenetic tree

Skip and go to line 786 to load it
```{r}

#get the sequences and save them as FASTA format
#ps_golden %>%
#refseq() %>%
#Biostrings::writeXStringSet("./07_consolidate_all_analysis/Results/Annotated_microbial_phylogenetic_tree/ITS/ITS_goldenASVs_sequences.fasta",
#             append=FALSE,compress=FALSE,
#             compression_level=NA, format="fasta")


#Go to MAFTT [...]
```

I followed the same approach as for Bacteria. 

30-08-24: I tried using the web server of MAFFT using `E-INS-i` as option for iterative refinement method,  leaving gapy regions and UPGMA guide tree. However, server was busy and instead used MAFFT as a tool in the EMBL's EBI server (https://www.ebi.ac.uk/jdispatcher/)

Job ID: mafft-I20240830-105421-0053-50505090-p1m

The alignment in fasta format  was used to run the following slurm script to run IQ-tree in the HPC: which means that iqtree will run with a DNA sequence type, 10,000 replicates in bootstrap analysis and that finds the best substitution model automatically.

```{bash}

#!/bin/bash -login

#SBATCH --time=4:00:00
#SBATCH --ntasks=1
#SBATCH --cpus-per-task=8
#SBATCH --mem=4G
#SBATCH --job-name=16S_tree
#SBATCH --output=ITS_tree.output


cd /lustre/nobackup/INDIVIDUAL/arago004/IQ_tree/ITS ##place where I had the .fasta alignment

iqtree2 -s ITS_goldenASVs_MAFFT_alignment_EBI.fasta -st DNA -m MFP -bb 10000 -nt AUTO ##

```

Once I got the tree (it took less than 5 minutes) I imported back to R

```{r}
#Import tree
tree_ITS <- read.tree("./07_consolidate_all_analysis/Results/Annotated_microbial_phylogenetic_tree/ITS/ITS_goldenASVs_MAFFT_alignment_EBI.fasta.treefile")

#check
tree_ITS
plot(tree_ITS) #good 
tree_ITS <- multi2di(tree_ITS) #root it

#save it in .tre format to be imported in iTOL
#ape::write.tree(tree_ITS, "./07_consolidate_all_analysis/Results/Annotated_microbial_phylogenetic_tree/ITS/ITS_tree.tre")

#change tree in ps object 
ps_golden@phy_tree <- tree_ITS

```

## Taxonomy dataframe

```{r}
#make a new df with two columns; one for MeJA and another one for SA
df_fun_sig_asvs <- as.data.frame(fun_sig_asvs) 
colnames(df_fun_sig_asvs) <- "asv"                   
df_fun_sig_asvs <- df_fun_sig_asvs %>%
                   mutate(MeJA=case_when(asv %in% MeJA_asvs ~ "1",
                                         TRUE ~ "-1"))%>% #0 is for empty shape, -1 is no shape
                   mutate(SA=case_when(asv %in% SA_asvs ~ "1",
                                         TRUE ~ "-1"))

#let's get the taxonomy
#clean it first
tax_table(ps_golden)[, colnames(tax_table(ps_golden))] <- gsub(tax_table(ps_golden)[, colnames(tax_table(ps_golden))], pattern = "[a-z]__", replacement = "")

#put it in a df 
df_taxonomy <- as.data.frame(ps_golden@tax_table)
order_asvs <- rownames(df_taxonomy) #get order of rows to keep using it for rest of df's

#re-order sig_asvs df 
df_fun_sig_asvs <- df_fun_sig_asvs %>%
                   column_to_rownames(var="asv") #first make new rownames
#now order it 
df_fun_sig_asvs <- df_fun_sig_asvs[order(match(rownames(df_fun_sig_asvs),
                                   order_asvs)), , drop=FALSE] 

#now you can merge df's
df_complete <- cbind(df_fun_sig_asvs, df_taxonomy)
df_complete <- df_complete %>% select(-Confidence) #remove non-informative columns 

```

## L2FC per Sp from Melissa [Not used for iTOL]

```{r}

#load df with log2-fold change
load("./07_consolidate_all_analysis/r_objects/MU_HeatmapsGoldenASVs/lfcres_GoldenASVs_FinalDF_l_replaceNA.RData")

#first, only MeJA
lfc_MeJA <- lfcres_GoldenASVs_FinalDF_l_replaceNA$lfcres_df_MeJA_Fun

#add MeJA suffix
lfc_MeJA <- lfc_MeJA %>% 
            mutate(across(everything(), ~ na_if(., 0.00)))%>% #first, change 0.0 to NA's
            rename_with(~ paste0(., "_MeJA"))

#let's make a column with the amount of times an ASV was significant
#(if 0.0 I asume it was NA and therefore not significant)
lfc_average <- lfc_MeJA %>% 
             mutate(LFC_average_MeJA = rowMeans(., na.rm=T))%>%  #get average
             select(LFC_average_MeJA)

lfc_ocurrences <- lfc_MeJA %>% 
             mutate(num_sp_sig_MeJA= rowSums(!is.na(.)))%>%
             mutate(num_sp_NA_MeJA=rowSums(is.na(.)))%>% 
             mutate(negative_lfc_MeJA=rowSums(lfc_MeJA < 0, na.rm=T))%>% 
             mutate(positive_lfc_MeJA=num_sp_sig_MeJA-negative_lfc_MeJA) 

#add the average value
lfc_average$ASV <- rownames(lfc_average)
lfc_ocurrences$ASV <- rownames(lfc_ocurrences)
lfc_MeJA <- left_join(lfc_ocurrences, lfc_average) 
lfc_MeJA <- lfc_MeJA %>% relocate(ASV)

#then, SA
lfc_SA <- lfcres_GoldenASVs_FinalDF_l_replaceNA$lfcres_df_SA_Fun

#add SA suffix
lfc_SA <- lfc_SA %>% 
            mutate(across(everything(), ~ na_if(., 0.00)))%>% #first, change 0.0 to NA's
            rename_with(~ paste0(., "_SA"))

#let's make a column with the amount of times an ASV was significant
#(if 0.0 I asume it was NA and therefore not significant)
lfc_average <- lfc_SA %>% 
             mutate(LFC_average_SA = rowMeans(., na.rm=T))%>%  #get average
             select(LFC_average_SA)

lfc_ocurrences <- lfc_SA %>% 
             mutate(num_sp_sig_SA= rowSums(!is.na(.)))%>%
             mutate(num_sp_NA_SA=rowSums(is.na(.)))%>% 
             mutate(negative_lfc_SA=rowSums(lfc_SA < 0, na.rm=T))%>% 
             mutate(positive_lfc_SA=num_sp_sig_SA-negative_lfc_SA) 

#add the average value
lfc_average$ASV <- rownames(lfc_average)
lfc_ocurrences$ASV <- rownames(lfc_ocurrences)
lfc_SA <- left_join(lfc_ocurrences, lfc_average) 
lfc_SA <- lfc_SA %>% relocate(ASV)

#Now, let's join the df's
lfc <- full_join(lfc_MeJA, lfc_SA, by="ASV")

#Let's make a new df first with MeJA information
df_complete2 <- df_complete #our dataframe with metadata

#add column named ASV
df_complete2$ASV <- rownames(df_complete2) #make new column 'ASV'

#join df 
df_complete2 <- full_join(df_complete2, lfc, by="ASV")

#order
df_complete2 <- df_complete2 %>% 
                dplyr::relocate(ASV, MeJA, SA)

rownames(df_complete2) <- df_complete2$ASV
```


## LFC from Ancombc [Used for iTOL]

Let's try to use Ancom just to calculate the log-fold change per ASVs between Control and Stress-treatments. This will be an average value accross species (I think this was the way Pedro plotted the heattrees but needs confirmation)

```{r}

#This code comes from the psf_microbiome project

ps_golden #this is my ps object with golden asvs for Bac

library(ANCOMBC)

#Just for Bac-GoldenASVs but using 3 groups
set.seed(123)
ancom <- ANCOMBC::ancombc(phyloseq = ps_golden, 
                      formula = "Stress", 
                      p_adj_method = "fdr", #same as "BH" Benjamini & Hochberg
                      lib_cut = 0,
                      prv_cut=0,
                      group = "Stress", 
                      struc_zero = TRUE, #with structural zero's
                      neg_lb = FALSE, #classifies structural zero's if it's zero and not its asymptotic lower bound 
                      tol = 1e-5,               
                      max_iter = 100, 
                      conserve = TRUE, 
                      alpha = 0.05,
                      global=TRUE)

head(ancom$res$lfc) #LFC

#Get LFC values

lfc_ancom <- ancom$res$lfc #log-fold change according to control 

#### Now I have a single value of LFC per each golden ASV for both MeJA and SA vs Control in the lfc object ####
lfc_ancom <- lfc_ancom %>% 
             mutate(ASV=taxon)%>% 
             select(c(ASV, StressMeJA, StressSA))%>%
             rename_with(~ paste0(., "_ancombc"))
colnames(lfc_ancom)[1] <- "ASV"

#finally, join it with full df 
df_complete3 <- left_join(df_complete2, lfc_ancom)
head(df_complete3)

rownames(df_complete3) <- df_complete3$ASV

#save df
save(df_complete3, file="./07_consolidate_all_analysis/Results/Annotated_microbial_phylogenetic_tree/ITS/df_complete.Rdata")

```


## iTOL


### Metadata in iTOL format

First, arrange data and colors to be used 
```{r}

#First gather what you need 
df_complete #our dataframe with metadata
tree_ITS #our tree

#Set-up colors

#### colors for Treatment ####
color_MeJA <- "#E69F00"
color_SA <- "#009E73"

```


#### Treatment

MeJA
```{r}

Treatment_MeJA <- df_complete3 %>%
                  dplyr::select(ASV,MeJA)%>%
                  mutate(id=ASV)%>% #make new column named "ID" for iTOL
                  relocate(id)%>% #put it first
                  select(-ASV) #remove repeated column
  
DF1 <- itol.toolkit::create_unit(data = Treatment_MeJA, 
                         key = "DF1_Treatment-MeJA", 
                         type = "DATASET_BINARY", #type of dataset
                         tree = tree_ITS)

DF1@field$colors <- color_MeJA #change colors
DF1@field$shapes #circle is 2, so ok

#save it
#itol.toolkit::write_unit(DF1, "./07_consolidate_all_analysis/Results/Annotated_microbial_phylogenetic_tree/ITS/")

```

SA
```{r}
Treatment_SA <- df_complete3 %>%
                  dplyr::select(ASV,SA)%>%
                  mutate(id=ASV)%>% #make new column named "ID" for iTOL
                  relocate(id)%>% #put it first
                  select(-ASV) #remove repeated column
  
DF2 <- itol.toolkit::create_unit(data = Treatment_SA, 
                         key = "DF2_Treatment-SA", 
                         type = "DATASET_BINARY", #type of dataset
                         tree = tree_ITS)

DF2@field$colors <- color_SA #change colors
DF2@field$shapes #check

#save it
#itol.toolkit::write_unit(DF2, "./07_consolidate_all_analysis/Results/Annotated_microbial_phylogenetic_tree/ITS/")

```

#### Genus

```{r}

df_Genus <- df_complete3 %>%
             dplyr::select(Genus, ASV)%>%
             dplyr::rename(id=ASV)%>% #make new column named "ID" for iTOL
             relocate(id) #put it first

#replace Genus names with names of ASV and add number to Genus so its clear that it's different ASVs (Marcel's suggestion)
#it also replaces NA's by the name of the ASV
df_Genus <- df_Genus %>% mutate(number=str_split_i(df_Genus$id, "_", 2))%>% #get number of ASV
                                mutate(Genus=paste(.$Genus, .$number, sep= "-"))%>% #make new with number and genus
                                mutate(Genus=gsub("NA", "fASV", .$Genus))%>%
                                dplyr::select(-number) #ok

#make df 
DF3 <- create_unit(data = df_Genus,
                       key = "DF3_label_Genus",
                       type = "LABELS",
                       tree = tree_ITS)

#itol.toolkit::write_unit(DF3, "./07_consolidate_all_analysis/Results/Annotated_microbial_phylogenetic_tree/ITS/")

```


#### Phylum
How to extend color palettes: #https://www.datanovia.com/en/blog/easy-way-to-expand-color-palettes-in-r/
```{r}

df_Phylum <- df_complete3 %>%
             dplyr::select(Phylum, ASV)%>%
             dplyr::rename(id=ASV)%>% #make new column named "ID" for iTOL
             relocate(id) #put it first
          
df_Phylum[is.na(df_Phylum)] <- "Not_assigned" #doesn't work if there's an NA
        
#add color to Phylum
Phylum <- unique(df_Phylum$Phylum) 
colorCount <- 7
color_phylum <- colorRampPalette(brewer.pal(colorCount, name = 'Pastel2'))(colorCount) #gives you a palette
color_phylum #copy-paste output so the colors can be seen
color_phylum <- c("#BEBADA","#000001","#B3E2CD", "#FDCDAC", "#CBD5E8", "#F4CAE4", "#FFF2AE")

#make it a df
df_Phylum_color <- data.frame(Phylum, color_phylum)

#merge with df_Phylum
df_Phylum <- df_Phylum%>%
             dplyr::inner_join(df_Phylum_color)

#option with circles at the end
DF4 <- itol.toolkit::create_unit(data = df_Phylum, 
                      key = "DF4_Phylum_colors", 
                      type = "DATASET_SYMBOL",
                      position=1,
                      tree = tree_ITS,
                      fill=1)

DF4@specific_themes$basic_plot$size_max <- 4

#change it for legend
DF4@common_themes$legend$shapes <- c("2", "2", "2", "2","2", "2", "2")
DF4@common_themes$legend$shape_scales <- c(1,1,1,1,1,1,1)
DF4@common_themes$legend$horizontal <- "1"

#save it
#itol.toolkit::write_unit(DF4, "./07_consolidate_all_analysis/Results/Annotated_microbial_phylogenetic_tree/ITS/")

```

#### Class

```{r}

df_Class<- df_complete3 %>%
             dplyr::select(Class, ASV, Phylum)%>%
             dplyr::rename(id=ASV)%>% #make new column named "ID" for iTOL
             relocate(id) #put it first
          
df_Class[is.na(df_Class)] <- "Not_assigned" #doesn't work if there's an NA
        
#merge with df_Phylum
df_Class2 <- df_Class%>%
             dplyr::inner_join(df_Phylum_color)%>%
             dplyr::rename(color_class=color_phylum)

#get unique combinations
df_Class3 <- df_Class2 %>% distinct(Phylum, Class, color_class)%>% arrange(Phylum, Class)

#Now, replace colors in those that have more than one class per Phylum
#the colors I got them from chat gpt by asking:
# "give me a set of X colors similar to to #examplecolor"

"#BEBADA"
Ascomycota <-  c("#AFA2C8","#D0C1E1", "#C6A7D3", "#B3A4CB", "#D4B8E4", "#A89FC0")
"#F4CAE4"
Basidiomycota <- c("#F6A8B8","#F2B7D4")
"#FFF2AE"
Chytridiomycota <- c("#F7E6A0","#FDF7B6", "#FDF0A0")

#replace values 
df_Class4 <- df_Class3 %>% mutate(color_class = case_when(
  Class == "Dothideomycetes" ~ Ascomycota[1],
  Class == "Eurotiomycetes" ~ Ascomycota[2],
  Class == "Leotiomycetes" ~ Ascomycota[3],
  Phylum == "Ascomycota" & Class == "Not_assigned" ~ Ascomycota[4],
  Class == "Pezizomycotina_cls_Incertae_sedis" ~ Ascomycota[5],
  Class == "Sordariomycetes" ~ Ascomycota[6],
  Class == "Agaricomycetes" ~ Basidiomycota[1],
  Class == "Tremellomycetes" ~ Basidiomycota[2],
  Class == "Chytridiomycetes" ~ Chytridiomycota[1],
  Class == "Rhizophlyctidomycetes" ~ Chytridiomycota[2],
  Class == "unidentified" ~ Chytridiomycota[3],
  TRUE ~ color_class
)) 

#Now, make it a single df 
df_Class5 <- df_Class %>% 
             inner_join(df_Class4)%>%
             select(-Phylum)

DF5 <- itol.toolkit::create_unit(data = df_Class5, 
                      key = "DF5_Class_colors", 
                      type = "TREE_COLORS", 
                      subtype = "range",
                      tree = tree_ITS)

#save it
#itol.toolkit::write_unit(DF5, "./07_consolidate_all_analysis/Results/Annotated_microbial_phylogenetic_tree/ITS/")

```

####LFC-1: Values from Meli

```{r}

#For this, first you'll take the summary values of the L2FC value per ASV per stress and this will be
#the input for iTOL (same as Pedro did)

df_lfc <- df_complete3 %>% 
          select(ASV,LFC_average_MeJA, LFC_average_SA)%>%
          dplyr::rename(id=ASV)

#make df 
DF6 <- create_unit(data = df_lfc,
                       key = "DF6_LFC_MeJA-SA_Heatmap",
                       type = "DATASET_HEATMAP",
                       tree = tree_ITS)

DF6@data$tip[DF6@data$tip == 0] <- NA #replace 0's for NAs

#select colors 
DF6@specific_themes$heatmap$color$min <- "#2986cc"
DF6@specific_themes$heatmap$color$max <- "#ff0000"
DF6@specific_themes$heatmap$color$mid <- "#ebeceb"
DF6@specific_themes$heatmap$value$mid <- 0
DF6@specific_themes$heatmap$color$nan <- "#FFFFFF" #if NA, it will be blank

DF6@specific_themes$heatmap$tree$tree_display <- 0 #no display of tree in heatmap

#final check
DF6

#save it
#itol.toolkit::write_unit(DF6, "./07_consolidate_all_analysis/Results/Annotated_microbial_phylogenetic_tree/ITS/")

```

####LFC-2: Values from ANCOMBC [Used]

The difference is that it calculates the overall LFC of stress vs control per ASV, without taking species into account (in order to have an average value)

```{r}

#For this, first you'll take the summary values of the L2FC value per ASV per stress and this will be
#the input for iTOL (same as Pedro did)

colnames(df_complete3)

df_lfc_ancom <- df_complete3 %>% 
                select(ASV,StressMeJA_ancombc, StressSA_ancombc)%>%
                dplyr::rename(id=ASV)

#make df 
DF7 <- create_unit(data = df_lfc_ancom,
                       key = "DF7_LFC_MeJA-SA_Heatmap_ancombc",
                       type = "DATASET_HEATMAP",
                       tree = tree_ITS)

#select colors 
DF7@specific_themes$heatmap$color$min <- "#2986cc"
DF7@specific_themes$heatmap$color$max <- "#ff0000"
DF7@specific_themes$heatmap$color$mid <- "#FFFFFF"
DF7@specific_themes$heatmap$value$mid <- 0
#DF7@specific_themes$heatmap$color$nan <- "#FFFFFF" #if NA, it will be blank

DF7@specific_themes$heatmap$tree$tree_display <- 0 #no display of tree in heatmap

#final check
DF7

#save it
#itol.toolkit::write_unit(DF7, "./07_consolidate_all_analysis/Results/Annotated_microbial_phylogenetic_tree/ITS/")

```

####LFC: Direction 
MeJA
```{r}

colnames(df_complete3)

#MeJA
df_lfc_direction <- df_complete3 %>% 
                    select(ASV, negative_lfc_MeJA, positive_lfc_MeJA)%>%
                    dplyr::rename(id=ASV)
                     
#make df 
DF8 <- create_unit(data = df_lfc_direction,
                       key = "DF8_LFC_MeJA-Direction",
                       type = "DATASET_MULTIBAR",
                       tree = tree_ITS)

DF8@field$labels <- c("#Sp with negative LFC", "#Sp with positive LFC")

#choose colors
display.brewer.pal(n=8, name='RdBu') #see palette
color_heatmap <- brewer.pal(n = 8, name = "RdBu") #get colors
negative <- color_heatmap[7]
positive <- color_heatmap[2]

DF8@field$colors <- c(negative, positive)


#final check
DF8

#save it
#itol.toolkit::write_unit(DF8, "./07_consolidate_all_analysis/Results/Annotated_microbial_phylogenetic_tree/ITS/")

```
SA
```{r}

#SA
df_lfc_direction <- df_complete3 %>% 
                    select(ASV, negative_lfc_SA, positive_lfc_SA)%>%
                    dplyr::rename(id=ASV)
                     
#make df 
DF9 <- create_unit(data = df_lfc_direction,
                       key = "DF9_LFC_SA-Direction",
                       type = "DATASET_MULTIBAR",
                       tree = tree_ITS)

DF9@field$labels <- c("#Sp with negative LFC", "#Sp with positive LFC")

#choose colors
display.brewer.pal(n=8, name='RdBu') #see palette
color_heatmap <- brewer.pal(n = 8, name = "RdBu") #get colors
negative <- color_heatmap[7]
positive <- color_heatmap[2]

DF9@field$colors <- c(negative, positive)


#final check
DF9

#save it
#itol.toolkit::write_unit(DF9, "./07_consolidate_all_analysis/Results/Annotated_microbial_phylogenetic_tree/ITS/")

```

Now it's finished, and we'll do the same with the fungal tree. For now let's save this environment so we don't have to run it again.

```{r}

#save.image(file="./07_consolidate_all_analysis/Results/Annotated_microbial_phylogenetic_tree/ITS/ITS_R-environment.Rdata")

#ready!
```


