---
title: "MU_10_Machine_learning_ReRunningCode"
author: "Pedro"
date: "2025-07-02"
output: html_document
editor_options: 
  chunk_output_type: console
---

This script is a step by step copy of https://github.com/PedroBeschoren/Family_experiment/blob/master/10_Machine_learning/code/10_check_RF_precision_metrics_MeJAvsControl_SAvsControl.Rmd made by Melissa to double checking of the code and identifying parts that could be cleaned.

##Start of the code from Pedro
this script will load 400k MeJa-VS-Control and SA-VS-Control boruta objects, with bacteria and fungi
then it extracts fixed_boruta ASVs names, removing all species absent from the MY tree
Then it runs the fisher analysis pipeline, and indicates the taxons highlighted

Then, it can extract importan ASV features
in the end of the script, we compare 20k boruta trees to 400k boruta trees



# 10.0 -  load libraries 
```{r results="hide"}
# this will load all key libraries for the whole project
source("./renv/0_load_libraries.R")



# this will load all key libraries for the whole project
library(metagMisc)
library(dplyr)
library(tibble)
library(tidyverse)
library(Boruta)
library(phyloseq)
library(caret)
library(randomForest)
library(parallel) # use parallel computing to speed up RF calculations
library(pROC)

# loads the custom functions used in this script
source("./10_Machine_learning/code/Machine_Learning_custom_functions.R")




# increases memory limit used by R
memory.limit(size = 350000)


```



# 10.1 - load phyloseq objects
```{r}

# this will load the CSS 16S and ITS data as a list of two objects
load("./Data/phyloseq_objects/CSS_BacFun_rhizoplane_HighFilt_ps_l.RData")

######### split ps objects

# split the original object by species
ps_sp_l<-lapply(CSS_BacFun_rhizoplane_HighFilt_ps_l, function (x) 
  phyloseq_sep_variable(x,variable = c("Sp_abb_name")))

# filter  the original object to keep only brassicacea present in the MY tree
ps_brassicaceae_BacFun_l<-lapply(CSS_BacFun_rhizoplane_HighFilt_ps_l, function (x) 
  subset_samples(x, MY_tree_presence =="inside_MY_tree"))

```

# 10.2 - Define and extract RF-important ASVs

Load boruta objects, fix them, apply Ds cutoff, define RF ASVs for 22sp, define RF ASVs for microiomes Type LI and TYpe LII (14sp+8sp)


## 10.2.1 - load 400k boruta objects, put in single list
```{r}
#load MejA data
load(file = "./10_Machine_learning/r_objects/Boruta_MeJA_bac_33sp_400k.RData")
load(file = "./10_Machine_learning/r_objects/Boruta_MeJA_fun_33sp_400k.RData")

# rename objects, remove old ones for memory saving
Boruta_MeJA_bac_33sp<-Boruta_stress_bac_33sp
Boruta_MeJA_fun_33sp<-Boruta_stress_fun_33sp
rm(Boruta_stress_bac_33sp, Boruta_stress_fun_33sp)


#load SA data
load(file = "./10_Machine_learning/r_objects/Boruta_SA_bac_33sp_400k.RData")
load(file = "./10_Machine_learning/r_objects/Boruta_SA_fun_33sp_400k.RData")

# rename objects, remove old ones for memory saving
Boruta_SA_bac_33sp<-Boruta_stress_bac_33sp
Boruta_SA_fun_33sp<-Boruta_stress_fun_33sp
rm(Boruta_stress_bac_33sp, Boruta_stress_fun_33sp)

#put these objects into a 4-object list
Boruta_BacFun_MeJASA_33sp_l<-list("Boruta_MeJA_bac_22sp" = Boruta_MeJA_bac_33sp,
                                 "Boruta_SA_bac_22sp" = Boruta_SA_bac_33sp,
                                 "Boruta_MeJA_fun_22sp" = Boruta_MeJA_fun_33sp,
                                 "Boruta_SA_fun_22sp" = Boruta_SA_fun_33sp)

#remove old objects
rm(Boruta_MeJA_bac_33sp, Boruta_MeJA_fun_33sp, Boruta_SA_bac_33sp,Boruta_SA_fun_33sp)


```

## 10.2.2 - fix boruta objects and remove plant sp outside MY tree
```{r}
#fix boruta objects 
#Note from MU: This is a function that modifies the finalDecision attribute of the Boruta object: https://search.r-project.org/CRAN/refmans/Boruta/html/TentativeRoughFix.html. "finalDecision" displays for each ASV one of three options, Confirmedd, Rejected or Tentative. Then this function Tentative rough fix uses median importance and importance of each ASV to do a strict cut to either confirmed or tentative
Fixed_Boruta_BacFun_MeJASA_33sp_l<-lapply(Boruta_BacFun_MeJASA_33sp_l, function(x)
  lapply(x, function(z)
    TentativeRoughFix(z)))

#define plant species inside MY tree
sp_inside_MY_tree<-ps_brassicaceae_BacFun_l$Bac@sam_data$Sp_abb_name%>%unique

# remove plant species outside MY tree
Fixed_Boruta_BacFun_MeJASA_22sp_l<-lapply(Fixed_Boruta_BacFun_MeJASA_33sp_l, function(x)
  x[as.character(sp_inside_MY_tree)])

# check length and sp names
lapply(Fixed_Boruta_BacFun_MeJASA_22sp_l, length)
lapply(Fixed_Boruta_BacFun_MeJASA_22sp_l, names)

# extract important ASV IDs
Boruta_ASVs_22sp<-lapply(Fixed_Boruta_BacFun_MeJASA_22sp_l, function (z)
  lapply(z, function(x)
    Boruta::getSelectedAttributes(x)))

# check histogram of ASVs IDs frequency
lapply(Boruta_ASVs_22sp, function (x) x%>%unlist%>%table%>%hist)

### attention: filter Ds here ####
# Ds has too many bASVs on SA because of the strong SA effects. keep only to top quartile to reduce wegihts of this species in cross-comparisons
Ds_bASVs_SA<-dplyr::filter(attStats(Fixed_Boruta_BacFun_MeJASA_22sp_l$Boruta_SA_bac_22sp$Ds), decision == "Confirmed") # important Ds bASVs in SA
Ds_bASVs_SA_top_25<- dplyr::filter(Ds_bASVs_SA, medianImp >= summary(Ds_bASVs_SA$medianImp)[5]) # keep bASVs on the top quartile of importance

Boruta_ASVs_22sp$Boruta_SA_bac_22sp$Ds<-rownames(Ds_bASVs_SA_top_25) # substitute the ~ 160 bASVs for the 40 most important bASVs

# put all ASVs into a flat list without repetitions
RF_ControlMeJA_ControlSA_ASVs_22sp<-lapply(Boruta_ASVs_22sp, function (x) x%>%unlist%>%unique)
```

## 10.2.3 - make a boxplot showing the Ds cutoff in imporatnt ASVs
```{r}
#get RF ASv importance into a list of dfs
RF_importance_df_l_l<-
lapply(Fixed_Boruta_BacFun_MeJASA_22sp_l, function(z)
  lapply(z, function(x)
  dplyr::filter(attStats(x), decision == "Confirmed")))

#merge the list
RF_importance_df_l<- 
  lapply(RF_importance_df_l_l, function(x)
    do.call("rbind", x)%>%
      rownames_to_column(var = "Sp_abb"))

# Bac_SA_Br is failing to add a on the snippet above here, I can't see why. bASV_1790 is present in the input object. this fix it via code
RF_importance_df_l$Boruta_SA_bac_22sp[RF_importance_df_l$Boruta_SA_bac_22sp$Sp_abb=="Br",]$Sp_abb<-
  paste0(RF_importance_df_l$Boruta_SA_bac_22sp[RF_importance_df_l$Boruta_SA_bac_22sp$Sp_abb=="Br",]$Sp_abb,
         ".",
         rownames(RF_importance_df_l_l$Boruta_SA_bac_22sp$Br)) 



RF_importance_df<-
    do.call("rbind", RF_importance_df_l)%>%
      rownames_to_column(var = "Stress")

#adjust column names and content
RF_importance_df<-  
  tidyr::separate(data = RF_importance_df,
                  col = Sp_abb,
                  into = c("Sp_abb", "ASV_ID"), 
                  sep = "\\.",
                  remove = FALSE)

RF_importance_df$Stress<-gsub("Boruta_", "",RF_importance_df$Stress)
RF_importance_df$Stress<-gsub('_22sp.*','',RF_importance_df$Stress)

RF_importance_df<-  
  tidyr::separate(data = RF_importance_df,
                  col = Stress,
                  into = c("Stress", "Kingdom"), 
                  sep = "_",
                  remove = FALSE)

# adjust species names as a factor that mirrors phylogeny
RF_importance_df$Sp_abb<-factor(RF_importance_df$Sp_abb, levels = levels(sp_inside_MY_tree ))

#colour vector
colour_v <- c("#F0E442", "#88CCEE", "#E69F00", "#009E73", "#999999")

# boxplot of number of important ASVs with labels
Ds_RF_imp_cutoff_NimpASVs<-
RF_importance_df%>%
  group_by(Sp_abb, Stress, Kingdom)%>%
  dplyr::summarise(count = n())%>%
  ggplot(aes(x = Stress, y = count, fill = Stress ))+
  geom_boxplot(outlier.shape = NA, 
               position=position_dodge(0.8), alpha = 0.5)+
  geom_text_repel(aes(label = Sp_abb, color = Stress, fill = NULL),
                   max.overlaps = 18,
                   position = position_dodge(width = 0.75))+
  theme_bw()+
  facet_wrap(~Kingdom, ncol = 1, nrow = 2 )+
    scale_discrete_manual(aesthetics = c("color", "fill"),
                        values = colour_v[3:4])+
    xlab("Treatment")+
  labs(fill = "Treatment", color ="Treatment" )+
  ylab("Number of ASVs tagged as important by Boruta")


# here we see that Ds is almost 20 times the median value
RF_importance_df_summary_count<-
RF_importance_df%>%
  group_by(Sp_abb, Stress, Kingdom)%>%
  dplyr::summarise(count = n())

RF_importance_df_summary_count%>%
  group_by(Stress, Kingdom)%>%
  dplyr::summarise(median_count = median(count))



# boxplot as ASV importance per sp with jitter
Ds_RF_imp_cutoff_topquartile<-
ggplot(data = RF_importance_df, aes(x = Sp_abb, y = meanImp, fill = Stress ))+
  geom_boxplot(outlier.shape = NA, 
               position=position_dodge(0.8), alpha = 0.5)+
 geom_point(position = position_jitterdodge(jitter.width = 0.2,
                                             dodge.width = 0.7),
             aes(fill = Stress),  alpha = 0.3,
             pch = 21)+
  scale_discrete_manual(aesthetics = "fill",
                        values = colour_v[3:4])+
  theme_bw()+
  labs(fill = "Treatment")+
  xlab("Plant Species")+
  ylab("Mean importance of ASVs tagged as important by Boruta")+
  facet_wrap(~Kingdom, ncol = 1, nrow = 2 )


# put the two plots above in a panel and save
ds_imp_cutooff_pannel<-
ggarrange(Ds_RF_imp_cutoff_NimpASVs, 
          Ds_RF_imp_cutoff_topquartile, 
          ncol = 2, 
          labels = "AUTO", common.legend = TRUE, legend = "top",
          widths = c(0.4, 1.6))
```


```{r}
#export figure
ggsave(ds_imp_cutooff_pannel,
       filename = "./10_Machine_learning/results/Ds_RF_imp_cutoff_pannel.svg", 
       width = 260,
       height = 260, 
       units = "mm")

```

# 10.3 - Check and Plot Model Precision

## 10.3.1 - Load Cross-Validation Models (25 Replicates), Define Functions and Vectors
```{r}
# Load MeJA models into the environment (can take a while)
load(file = "./10_Machine_learning/r_objects/CV_25replicated_MeJA_400k_l_l_33sp.RData")

# Load SA models into the environment (can take a while)
load(file = "./10_Machine_learning/r_objects/CV_25replicated_SA_400k_l_l_33sp.RData")

# Define vectors of plant species and phylogenies
Sp_Lineage_Walden <- CSS_BacFun_rhizoplane_HighFilt_ps_l$Bac@sam_data$Sp_Lineage_Walden
sp_vector <- CSS_BacFun_rhizoplane_HighFilt_ps_l$Bac@sam_data$Sp_abb_name
MY_tree_vector <- CSS_BacFun_rhizoplane_HighFilt_ps_l$Bac@sam_data$MY_tree_presence
lineage_sp_df <- as.data.frame(Sp_Lineage_Walden)
lineage_sp_df$Sp_abb_name <- sp_vector
lineage_sp_df$MY_tree_vector <- MY_tree_vector
lineage_sp_df <- lineage_sp_df %>% unique()

# Function to extract CV metrics
extract_accuracy <- function(CV_output_l) {
  cv_metrics <- map(CV_output_l, 3)
  accuracy <- unlist(map(cv_metrics, 1)) # Accuracy
  kappa <- unlist(map(cv_metrics, 2))   # Kappa
  AccuracyPValue <- unlist(map(cv_metrics, 6)) # Accuracy P-value
  model_metrics <- as.data.frame(accuracy)
  model_metrics$kappa <- kappa
  model_metrics$AccuracyPValue <- AccuracyPValue
  return(model_metrics)
}

# Function to extract metrics from 25 replicates, returning a data frame
replicated_CV_metrics_extraction<-function(CV_output_l){
list_precison<-list()
for(i in 1:25){
list_precison[i]<-lapply(CV_output_l[2,i], function(x)
  extract_accuracy(CV_output_l = x))
  }
list_precison

precison_df<-do.call(rbind, list_precison)
precison_df<-rownames_to_column(precison_df, var = "Sp_abb_name")
precison_df$Sp_abb_name<-as.factor(gsub(pattern = "[0-9]", replacement = "",  precison_df$Sp_abb_name))

precison_df<-left_join(x = precison_df, 
                       y = lineage_sp_df,
                       by = "Sp_abb_name" )

precison_df$Sp_abb_name<-factor(x = precison_df$Sp_abb_name, levels = levels(lineage_sp_df$Sp_abb_name))

return(precison_df)

}
```


## 10.3.2 - Extract Key CV Metrics of MeJA Models
```{r}
# Run the custom functions to extract accuracy, kappa, and p-values
RF_accuracy_MeJA_df <- lapply(CV_25replicated_MeJA_400k_l_l_33sp, replicated_CV_metrics_extraction)

# Get a table with summarized results
RF_accuracy_MeJA_df$replicated_model_400k_bac_MeJA %>%
  group_by(Sp_abb_name) %>%
  summarise(mean_accuracy = mean(accuracy),
            median_accuracy = median(accuracy),
            sd_accuracy = sd(accuracy))

# Remove MeJA models from environment
rm(CV_25replicated_MeJA_400k_l_l_33sp)
```


## 10.3.3 - Extract Key CV Metrics of SA Models
```{r}
# Run the custom functions to extract accuracy, kappa, and p-values
RF_accuracy_SA_df <- lapply(CV_25replicated_SA_400k_l_l_33sp, replicated_CV_metrics_extraction)

# Get a table with summarized results
RF_accuracy_SA_df$replicated_model_400k_bac_SA %>%
  group_by(Sp_abb_name) %>%
  summarise(mean_accuracy = mean(accuracy),
            median_accuracy = median(accuracy),
            sd_accuracy = sd(accuracy))

# Remove SA models from environment
rm(CV_25replicated_SA_400k_l_l_33sp)
```

### 10.3.4 -  Merge MeJA and SA Data Frames for a Single Data Frame
```{r}
# Add Stress column to each data frame
RF_accuracy_MeJA_df <- lapply(RF_accuracy_MeJA_df, function(x) {
  x$Stress <- "MeJA"
  return(x)
})
RF_accuracy_SA_df <- lapply(RF_accuracy_SA_df, function(x) {
  x$Stress <- "SA"
  return(x)
})

# Merge data frames
RF_accuracy_MeJASA_df <- mapply(function(x, y) rbind(x, y),
                                x = RF_accuracy_MeJA_df,
                                y = RF_accuracy_SA_df,
                                SIMPLIFY = FALSE)

# Add a kappa classification as a categorical result
RF_accuracy_MeJASA_df <- lapply(RF_accuracy_MeJASA_df, function(x)
  mutate(.data = x,
         kappa_class = factor(ifelse(
           kappa >= 0.81, "Almost_perfect",
           ifelse(kappa >= 0.61 & kappa < 0.81, "Substantial",
                  ifelse(kappa >= 0.41 & kappa < 0.61, "Moderate",
                         ifelse(kappa >= 0.21 & kappa < 0.41, "Fair",
                                ifelse(kappa > 0 & kappa < 0.21, "Slight", "No_agreement"))))),
           levels = c("Almost_perfect", "Substantial", "Moderate", "Fair", "Slight", "No_agreement"))
  ))

# Adjust p-values
RF_accuracy_MeJASA_df <- lapply(RF_accuracy_MeJASA_df, function(x) {
  x$AccuracyPValue.adjusted <- p.adjust(x$AccuracyPValue, method = "fdr")
  return(x)
})

# Calculate mean accuracy of valid models for barplot labels
RF_accuracy_MeJASA_df <- lapply(RF_accuracy_MeJASA_df, function(x)
  x %>%
    group_by(Sp_abb_name, Stress) %>%
    mutate(mean_accuracy_all_models = mean(accuracy)))

# Specify Bac/Fun data frames
RF_accuracy_MeJASA_df$replicated_model_400k_bac_MeJA$Kingdom <- "Bacteria"
RF_accuracy_MeJASA_df$replicated_model_400k_fun_MeJA$Kingdom <- "Fungi"
```



### 10.3.5 -  visualize model's precison, p and kappa

```{r}

# barplot showing classes of kappa agreement of valid models, number of models with pvalue for accuracy>0.05, and precision on bartop label
plot_RF_22sp_precison<-lapply(RF_accuracy_MeJASA_df, function (x)
  ggplot(filter(x, MY_tree_vector == "inside_MY_tree", AccuracyPValue.adjusted <0.057) , # adjusted p values of 0.566 include fungal models with a single misclassification. if we remove the species outside the MY, this won't be a problem
                                  aes(y = Sp_abb_name ))+
    geom_bar(position = "stack", aes(fill = kappa_class))+
    scale_y_discrete(limits=rev)+
    scale_fill_viridis(discrete = TRUE, direction = -1)+
    facet_wrap("Stress")+
     geom_text(stat = "count", aes(label = round(mean_accuracy_all_models, digits = 2)))+
    theme_bw())

#put bac and fun side by side
RF_precison_plot<-
ggarrange(plot_RF_22sp_precison$replicated_model_400k_bac_MeJA,
          plot_RF_22sp_precison$replicated_model_400k_fun_MeJA,
          ncol = 2,labels = c("Bacteria", "Fungi"),common.legend = TRUE, hjust = 0, legend="bottom")




# save plot externally
ggsave(RF_precison_plot,
       filename = "./10_Machine_learning/results/RF_22sp_kappa_precision_p_MeJa_SA.pdf",
       width = 20,
       height = 20,
       units = "cm")

```

