---
title: "2_beta_diversity_CLEAN"
author: "Zulema Carracedo Lorenzo"
date: "March 31, 2023"
output: html_document
editor_options: 
  chunk_output_type: console
---
## Load libraries

```{r}
#Load libraries
library(phyloseq)
library(microbiome)
library(vegan)
library(ggplot2)
library(tidyr)
library(dplyr)
library(ggpubr)
library(tibble)
library(car)
library(purrr)
library(umap) 
library(vegan)
library(spaa)
library(EcolUtils)
library(pairwiseAdonis)
```


## Load the full data

You will likely only need one type of normalization for the analysis tool you are using

```{r results="hide"}
# this will load the CSS-normalized 16S and ITS data as a list of two objects
load("./data/phyloseq_objects/CSS_BacFun_rhizoplane_HighFilt_ps_l.RData")

#Subset for the Brassica species that are included in the phylogenetic tree
CSS_BacFun_tree_l <- lapply(CSS_BacFun_rhizoplane_HighFilt_ps_l, function(x) 
                      subset_samples(x, MY_tree_presence == "inside_MY_tree"))
  
CSS_BacFun_tree_l <- lapply(CSS_BacFun_tree_l, function(x) prune_taxa(taxa_sums(x) >0,x)) #we need to prune the ASV that are not present in the subset

#Now we can remove the original phyloseq object
rm(CSS_BacFun_rhizoplane_HighFilt_ps_l)

phylogeny_vector <- c("Arabidopsis_thaliana", "Arabidopsis_lyrata", "Camelina_sativa", "Turritis_glabra","Erysium_cheirantoides","Malcomia_maritima", "Barbarea_vulgaris","Lepidium_ruderale", "Descurainia_sophiades", "Brassica_nigra", "Sinapis_arvensis", "Hirschfeldia_incana", "Brassica_rapa",  "Brassica_oleraceae_Rivera_Medium", "Diplotaxis_tenuifolia", "Sysimbrium_irio", "Isatis_tinctoria", "Cochlearia_officinalis", "Lobularia_maritima", "Iberis_amara", "Berteroa_incana", "Euclydium_syriacum") #needed for the plots

# add extended lineage as metadata
 
   CSS_BacFun_tree_l<- lapply(CSS_BacFun_tree_l, function(x){

  x@sam_data$Sp_Lineage_Walden_extended<-
    if_else(condition = x@sam_data$Sp_abb_name %in% c("Ia", "Lm", "Co", "Bi"), 
            true ="lineage_II_extended",
            false =  x@sam_data$Sp_Lineage_Walden)

    return(x)

  })  

# add supertribe informationa as metadara 
   
supertribe <- read.csv("./16_Plant_species_information/data/df_full_sp_info_clean.csv") #dataframe with supertribe info

#keep only the information needed
supertribe <- supertribe %>% select(Sp_abb_name, SuperTribe, Color_HEX)

# extract the current sample data
metadata_l <- list(Bac=(data.frame(CSS_BacFun_tree_l$Bac@sam_data)),           Fun=(data.frame(CSS_BacFun_tree_l$Fun@sam_data))) #list with dataframes from currecnt original sample data phyloseq for Bac and fun

#Merge both dataframes

merged_sample_data_l <- lapply(metadata_l, function(x){

  merge(
    x,
    supertribe,
  by = "Sp_abb_name",  # Adjust the column name as needed
  all.x = TRUE)  # Keeps all rows from the current_sample_data
  })  


# Set rownames back to sample IDs

merged_sample_data_l$Bac$Harvest_ID_2 <- paste(merged_sample_data_l$Bac$Harvest_ID, merged_sample_data_l$Bac$Target, sep = "_")

merged_sample_data_l$Fun$Harvest_ID_2 <- paste(merged_sample_data_l$Fun$Harvest_ID, merged_sample_data_l$Fun$Target, sep = "_")

rownames(merged_sample_data_l$Bac) <- merged_sample_data_l$Bac$Harvest_ID_2
rownames(merged_sample_data_l$Fun) <- merged_sample_data_l$Fun$Harvest_ID_2

# Update the phyloseq object
sample_data(CSS_BacFun_tree_l$Bac) <- sample_data(merged_sample_data_l$Bac)
sample_data(CSS_BacFun_tree_l$Fun) <- sample_data(merged_sample_data_l$Fun)

# Verify
sample_data(CSS_BacFun_tree_l$Bac)
sample_data(CSS_BacFun_tree_l$Fun)

```

## X.03 Subsets of data needed for the script

```{r}
CSS_Stress_ps_l <- lapply(CSS_BacFun_tree_l, function (x) 
  metagMisc::phyloseq_sep_variable(x, variable = "Stress")) #divides the phyloseq object according to the treatment (stress). We get a list of 2 objects (bacteria and fungi) each of the with 3 objects (3 stresses)

CSS_species_ps_l <- lapply(CSS_BacFun_tree_l, function (x) 
  metagMisc::phyloseq_sep_variable(x, variable = "Sp_full_name")) #divides the phyloseq object according to the treatment (stress). We get a list of 2 objects (bacteria and fungi) each of the with 3 objects (3 stresses)

CSS_Bac_ps <- CSS_BacFun_tree_l$Bac #only 16S samples

CSS_Fun_ps <- CSS_BacFun_tree_l$Fun #only ITS samples


```


# 2.1. Beta diversity plots

### 2.1.1. UMAP (Suplementary figure 6 C and D)

```{r}

# First transpose OTU table so that rows are samples and OTUs columns 

otu_table_BacFun_tree_l <- lapply(CSS_BacFun_tree_l, function(x)
                                t(otu_table(x)))

# Calculate dissimilarity indexes with bray curtis method 

bray_l <- lapply(otu_table_BacFun_tree_l, function (x)
    vegdist(x, method = "bray"))

bray_matrix_l <- lapply(bray_l, function (x)
    as.matrix(x))

# Run UMAP on bray curtis dissimilarity matrix

UMAP_bray_l <- lapply(bray_matrix_l, function (x)
      umap(x))

save(UMAP_bray_l, file = "./02_Beta_Diversity/r_objects/UMAP_bray_l.RData")
#load(file = "./02_Beta_Diversity/r_objects/UMAP_bray_l.RData")

head(UMAP_bray_l$Bac$layout, 3) #layout is the matrix in which we are interested

# Extract UMAP axis to test for phylogenetic signal

UMAP_axis_Bac <- UMAP_bray_l$Bac$layout
UMAP_axis_Fun <- UMAP_bray_l$Fun$layout

save(UMAP_axis_Bac, file = "./02_Beta_Diversity/r_objects/UMAP_axis_Bac.RData")
#load(file = "./02_Beta_Diversity/r_objects/Final_paper/UMAP_axis_Bac.RData")

save(UMAP_axis_Fun, file = "./02_Beta_Diversity/r_objects/UMAP_axis_Fun.RData")
#load(file = "./02_Beta_Diversity/r_objects/Final_paper/UMAP_axis_Fun.RData")

# plot umap

layoutBray <- lapply(UMAP_bray_l, function (x)
      as.data.frame(x$layout))

colnames(layoutBray$Bac) <-  c("UMAP1","UMAP2")
colnames(layoutBray$Fun) <-  c("UMAP1","UMAP2")


umap_meta_bac <- cbind(layoutBray$Bac, sample_data(CSS_BacFun_tree_l$Bac))
umap_meta_fun <- cbind(layoutBray$Fun, sample_data(CSS_BacFun_tree_l$Fun))

# we calculate centroids per species/treatment combination by averaging X and Y coordinates of all samples from that species/treatment

# group by species/treatment and calculate average UMAP1 and UMAP2 coordinates


umap_df_avarage_funtion <- function(x){

  umap_df_average <- x %>%

    group_by(Sp_abb_name2, Stress, Sp_Lineage_Walden_extended) %>%

    summarise(UMAP1 = mean(UMAP1), UMAP2 = mean(UMAP2))

  return(umap_df_average)

}


#For bacteria (Supplementary Figure 6-C)

umap_meta_bac <- umap_meta_bac %>%
                                        mutate(Sp_abb_name2 = 
                                        case_when(Sp_abb_name  == "Bo_M" ~ "Bo",
                                        TRUE ~ Sp_abb_name)) #rename Bo_M to Bo for the plots
  
umap_df_avarage_Bac <-  umap_df_avarage_funtion(umap_meta_bac) 
colnames(umap_df_avarage_Bac) <- c('Sp_abb_name2', 'Treatment', 'Sp_Lineage_Walden_extended', 'UMAP1', 'UMAP2') #rename "stress" to "treatment"


UMAP_plot_centroids_stress2_Bac <-ggplot(umap_df_avarage_Bac, aes(x=UMAP1,y=UMAP2, color=Treatment)) +
   scale_color_manual(values=c('#999999','#E69F00', '#009E73'))+
    ggtitle("UMAP - Bacteria") +
    geom_text(aes(label = Sp_abb_name2), size=12) +
    theme_bw() + 
    theme(text = element_text(size = 20)) 

  ggsave("UMAP_plot_centroids_stress_Bac.pdf", path = "./02_Beta_Diversity/results",
         width = 30, height = 20, units = "cm")
    ggsave("UMAP_plot_centroids_Bac.png", path = "./02_Beta_Diversity/results",
         width = 30, height = 20, units = "cm")

 
 #For fungi (Supplementary Figure 6-D)
umap_meta_fun <- umap_meta_fun %>%
                                        mutate(Sp_abb_name2 = 
                                        case_when(Sp_abb_name  == "Bo_M" ~ "Bo",
                                        TRUE ~ Sp_abb_name)) #rename Bo_M to Bo for the plots  
  
umap_df_avarage_Fun <-  umap_df_avarage_funtion(umap_meta_fun) 
colnames(umap_df_avarage_Fun) <- c('Sp_abb_name2', 'Treatment', 'Sp_Lineage_Walden_extended', 'UMAP1', 'UMAP2') #rename" stress " to "treatment"

  UMAP_plot_centroids_stress2_Fun <-ggplot(umap_df_avarage_Fun, aes(x=UMAP1,y=UMAP2, color=Treatment)) +
   scale_color_manual(values=c('#999999','#E69F00', '#009E73'))+
    ggtitle("UMAP - Fungi") +
    geom_text(aes(label = Sp_abb_name2), size=12) +
    theme_bw() + 
    theme(text = element_text(size = 20)) 

  ggsave("UMAP_plot_centroids_stress_Fun.pdf", path = "./02_Beta_Diversity/results",
         width = 30, height = 20, units = "cm")
   ggsave("UMAP_plot_centroids_stress_Fun.png", path = "./02_Beta_Diversity/results",
         width = 30, height = 20, units = "cm")

```


### 2.1.2. PCoA per supertribe (Figure 1 and Supplementary Figure S8 A, B, C, D E, F)

```{r}

# calculate PCoA ordination per stress

CSS_Stress_ps_l <- lapply(CSS_BacFun_tree_l, function (x) 
  metagMisc::phyloseq_sep_variable(x, variable = "Stress")) #divides the phyloseq object according to the treatment (stress). We get a list of 2 objects (bacteria and fungi) each of the with 3 objects (3 stresses)


  set.seed(300)
  PCoA_stress_BacFun_l <- lapply(CSS_Stress_ps_l, function (list_of_two)
    lapply(list_of_two, function (list_of_3)
          ordinate(list_of_3,
                   method="PCoA",
                   distance="bray",
                   try=200,
                   autotransform=TRUE)))
  
save(PCoA_stress_BacFun_l, file = "./02_Beta_Diversity/r_objects/PCoA_stress_BacFun_l.RData")
#load(file = "./02_Beta_Diversity/r_objects/PCoA_stress_BacFun_l.RData")

#extract variance for plots
  
 # Percentage of variance explained
variance_Bac_cntrl <- PCoA_stress_BacFun_l$Bac$Control$values$Relative_eig * 100
variance_Bac_MeJA <- PCoA_stress_BacFun_l$Bac$MeJA$values$Relative_eig * 100
variance_Bac_SA <- PCoA_stress_BacFun_l$Bac$SA$values$Relative_eig * 100

variance_Fun_cntrl <- PCoA_stress_BacFun_l$Fun$Control$values$Relative_eig * 100
variance_Fun_MeJA <- PCoA_stress_BacFun_l$Fun$MeJA$values$Relative_eig * 100
variance_Fun_SA <- PCoA_stress_BacFun_l$Fun$SA$values$Relative_eig * 100 


  ### Repeat so that we don't get the plots mirror ###
  

  set.seed(300)
  PCoA_BacFun_l <- lapply(CSS_BacFun_tree_l, function (list_of_two)
          ordinate(list_of_two,
                   method="PCoA",
                   distance="bray",
                   try=200,
                   autotransform=TRUE))
  
save(PCoA_BacFun_l, file = "./02_Beta_Diversity/r_objects/PCoA_BacFun_l.RData")
#load(file = "./02_Beta_Diversity/r_objects/PCoA_BacFun_l.RData")

# Coordinates of samples
coordinates_Bac <- data.frame(PCoA_BacFun_l$Bac$vectors)
colnames(coordinates_Bac)[1:2] <- c("PC1", "PC2")  # Adjust if using more axes

coordinates_Fun <- data.frame(PCoA_BacFun_l$Fun$vectors)
colnames(coordinates_Fun)[1:2] <- c("PC1", "PC2")  # Adjust if using more axes


# Combine coordinates with metadata
pcoa_data_Bac <- cbind(CSS_BacFun_tree_l$Bac@sam_data, coordinates_Bac)

pcoa_data_Fun <- cbind(CSS_BacFun_tree_l$Fun@sam_data, coordinates_Fun)


#ggplot

# Create the PCoA plots
colors <- c("Camelinodae_(I)"="#de70a8", #same colors as BrassiTol
            "Arabodae_(IV)"="#9474e1",
            "Hesperodae_(III)"="#5ab2eb",
            "Brassicodae_(II)"="#10a072",
            "Unplaced"="#e99c00",
            "Heliophilodae_(V)"="#e05a00")


  
  # If I represent the centroids and merge it with the   individual samples?


# divide by stress
pcoa_data_Bac_cntrl <- filter(pcoa_data_Bac, Stress == "Control") #Bac control
pcoa_data_Bac_MeJA <- filter(pcoa_data_Bac, Stress == "MeJA") #Bac MeJA
pcoa_data_Bac_SA <- filter(pcoa_data_Bac, Stress == "SA") #Bac SA

pcoa_data_Fun_cntrl <- filter(pcoa_data_Fun, Stress == "Control") #Fun control
pcoa_data_Fun_MeJA <- filter(pcoa_data_Fun, Stress == "MeJA") #Fun MeJA
pcoa_data_Fun_SA <- filter(pcoa_data_Fun, Stress == "SA") #Fun SA

#calculate centroids  
avg_pcoa_data_Bac_cntrl <- pcoa_data_Bac_cntrl %>% 
                            group_by(Sp_abb_name) %>%
                               summarise(mean_PC1 = mean(PC1), mean_PC2 = mean(PC2)) %>%
                                mutate(Stress = "Control")
  
avg_pcoa_data_Bac_MeJA <- pcoa_data_Bac_MeJA %>% 
                            group_by(Sp_abb_name) %>%
                               summarise(mean_PC1 = mean(PC1), mean_PC2 = mean(PC2)) %>%
                                mutate(Stress = "MeJA")

avg_pcoa_data_Bac_SA <- pcoa_data_Bac_SA %>% 
                            group_by(Sp_abb_name) %>%
                               summarise(mean_PC1 = mean(PC1), mean_PC2 = mean(PC2))  %>%
                                mutate(Stress = "SA")

avg_pcoa_data_Fun_cntrl <- pcoa_data_Fun_cntrl %>% 
                            group_by(Sp_abb_name) %>%
                               summarise(mean_PC1 = mean(PC1), mean_PC2 = mean(PC2)) %>%
                                mutate(Stress = "Control")
  
avg_pcoa_data_Fun_MeJA <- pcoa_data_Fun_MeJA %>% 
                            group_by(Sp_abb_name) %>%
                               summarise(mean_PC1 = mean(PC1), mean_PC2 = mean(PC2))  %>%
                                mutate(Stress = "MeJA")

avg_pcoa_data_Fun_SA <- pcoa_data_Fun_SA %>% 
                            group_by(Sp_abb_name) %>%
                               summarise(mean_PC1 = mean(PC1), mean_PC2 = mean(PC2))  %>%
                                mutate(Stress = "SA")



    # Bac control 

pcoa_data_Bac_cntrl_2 <- merge(pcoa_data_Bac_cntrl, avg_pcoa_data_Bac_cntrl, by = "Sp_abb_name")

# Replace "Lr" with "Lv" in the Sp_abb_name column
pcoa_data_Bac_cntrl_2 <- pcoa_data_Bac_cntrl_2 %>%
  mutate(Sp_abb_name = case_when(
    Sp_abb_name == "Lr" ~ "Lv",  # Replace "Lr" with "Lv"
    TRUE ~ Sp_abb_name           # Keep other values unchanged
  ))

# Replace "Bo_M" with "Bo" in the Sp_abb_name column
pcoa_data_Bac_cntrl_2 <- pcoa_data_Bac_cntrl_2 %>%
  mutate(Sp_abb_name = case_when(
    Sp_abb_name == "Bo_M" ~ "Bo",  # Replace "Bo_M" with "Bo"
    TRUE ~ Sp_abb_name           # Keep other values unchanged
  ))
# plot

pcoa_data_Bac_cntrl_2$SuperTribe <- factor(pcoa_data_Bac_cntrl_2$SuperTribe, levels = c("Camelinodae_(I)", "Brassicodae_(II)", "Hesperodae_(III)", "Arabodae_(IV)", "Heliophilodae_(V)", "Unplaced"))

p1 <- ggplot(pcoa_data_Bac_cntrl_2,aes(PC1,PC2,color=SuperTribe)) +
  geom_point(size=2, alpha=0.3) + 
  geom_text(data=pcoa_data_Bac_cntrl_2, aes(x=mean_PC1, y=mean_PC2,label= Sp_abb_name),size=10) +  
  geom_segment(aes(x=mean_PC1, y=mean_PC2, xend=PC1, yend=PC2), alpha = 0.3) +
  labs(
    x = paste0("PC1 (", round(variance_Bac_cntrl[1], 2), "%)"),
    y = paste0("PC2 (", round(variance_Bac_cntrl[2], 2), "%)"),
    title = "PCoA 16S control"
  ) +
  scale_color_manual(values=colors)+
    theme_bw() + 
    theme(text = element_text(size = 25))

ggsave("PCoA_SuperTribe_centroid_Bac_cntrl.png", path = "./02_Beta_Diversity/results",
         width = 30, height = 20, units = "cm")
 
  ggsave("PCoA_SuperTribe_centroid_Bac_cntrl.pdf", path = "./02_Beta_Diversity/results",
         width = 30, height = 20, units = "cm")
  

  
    # Bac MeJA 
pcoa_data_Bac_MeJA_2 <- merge(pcoa_data_Bac_MeJA, avg_pcoa_data_Bac_MeJA, by = "Sp_abb_name")

# Replace "Lr" with "Lv" in the Sp_abb_name column
pcoa_data_Bac_MeJA_2 <- pcoa_data_Bac_MeJA_2 %>%
  mutate(Sp_abb_name = case_when(
    Sp_abb_name == "Lr" ~ "Lv",  # Replace "Lr" with "Lv"
    TRUE ~ Sp_abb_name           # Keep other values unchanged
  ))

# Replace "Bo_M" with "Bo" in the Sp_abb_name column
pcoa_data_Bac_MeJA_2 <- pcoa_data_Bac_MeJA_2 %>%
  mutate(Sp_abb_name = case_when(
    Sp_abb_name == "Bo_M" ~ "Bo",  # Replace "Bo_M" with "Bo"
    TRUE ~ Sp_abb_name           # Keep other values unchanged
  ))

# plot
pcoa_data_Bac_MeJA_2$SuperTribe <- factor(pcoa_data_Bac_MeJA_2$SuperTribe, levels = c("Camelinodae_(I)", "Brassicodae_(II)", "Hesperodae_(III)", "Arabodae_(IV)", "Heliophilodae_(V)", "Unplaced"))

p2 <-ggplot(pcoa_data_Bac_MeJA_2,aes(PC1,PC2,color=SuperTribe)) +
  geom_point(size=2, alpha=0.3) + 
  geom_text(data=pcoa_data_Bac_MeJA_2, aes(x=mean_PC1, y=mean_PC2,label= Sp_abb_name),size=5) +  
  geom_segment(aes(x=mean_PC1, y=mean_PC2, xend=PC1, yend=PC2), alpha = 0.3) +
  labs(
    x = paste0("PC1 (", round(variance_Bac_MeJA[1], 2), "%)"),
    y = paste0("PC2 (", round(variance_Bac_MeJA[2], 2), "%)"),
    title = "PCoA 16S MeJA"
  ) +
  scale_color_manual(values=colors)+
    theme_bw() + 
    theme(text = element_text(size = 20))

ggsave("PCoA_SuperTribe_centroid_Bac_MeJA.png", path = "./02_Beta_Diversity/results",
         width = 30, height = 20, units = "cm")
 
  ggsave("PCoA_SuperTribe_centroid_Bac.pdf", path = "./02_Beta_Diversity/results",
         width = 30, height = 20, units = "cm")    
  

      # Bac SA 
pcoa_data_Bac_SA_2 <- merge(pcoa_data_Bac_SA, avg_pcoa_data_Bac_SA, by = "Sp_abb_name")

# Replace "Lr" with "Lv" in the Sp_abb_name column
pcoa_data_Bac_SA_2 <- pcoa_data_Bac_SA_2 %>%
  mutate(Sp_abb_name = case_when(
    Sp_abb_name == "Lr" ~ "Lv",  # Replace "Lr" with "Lv"
    TRUE ~ Sp_abb_name           # Keep other values unchanged
  ))

# Replace "Bo_M" with "Bo" in the Sp_abb_name column
pcoa_data_Bac_SA_2 <- pcoa_data_Bac_SA_2 %>%
  mutate(Sp_abb_name = case_when(
    Sp_abb_name == "Bo_M" ~ "Bo",  # Replace "Bo_M" with "Bo"
    TRUE ~ Sp_abb_name           # Keep other values unchanged
  ))

# plot
pcoa_data_Bac_SA_2$SuperTribe <- factor(pcoa_data_Bac_SA_2$SuperTribe, levels = c("Camelinodae_(I)", "Brassicodae_(II)", "Hesperodae_(III)", "Arabodae_(IV)", "Heliophilodae_(V)", "Unplaced"))

p3 <- ggplot(pcoa_data_Bac_SA_2,aes(PC1,PC2,color=SuperTribe)) +
  geom_point(size=2, alpha=0.3) + 
  geom_text(data=pcoa_data_Bac_SA_2, aes(x=mean_PC1, y=mean_PC2,label= Sp_abb_name),size=10) +  
  geom_segment(aes(x=mean_PC1, y=mean_PC2, xend=PC1, yend=PC2), alpha = 0.3) +
  labs(
    x = paste0("PC1 (", round(variance_Bac_SA[1], 2), "%)"),
    y = paste0("PC2 (", round(variance_Bac_SA[2], 2), "%)"),
    title = "PCoA 16S SA"
  ) +
  scale_color_manual(values=colors)+
    theme_bw() + 
    theme(text = element_text(size = 25))

ggsave("PCoA_SuperTribe_centroid_Bac_SA.png", path = "./02_Beta_Diversity/results",
         width = 30, height = 20, units = "cm")
 
  ggsave("PCoA_SuperTribe_centroid_Bac_SA.pdf", path = "./02_Beta_Diversity/results",
         width = 30, height = 20, units = "cm")    
  
   #Fun control
  
pcoa_data_Fun_cntrl_2 <- merge(pcoa_data_Fun_cntrl, avg_pcoa_data_Fun_cntrl, by = "Sp_abb_name")

# Replace "Lr" with "Lv" in the Sp_abb_name column
pcoa_data_Fun_cntrl_2 <- pcoa_data_Fun_cntrl_2 %>%
  mutate(Sp_abb_name = case_when(
    Sp_abb_name == "Lr" ~ "Lv",  # Replace "Lr" with "Lv"
    TRUE ~ Sp_abb_name           # Keep other values unchanged
  ))

# Replace "Bo_M" with "Bo" in the Sp_abb_name column
pcoa_data_Fun_cntrl_2 <- pcoa_data_Fun_cntrl_2 %>%
  mutate(Sp_abb_name = case_when(
    Sp_abb_name == "Bo_M" ~ "Bo",  # Replace "Bo_M" with "Bo"
    TRUE ~ Sp_abb_name           # Keep other values unchanged
  ))

# plot  
pcoa_data_Fun_cntrl_2$SuperTribe <- factor(pcoa_data_Fun_cntrl_2$SuperTribe, levels = c("Camelinodae_(I)", "Brassicodae_(II)", "Hesperodae_(III)", "Arabodae_(IV)", "Heliophilodae_(V)", "Unplaced"))

p4 <- ggplot(pcoa_data_Fun_cntrl_2,aes(PC1,PC2,color=SuperTribe)) +
  geom_point(size=2, alpha=0.3) + 
  geom_text(data=pcoa_data_Fun_cntrl_2, aes(x=mean_PC1, y=mean_PC2,label= Sp_abb_name),size=5) +  
  geom_segment(aes(x=mean_PC1, y=mean_PC2, xend=PC1, yend=PC2), alpha = 0.3) +
  labs(
    x = paste0("PC1 (", round(variance_Fun_cntrl[1], 2), "%)"),
    y = paste0("PC2 (", round(variance_Fun_cntrl[2], 2), "%)"),
    title = "PCoA ITS control"
  ) +
  scale_color_manual(values=colors)+
    theme_bw() + 
    theme(text = element_text(size = 20))

ggsave("PCoA_SuperTribe_centroid_Fun_cntrl.png", path = "./02_Beta_Diversity/results",
         width = 30, height = 20, units = "cm")
 
  ggsave("PCoA_SuperTribe_centroid_Fun_cntrl.pdf", path = "./02_Beta_Diversity/results",
         width = 30, height = 20, units = "cm")

  
    # Fun MeJA 
pcoa_data_Fun_MeJA_2 <- merge(pcoa_data_Fun_MeJA, avg_pcoa_data_Fun_MeJA, by = "Sp_abb_name")

# Replace "Lr" with "Lv" in the Sp_abb_name column
pcoa_data_Fun_MeJA_2 <- pcoa_data_Fun_MeJA_2 %>%
  mutate(Sp_abb_name = case_when(
    Sp_abb_name == "Lr" ~ "Lv",  # Replace "Lr" with "Lv"
    TRUE ~ Sp_abb_name           # Keep other values unchanged
  ))

# Replace "Bo_M" with "Bo" in the Sp_abb_name column
pcoa_data_Fun_MeJA_2 <- pcoa_data_Fun_MeJA_2 %>%
  mutate(Sp_abb_name = case_when(
    Sp_abb_name == "Bo_M" ~ "Bo",  # Replace "Bo_M" with "Bo"
    TRUE ~ Sp_abb_name           # Keep other values unchanged
  ))

# plot
pcoa_data_Fun_MeJA_2$SuperTribe <- factor(pcoa_data_Fun_MeJA_2$SuperTribe, levels = c("Camelinodae_(I)", "Brassicodae_(II)", "Hesperodae_(III)", "Arabodae_(IV)", "Heliophilodae_(V)", "Unplaced"))

p5 <- ggplot(pcoa_data_Fun_MeJA_2,aes(PC1,PC2,color=SuperTribe)) +
  geom_point(size=2, alpha=0.3) + 
  geom_text(data=pcoa_data_Fun_MeJA_2, aes(x=mean_PC1, y=mean_PC2,label= Sp_abb_name),size=5) +  
  geom_segment(aes(x=mean_PC1, y=mean_PC2, xend=PC1, yend=PC2), alpha = 0.3) +
  labs(
    x = paste0("PC1 (", round(variance_Fun_MeJA[1], 2), "%)"),
    y = paste0("PC2 (", round(variance_Fun_MeJA[2], 2), "%)"),
    title = "PCoA ITS MeJA"
  ) +
  scale_color_manual(values=colors)+
    theme_bw() + 
    theme(text = element_text(size = 20))

ggsave("PCoA_SuperTribe_centroid_Fun_MeJA.png", path = "./02_Beta_Diversity/results",
         width = 30, height = 20, units = "cm")
 
  ggsave("PCoA_SuperTribe_centroid_Fun_MeJA.pdf", path = "./02_Beta_Diversity/results",
         width = 30, height = 20, units = "cm")    
  

      # Fun SA 
pcoa_data_Fun_SA_2 <- merge(pcoa_data_Fun_SA, avg_pcoa_data_Fun_SA, by = "Sp_abb_name")

# Replace "Lr" with "Lv" in the Sp_abb_name column
pcoa_data_Fun_SA_2 <- pcoa_data_Fun_SA_2 %>%
  mutate(Sp_abb_name = case_when(
    Sp_abb_name == "Lr" ~ "Lv",  # Replace "Lr" with "Lv"
    TRUE ~ Sp_abb_name           # Keep other values unchanged
  ))

# Replace "Bo_M" with "Bo" in the Sp_abb_name column
pcoa_data_Fun_MeJA_2 <- pcoa_data_Fun_MeJA_2 %>%
  mutate(Sp_abb_name = case_when(
    Sp_abb_name == "Bo_M" ~ "Bo",  # Replace "Bo_M" with "Bo"
    TRUE ~ Sp_abb_name           # Keep other values unchanged
  ))

# plot
pcoa_data_Fun_SA_2$SuperTribe <- factor(pcoa_data_Fun_SA_2$SuperTribe, levels = c("Camelinodae_(I)", "Brassicodae_(II)", "Hesperodae_(III)", "Arabodae_(IV)", "Heliophilodae_(V)", "Unplaced"))

p6 <- ggplot(pcoa_data_Fun_SA_2,aes(PC1,PC2,color=SuperTribe)) +
  geom_point(size=2, alpha=0.3) + 
  geom_text(data=pcoa_data_Fun_SA_2, aes(x=mean_PC1, y=mean_PC2,label= Sp_abb_name),size=5) +  
  geom_segment(aes(x=mean_PC1, y=mean_PC2, xend=PC1, yend=PC2), alpha = 0.3) +
  labs(
    x = paste0("PC1 (", round(variance_Fun_SA[1], 2), "%)"),
    y = paste0("PC2 (", round(variance_Fun_SA[2], 2), "%)"),
    title = "PCoA ITS SA"
  ) +
  scale_color_manual(values=colors)+
    theme_bw() + 
    theme(text = element_text(size = 20))

ggsave("PCoA_SuperTribe_centroid_Fun_SA.png", path = "./02_Beta_Diversity/results",
         width = 30, height = 20, units = "cm")
 
  ggsave("PCoA_SuperTribe_centroid_Fun_SA.pdf", path = "./02_Beta_Diversity/results",
         width = 30, height = 20, units = "cm")    
  
#Merge all the plot together 
  library(patchwork)

p1 <- p1 + theme(legend.position = "none")
p2 <- p2 + theme(legend.position = "none")
p3 <- p3 + theme(legend.position = "none")
p4 <- p4 + theme(legend.position = "none")
p5 <- p5 + theme(legend.position = "none")
p6 <- p6 + theme(legend.position = "none")
    
p_merged <- (p1 | p4) /
            (p2 | p5) /
            (p3 | p6)   #figure used for Supplementary SX7

ggsave("PCoA_SuperTribe_centroid_all.png", path = "./02_Beta_Diversity/results",
         width = 35, height = 45, units = "cm")
 
  ggsave("PCoA_SuperTribe_centroid_all.pdf", path = "./02_Beta_Diversity/results",
         width = 35, height = 45, units = "cm")   
  
  
  
  
#adding the PC1 values to the x axis
  
p_Bac_cntrl_vs_SA <- (p1 | p3) #plots used for new Figure 1

ggsave("PCoA_SuperTribe_centroid_Bac_cntrl_vs_SA.png", path = "./02_Beta_Diversity/results",
         width = 55, height =25 ,units = "cm")
 
  ggsave("PCoA_SuperTribe_centroid_Bac_cntrl_vs_SA.pdf", path = "./02_Beta_Diversity/results",
         width = 55, height = 25, units = "cm")   
  
  #control bac
  p1_B<-
  ggplot(pcoa_data_Bac_cntrl_2,aes(PC1,PC2,color=SuperTribe)) +
  geom_point(size=2, alpha=0.3) + 
    geom_point(aes(x=mean_PC1,y=-Inf), size = 5) +
  coord_cartesian(clip = 'off') +
  geom_text(data=pcoa_data_Bac_cntrl_2, aes(x=mean_PC1, y=mean_PC2,label= Sp_abb_name),size=5) +  
  geom_segment(aes(x=mean_PC1, y=mean_PC2, xend=PC1, yend=PC2), alpha = 0.3) +

  labs(
    x = paste0("PC1 (", round(variance_Bac_cntrl[1], 2), "%)"),
    y = paste0("PC2 (", round(variance_Bac_cntrl[2], 2), "%)"),
    title = "PCoA 16S control"
  ) +
  scale_color_manual(values=colors)+
    theme_bw() + 
    theme(text = element_text(size = 20))
  
ggsave("PCoA_SuperTribe_centroid_Bac_cntrl_PC1axis.png", path = "./02_Beta_Diversity/results",
         width = 30, height = 20, units = "cm")
 
  ggsave("PCoA_SuperTribe_centroid_Bac_cntrl_PC1axis.pdf", path = "./02_Beta_Diversity/results",
         width = 30, height = 20, units = "cm")
  
  #SA bac
  p3_B <- 
  ggplot(pcoa_data_Bac_SA_2,aes(PC1,PC2,color=SuperTribe)) +
  geom_point(size=2, alpha=0.3) + 
    geom_point(aes(x=mean_PC1,y=-Inf), size = 5) +
  coord_cartesian(clip = 'off') +
  geom_text(data=pcoa_data_Bac_SA_2,aes(x=mean_PC1, y=mean_PC2,label= Sp_abb_name),size=5) +  
  geom_segment(aes(x=mean_PC1, y=mean_PC2, xend=PC1, yend=PC2), alpha = 0.3) +

  labs(
    x = paste0("PC1 (", round(variance_Bac_SA[1], 2), "%)"),
    y = paste0("PC2 (", round(variance_Bac_SA[2], 2), "%)"),
    title = "PCoA 16S SA"
  ) +
  scale_color_manual(values=colors)+
    theme_bw() + 
    theme(text = element_text(size = 20))
  
ggsave("PCoA_SuperTribe_centroid_Bac_SA_PC1axis.png", path = "./02_Beta_Diversity/results",
         width = 30, height = 20, units = "cm")
 
  ggsave("PCoA_SuperTribe_centroid_Bac_SA_PC1axis.pdf", path = "./02_Beta_Diversity/results",
         width = 30, height = 20, units = "cm")  
  
  #join control and SA plots
  
  p1_B <- p1_B + theme(legend.position = "none")
  p3_B <- p3_B + theme(legend.position = "none")
  (p1_B | p3_B) #figure used to get the dots for Axis 1
  
  ggsave("PCoA_SuperTribe_centroid_Bac_cntrl_vs_SA_PC1axis.png", path = "./02_Beta_Diversity/results",
         width = 55, height =25 ,units = "cm")
 
  ggsave("PCoA_SuperTribe_centroid_Bac_cntrl_vs_SA_PC1axis.pdf", path = "./02_Beta_Diversity/results",
         width = 55, height = 25, units = "cm")  
  
 

  
```
 


# 2.3. Permanovas

### 2.3.1. PERMANOVAS all samples (Supplementary table S4 A and B)
```{r}
#testing permanovas 16S 
CSS_Bac_ps <- CSS_BacFun_tree_l$Bac

set.seed(303848)
permanova_Bac_all <- adonis2(formula = phyloseq::distance(t(otu_table(CSS_Bac_ps)), method="bray") ~ Sp_full_name  + Stress + Sp_full_name*Stress + Block,
              data = as(sample_data(CSS_Bac_ps),"data.frame"), # changing with as.data.frame is insufficient
              permutations = 999, by = "terms")

save(permanova_Bac_all, 
     file = "./02_Beta_Diversity/r_objects/permanova_Bac_all.RData")


#testing permanovas ITS 
CSS_Fun_ps <- CSS_BacFun_tree_l$Fun

set.seed(303848)
permanova_Fun_all <- adonis2(formula = phyloseq::distance(t(otu_table(CSS_Fun_ps)), method="bray") ~ Sp_full_name  + Stress + Sp_full_name*Stress + Block,
              data = as(sample_data(CSS_Fun_ps),"data.frame"), # changing with as.data.frame is insufficient
              permutations = 999, by = "terms")

save(permanova_Fun_all, 
     file = "./02_Beta_Diversity/r_objects/permanova_Fun_all.RData")

```

### 2.3.2. Permanova per plant species (Supplementary Figure S6-B)
Here we check the pairwise effect of the treatment (control vs MeJA, control vs SA) in eachplant species

#### For 16S samples

```{r}

# list of lists per stress
CSS_species_ps_l #input list 

set.seed(303848)
permanova_Bac_results <- lapply(CSS_species_ps_l$Bac, function (x)
       adonis2(formula = phyloseq::distance(t(otu_table(x)), method="bray") 
               ~ Stress  + Block,
              data = as(sample_data(x),"data.frame"), # changing with as.data.frame is insufficient
              permutations = 999,  by = "margin")) #when we use by="margin" order of factors doesn't matter

#Save permanova results as an external object 
save(permanova_Bac_results, 
     file = "./02_Beta_Diversity/r_objects/permanova_Bac_per_species.RData")

#load("./02_Beta_Diversity/r_objectspermanova_Bac_per_species.RData")


# Pairwise PERMANOVA 16S Stress factor
set.seed(303848)
pairwise_permanova_bac <- lapply(CSS_species_ps_l$Bac, function (x)
  adonis.pair(dist.mat= phyloseq::distance(otu_table(x), method="bray"),
              Factor= as.factor(as(phyloseq::sample_data(x),"data.frame")$Stress)))

save(pairwise_permanova_bac, 
     file = "./02_Beta_Diversity/results/pairwise_permanova_bac.RData")

#load("./02_Beta_Diversity/results/pairwise_permanova_bac.RData")

#Extracting the p-values

p_values_pairwise <- function(pairwise_permanova_bac)
  
{
  
  # First we select the corrected pvalues of the pairwise analysis
  
  p_values <-map(pairwise_permanova_bac, 7)
  r_values <-map(pairwise_permanova_bac, 5)
  species_name <- names(c(p_values))
  
  # We create a data frame for each pairwise comparison
  
  Cntrl_MeJA_pvalue <- as.vector(unlist(map(p_values,1)))
  Cntrl_MeJA_R2 <-  as.vector(unlist(map(r_values,1)))
  pair1 <- rep("Cntrl vs MeJA")
  df_1 <- data.frame(species_name, pair1, Cntrl_MeJA_pvalue, Cntrl_MeJA_R2)
  names(df_1)<-c("species","treatment", "corrected_pvalue", "R2")
  
  
  Cntrl_SA_pvalue <- as.vector(unlist(map(p_values,2)))
  Cntrl_SA_R2 <- as.vector(unlist(map(r_values,2)))
  pair2 <- rep("Cntrl vs SA")
  df_2 <- data.frame(species_name, pair2, Cntrl_SA_pvalue, Cntrl_SA_R2)
  names(df_2)<-c("species","treatment", "corrected_pvalue", "R2")
  
  
  MeJA_SA_pvalue <- as.vector(unlist(map(p_values,3)))
  MeJA_SA_rvalue <- as.vector(unlist(map(r_values,3)))
  pair3 <- rep("MeJA vs SA")
  df_3 <- data.frame(species_name, pair3, MeJA_SA_pvalue, MeJA_SA_rvalue)
  names(df_3)<-c("species","treatment", "corrected_pvalue", "R2")
  
# We merge the 3 data frames in a single one
  
  df_list <- list(df_1, df_2, df_3)
  
  output_df <-Reduce(function(x, y) merge(x, y, all=TRUE), df_list)

  print(output_df)
}

# Pairwise p-values and plots for 16S

p_values_pairwise_bac_df <-  p_values_pairwise(pairwise_permanova_bac)

p_values_pairwise_bac_df <- p_values_pairwise_bac_df %>%
                                        mutate(Significance = 
                                        case_when(corrected_pvalue <= 0.001 ~ "< 0.001",
                                        corrected_pvalue <= 0.01 ~ "< 0.01",
                                        corrected_pvalue <= 0.05 ~ "< 0.05",
                                        TRUE ~ as.character(corrected_pvalue)))


p_values_pairwise_bac_df <- p_values_pairwise_bac_df%>%
    arrange(match(species, phylogeny_vector)) 

write.csv(p_values_pairwise_bac_df, "./02_Beta_diversity/results/p_values_pairwise_bac_df.csv", row.names=FALSE)

p_values_pairwise_MeJA_bac_df <- p_values_pairwise_bac_df %>% filter(treatment == "Cntrl vs MeJA")
write.csv(p_values_pairwise_MeJA_bac_df, "./02_Beta_diversity/results/p_values_pairwise_bac_df_MeJA.csv", row.names=FALSE)

p_values_pairwise_SA_bac_df <- p_values_pairwise_bac_df %>% filter(treatment == "Cntrl vs SA")
write.csv(p_values_pairwise_SA_bac_df, "./02_Beta_diversity/results/p_values_pairwise_bac_df_SA.csv", row.names=FALSE)

```

#### For ITS samples

```{r}
# list of lists per stress
CSS_species_ps_l #input list 

# Pairwise PERMANOVA ITS Stress factor
set.seed(303848)
pairwise_permanova_fun <- lapply(CSS_species_ps_l$Fun, function (x)
  adonis.pair(dist.mat= phyloseq::distance(otu_table(x), method="bray"),
              Factor= as.factor(as(phyloseq::sample_data(x),"data.frame")$Stress)))

save(pairwise_permanova_fun, 
     file = "./02_Beta_Diversity/results/pairwise_permanova_fun.RData")

#load("./02_Beta_Diversity/results/pairwise_permanova_fun.RData")


# Pairwise p-values and plots for 16S

p_values_pairwise_fun_df <-  p_values_pairwise(pairwise_permanova_fun)

p_values_pairwise_fun_df <- p_values_pairwise_fun_df %>%
                                        mutate(Significance = 
                                        case_when(corrected_pvalue <= 0.001 ~ "< 0.001",
                                        corrected_pvalue <= 0.01 ~ "< 0.01",
                                        corrected_pvalue <= 0.05 ~ "< 0.05",
                                        TRUE ~ as.character(corrected_pvalue)))

p_values_pairwise_fun_df <- p_values_pairwise_fun_df%>%
    arrange(match(species, phylogeny_vector))

write.csv(p_values_pairwise_fun_df, "./02_Beta_diversity/results/p_values_pairwise_fun_df.csv", row.names=FALSE)

p_values_pairwise_MeJA_fun_df <- p_values_pairwise_fun_df %>% filter(treatment == "Cntrl vs MeJA")
write.csv(p_values_pairwise_MeJA_fun_df, "./02_Beta_diversity/results/p_values_pairwise_fun_df_MeJA.csv", row.names=FALSE)

p_values_pairwise_SA_fun_df <- p_values_pairwise_fun_df %>% filter(treatment == "Cntrl vs SA")
write.csv(p_values_pairwise_SA_fun_df, "./02_Beta_diversity/results/p_values_pairwise_fun_df_SA.csv", row.names=FALSE)



```


### 2.3.3.PERMANOVA per stess (Statistics supplementary Figure S8)

```{r}

# Here we want to test the effect of the pant species in each treatment 
CSS_Stress_ps_l #we need this CSS object

#Permanova 16S control 
set.seed(303848)
permanova_Bac_control <- adonis2(formula = phyloseq::distance(t(otu_table(CSS_Stress_ps_l$Bac$Control)), method="bray") ~ Sp_full_name,
              data = as(sample_data(CSS_Stress_ps_l$Bac$Control),"data.frame"), # changing with as.data.frame is insufficient
              permutations = 999, by = "margin")

save(permanova_Bac_control, 
     file = "./02_Beta_Diversity/r_objects/permanova_Bac_control.RData") #supplementary Figure S8-A

#load("./02_Beta_Diversity/r_objects/permanova_Bac_control.RData")

table <- nice_table(permanova_Bac_control,
              title = c("Table X. PERMANOVA of Bray curtis distances of 16S samples in control treatment"),
              note = "~ Pant species")

table <- table %>%
         flextable::colformat_num(big.mark = "", digits = 3) 

flextable::save_as_docx(table, path = "./02_Beta_Diversity/results/permanova_Bac_control.docx") #supplementary Figure S8-A


#Permanova 16S MeJA 
set.seed(303848)
permanova_Bac_MeJA <- adonis2(formula = phyloseq::distance(t(otu_table(CSS_Stress_ps_l$Bac$MeJA)), method="bray") ~ Sp_full_name,
              data = as(sample_data(CSS_Stress_ps_l$Bac$MeJA),"data.frame"), # changing with as.data.frame is insufficient
              permutations = 999, by = "margin")

save(permanova_Bac_MeJA, 
     file = "./02_Beta_Diversity//r_objects/permanova_Bac_MeJA.RData")

#load("./02_Beta_Diversity/r_objects/permanova_Bac_MeJA.RData")

table <- nice_table(permanova_Bac_MeJA,
              title = c("Table X. PERMANOVA of Bray curtis distances of 16S samples in MeJA treatment"),
              note = "~ Pant species")

table <- table %>%
         flextable::colformat_num(big.mark = "", digits = 3) 

flextable::save_as_docx(table, path = "./02_Beta_Diversity/results//permanova_Bac_MeJA.docx") #supplementary Figure S8-B



#Permanova 16S SA 
set.seed(303848)
permanova_Bac_SA <- adonis2(formula = phyloseq::distance(t(otu_table(CSS_Stress_ps_l$Bac$SA)), method="bray") ~ Sp_full_name,
              data = as(sample_data(CSS_Stress_ps_l$Bac$SA),"data.frame"), # changing with as.data.frame is insufficient
              permutations = 999, by = "margin")

save(permanova_Bac_SA, 
     file = "./02_Beta_Diversity/r_objects/permanova_Bac_SA.RData")

#load("./02_Beta_Diversity/r_objects/permanova_Bac_SA.RData")

table <- nice_table(permanova_Bac_SA,
              title = c("Table X. PERMANOVA of Bray curtis distances of 16S samples in SA treatment"),
              note = "~ Pant species")

table <- table %>%
         flextable::colformat_num(big.mark = "", digits = 3) 

flextable::save_as_docx(table, path = "./02_Beta_Diversity/results/permanova_Bac_SA.docx") #supplementary Figure S8-C


#Permanova ITS control 
set.seed(303848)
permanova_fUN_control <- adonis2(formula = phyloseq::distance(t(otu_table(CSS_Stress_ps_l$Fun$Control)), method="bray") ~ Sp_full_name,
              data = as(sample_data(CSS_Stress_ps_l$Fun$Control),"data.frame"), # changing with as.data.frame is insufficient
              permutations = 999, by = "margin")

save(permanova_fUN_control, 
     file = "./02_Beta_Diversity/r_objects/permanova_Fun_control.RData")

#load("./02_Beta_Diversity/r_objects/permanova_Fun_control.RData")

table <- nice_table(permanova_Fun_control,
              title = c("Table X. PERMANOVA of Bray curtis distances of ITS samples in control treatment"),
              note = "~ Pant species")

table <- table %>%
         flextable::colformat_num(big.mark = "", digits = 3) 

flextable::save_as_docx(table, path = "./02_Beta_Diversity/results/permanova_Fun_control.docx") #supplementary Figure S8-D



#Permanova ITS MeJA 
set.seed(303848)
permanova_Fun_MeJA <- adonis2(formula = phyloseq::distance(t(otu_table(CSS_Stress_ps_l$Fun$MeJA)), method="bray") ~ Sp_full_name,
              data = as(sample_data(CSS_Stress_ps_l$Fun$MeJA),"data.frame"), # changing with as.data.frame is insufficient
              permutations = 999, by = "margin")

save(permanova_Fun_MeJA, 
     file = "./02_Beta_Diversity/r_objects/permanova_Fun_MeJA.RData")

#load("./02_Beta_Diversity/r_objects/permanova_Fun_MeJA.RData")

table <- nice_table(permanova_Fun_MeJA,
              title = c("Table X. PERMANOVA of Bray curtis distances of ITS samples in MeJA treatment"),
              note = "~ Pant species")

table <- table %>%
         flextable::colformat_num(big.mark = "", digits = 3) 

flextable::save_as_docx(table, path = "./02_Beta_Diversity/results/permanova_Fun_MeJA.docx") #supplementary Figure S8-E

#Permanova ITS SA 
set.seed(303848)
permanova_Fun_SA <- adonis2(formula = phyloseq::distance(t(otu_table(CSS_Stress_ps_l$Fun$SA)), method="bray") ~ Sp_full_name,
              data = as(sample_data(CSS_Stress_ps_l$Fun$SA),"data.frame"), # changing with as.data.frame is insufficient
              permutations = 999, by = "margin")

save(permanova_Fun_SA, 
     file = "./02_Beta_Diversity/r_objects/permanova_Fun_SA.RData")

#load("./02_Beta_Diversity/r_objects/permanova_Fun_SA.RData")


table <- nice_table(permanova_Fun_SA,
              title = c("Table X. PERMANOVA of Bray curtis distances of ITS samples in SA treatment"),
              note = "~ Pant species")

table <- table %>%
         flextable::colformat_num(big.mark = "", digits = 3) 

flextable::save_as_docx(table, path = "./02_Beta_Diversity/results//permanova_Fun_SA.docx") #supplementary Figure S8-F




```


\`\`\`
