---
title: "2_five_families_beta_diversity"
author: "Melissa"
date: "2026-05-01"
output: html_document
editor_options: 
  chunk_output_type: console
---

#Load libraries
```{r}
library(phyloseq)
library(vegan)
library(ggplot2)
library(tidyr)
library(dplyr)

set.seed(33)
theme_set(theme_bw())
```

#Load the CSS-normalized phyloseq object and filter to only the Phylogeny Brassicaceae species and the guest species
##Load the pbject and extract the names of the Guest species
```{r}
# this will load the CSS-normalized 16S and ITS data as a list of two objects
load("./data/phyloseq_objects/CSS_BacFun_rhizoplane_HighFilt_ps_l.RData")

#Extract the names of the 5 guest species
sam_data <- CSS_BacFun_rhizoplane_HighFilt_ps_l$Bac@sam_data
Guest_sp <- subset(sam_data, Sp_Guest == "yes")
Guest_sp_vector <- unique(Guest_sp$Sp_full_name)

Guest_sp_vector

#Remove Brassica_oleraceae_Rivera_Slow and Arabidopsis_thaliana
Guest_sp_vector <- Guest_sp_vector[!(Guest_sp_vector %in% c("Brassica_oleraceae_Rivera_Slow", "Arabidopsis_thaliana"))]

print(Guest_sp_vector)

#Extract the phylogeny Brassicaceae
Phylogeny <- subset(sam_data, MY_tree_presence == "inside_MY_tree")
Phylogeny_vector <- unique(Phylogeny$Sp_full_name)

Phylogeny_vector

#Merge the vectors
Phylogeny_Guest_vector <- c(Phylogeny_vector, Guest_sp_vector)
as.vector(Phylogeny_Guest_vector)


CSS_BacFun_tree_Guests_l <- lapply(CSS_BacFun_rhizoplane_HighFilt_ps_l, function(x) 
                      subset_samples(x, Sp_full_name %in% Phylogeny_Guest_vector))

#Check that you have 22 + 7 species
length(unique(CSS_BacFun_tree_Guests_l$Bac@sam_data$Sp_full_name))

#29 is right!
```

#Calculate and store Bray-Curtis dissimilarity indices
```{r}
physeq_bray_l <- lapply(CSS_BacFun_tree_Guests_l, function(x) 
  phyloseq::distance(x, method = "bray"))
```

###Export
```{r}
save(physeq_bray_l, file = "C:/RProjects/Family_experiment/02_Beta_diversity/r_objects/Guest_species/physeq_bray_l.RData")
```

#Calculate adonis
##Species and stress
```{r}
#Load the bray-curtis matrix
load("C:/RProjects/Family_experiment/02_Beta_diversity/r_objects/Guest_species/physeq_bray_l.RData")
adonis_bray_l <- mapply(function(x, y) {
  adonis2(x ~ sample_data(y)$Sp_full_name * sample_data(y)$Stress)
},
x = physeq_bray_l,
y = CSS_BacFun_tree_Guests_l,
SIMPLIFY = FALSE)

adonis_bray_l
```

> adonis_bray_l
$Bac
Permutation test for adonis under reduced model
Terms added sequentially (first to last)
Permutation: free
Number of permutations: 999

adonis2(formula = x ~ sample_data(y)$Sp_full_name * sample_data(y)$Stress)
                                                    Df SumOfSqs      R2
sample_data(y)$Sp_full_name                         28   70.077 0.27339
sample_data(y)$Stress                                2    0.620 0.00242
sample_data(y)$Sp_full_name:sample_data(y)$Stress   56   10.281 0.04011
Residual                                          1237  175.348 0.68408
Total                                             1323  256.326 1.00000
                                                        F Pr(>F)    
sample_data(y)$Sp_full_name                       17.6558  0.001 ***
sample_data(y)$Stress                              2.1876  0.001 ***
sample_data(y)$Sp_full_name:sample_data(y)$Stress  1.2951  0.001 ***
Residual                                                            
Total                                                               
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

$Fun
Permutation test for adonis under reduced model
Terms added sequentially (first to last)
Permutation: free
Number of permutations: 999

adonis2(formula = x ~ sample_data(y)$Sp_full_name * sample_data(y)$Stress)
                                                    Df SumOfSqs      R2
sample_data(y)$Sp_full_name                         28    48.96 0.12883
sample_data(y)$Stress                                2     1.20 0.00317
sample_data(y)$Sp_full_name:sample_data(y)$Stress   56    18.51 0.04870
Residual                                          1219   311.34 0.81930
Total                                             1305   380.01 1.00000
                                                       F Pr(>F)    
sample_data(y)$Sp_full_name                       6.8458  0.001 ***
sample_data(y)$Stress                             2.3576  0.001 ***
sample_data(y)$Sp_full_name:sample_data(y)$Stress 1.2940  0.001 ***
Residual                                                           
Total                                                              
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

###Export
```{r}
save(adonis_bray_l, file = "C:/RProjects/Family_experiment/02_Beta_diversity/r_objects/Guest_species/adonis_bray_l.RData")
```

##Plant family, species and stress
```{r}
adonis_bray_PlantFamily_l <- mapply(function(x, y) {
  adonis2(x ~ sample_data(y)$Sp_Family*sample_data(y)$Sp_full_name * sample_data(y)$Stress)
},
x = physeq_bray_l,
y = CSS_BacFun_tree_Guests_l,
SIMPLIFY = FALSE)

adonis_bray_PlantFamily_l
```

###Export
```{r}
save(adonis_bray_PlantFamily_l, file = "C:/RProjects/Family_experiment/02_Beta_diversity/r_objects/Guest_species/adonis_bray_PlantFamily_l.RData")
```

##Only plant family
```{r}
adonis_bray_PlantFamily2_l <- mapply(function(x, y) {
  adonis2(x ~ sample_data(y)$Sp_Family)
},
x = physeq_bray_l,
y = CSS_BacFun_tree_Guests_l,
SIMPLIFY = FALSE)

adonis_bray_PlantFamily2_l
```

###Export
```{r}
save(adonis_bray_PlantFamily2_l, file = "C:/RProjects/Family_experiment/02_Beta_diversity/r_objects/Guest_species/adonis_bray_PlantFamily2_l.RData")
```

##Plant family * stress
```{r}
adonis_bray_PlantFamily3_l <- mapply(function(x, y) {
  adonis2(x ~ sample_data(y)$Sp_Family* sample_data(y)$Stress)
},
x = physeq_bray_l,
y = CSS_BacFun_tree_Guests_l,
SIMPLIFY = FALSE)

adonis_bray_PlantFamily3_l
```

###Export
```{r}
save(adonis_bray_PlantFamily3_l, file = "C:/RProjects/Family_experiment/02_Beta_diversity/r_objects/Guest_species/adonis_bray_PlantFamily3_l.RData")
```

#PCoAs
##Ordination
```{r}
physeq_bray_pcoa_l <- mapply(function(x, y){
  ordinate(x, "PCoA", y)
},
x = CSS_BacFun_tree_Guests_l,
y = physeq_bray_l,
SIMPLIFY = FALSE)

```

###Store ordination
```{r}
save(physeq_bray_pcoa_l, file = "C:/RProjects/Family_experiment/02_Beta_diversity/r_objects/Guest_species/physeq_bray_pcoa_l.RData")
```

#Create PCoA df
```{r}
#Make a df out of the ordination
df_pcoa_l <- lapply(physeq_bray_pcoa_l, function(x) 
  as.data.frame.matrix(x$vectors))

#Add first and second axis to the metadata
metadata_l <- lapply(CSS_BacFun_tree_Guests_l, function(x)
  sample_data(x))

pcoa_metadata_l <- mapply(function(x, y) {
  merge(x[1:2],y, by=0, all=TRUE)
},
x = df_pcoa_l,
y = metadata_l,
SIMPLIFY = FALSE)

#with labels for each genotype
plot_metadata_l <- lapply(pcoa_metadata_l, function(x)
  ggplot(x,aes(Axis.1,Axis.2,color=factor(Sp_Family), shape=factor(Stress))) +
  geom_point(size=2) +
  scale_shape_manual(values= c(17, 15, 16)) + 
  scale_color_manual(values = c("Brassicaceae" = "#548B54", "Cucurbitacea" = "lightgray", "Fabaceae" = "gray", "Poaceae" = "darkgray", "Solanaceae" = "black")))
```

#Now with plot_ordination from phyloseq
```{r}
plot_metadata_plot_ordination_l <- mapply(function(x,y) { 
  plot_ordination(x, y, type="samples", color="Sp_Family", shape = "Stress", title="PCoA on Bray-Curtis dissimilarity indices") + 
  geom_point(size=0.5) + 
  scale_color_manual(values = c("Brassicaceae" = "#33A02C", "Cucurbitacea" = "#1F78B4", "Fabaceae" = "#B15928", "Poaceae" = "#FF7F00", "Solanaceae" = "#6A3D9A")) +
  theme(panel.background = element_rect(fill = "white", colour = "azure3",
                                        size = 2, linetype = "solid"))
},
  x = CSS_BacFun_tree_Guests_l,
  y = physeq_bray_pcoa_l,
SIMPLIFY = FALSE)

plot_metadata_plot_ordination_l$Bac
plot_metadata_plot_ordination_l$Fun

#Alternative colouring
#scale_colour_brewer(palette = "Paired") +
#scale_color_manual(values = c("Brassicaceae" = "#33A02C", "Cucurbitacea" = "#1F78B4", "Fabaceae" = "#B15928", "Poaceae" = "#FF7F00", "Solanaceae" = "#6A3D9A")
#scale_color_manual(values = c("Brassicaceae" = "#33A02C", "Cucurbitacea" = "#A6CEE3", "Fabaceae" = "#969696", "Poaceae" = "#FDBF6F", "Solanaceae" = "#9E9AC8"))
```

###Export
```{r}
ggsave("C:/RProjects/Family_experiment/02_Beta_diversity/results/Guest_species/physeq_bray_pcoa_l_Bac.pdf", plot_metadata_plot_ordination_l$Bac, width = 8, height = 6, units = "in", useDingbats = FALSE)
ggsave("C:/RProjects/Family_experiment/02_Beta_diversity/results/Guest_species/physeq_bray_pcoa_l_Fun.pdf", plot_metadata_plot_ordination_l$Fun, width = 8, height = 6, units = "in", useDingbats = FALSE)
```

```{r}
sessionInfo()
```

R version 4.5.2 (2025-10-31 ucrt)
Platform: x86_64-w64-mingw32/x64
Running under: Windows 11 x64 (build 22631)

Matrix products: default
  LAPACK version 3.12.1

locale:
[1] LC_COLLATE=Dutch_Netherlands.utf8  LC_CTYPE=Dutch_Netherlands.utf8    LC_MONETARY=Dutch_Netherlands.utf8
[4] LC_NUMERIC=C                       LC_TIME=Dutch_Netherlands.utf8    

time zone: Europe/Amsterdam
tzcode source: internal

attached base packages:
[1] stats     graphics  grDevices datasets  utils     methods   base     

other attached packages:
[1] dplyr_1.2.1         tidyr_1.3.2         ggplot2_4.0.3       vegan_2.6-2         lattice_0.22-7      permute_0.9-10     
[7] phyloseq_1.54.2     BiocManager_1.30.27

loaded via a namespace (and not attached):
 [1] generics_0.1.4      renv_1.2.2          stringi_1.8.7       digest_0.6.39       magrittr_2.0.5      grid_4.5.2         
 [7] RColorBrewer_1.1-3  iterators_1.0.14    foreach_1.5.2       plyr_1.8.9          jsonlite_2.0.0      Matrix_1.7-4       
[13] ape_5.8-1           survival_3.8-3      mgcv_1.9-4          purrr_1.2.2         scales_1.4.0        Biostrings_2.78.0  
[19] codetools_0.2-20    ade4_1.7-24         cli_3.6.6           rlang_1.2.0         crayon_1.5.3        XVector_0.50.0     
[25] Biobase_2.70.0      splines_4.5.2       withr_3.0.2         tools_4.5.2         parallel_4.5.2      reshape2_1.4.5     
[31] multtest_2.66.0     BiocGenerics_0.56.0 vctrs_0.7.3         R6_2.6.1            stats4_4.5.2        lifecycle_1.0.5    
[37] Seqinfo_1.0.0       stringr_1.6.0       biomformat_1.38.3   S4Vectors_0.48.1    IRanges_2.44.0      MASS_7.3-65        
[43] cluster_2.1.8.1     pkgconfig_2.0.3     pillar_1.11.1       gtable_0.3.6        data.table_1.18.2.1 glue_1.8.1         
[49] Rcpp_1.1.1-1.1      tidyselect_1.2.1    tibble_3.3.1        farver_2.1.2        nlme_3.1-168        igraph_2.3.1       
[55] labeling_0.4.3      compiler_4.5.2      S7_0.2.2   


#Only 2 representatives from the BRassicaceae family
```{r}
#Extract the names of the 5 guest species
sam_data <- CSS_BacFun_rhizoplane_HighFilt_ps_l$Bac@sam_data
Guest_sp <- subset(sam_data, Sp_Guest == "yes")
Guest_sp_vector <- unique(Guest_sp$Sp_full_name)

Guest_sp_vector

print(Guest_sp_vector)

#subset
CSS_BacFun_Guests_l <- lapply(CSS_BacFun_rhizoplane_HighFilt_ps_l, function(x) 
                      subset_samples(x, Sp_full_name %in% Guest_sp_vector))

length(unique(CSS_BacFun_Guests_l$Bac@sam_data$Sp_full_name))
```

#Calculate and store Bray-Curtis dissimilarity indices
```{r}
physeq_bray_OnlyGuests_l <- lapply(CSS_BacFun_Guests_l, function(x) 
  phyloseq::distance(x, method = "bray"))
```

###Export
```{r}
save(physeq_bray_OnlyGuests_l, file = "C:/RProjects/Family_experiment/02_Beta_diversity/r_objects/Guest_species/physeq_bray_OnlyGuests_l.RData")
```

##Plant family, species and stress
```{r}
adonis_bray_OnlyGuests_PlantFamily_l <- mapply(function(x, y) {
  adonis2(x ~ sample_data(y)$Sp_Family*sample_data(y)$Sp_full_name * sample_data(y)$Stress)
},
x = physeq_bray_OnlyGuests_l,
y = CSS_BacFun_Guests_l,
SIMPLIFY = FALSE)

adonis_bray_OnlyGuests_PlantFamily_l
```

###Export
```{r}
save(adonis_bray_OnlyGuests_PlantFamily_l, file = "C:/RProjects/Family_experiment/02_Beta_diversity/r_objects/Guest_species/adonis_bray_OnlyGuests_PlantFamily_l.RData")
```

#PCoAs
##Ordination
```{r}
physeq_bray_OnlyGuests_pcoa_l <- mapply(function(x, y){
  ordinate(x, "PCoA", y)
},
x = CSS_BacFun_Guests_l,
y = physeq_bray_OnlyGuests_l,
SIMPLIFY = FALSE)

head(physeq_bray_OnlyGuests_pcoa_l)
```

###Store ordination
```{r}
save(physeq_bray_OnlyGuests_pcoa_l, file = "C:/RProjects/Family_experiment/02_Beta_diversity/r_objects/Guest_species/physeq_bray_OnlyGuests_pcoa_l.RData")
```

#Now with plot_ordination from phyloseq
```{r}
load("C:/RProjects/Family_experiment/02_Beta_diversity/r_objects/Guest_species/physeq_bray_OnlyGuests_pcoa_l.RData")
plot_physeq_bray_OnlyGuests_pcoa_l <- mapply(function(x,y) { 
  plot_ordination(x, y, type="samples", color="Sp_Family", shape = "Stress", title="PCoA on Bray-Curtis dissimilarity indices") + 
  geom_point(size=3) + 
  scale_color_manual(values = c("Brassicaceae" = "#33A02C", "Cucurbitacea" = "#1F78B4", "Fabaceae" = "#B15928", "Poaceae" = "#FF7F00", "Solanaceae" = "#6A3D9A")) +
  theme(panel.background = element_rect(fill = "white", colour = "azure3",
                                        size = 2, linetype = "solid"))
},
  x = CSS_BacFun_Guests_l,
  y = physeq_bray_OnlyGuests_pcoa_l,
SIMPLIFY = FALSE)

plot_physeq_bray_OnlyGuests_pcoa_l$Bac
plot_physeq_bray_OnlyGuests_pcoa_l$Fun

#Alternative colouring
#scale_colour_brewer(palette = "Paired") +
#scale_color_manual(values = c("Brassicaceae" = "#33A02C", "Cucurbitacea" = "#1F78B4", "Fabaceae" = "#B15928", "Poaceae" = "#FF7F00", "Solanaceae" = "#6A3D9A")
#scale_color_manual(values = c("Brassicaceae" = "#33A02C", "Cucurbitacea" = "#A6CEE3", "Fabaceae" = "#969696", "Poaceae" = "#FDBF6F", "Solanaceae" = "#9E9AC8"))
```

###Export
```{r}
ggsave("C:/RProjects/Family_experiment/02_Beta_diversity/results/Guest_species/plot_physeq_bray_OnlyGuests_pcoa_l_Bac.pdf", plot_physeq_bray_OnlyGuests_pcoa_l$Bac, width = 8, height = 6, units = "in", useDingbats = FALSE)
ggsave("C:/RProjects/Family_experiment/02_Beta_diversity/results/Guest_species/plot_physeq_bray_OnlyGuests_pcoa_l_Fun.pdf", plot_physeq_bray_OnlyGuests_pcoa_l$Fun, width = 8, height = 6, units = "in", useDingbats = FALSE)
```
