---
title: "Phylogenetic_signal_v2"
author: "Marcela Aragon"
date: "2024-11-18"
output: html_document
---

The goal of this script is to test whether different traits we've analyzed so far are more similar between closely related species than those that are father away. For this, I will analyze:

* Plant biomass 
* Change in plant biomass treatment vs control (Cohen´s D)
* Alpha diversity metrics
* R2 values from PERMANOVA
* R2 value from NM fit 

And calculate (1) phylogenetic signal as *Bloomberg's K* and *Pagel's lambda* as well as (2) a Mantel test to test hierarchical correlations between plant phylogeny and a given trait's distance matrix.


# 1.0 set up working environemnt

## 1.01 load libraries with renv, improve memory use

```{r}
# set working directory to project directory 

library("phyloseq")
library("ggplot2")
library("dplyr")
library("tidyr")
library("tibble")
library("car") # for Levene's test
library("purrr")
library("ape") #handles phylogenetic objects
library("TreeTools")
library("tidytree")
library("lattice")
library("phangorn") #UPGMA & NJ
library("ade4")
library("distory") #to compare trees
library("geiger")
library("phytools") 
library("stats") #cophenetic distances 
library("heatmaply")
library("pheatmap")
library("RColorBrewer")
library("viridis")
library("grid") #change as.grob
library("ggplotify") #change as.grob
#library("ggpmisc") #plots R2 and regression lines 
library("picante")#phylogenetic signal
library("adephylo")
library("phylobase")
library("phylosignal")
library("vegan")
library("ggpubr")

#load lapply_lapply
source("./renv/functions/lapply_lapply.R")


```


## 1.02 Settings for plots

```{r}
# Set axis looks
theme_set(theme_bw())
axis_looks <- theme(axis.text.x = element_text(colour = "black", size = 10,
                                               face = "bold"))+
  theme(axis.text.y = element_text(colour = "black", size = 12, face = "bold"))+
  theme(axis.title=element_text(size=15, face = "bold"))
```


## 1.02 Load data 
```{r}

#---- load ps object ----
load("./Data/phyloseq_objects/CSS_BacFun_rhizoplane_HighFilt_ps_l.RData")

#---- load tree ----
#Final plant phylogeny from Kasper
Bra_tree <- read.tree("./17_Plant_phylogeny/Results/Supermatrix_Kasper_tree_renamed_rotated_rooted.tre")
#quick look
plot(Bra_tree, type="phylogram", align.tip.label=TRUE)
phylo_tips <- rev(Bra_tree$tip.label) #vector with phylogeny order of 22 final species with Sp abb name

#---- alpha diversity info ----
#load("./03_Alpha_diversity/r_objects/Final_paper/Diversity_all_Nov23.RData")
alpha.Bac <- readRDS(file="./18_Sequencing_depth/results/Bac_alpha_Filt_and_Unfilt_df.rds") #to-do later
alpha.Fun <- readRDS(file="./18_Sequencing_depth/results/Fun_alpha_Filt_and_Unfilt_df.rds") #to-do later
diversity_all <- list(alpha.Bac, alpha.Fun)

#add names
names(diversity_all) <- c("Bac", "Fun")
diversity_all$Bac #ok

#check how many species
levels(diversity_all$Bac$Sp_abb_name) #22, ok

#---- Cohen's D information ----
#load("./15_Plant_Biomass/Objects/shoot_biomass_CohenD.Rdata")
#all_cohen_df 
#all_cohen_df$estimate #this is the value of treatment vs control 

#---- PCoA information ----
load("./14_Phylogenetic_distance/r_objects/pcoa_coordinates_Bac_l.RData")
load("./14_Phylogenetic_distance/r_objects/pcoa_coordinates_Fun_l.RData")

#list with PC1 and PC2 per stress treatment 
pcoa_coordinates_Bac_l
pcoa_coordinates_Fun_l

```



# Make DFs
from the metadata info it would be important to check: 

*Amplicon_concentation_ngul
*Plastid_reads
*Host_DNA_contamination_pct
*Library size (updated)
*Biomass
*PC1 and PC2


```{r}
#---- Library size new ----
#library size new means the final number of reads after Chrat's filtering (HighFilt)

#First add library size 
library_size_new <- sample_sums(CSS_BacFun_rhizoplane_HighFilt_ps_l$Bac)
sample_data(CSS_BacFun_rhizoplane_HighFilt_ps_l$Bac)$library_size_new <- library_size_new

library_size_newF <- sample_sums(CSS_BacFun_rhizoplane_HighFilt_ps_l$Fun)
sample_data(CSS_BacFun_rhizoplane_HighFilt_ps_l$Fun)$library_size_new <- library_size_newF

#now get df 
metadata_B <- data.frame(sample_data(CSS_BacFun_rhizoplane_HighFilt_ps_l$Bac))
metadata_F <- data.frame(sample_data(CSS_BacFun_rhizoplane_HighFilt_ps_l$Fun))

colnames(metadata_F)

#---- Sequencing traits ----

traits <- c("Amplicon_concentation_ngul", "Plastid_reads", "Host_DNA_contamination_pct", "library_size_new")
traitsB <- metadata_B %>% 
                group_by(Sp_abb_name,Stress)%>%
                summarise(across(
      all_of(traits),  # Use the vector of column names
      ~ mean(.x, na.rm = TRUE)))       # Keep original column name and add "_Bac"

traitsB <- traitsB %>% filter(Sp_abb_name %in% phylo_tips) #66 (22 x 3 treatments)


traitsF <- c("Amplicon_concentation_ngul",  "library_size_new")
traitsF <- metadata_F %>% 
                group_by(Sp_abb_name,Stress)%>%
                summarise(across(
      all_of(traitsF),  # Use the vector of column names
      ~ mean(.x, na.rm = TRUE)))      # Keep original column name and add "_Bac"

traitsF <- traitsF %>% filter(Sp_abb_name %in% phylo_tips) #66 (22 x 3 treatments)


#---- Biomass ----
biomass <- metadata_B %>% 
                group_by(Sp_abb_name,Stress)%>%
                summarise(biomass=mean(Shoot_.DryWeight, na.rm=TRUE))

biomass <- biomass %>% filter(Sp_abb_name %in% phylo_tips) #66 (22 x 3 treatments)

#---- Alpha diversity ----
str(diversity_all$Bac)

alphaB <-  diversity_all$Bac %>% 
  group_by(Sp_abb_name, Stress, filter) %>%
  summarise(
    across(
      Observed:Fisher,  # Replace with your consecutive columns
      ~ mean(.x, na.rm = TRUE)))        # Apply mean function ignoring NAs
                 # Add "_Bac" suffix to the column name

#For now, lets only keep the highFilt
alphaB <- alphaB %>% filter(filter == "HighFilt")

alphaF <-  diversity_all$Fun %>% 
  group_by(Sp_abb_name, Stress, filter) %>%
  summarise(
    across(
      Observed:Fisher,  # Replace with your consecutive columns
      ~ mean(.x, na.rm = TRUE)))        # Apply mean function ignoring NAs

#For now, lets only keep the highFilt
alphaF <- alphaF %>% filter(filter == "HighFilt")
       
#---- Community similarity info (PCoA) ----

pcoa_coordinates_Bac_l

#First I need to calculate the average per sp and stress
pcoa_Bac <- lapply(pcoa_coordinates_Bac_l, function(df){
  df <- df %>% 
    group_by(Sp_abb_name) %>% 
      summarise(PC1=mean(PC1),
                PC2=mean(PC2))
  return(df)
})

pcoa_Bac$control <- pcoa_Bac$control %>% mutate(Stress="Control")
pcoa_Bac$MeJA <- pcoa_Bac$MeJA %>% mutate(Stress="MeJA")
pcoa_Bac$SA <- pcoa_Bac$SA %>% mutate(Stress="SA")

pcoa_Bac <- bind_rows(pcoa_Bac) #ok

#Now the same but for fungi
pcoa_Fun <- lapply(pcoa_coordinates_Fun_l, function(df){
  df <- df %>% 
    group_by(Sp_abb_name) %>% 
      summarise(PC1=mean(PC1),
                PC2=mean(PC2))
  return(df)
})

pcoa_Fun$control <- pcoa_Fun$control %>% mutate(Stress="Control")
pcoa_Fun$MeJA <- pcoa_Fun$MeJA %>% mutate(Stress="MeJA")
pcoa_Fun$SA <- pcoa_Fun$SA %>% mutate(Stress="SA")

pcoa_Fun <- bind_rows(pcoa_Fun) #ok


#---- Full Bac ----
#traitsB, biomass, alphaB, pcoa_Bac

Bac <- biomass %>%
  inner_join(traitsB, by = c("Sp_abb_name", "Stress")) %>%
  inner_join(alphaB, by = c("Sp_abb_name", "Stress"))%>%
  inner_join(pcoa_Bac, by = c("Sp_abb_name", "Stress"))%>%
  as.data.frame()

#Bac <- full_join(Bac, Cohen4, by = c("Sp_abb_name", "Stress")) #works

#let's only keep those that we are interested now (16/12/2024):
traits_albouheif <- c("Sp_abb_name","Stress","biomass","PC1", "PC2", "Observed",
                      "Shannon", "library_size_new")

Bac <- Bac %>% select(all_of(traits_albouheif))

#rename
Bac <- Bac %>%
  rename_with(~ paste0(., "_Bac"), PC1:library_size_new)

#split by stress
#Bac_l <- split(Bac, Bac$Stress)


```

Now for Fungi
```{r}
#---- Full Fun ----
#traitsF, alphaF, pcoa_Fun

Fun <- traitsF %>%
  inner_join(alphaF, by = c("Sp_abb_name", "Stress")) %>%
  inner_join(pcoa_Fun, by = c("Sp_abb_name", "Stress"))%>%
  as.data.frame()

#let's only keep those that we are interested now (16/12/2024):
traits_albouheif <- c("Sp_abb_name","Stress", "PC1", "PC2", "Observed",
                      "Shannon", "library_size_new")

#let's delete columns that we don't need 
Fun <- Fun %>% select(all_of(traits_albouheif))

#let's add a Fun sufix to PC1:library_size_new
Fun <- Fun %>%
  rename_with(~ paste0(., "_Fun"), PC1:library_size_new)

#let's join both df 
Full_df <- left_join(Bac, Fun)

#split by stress
Full_df_l <- split(Full_df, Full_df$Stress)

#clean
rm(CSS_BacFun_rhizoplane_HighFilt_ps_l) #too large, no needed now

```

Fix df's
```{r}

#First, modify df's by removing columns and using rownames as species names so it matches with the phylo tree
#Bac_l <- lapply(Bac_l, function(df){
#          rownames(df) <- NULL
#          rownames(df) <- df$Sp_abb_name
#          df <- df[, !names(df) %in% c("Sp_abb_name", "Stress")]
 #         return(df)})

Full_df_l <- lapply(Full_df_l , function(df){
          rownames(df) <- NULL
          rownames(df) <- df$Sp_abb_name
          df <- df[, !names(df) %in% c("Sp_abb_name", "Stress")]
          return(df)})
  

#Bac_l$Control <- Bac_l$Control[, !names(Bac_l$Control) %in% "cohens"] #remove last column of cohens that is all NAs in controls

#check
Full_df_l #ok

#let's order the species by how they are in the phylogenetic tree just in case
order_tree <- Bra_tree$tip.label
Full_df_l <- lapply(Full_df_l, function(df){
  df <- df[order(match(rownames(df), order_tree)),]
  return(df)
}) #ok

```


# Phylogenetic signal (per stress)

##Abouheif's Cmean

```{r}

#let's make a function to get a df out of abouheif.moran() result to avoid repeating code 
abouheif_result_to_df <- function(abouheif_result, data_from = ""){
  df <- data.frame(
  #Method = "Abouheif",              # Name of the method
  Trait = abouheif_result$names,      # Name of the trait 
  Cmean = abouheif_result$obs,      # Observed Cmean value
  P_value = abouheif_result$pvalue, # P-value
  Stress = data_from
)
  return(df)
}

```

```{r}

#[DEPRECATED 17-12-24]let's use phylosignal package that calculates all of them at once
#let's use the abouheif.moran() function so it's reproducible with Pedro's results
set.seed(12345)

#---- Control ----
pd4_control <- phylo4d(Bra_tree, Full_df_l$Control)
#df_controlB <- phyloSignal(p4d = pd4_controlB, method = "all")
control <- abouheif.moran(pd4_control, method = "oriAbouheif", nrepet = 9999)
df_control <- abouheif_result_to_df(control, "Control")
#df_controlB <- lapply(df_controlB, function(df){ df %>% mutate(Stress="Control")})

#---- MeJA ----
pd4_MeJA <- phylo4d(Bra_tree, Full_df_l$MeJA)
MeJA <- abouheif.moran(pd4_MeJA, method = "oriAbouheif", nrepet = 9999)
df_MeJA <- abouheif_result_to_df(MeJA, "MeJA")

#---- SA----
pd4_SA <- phylo4d(Bra_tree, Full_df_l$SA)
SA <- abouheif.moran(pd4_SA, method = "oriAbouheif", nrepet = 9999)
df_SA <- abouheif_result_to_df(SA, "SA")


```

make a single df, for now it says 'Bac' but its actually both Bac and Fun together
```{r}

#make it a single df 
abouheif_full <- rbind(df_control, df_MeJA, df_SA) 

#re-order
abouheif_full$Trait <- factor(abouheif_full$Trait, levels = c("biomass", "PC1_Bac", "PC2_Bac",
                                                                    "Observed_Bac", "Shannon_Bac",
                                                                    "library_size_new_Bac",
                                                                    "PC1_Fun", "PC2_Fun","Observed_Fun",
                                                                    "Shannon_Fun",
                                                                    "library_size_new_Fun"))

#save full df 
saveRDS(abouheif_full, "./14_Phylogenetic_distance/results/New_S7Fig_Abouheif_18-02/abouheif_full.rds")

#select only those for new S7 figure (18-02-2025)
selected_traits <- c("PC1_Bac", "PC2_Bac", "Observed_Bac", "Shannon_Bac", "PC1_Fun", "PC2_Fun",
                     "Observed_Fun", "Shannon_Fun")

abuoheif_full_selection <- abouheif_full %>% filter(Trait %in% selected_traits) 

```


##Plot

```{r}

abouheif_full$Trait  <- factor(abouheif_full$Trait , levels=rev(levels(abouheif_full$Trait))) #reverse order for plot 

#make an extra column if P value is significant
abouheif_full$signif <- ifelse(abouheif_full$P_value < 0.05, "*", "")

#---- Abouheif's ---- 
p.abouheif <- ggplot(data= abouheif_full,
       mapping = aes(x = Trait, y = Cmean, fill=Stress))+
    geom_bar(stat='identity', position = "dodge")+
    scale_fill_manual(values=c('#999999','#E69F00', '#009E73'))+
    geom_hline(aes(yintercept=0), linetype="solid") +
    coord_flip()+
    facet_wrap(~Stress, ncol=3)+
     geom_text(aes(label = signif, y = Cmean + 0.02), # Position above bars
              color = "black", size = 8, vjust = 0.7)+
  labs(y="Abouheif's Cmean")+
   axis_looks+
  theme(legend.position = "none") 
   

#save plot
mapply(function(x)
  ggsave(
  plot = p.abouheif,
  filename =x,
  path = "./14_Phylogenetic_distance/results/plots_abouheif_16-12",
  scale = 2,
  width = 1500,
  height = 700,
  units = "px",
  dpi = 300),
x=c("Traits_abouheif_abouheif.moran-function.pdf",
    "Traits_abouheif_abouheif.moran-function.png",
    "Traits_abouheif_abouheif.moran-function.svg"))

```


# Phylogenetic signal (without spliting first)

Let's test whether the value changes if you use as input a matrix in which the 3 treatments were included from the beggingin (in the previous case, the ps object was first split by stress, Bray curtis distances calculated and then 3 independent PCoA plots were made). In this case the ps object was not split by stress to calculate the Bray curtis distances and to make the PCoAs where the coordenates come from [as we think Pedro did for his S7 figure]

##Load new data
```{r}

load("./14_Phylogenetic_distance/r_objects/pcoa_coordinates_l.RData") #new matrix given by Zulema 17-12
pcoa_coordinates_l #Bac and Fun

#let's first get it split by stress
pcoa_Bac2_df <- pcoa_coordinates_l$Bac
pcoa_Bac2_l <- split(pcoa_Bac2_df, pcoa_Bac2_df$Stress)

#---- Community similarity info (PCoA) [full community first] ----

pcoa_Bac2_l

#First I need to calculate the average per sp and stress
pcoa_Bac2 <- lapply(pcoa_Bac2_l, function(df){
  df <- df %>% 
    group_by(Sp_abb_name) %>% 
      summarise(PC1=mean(PC1),
                PC2=mean(PC2))
  return(df)
})

pcoa_Bac2$Control <- pcoa_Bac2$Control %>% mutate(Stress="Control")
pcoa_Bac2$MeJA <- pcoa_Bac2$MeJA %>% mutate(Stress="MeJA")
pcoa_Bac2$SA <- pcoa_Bac2$SA %>% mutate(Stress="SA")

pcoa_Bac2 <- bind_rows(pcoa_Bac2) #ok
colnames(pcoa_Bac2) <- c("Sp_abb_name", "PC1_Bac_full" ,  "PC2_Bac_full", "Stress")

#Now the same but for fungi

#let's first get it split by stress
pcoa_Fun2_df <- pcoa_coordinates_l$Fun
pcoa_Fun2_l <- split(pcoa_Fun2_df, pcoa_Fun2_df$Stress)

pcoa_Fun2 <- lapply(pcoa_Fun2_l , function(df){
  df <- df %>% 
    group_by(Sp_abb_name) %>% 
      summarise(PC1=mean(PC1),
                PC2=mean(PC2))
  return(df)
})

pcoa_Fun2$Control <- pcoa_Fun2$Control %>% mutate(Stress="Control")
pcoa_Fun2$MeJA <- pcoa_Fun2$MeJA %>% mutate(Stress="MeJA")
pcoa_Fun2$SA <- pcoa_Fun2$SA %>% mutate(Stress="SA")

pcoa_Fun2 <- bind_rows(pcoa_Fun2) #ok
colnames(pcoa_Fun2) <- c("Sp_abb_name", "PC1_Fun_full" ,  "PC2_Fun_full", "Stress")


```

Let's merge it with the previous Bac and Fun df 
```{r}
Bac
Fun

Bac2 <- Bac %>%
  inner_join(pcoa_Bac2, by = c("Sp_abb_name", "Stress")) %>%
  as.data.frame()

Fun2 <- Fun %>% 
        inner_join(pcoa_Fun2, by = c("Sp_abb_name", "Stress")) %>%
        as.data.frame()


#let's join both df 
Full_df2 <- left_join(Bac2, Fun2)

#split by stress
Full_df_l2 <- split(Full_df2, Full_df2$Stress)

#let's arrange the df so it works with the function
Full_df_l2 <- lapply(Full_df_l2 , function(df){
          rownames(df) <- NULL
          rownames(df) <- df$Sp_abb_name
          df <- df[, !names(df) %in% c("Sp_abb_name", "Stress")]
          return(df)})

#check
Full_df_l2 #ok

#let's order the species by how they are in the phylogenetic tree just in case
order_tree <- Bra_tree$tip.label
Full_df_l2 <- lapply(Full_df_l2, function(df){
  df <- df[order(match(rownames(df), order_tree)),]
  return(df)
}) #ok

```


##Abouheif's Cmean
```{r}

set.seed(12345)

#---- Control ----
pd4_control2 <- phylo4d(Bra_tree, Full_df_l2$Control)
control2 <- abouheif.moran(pd4_control2, method = "oriAbouheif", nrepet = 9999)
df_control2 <- abouheif_result_to_df(control2, "Control")

#---- MeJA ----
pd4_MeJA2 <- phylo4d(Bra_tree, Full_df_l2$MeJA)
MeJA2 <- abouheif.moran(pd4_MeJA2, method = "oriAbouheif", nrepet = 9999)
df_MeJA2 <- abouheif_result_to_df(MeJA2, "MeJA")

#---- SA----
pd4_SA2 <- phylo4d(Bra_tree, Full_df_l2$SA)
SA2 <- abouheif.moran(pd4_SA2, method = "oriAbouheif", nrepet = 9999)
df_SA2 <- abouheif_result_to_df(SA2, "SA")

```

##Plot

```{r}
#make it a single df 
abouheif_full2 <- rbind(df_control2, df_MeJA2, df_SA2) 

```


##Plot

```{r}

PC_comparison <- c("PC1_Bac", "PC1_Bac_full", "PC2_Bac", "PC2_Bac_full",
                   "PC1_Fun", "PC1_Fun_full","PC2_Fun", "PC2_Fun_full")

#select only PC values for new plot
abouheif_full2_PC <- abouheif_full2 %>% filter(Trait %in% PC_comparison) %>% droplevels()

#order them
abouheif_full2_PC$Trait <- factor(abouheif_full2_PC$Trait, levels=PC_comparison)

#reverse order for plot 
abouheif_full2_PC$Trait  <- factor(abouheif_full2_PC$Trait , levels=rev(levels(abouheif_full2_PC$Trait))) 

#make an extra column if P value is significant
abouheif_full2_PC$signif <- ifelse(abouheif_full2_PC$P_value < 0.05, "*", "")

#---- Abouheif's ---- 
p.abouheif2 <- ggplot(data= abouheif_full2_PC,
       mapping = aes(x = Trait, y = Cmean, fill=Stress))+
    geom_bar(stat='identity', position = "dodge")+
    scale_fill_manual(values=c('#999999','#E69F00', '#009E73'))+
    geom_hline(aes(yintercept=0), linetype="solid") +
    coord_flip()+
    facet_wrap(~Stress, ncol=3)+
     geom_text(aes(label = signif, y = Cmean + 0.02), # Position above bars
              color = "black", size = 8, vjust = 0.7)+
  labs(y="Abouheif's Cmean")+
   axis_looks+
  theme(legend.position = "none") 
   

#save plot
mapply(function(x)
  ggsave(
  plot = p.abouheif2,
  filename =x,
  path = "./14_Phylogenetic_distance/results/plots_abouheif_16-12",
  scale = 2,
  width = 1500,
  height = 700,
  units = "px",
  dpi = 300),
x=c("Traits_abouheif_abouheif.moran-function_comparisonSplitvsFull.pdf",
    "Traits_abouheif_abouheif.moran-function_comparisonSplitvsFull.png",
    "Traits_abouheif_abouheif.moran-function_comparisonSplitvsFull.svg"))

```


#Table with libsize[NEW with unnormalized]

Let's also get a table with libsize per species 
```{r}

#First, load the unnormalized ps object with the HighFilter of Chrats applied.
load("C:/RProjects/Family/Data/phyloseq_objects/unnormalized_BacFun_rhizoplane_HighFilt_ps_l.RData")

unnormalized_BacFun_rhizoplane_HighFilt_ps_l

#First add library size 
library_size_new <- sample_sums(unnormalized_BacFun_rhizoplane_HighFilt_ps_l$Bac)
sample_data(unnormalized_BacFun_rhizoplane_HighFilt_ps_l$Bac)$library_size_new <- library_size_new

library_size_newF <- sample_sums(unnormalized_BacFun_rhizoplane_HighFilt_ps_l$Fun)
sample_data(unnormalized_BacFun_rhizoplane_HighFilt_ps_l$Fun)$library_size_new <- library_size_newF


#now get new df 
metadata_B <- data.frame(sample_data(unnormalized_BacFun_rhizoplane_HighFilt_ps_l$Bac))
metadata_F <- data.frame(sample_data(unnormalized_BacFun_rhizoplane_HighFilt_ps_l$Fun))


#first let's go back to get libsize from all species so we can also calculate SD and SE for Fungi

traitsF <- c("library_size_new")
libsizeF <- metadata_F %>% 
                group_by(Sp_abb_name)%>%
                summarise(across(
      all_of(traitsF),  # Use the vector of column names
      list(
      mean = ~ mean(.x, na.rm = TRUE),
      sd = ~ sd(.x, na.rm=TRUE),
      se= ~sd(.x, na.rm=TRUE) / sqrt(sum(!is.na(.x)))
      ),
      .names = "{.col}_{.fn}"),
      n=n(),
      .groups= "drop")%>% 
    mutate(community="Fun")

libsizeF  <- libsizeF  %>% filter(Sp_abb_name %in% phylo_tips) #(22 species)
colnames(libsizeF) <- c("Sp_abb_name","mean", "SD","se","N","community")

#Now for bacteria
libsizeB <- metadata_B %>% 
                group_by(Sp_abb_name)%>%
                summarise(across(
      all_of(traitsF),  # Use the vector of column names
      list(
      mean = ~ mean(.x, na.rm = TRUE),
      sd = ~ sd(.x, na.rm=TRUE),
      se= ~sd(.x, na.rm=TRUE) / sqrt(sum(!is.na(.x)))
      ),
      .names = "{.col}_{.fn}"),
      n=n(),
      .groups= "drop")%>% 
    mutate(community="Bac")

libsizeB  <- libsizeB  %>% filter(Sp_abb_name %in% phylo_tips) #(22 species)
colnames(libsizeB) <- c("Sp_abb_name","mean", "SD","se","N","community")


#Now let's put it into table format

# Combine mean and SD into "mean ± SD" format
table_with_mean_sd <- libsizeB %>%
  mutate(libzise_Bac = sprintf("%.1f ± %.1f", mean, SD)) %>%
  mutate(N_Bac=N) %>% 
  select(Sp_abb_name, libzise_Bac, N_Bac)

table_with_mean_sdF <- libsizeF %>%
  mutate(libzise_Fun = sprintf("%.1f ± %.1f", mean, SD)) %>%
  mutate(N_Fun=N) %>% 
  select(Sp_abb_name, libzise_Fun, N_Fun)

#make it into a single df 
final_table_libsize <- left_join(table_with_mean_sd, table_with_mean_sdF)

#re-order as in phylogeny
final_table_libsize <- final_table_libsize %>% 
                        mutate(Sp_abb_name= factor(Sp_abb_name, levels=phylo_tips)) %>% 
                        arrange(Sp_abb_name)

#save it into a word file 
library(rempsyc)
table <- nice_table(final_table_libsize ,
              title = c("Table X. Library size +/- SD for both Bac and Fun"))
flextable::save_as_docx(table,
                        path = "./18_Sequencing_depth/results/LibrarySize_BySpecies_unnormalized.docx")

#save it as .csv file
write.csv(final_table_libsize, "./18_Sequencing_depth/results/LibrarySize_BySpecies_unnormalized.csv", row.names = FALSE)

```


#Table with libsize[OLD with CSS ps]

Let's also get a table with libsize per species 
```{r}
#first let's go back to get libsize from all species so we can also calculate SD and SE

traitsF <- c("library_size_new")
libsizeF <- metadata_F %>% 
                group_by(Sp_abb_name)%>%
                summarise(across(
      all_of(traitsF),  # Use the vector of column names
      list(
      mean = ~ mean(.x, na.rm = TRUE),
      sd = ~ sd(.x, na.rm=TRUE),
      se= ~sd(.x, na.rm=TRUE) / sqrt(sum(!is.na(.x)))
      ),
      .names = "{.col}_{.fn}"),
      n=n(),
      .groups= "drop")%>% 
    mutate(community="Fun")

libsizeF  <- libsizeF  %>% filter(Sp_abb_name %in% phylo_tips) #(22 species)
colnames(libsizeF) <- c("Sp_abb_name","mean", "SD","se","N","community")

#Now for bacteria
libsizeB <- metadata_B %>% 
                group_by(Sp_abb_name)%>%
                summarise(across(
      all_of(traitsF),  # Use the vector of column names
      list(
      mean = ~ mean(.x, na.rm = TRUE),
      sd = ~ sd(.x, na.rm=TRUE),
      se= ~sd(.x, na.rm=TRUE) / sqrt(sum(!is.na(.x)))
      ),
      .names = "{.col}_{.fn}"),
      n=n(),
      .groups= "drop")%>% 
    mutate(community="Bac")

libsizeB  <- libsizeB  %>% filter(Sp_abb_name %in% phylo_tips) #(22 species)
colnames(libsizeB) <- c("Sp_abb_name","mean", "SD","se","N","community")


#Now let's put it into table format

# Combine mean and SD into "mean ± SD" format
table_with_mean_sd <- libsizeB %>%
  mutate(libzise_Bac = sprintf("%.1f ± %.1f", mean, SD)) %>%
  mutate(N_Bac=N) %>% 
  select(Sp_abb_name, libzise_Bac, N_Bac)

table_with_mean_sdF <- libsizeF %>%
  mutate(libzise_Fun = sprintf("%.1f ± %.1f", mean, SD)) %>%
  mutate(N_Fun=N) %>% 
  select(Sp_abb_name, libzise_Fun, N_Fun)

#make it into a single df 
final_table_libsize <- left_join(table_with_mean_sd, table_with_mean_sdF)

#re-order as in phylogeny
final_table_libsize <- final_table_libsize %>% 
                        mutate(Sp_abb_name= factor(Sp_abb_name, levels=phylo_tips)) %>% 
                        arrange(Sp_abb_name)

#save it into a word file 
library(rempsyc)
table <- nice_table(final_table_libsize ,
              title = c("Table X. Library size +/- SD for both Bac and Fun"))
flextable::save_as_docx(table,
                        path = "./18_Sequencing_depth/results/LibrarySize_BySpecies.docx")

#save it as .csv file
write.csv(final_table_libsize, "./18_Sequencing_depth/results/LibrarySize_BySpecies.csv", row.names = FALSE)

```

```{r}
unnormalized_BacFun_rhizoplane_HighFilt_ps_l

MY_tree

ps <- lapply(unnormalized_BacFun_rhizoplane_HighFilt_ps_l, function(ps){
  ps <-subset_samples(ps, Sp_abb_name %in% MY_tree)
  ps <- prune_taxa(taxa_sums(ps) > 0, ps)
  print(mean(sample_sums(ps)))
  return(ps)
})



```



#[DELETE]
## Fun
```{r}

#---- Fun ---- 
indexes_full_Fun$Trait <- factor(indexes_full_Fun$Trait, levels = c("PC1", "PC2",
                                                                    "Observed", "Chao1", "Shannon", "Simpson", "InvSimpson", "ACE", "Fisher",
                                                                    "library_size_new", "Amplicon_concentation_ngul"))

indexes_full_Fun$Trait <- factor(indexes_full_Fun$Trait, levels=rev(levels(indexes_full_Fun$Trait))) #reverse order for plot 

#plot index values  

#---- Abouheif's ---- 
p.abouheif_F <- ggplot(data= subset(indexes_full_Fun, index %in% "Cmean"),
       mapping = aes(x = Trait, y = index_value, fill=Stress))+
    geom_bar(stat='identity', position = "dodge")+
    scale_fill_manual(values=c('#999999','#E69F00', '#009E73'))+
    geom_hline(aes(yintercept=0), linetype="solid") +
    coord_flip()+
    facet_wrap(~Stress, ncol=3)+
     geom_text(data = subset(indexes_full_Fun, index %in% "Cmean" & p_value < 0.05),
              aes(label = "*", y = index_value+0.02),  # Position above bars
              color = "black", size = 8, vjust = 0.7)+
  labs(y="Abouheif's Cmean")+
  axis_looks+
  theme(legend.position = "none") 


#save plot
mapply(function(x)
  ggsave(
  plot = p.abouheif_F,
  filename =x,
  path = "./14_Phylogenetic_distance/results/plots_phylosignal_19-11",
  scale = 2,
  width = 1500,
  height = 700,
  units = "px",
  dpi = 300),
x=c("Fun_abouheif.pdf",
    "Fun_abouheif.png"))

#---- Moran's I ---- 

p.moran_F <- ggplot(data= subset(indexes_full_Fun, index %in% "I"),
       mapping = aes(x = Trait, y = index_value, fill=Stress))+
    geom_bar(stat='identity', position = "dodge")+
    scale_fill_manual(values=c('#999999','#E69F00', '#009E73'))+
    geom_hline(aes(yintercept=0), linetype="solid") +
    coord_flip()+
    facet_wrap(~Stress, ncol=3)+
     geom_text(data = subset(indexes_full_Fun, index %in% "I" & p_value < 0.05),
              aes(label = "*", y = index_value+0.006),  # Position above bars
              color = "black", size = 8, vjust = 0.8)+
  labs(y="Moran's I")+
    axis_looks+
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank(),  # Remove x-axis title
        axis.text.y = element_blank(),   # Remove x-axis text
        axis.ticks.y = element_blank()) # Remove x-axis ticks

#save plot
mapply(function(x)
  ggsave(
  plot = p.moran_F,
  filename =x,
  path = "./14_Phylogenetic_distance/results/plots_phylosignal_19-11",
  scale = 2,
  width = 1500,
  height = 700,
  units = "px",
  dpi = 300),
x=c("Fun_moran.pdf",
    "Fun_moran.png"))

#---- Pagel's Lambda ---- 
p.lambda_F <- ggplot(data= subset(indexes_full_Fun, index %in% "Lambda"),
       mapping = aes(x = Trait, y = index_value, fill=Stress))+
    geom_bar(stat='identity', position = "dodge")+
    scale_fill_manual(values=c('#999999','#E69F00', '#009E73'))+
    geom_hline(aes(yintercept=0), linetype="solid") +
    coord_flip()+
    facet_wrap(~Stress, ncol=3)+
     geom_text(data = subset(indexes_full_Fun, index %in% "Lambda" & p_value < 0.05),
              aes(label = "*", y = index_value+0.006),  # Position above bars
              color = "black", size = 8, vjust = 0.8)+
  labs(y="Pagel's lambda")+
    axis_looks+
  theme(legend.position = "none") +
    theme(axis.title.y = element_blank(),  # Remove x-axis title
        axis.text.y = element_blank(),   # Remove x-axis text
        axis.ticks.y = element_blank()) # Remove x-axis ticks

#save plot
mapply(function(x)
  ggsave(
  plot = p.lambda_F,
  filename =x,
  path = "./14_Phylogenetic_distance/results/plots_phylosignal_19-11",
  scale = 2,
  width = 1500,
  height = 700,
  units = "px",
  dpi = 300),
x=c("Fun_lambda.pdf",
    "Fun_lambda.png"))


#---- Bloomberg's K ---- 

p.bloomberg_F <- ggplot(data= subset(indexes_full_Fun, index %in% "K"),
       mapping = aes(x = Trait, y = index_value, fill=Stress))+
    geom_bar(stat='identity', position = "dodge")+
    scale_fill_manual(values=c('#999999','#E69F00', '#009E73'))+
    geom_hline(aes(yintercept=0), linetype="solid") +
    coord_flip()+
    facet_wrap(~Stress, ncol=3)+
     geom_text(data = subset(indexes_full_Fun, index %in% "K" & p_value < 0.05),
              aes(label = "*", y = index_value+0.006),  # Position above bars
              color = "black", size = 8, vjust = 0.8)+
  labs(y="Bloomberg's K")+
    axis_looks+
    theme(axis.title.y = element_blank(),  # Remove x-axis title
        axis.text.y = element_blank(),   # Remove x-axis text
        axis.ticks.y = element_blank()) # Remove x-axis ticks

#save plot
mapply(function(x)
  ggsave(
  plot = p.bloomberg_F,
  filename =x,
  path = "./14_Phylogenetic_distance/results/plots_phylosignal_19-11",
  scale = 2,
  width = 1500,
  height = 700,
  units = "px",
  dpi = 300),
x=c("Fun_bloomberg.pdf",
    "Fun_bloomberg.png"))

#single plot
#go Funk and add no legend in y axis and no legend of colors 
full.p_F <- (p.abouheif_F | p.moran_F | p.lambda_F | p.bloomberg_F)

#save plot
mapply(function(x)
  ggsave(
  plot = full.p,
  filename =x,
  path = "./14_Phylogenetic_distance/results/plots_phylosignal_19-11",
  scale = 4,
  width = 1500,
  height = 500,
  units = "px",
  dpi = 300),
x=c("Fun_all_indexes.pdf",
    "Fun_all_indexes.png"))

```

# small df for Zulema
```{r}

indexes_full_Fun
indexes_full_Bac

PC1_Bac <- subset(indexes_full_Bac, Trait %in% "PC1")
PC1_Fun <- subset(indexes_full_Fun, Trait %in% "PC1")

index_values_PC1_l <- list(Bac=PC1_Bac, Fun=PC1_Fun)

saveRDS(index_values_PC1_l, file="./14_Phylogenetic_distance/results/index_values_PC1_l.rds")

```

#Mantel test
```{r}

#---- load ps object ----
load("./Data/phyloseq_objects/CSS_BacFun_rhizoplane_HighFilt_ps_l.RData")

#keep only those in the tree
# we will select only the Brassica samples from the million year tree
#define species present in the MY tree
MY_tree <- c("At", "Al", "Cas", "Tg", "Ec", "Mm", "Bv",
                "Lr", "Ds", "Bi", "Bn", "Sar", "Hi", "Br", "Bo_M", 
                "Dt", "Si", "It", "Co", "Lm", "Ia", "Es")

#check sample data info
head(sample_data(CSS_BacFun_rhizoplane_HighFilt_ps_l$Bac)) #already has MY presence
head(sample_data(CSS_BacFun_rhizoplane_HighFilt_ps_l$Fun)) #already has MY presence 


#now keep only those inside MY tree
CSS_BacFun_rhizoplane_HighFilt_ps_l$Bac <- subset_samples(CSS_BacFun_rhizoplane_HighFilt_ps_l$Bac,
                                                          MY_tree_presence == "inside_MY_tree")
CSS_BacFun_rhizoplane_HighFilt_ps_l$Fun <- subset_samples(CSS_BacFun_rhizoplane_HighFilt_ps_l$Fun,
                                                          MY_tree_presence == "inside_MY_tree")

#remove remaining ASVs with zero reads
CSS_BacFun_rhizoplane_HighFilt_ps_l <- lapply(CSS_BacFun_rhizoplane_HighFilt_ps_l, function(ps){
  prune_taxa(taxa_sums(ps) >0, ps)})


```

```{r}
Bra_tree
phylo_tips

#calculate plant distance
phylo_dist <- cophenetic(Bra_tree)

#calculate microbiome distance
otu_table <- lapply(CSS_BacFun_rhizoplane_HighFilt_ps_l, function(ps){
                    otu_table <-otu_table(ps)
                    otu_table <- t(otu_table)})

microbiome_dist <- lapply(otu_table, function(x){
                  microbiome_dist <- vegdist(x, method = "bray")})

#save
save(microbiome_dist, file="./14_Phylogenetic_distance/results/Results_Dec2024_mantel/BacFun_dist_Bray.Rdata")

```

```{r}

#load data
load(file="./14_Phylogenetic_distance/results/Results_Dec2024_mantel/BacFun_dist_Bray.Rdata")


#Prepare matrix for mantel test using Roland's function (adapted)
#which takes distance within species as well (diagonal is not 0)
distance_to_Mantel <- function(dist, ps){
  
  #dist <- microbiome_dist$Bac
  #dist <- dist.bray$Control
  
  #This function takes a distance object and makes it a distance matrix ready
  
  # change object to matrix
  dist <- as.matrix(dist)
  
  # change object to data frame
  dist <- as.data.frame(dist)
  
  # row names to column "sample1"
  dist <- tibble::rownames_to_column(dist, "sample1")
  
  # transform wide to long
  dist <- dist %>% pivot_longer(cols=colnames(dist)[-1],
                                names_to='sample2',
                                values_to='distance')
  
  ## replace sample names with plant species-stress treatment combination
  
  # get simplified metadata table with sample names, species codes and treatments
  metadata <- sample_data(ps)
  metadata <- as.matrix(metadata)
  metadata <- as.data.frame(metadata)
  metadata$Sp_abb_name[metadata$Sp_abb_name == "Bo_M"] <- 'Bo' #changing this so no problem with the "_" from Bo_M 
  
  metadata <- metadata[,c("Sp_abb_name", "Stress")]
  metadata <- tibble::rownames_to_column(metadata, "sample_ID")  
  metadata$species_treatment <- paste0(metadata$Sp_abb_name, "_", metadata$Stress)
  metadata <- metadata[,c("sample_ID", "species_treatment")]
  
  # merge dist with metadata
  dist <- merge(dist, metadata, by.x = "sample1", by.y = "sample_ID")
  dist$sample1 <- dist$species_treatment
  dist <- dist[,-4]
  dist <- merge(dist, metadata, by.x = "sample2", by.y = "sample_ID")
  dist$sample2 <- dist$species_treatment
  dist <- dist[,-4]
  
  #check how many rows have 0 in pairwise comparison (1,031 data points)
  nrow(dist[dist$distance == 0,]) #check how many data points

  t <- as.data.frame(with(dist, table(sample1, sample2))) #4,356 pairwise comparisons
  #lowest has 182 data points and highest has 256 (16*16)
  
  # calculate mean distances with the function "summarize"
  dist <- dist %>%
    filter(distance != 0) %>% #remove all those that have 0 (self comparison by sample)
    group_by(sample1, sample2) %>%
    dplyr::summarize(mean_distance = mean(distance))
  
  # [RB NOTE]: there is also a distance of a certain species/treatment combination with itself, which is the average distance between the replicates. we will later exclude the self-comparisons, but it is also good to check how the correlation looks if we include them.
  
  # subset by combination of sp and treatment 
  dist <- separate(data = dist, col = sample1, into = c("species_code_1", "treatment_1"), sep = "_")
  dist <- separate(data = dist, col = sample2, into = c("species_code_2", "treatment_2"), sep = "_")
  
  # subset only rows that have Control for both
  dist_Control <- subset(dist, (treatment_1 == "Control" & treatment_2 == "Control"))
  # subset only rows that have MeJA for both
  dist_MeJA <- subset(dist, (treatment_1 == "MeJA" & treatment_2 == "MeJA"))
  # subset only rows that have SA for both
  dist_SA <- subset(dist, (treatment_1 == "SA" & treatment_2 == "SA"))
  
  # drop columns "treatment_1" and "treatment_2"
  dist_Control = subset(dist_Control, select = -c(treatment_1, treatment_2) ) 
  dist_MeJA = subset(dist_MeJA, select = -c(treatment_1, treatment_2) ) 
  dist_SA = subset(dist_SA, select = -c(treatment_1, treatment_2)) 
  
  # Call function that uses "dist_X", which can then be either "dist_Control" or "dist_MeJA" or "dist_SA"
  # to make it the final distance matrix (to be used later)
  
  format_dist <- function(dist_X){
    
    # transform long to wide [MA] important
    dist_X <- dist_X %>% pivot_wider(names_from="species_code_2",
                                     values_from="mean_distance")
    
    # change to matrix, assign row names
    dist_X <- as.matrix(dist_X)
    rownames(dist_X) <- dist_X[,1] # change row 1 to row names
    dist_X <- dist_X[,-1] # remove the column that contained the row names
    
    return(dist_X) #[MA] returns distance matrix
  }
  
  
  # run the function on the Stress- bray-curtis distance matrices
  dist_Control <- format_dist(dist_Control)
  dist_MeJA <- format_dist(dist_MeJA)
  dist_SA <- format_dist(dist_SA)

  
  # make it into a list and return it out of the function
  dist_ByStress_list <- list(dist_Control, dist_MeJA, dist_SA)
  names(dist_ByStress_list)<-c("Control","MeJA", "SA")
  
  #change to numeric
  numeric_dist_ByStress_list <- lapply(dist_ByStress_list, function(x){
      class(x) <- "numeric"  
      return(x)
  }) 
  
  return(numeric_dist_ByStress_list)
  
}

```

Run function
```{r}

matrix.Bac <- distance_to_Mantel(microbiome_dist$Bac, CSS_BacFun_rhizoplane_HighFilt_ps_l$Bac)
matrix.Fun <- distance_to_Mantel(microbiome_dist$Fun, CSS_BacFun_rhizoplane_HighFilt_ps_l$Fun)

```

Fix names in matrix so it matches 
```{r}

print(rownames(phylo_dist) == rownames(matrix.Bac$Control)) #not same order
print(colnames(phylo_dist) == colnames(matrix.Bac$Control))#function to change the order in matrix 

#first, let's change Bo to Bo_M
rownames(matrix.Bac$Control) <- c("Al", "At", "Bi", "Bn", "Bo_M", "Br", "Bv", "Cas", "Co", "Ds", "Dt", "Ec", "Es", "Hi", "Ia", "It", "Lm", "Lr", "Mm", "Sar", "Si", "Tg")
colnames(matrix.Bac$Control) <- c("Al", "At", "Bi", "Bn", "Bo_M", "Br", "Bv", "Cas", "Co", "Ds", "Dt", "Ec", "Es", "Hi", "Ia", "It", "Lm", "Lr", "Mm", "Sar", "Si", "Tg")

#Now, let's change phylo_dist order to match matrix.Bac

#reorder matrix
phylo_dist <- phylo_dist[rownames(matrix.Bac$Control),colnames(matrix.Bac$Control)]

#check
common_samples <- intersect(rownames(matrix.Bac$Control), rownames(phylo_dist)) #ok

#Run mantel
mantel_Bac <- lapply(matrix.Bac, function(ps_dist){
                mantel(ps_dist, phylo_dist, method = "pearson", permutations = 999)})

mantel_Fun <- lapply(matrix.Fun, function(ps_dist){
                mantel(ps_dist, phylo_dist, method = "pearson", permutations = 999)})


#make into a df
df_Mantel <- function(x){
  
 Mantel_R <- as.data.frame(x$statistic) 
 pvalue <- as.data.frame(x$signif)
 #Distance <- "Bray-Curtis"
 Treatment <- deparse(substitute(x)) #gets name of the object as a string character
 
 df <- data.frame(Treatment, Mantel_R, pvalue)
 
 return(df)
}

#Bacteria
mantel_Bac_df <- lapply(mantel_Bac, function(mantel){df_Mantel(mantel)})
mantel_Bac_df <- map_dfr(mantel_Bac_df , ~ ., .id = "ListName")
mantel_Bac_df <- mantel_Bac_df %>% mutate(community="Bac")

#Fungi
mantel_Fun_df <- lapply(mantel_Fun, function(mantel){df_Mantel(mantel)})
mantel_Fun_df <- map_dfr(mantel_Fun_df , ~ ., .id = "ListName")
mantel_Fun_df <- mantel_Fun_df %>% mutate(community="Fun")

#join df
mantel_df <- rbind(mantel_Bac_df, mantel_Fun_df)

#save result
saveRDS(mantel_df, file="./14_Phylogenetic_distance/results/Results_Dec2024_mantel/mantel_df.rds")
```



#Plot (later)
```{r}

plot(as.vector(phylo_dist), as.vector(matrix.Bac$Control),
     xlab = "Phylogenetic Distance",
     ylab = "Microbiome Distance",
     main = "Mantel Test Correlation",
     pch = 19, col = "blue")
abline(lm(as.vector(matrix.Bac$Control) ~ as.vector(phylo_dist)), col = "red")

```

