---
title: "14a_split_calculate_phylosigs_PERMANOVAS_PCoAs"
author: "Pedro_Beschoren"
date: "20/04/2023"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: consolehtml_document
---

# 14.1 - set up working environemnt


```{r}
# this will load all key libraries for the whole project
source("./renv/0_load_libraries.R")
library(plyr)
library("umap")
library(RColorBrewer)
library(cowplot) 
library(multiblock)
library(data.table)
library(ggExtra)
library(ggplot2)
# IMPORTANT: load the custom functions necessary to run this script
source("./14_Phylogenetic_distance/code/custom_functions_PB.R")


library(picante)
library(ape)
library(adephylo)
library(ade4)
library(phylobase)
library(geiger)
library(phytools)
library(vegan)
library(phyloseq)
library(dplyr)


# increases memory limit used by R
memory.limit(size = 350000)
```

# 14.2 - load ps, removie species outside tree 

```{R}
# load PS data
load("./Data/phyloseq_objects/CSS_BacFun_rhizoplane_HighFilt_ps_l.RData")

# remove all species outside the MY trees
CSS_BacFun_rhizoplane_HighFilt_ps_l <- 
  lapply(CSS_BacFun_rhizoplane_HighFilt_ps_l, function(x) 
    subset_samples(x, MY_tree_presence == "inside_MY_tree"))


```



# 14.3 - calculate phylogenetic signal with a  single function


Bacterial and fungal phylosig, kingdom level only
```{r}

#calculate phylosig on full bacteria and fungal communities
pipeline_output_full<- phylogenetic_signal_pipeline(CSS_BacFun_rhizoplane_HighFilt_ps_l,
                                                         return_plot = TRUE,
                                                         return_lambda = TRUE,
                                                         phylogenetic_test = "Abouheif's Cmean",
                                                         ordination_method = "PCoA",
                                                         significant_signal_only = FALSE,
                                                         higher_PC_only = FALSE)



# mantel pipeline
mantel_pipeline<-ps_to_mantel_pipeline(ps_obj_l = CSS_BacFun_rhizoplane_HighFilt_ps_l)







```

# ------------------
# BACTERIAL phylogenetic signal
# -----------------

# 14.4 - individual phylosig of all bacterial tonomic groups 
Here we split the microbiome data, with one phyloseq object for each taxonomic level, and then calculate the phylogenetic signal of each taxonomic level. This means we have one phylogenetic signal for c__Alphaproteobacteria, one phylogenetic signal for c__Gammaproteobacteria...


## 14.4.1 - prepare a list of ps objects with 1 ps_object per taxon
```{r}

#make ps list of bac
ps_taxa_l<-c(phyloseq_sep_tax(CSS_BacFun_rhizoplane_HighFilt_ps_l$Bac, TaxRank = "Phylum"),
               phyloseq_sep_tax(CSS_BacFun_rhizoplane_HighFilt_ps_l$Bac, TaxRank = "Class"),
               phyloseq_sep_tax(CSS_BacFun_rhizoplane_HighFilt_ps_l$Bac, TaxRank = "Order"),
               phyloseq_sep_tax(CSS_BacFun_rhizoplane_HighFilt_ps_l$Bac, TaxRank = "Family"),
               phyloseq_sep_tax(CSS_BacFun_rhizoplane_HighFilt_ps_l$Bac, TaxRank = "Genus"))



# remove taxa wihtout least 5 different ASVs (filter, prevents artefacts, prevents UMAP crash)
ps_taxa_l<-ps_taxa_l[lapply(ps_taxa_l, ntaxa)>5]


# taxa must be present at least once in control, once SA and once in  MeJA on every plant species, or lese we ccannot calculate a phylognetic signal (I think that at some point data is divided by zero)

# check if a taxon has zero ASVs across a species_stress combination
has_bad_zero<-lapply(ps_taxa_l, function(x){
  
  melted<-psmelt(x)

melted_summrary<-melted%>%
  dplyr::group_by(Stress, Sp_abb_name)%>%
  dplyr::summarise(abu_sum = sum(Abundance))



taxon_missing_in_sp_Stress<-any(melted_summrary[,3]==0)

  
})

# remove taxa with zero ASVs in any species_stress combination
ps_taxa_l<-ps_taxa_l[!unlist(has_bad_zero)]

#remove taxonomies with NA, such as p__NA
ps_taxa_l<-ps_taxa_l[names(ps_taxa_l)!= "NAs"]

# remove samples with zero ASVs
samples_to_keep<-lapply(ps_taxa_l, function(x)
  colSums(otu_table(x))>0)

ps_taxa_l<-mapply(function(x, y)
  prune_samples(y,x),
  x = ps_taxa_l,
  y = samples_to_keep,
  SIMPLIFY = FALSE)

#add full bac community to top of the list as kingdom_bacteria
ps_taxa_l<-append(ps_taxa_l, values = CSS_BacFun_rhizoplane_HighFilt_ps_l$Bac, after = 0)
names(ps_taxa_l)[1]<-"k__Bacteria"

# save list of ps objects, which takes  while to compute
save(ps_taxa_l, file = "./14_Phylogenetic_distance/r_objects/Bac_alltaxa_pre_phylosig_pipeline_ps_l.RData") 
```



## 14.4.2 - calculate phylosig (PCoA, abouheif), save rdata
```{r}

#load the per-taxa ps object from  chunk 14.4.1
load(file = "./14_Phylogenetic_distance/r_objects/Bac_alltaxa_pre_phylosig_pipeline_ps_l.RData") 




# run pipeline; ~ 40 min on all bac PCoA - ABOUHEIF
bac_alltaxa_phylosig_output_PCoA_nonsig<- phylogenetic_signal_pipeline(ps_l = ps_taxa_l,
                                                                           phylogenetic_test = "Abouheif's Cmean",
                                                                           ordination_method = "PCoA",
                                                                           significant_signal_only = FALSE,
                                                                           return_plot = TRUE, 
                                                                           higher_PC_only = FALSE,
                                                                           return_lambda = FALSE) #!

# save object with full phylosig pipeline output
save(bac_alltaxa_phylosig_output_PCoA_nonsig, file = "./14_Phylogenetic_distance/r_objects/bac_alltaxa_phylosig_output_PCoA_nonsig.RData")
load(file = "./14_Phylogenetic_distance/r_objects/bac_alltaxa_phylosig_output_PCoA_nonsig.RData") 


```



# Mantel pipeline

```{r}

# run mantel pipeline on all taxonomic groups
bac_alltaxa_mantel_pipeline<-ps_to_mantel_pipeline(ps_obj_l = ps_taxa_l)

# remove p/r mantel plots as 500 of them simultaneously are not informative
bac_alltaxa_mantel_pipeline$plot_output<-NULL

# remove  g__env.OPS_17, f__env.OPS_17 - the taxon names are incompatible with the code because of the "." and generate errors
# these taxa fail the permanova/PCOA quality control and have no significant abouheif. they also do not show a significant mantel result.let's just discard them
bac_alltaxa_mantel_pipeline$mantel_df_output<-
  dplyr::filter(bac_alltaxa_mantel_pipeline$mantel_df_output, !Community %in% c("g__env", "f__env"))%>% # removes 2 bad taxa
    dplyr::select(Community,Treatment,distance,r2_Mantel,p_value,significant) # removes bad columns added by bad taxad ue to "." in name

#adjsut name of "Community" to taxonomic group
names(bac_alltaxa_mantel_pipeline$mantel_df_output)[1]<-"taxonomic_group"

#save 500 plots
bac_alltaxa_mantel_pipeline_bray_correlation_plot<-bac_alltaxa_mantel_pipeline$bray_correlation_plot
save(bac_alltaxa_mantel_pipeline_bray_correlation_plot, 
     file = "./14_Phylogenetic_distance/r_objects/bac_alltaxa_mantel_pipeline_bray_correlation_plot.RData")

#save df
bac_alltaxa_mantel_pipeline_df<-bac_alltaxa_mantel_pipeline$mantel_df_output 
save(bac_alltaxa_mantel_pipeline_df, 
     file = "./14_Phylogenetic_distance/r_objects/bac_alltaxa_mantel_pipeline_df.RData")


```




## Calculate PCoA (and % var explaned) for each taxnomic group, bac and fun 
```{r}
# load output of pipeline, Bac 
load(file ="./14_Phylogenetic_distance/r_objects/bac_alltaxa_phylosig_output_PCoA_nonsig.RData")


# load output of pipeline, Fun 
load(file ="./14_Phylogenetic_distance/r_objects/fun_alltaxa_phylosig_PCoA_nonsig.RData")


# run the PCoA axis pct and plot, Bac
PCoA_528_plots_axis_pct<-
explained_PCoA_Axis_per_taxonomic_group(ps_obj_l = bac_alltaxa_phylosig_output_PCoA_nonsig$ordination_coordinates_l)

#add names to plots, according taoxnomic group
PCoA_528_plots_axis_pct$ploted_ordination<-
mapply(function (x,y)
  x+ggtitle(label = y),
  x = PCoA_528_plots_axis_pct$ploted_ordination,
  y = as.list(names(PCoA_528_plots_axis_pct$ploted_ordination)),
  SIMPLIFY = FALSE)

#save 528 PCoA plots & ordiantions, Bac
save(PCoA_528_plots_axis_pct, file = "./14_Phylogenetic_distance/r_objects/PCoA_528_plots_axis_pct.RData")

# run the 70  PCoA axis pct and plot, Fun
PCoA_Fun_70_plots_axis_pct<-
  explained_PCoA_Axis_per_taxonomic_group(ps_obj_l = fun_alltaxa_phylosig_PCoA_nonsig$ordination_coordinates_l)


#add names to plots, according taxonomic group
PCoA_Fun_70_plots_axis_pct$ploted_ordination<-
mapply(function (x,y)
  x+ggtitle(label = y),
  x = PCoA_Fun_70_plots_axis_pct$ploted_ordination,
  y = as.list(names(PCoA_Fun_70_plots_axis_pct$ploted_ordination)),
  SIMPLIFY = FALSE)

#save 70 PCoA plots & ordiantions, Fun
save(PCoA_Fun_70_plots_axis_pct, file = "./14_Phylogenetic_distance/r_objects/PCoA_Fun_70_plots_axis_pct.RData")

 


#split by stress, re-run function, Bac
Control_ps_l<-
  lapply(bac_alltaxa_phylosig_output_PCoA_nonsig$ordination_coordinates_l, function (x)
    subset_samples(x, Stress =="Control"))

MeJA_ps_l<-
  lapply(bac_alltaxa_phylosig_output_PCoA_nonsig$ordination_coordinates_l, function (x)
    subset_samples(x, Stress =="MeJA"))

SA_ps_l<-
  lapply(bac_alltaxa_phylosig_output_PCoA_nonsig$ordination_coordinates_l, function (x)
    subset_samples(x, Stress =="SA"))

# run the PoA axis pct and plot, bac
PCoA_528_plots_axis_pct_Control<-
  explained_PCoA_Axis_per_taxonomic_group(ps_obj_l = Control_ps_l)

PCoA_528_plots_axis_pct_MeJA<-
  explained_PCoA_Axis_per_taxonomic_group(ps_obj_l = MeJA_ps_l)

PCoA_528_plots_axis_pct_SA<-
  explained_PCoA_Axis_per_taxonomic_group(ps_obj_l = SA_ps_l)


#add names to plots, according taxonomic group
PCoA_528_plots_axis_pct_Control$ploted_ordination<-
mapply(function (x,y)
  x+ggtitle(label = y),
  x = PCoA_528_plots_axis_pct_Control$ploted_ordination,
  y = as.list(names(PCoA_528_plots_axis_pct_Control$ploted_ordination)),
  SIMPLIFY = FALSE)
names(PCoA_528_plots_axis_pct_Control)<-paste0(names(PCoA_528_plots_axis_pct_Control), "_Control")

PCoA_528_plots_axis_pct_MeJA$ploted_ordination<-
mapply(function (x,y)
  x+ggtitle(label = y),
  x = PCoA_528_plots_axis_pct_MeJA$ploted_ordination,
  y = as.list(names(PCoA_528_plots_axis_pct_MeJA$ploted_ordination)),
  SIMPLIFY = FALSE)
names(PCoA_528_plots_axis_pct_MeJA)<-paste0(names(PCoA_528_plots_axis_pct_MeJA), "_MeJA")

PCoA_528_plots_axis_pct_SA$ploted_ordination<-
mapply(function (x,y)
  x+ggtitle(label = y),
  x = PCoA_528_plots_axis_pct_SA$ploted_ordination,
  y = as.list(names(PCoA_528_plots_axis_pct_SA$ploted_ordination)),
  SIMPLIFY = FALSE)
names(PCoA_528_plots_axis_pct_SA)<-paste0(names(PCoA_528_plots_axis_pct_SA), "_SA")


# combine lists and save, bac
PCoA_528_3x_plots_axis_pct<-c(PCoA_528_plots_axis_pct_Control,
                              PCoA_528_plots_axis_pct_MeJA,
                              PCoA_528_plots_axis_pct_SA)

save(PCoA_528_3x_plots_axis_pct, file = "./14_Phylogenetic_distance/r_objects/PCoA_528_3x_plots_axis_pct.RData")





#split by stress, re-run function, Fun
Control_Fun_ps_l<-
  lapply(fun_alltaxa_phylosig_PCoA_nonsig$ordination_coordinates_l, function (x)
    subset_samples(x, Stress =="Control"))

MeJA_Fun_ps_l<-
  lapply(fun_alltaxa_phylosig_PCoA_nonsig$ordination_coordinates_l, function (x)
    subset_samples(x, Stress =="MeJA"))

SA_Fun_ps_l<-
  lapply(fun_alltaxa_phylosig_PCoA_nonsig$ordination_coordinates_l, function (x)
    subset_samples(x, Stress =="SA"))

# run the PoA axis pct and plot, Fun
PCoA_Fun_528_plots_axis_pct_Control<-
  explained_PCoA_Axis_per_taxonomic_group(ps_obj_l = Control_Fun_ps_l)
names(PCoA_Fun_528_plots_axis_pct_Control)<-paste0(names(PCoA_Fun_528_plots_axis_pct_Control), "_Control")

PCoA_Fun_528_plots_axis_pct_MeJA<-
  explained_PCoA_Axis_per_taxonomic_group(ps_obj_l = MeJA_Fun_ps_l)
names(PCoA_Fun_528_plots_axis_pct_MeJA)<-paste0(names(PCoA_Fun_528_plots_axis_pct_MeJA), "_MeJA")

PCoA_Fun_528_plots_axis_pct_SA<-
  explained_PCoA_Axis_per_taxonomic_group(ps_obj_l = SA_Fun_ps_l)
names(PCoA_Fun_528_plots_axis_pct_SA)<-paste0(names(PCoA_Fun_528_plots_axis_pct_SA), "_SA")


# combine lists and save, Fun
PCoA_Fun_70_3x_plots_axis_pct<-c(PCoA_Fun_528_plots_axis_pct_Control,
                              PCoA_Fun_528_plots_axis_pct_MeJA,
                              PCoA_Fun_528_plots_axis_pct_SA)

save(PCoA_Fun_70_3x_plots_axis_pct, file = "./14_Phylogenetic_distance/r_objects/PCoA_Fun_70_3x_plots_axis_pct.RData")


```






## Calcualte PERMANOVAS for the 528 taxonomic groups...

... with the HPC

*ATTENTION* The 528 permanova set was calculated in the anunnna HPC. this chunk will source the R code used for this step. It will fail if you run it natively on windows because of the parallel computing made by mcmapply and mclapply. The slurm file for submitng the job to the HPC is available on "./14_Phylogenetic_distance/code/528_permanovas.R" 

if you do not use the HPC for these calculations, ignore this chunk and run the next one that has adapted code for local windows machines

```{r}
source("./14_Phylogenetic_distance/code/528_permanovas.R")
```

Alternatively calculate PERMANOVAS for the 528 taxonomic groups locally, without an HPC

```{r}


# testing sp*stress per taxonomic group is a priority. this function will perform that
permanova_per_taxonomic_group_spXstress<-function(ps_obj_l){
  # ps_obj_l = a list of phyloseq objects, split by taxonomic group
  
  
  
  ps_obj_l<- lapply(ps_obj_l, function(x){
    
    x@sam_data$Sp_Lineage_Walden_extended<-
      if_else(condition = x@sam_data$Sp_abb_name %in% c("Ia", "Lm", "Co", "Bi"), 
              true ="lineage_II_extended",
              false =  x@sam_data$Sp_Lineage_Walden)
    
    return(x)
    
  })
  
  
  #calculate permanova at individual stress conditions
  set.seed(303848)
  permanova_l <- lapply(ps_obj_l, function (x)
    adonis2(formula = phyloseq::distance(t(otu_table(x)), method="bray") 
            ~ Sp_full_name + Stress + Sp_full_name*Stress + Block,
            data = as(sample_data(x),"data.frame"), # changing with as.data.frame is insufficient
            permutations = 999, by = "terms"))
  
  
  beta_disper_l <- lapply(ps_obj_l, function (x){
    
    beta_disper_sp <-  betadisper(phyloseq::distance(t(otu_table(x)), method = "bray"), 
                                  sample_data(x)$Sp_abb_name)
    beta_disper_stress <-  betadisper(phyloseq::distance(t(otu_table(x)), method = "bray"),
                                      sample_data(x)$Stress)
    beta_disper_block <-  betadisper(phyloseq::distance(t(otu_table(x)), method = "bray"),
                                     sample_data(x)$Block)
    
    output<-list("beta_disper_sp" = beta_disper_sp,
                 "beta_disper_stress" = beta_disper_stress,
                 "beta_disper_block" = beta_disper_block)
    
    return(output)     
  })
  
  
  
  output<-list("permanova_l" = permanova_l,
               "beta_disper_l" = beta_disper_l)
  
  return(output)
  
}

#run permanvoas
permanovas_spXstress_528<-
  permanova_per_taxonomic_group_spXstress(ps_obj_l = bac_alltaxa_phylosig_output_PCoA_nonsig$ordination_coordinates_l)


# save permanovas
save(permanovas_spXstress_528, file ="./14_Phylogenetic_distance/r_objects/permanovas_spXstress_528.RData")


```





# ------------------
# FUNGAL phylogenetic signal
# -----------------



# 14.10 - individual phylosig of all fungal tonomic groups 
Here we split the microbiome data, with one phyloseq object for each taxonomic level, and then calculate the phylogenetic signal of each taxonomic level. This means we have one phylogenetic signal for c__Eurotiomycetes, one phylogenetic signal for c__Sordariomycetes...



# 14.10.1 - prepare a list of ps objects with 1 ps object per taxon

```{r}


#make ps list of bac
ps_taxa_l_fun<-c( "k__Fungi" = CSS_BacFun_rhizoplane_HighFilt_ps_l$Fun,
              phyloseq_sep_tax(CSS_BacFun_rhizoplane_HighFilt_ps_l$Fun, TaxRank = "Phylum"),
               phyloseq_sep_tax(CSS_BacFun_rhizoplane_HighFilt_ps_l$Fun, TaxRank = "Class"),
               phyloseq_sep_tax(CSS_BacFun_rhizoplane_HighFilt_ps_l$Fun, TaxRank = "Order"),
              phyloseq_sep_tax(CSS_BacFun_rhizoplane_HighFilt_ps_l$Fun, TaxRank = "Family"),
              phyloseq_sep_tax(CSS_BacFun_rhizoplane_HighFilt_ps_l$Fun, TaxRank = "Genus"))


# remove taxa wihtout least 5 different ASVs, or UMAP crashes
# ps_taxa_l_fun<-ps_taxa_l_fun[lapply(ps_taxa_l, ntaxa)>5]

# taxa msut be present at least once in control, once SA and once in  MeJA on every plant species




# check if a taxon has zero ASVs across a species_stress combination; we can't calculate phylosig with that
has_bad_zero_fun<-lapply(ps_taxa_l_fun, function(x){
  
  melted<-psmelt(x)

melted_summrary<-melted%>%
  dplyr::group_by(Stress, Sp_abb_name)%>%
  dplyr::summarise(abu_sum = sum(Abundance))



taxon_missing_in_sp_Stress<-any(melted_summrary[,3]==0)


  
})


# remove taxa with zero ASVs in any species_stress combination
ps_taxa_l_fun<-ps_taxa_l_fun[!unlist(has_bad_zero_fun)]




#remove taxonomies with NA, such as p__NA
ps_taxa_l_fun<-ps_taxa_l_fun[names(ps_taxa_l_fun)!= "NAs"]

# remove samples with zero ASVs
samples_to_keep<-lapply(ps_taxa_l_fun, function(x)
  colSums(otu_table(x))>0)

ps_taxa_l_fun<-mapply(function(x, y)
  prune_samples(y,x),
  x = ps_taxa_l_fun,
  y = samples_to_keep,
  SIMPLIFY = FALSE)


# save list of ps objects, which takes  while to compute
save(ps_taxa_l_fun, file = "./14_Phylogenetic_distance/r_objects/ps_taxa_l_fun_pre_phylosig.RData") 



```





## 14.10.2 - calculate phylosig (PCoA, abouheif), save rdata


```{r}

# load list of fungal ps objects, which takes  while to compute
load(file = "./14_Phylogenetic_distance/r_objects/ps_taxa_l_fun_pre_phylosig.RData") 
ps_taxa_l_fun


# run pipeline
fun_alltaxa_phylosig_PCoA_nonsig<- phylogenetic_signal_pipeline(ps_l = ps_taxa_l_fun,
                                                     phylogenetic_test = "Abouheif's Cmean",
                                                     ordination_method = "PCoA",
                                                     significant_signal_only = FALSE,
                                                     return_plot = TRUE, 
                                                     higher_PC_only = FALSE,
                                                     return_lambda = FALSE)


save(fun_alltaxa_phylosig_PCoA_nonsig, 
     file = "./14_Phylogenetic_distance/r_objects/fun_alltaxa_phylosig_PCoA_nonsig.RData") 



```





# Mantel pipeline

```{r}

# run mantel pipeline on all taxonomic groups
fun_alltaxa_mantel_pipeline<-ps_to_mantel_pipeline(ps_obj_l = ps_taxa_l_fun)

# remove p/r mantel plots as 500 of them simultaneously are not informative
fun_alltaxa_mantel_pipeline$plot_output<-NULL

#adjsut name of "Community" to taxonomic group
names(fun_alltaxa_mantel_pipeline$mantel_df_output)[1]<-"taxonomic_group"

#save 500 plots
fun_alltaxa_mantel_pipeline_bray_correlation_plot<-fun_alltaxa_mantel_pipeline$bray_correlation_plot
save(fun_alltaxa_mantel_pipeline_bray_correlation_plot, 
     file = "./14_Phylogenetic_distance/r_objects/fun_alltaxa_mantel_pipeline_bray_correlation_plot.RData")

#save df
fun_alltaxa_mantel_pipeline_df<-fun_alltaxa_mantel_pipeline$mantel_df_output 
save(fun_alltaxa_mantel_pipeline_df, 
     file = "./14_Phylogenetic_distance/r_objects/fun_alltaxa_mantel_pipeline_df.RData")


```


