---
title: "Figure_3_import_calculate_plot"
author: "Pedro_Beschoren"
date: "20/04/2023"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: consolehtml_document
---


# 14.1 - set up working environemnt


```{r}
# this will load all key libraries for the whole project
source("./renv/0_load_libraries.R")
library(plyr)
library("umap")
library(RColorBrewer)
library(cowplot) 
library(multiblock)
library(data.table)
library(ggExtra)
library(multimode) # for testing multiple modes in dsitribution


# IMPORTANT: load the custom functions necessary to run this script
source("./14_Phylogenetic_distance/code/custom_functions_PB.R")


library(picante)
library(ape)
library(adephylo)
library(ade4)
library(phylobase)
library(geiger)
library(phytools)
library(vegan)
library(phyloseq)
library(dplyr)


# increases memory limit used by R
memory.limit(size = 350000)
```



# ------
# --- Bacteria --- 
# ------


# load processed R objects, BAC
```{r}

# load output of pipeline, Bac 
load(file ="./14_Phylogenetic_distance/r_objects/bac_alltaxa_phylosig_output_PCoA_nonsig.RData")



#load hundred of permanovas calculated on the HPC (output of permanova_per_taxonomic_group)
load(file = "./14_Phylogenetic_distance/r_objects/permanovas_spXstress_528.RData") 

# load hundreds of PCoA plots
load(file = "./14_Phylogenetic_distance/r_objects/PCoA_528_plots_axis_pct.RData") # one per tax group
load(file = "./14_Phylogenetic_distance/r_objects/PCoA_528_3x_plots_axis_pct.RData") # one per tax group and per stress

# load ps objects with overwrriten library size per taxonomic group (from 11h_library_size_on_phylosig.Rmd)
load(file = "./14_Phylogenetic_distance/r_objects/Bac_ps_taxa_l_overwrittenLibSize_ps_l.RData") 

#load output of mantel pipeline 
load(file = "./14_Phylogenetic_distance/r_objects/bac_alltaxa_mantel_pipeline_df.RData") #df
load(file = "./14_Phylogenetic_distance/r_objects/bac_alltaxa_mantel_pipeline_bray_correlation_plot.RData") #plots



```

# - 
# Prepare a df of results from 528 Permanovas + PCoAs


## extract p/r valeus from 528 permanovas, set base df
```{r}


#extract r/p values
perm_528_df<-
permanova_output_to_plot_input2(permanova_output_l = permanovas_spXstress_528$permanova_l)


# put % explained variaance per PC in a df
axis_pct_tax_group<-
  PCoA_plot_l_to_pct_variance_df(PCoA_528_plots_axis_pct)

axis_pct_tax_group_Control<-
  PCoA_plot_l_to_pct_variance_df(PCoA_528_3x_plots_axis_pct[1:3])
names(axis_pct_tax_group_Control)[2:3]<-paste0(names(axis_pct_tax_group_Control)[2:3], "_Control_only")

axis_pct_tax_group_MeJA<-
  PCoA_plot_l_to_pct_variance_df(PCoA_528_3x_plots_axis_pct[4:6])
names(axis_pct_tax_group_MeJA)[2:3]<-paste0(names(axis_pct_tax_group_MeJA)[2:3], "_MeJA_only")

axis_pct_tax_group_SA<-
  PCoA_plot_l_to_pct_variance_df(PCoA_528_3x_plots_axis_pct[7:9])
names(axis_pct_tax_group_SA)[2:3]<-paste0(names(axis_pct_tax_group_SA)[2:3], "_SA_only")

#put 528 PCoAs from the 3 stresses into 1 df
axis_pct_tax_group_per_stress<-
  cbind(axis_pct_tax_group_Control, axis_pct_tax_group_MeJA[,2:3], axis_pct_tax_group_SA[,2:3])


# widen the df
perm_528_df_wide<-
pivot_wider(data = perm_528_df,
            names_from = factor,
            values_from = c("pvalue", "R2"))

# widen phylosig df
df_phylosig_wide<-
pivot_wider(data = bac_alltaxa_phylosig_output_PCoA_nonsig$df_phylosig, 
            names_from = c("stress", "higher_lower_PC"),
            values_from = c("trait", "value", "pvalue", "pvalue_adjust"))

# remove unecessary info (phylosig test used, phylosig test adjsuted p)
df_phylosig_wide<-
  df_phylosig_wide[,-c(2, 21:26)]


#join 528 permanovas to widened df-phylosig
perm_528_phylosig_df_wide<-
  left_join(perm_528_df_wide, df_phylosig_wide, by = c("taxonomic_group" = "column_label"))

#add % explained varaince
perm_528_phylosig_pct_df<-
  left_join(perm_528_phylosig_df_wide, axis_pct_tax_group)

perm_528_phylosig_pct_df<-
  left_join(perm_528_phylosig_pct_df, axis_pct_tax_group_per_stress)



```

## Add library sizes to DF of 528 PERMANOVAS + PCoA
these updated library sizes objects were generated in script 11h_library_size_on_phylosig.Rmd
```{r}

#add library size (all 3 stresses together; that is the PERMANOVA input) to into a df 
libSizes<-
  lapply(ps_taxa_l_overwrittenLibSize, function(x) sum(x@sam_data$library_size))

libSizes_df<-
  as.data.frame(do.call("rbind", libSizes))%>%rownames_to_column()

names(libSizes_df)<-c("taxonomic_group", "Library_size")

#join with rest of 528 permanova df
perm_528_phylosig_pct_df<-
  left_join(perm_528_phylosig_pct_df, libSizes_df)

#add library size under each stress condition(split by taxonomic gorupto df 
libSizes_per_stress<-
  lapply(ps_taxa_l_overwrittenLibSize, function(x) {
    ps_per_stress<-phyloseq_sep_variable(x, var = "Stress")
      output<-lapply(ps_per_stress, function(z)
       sum(z@sam_data$library_size)) 
    return(output)
    })

libSizes_per_stress<-
  lapply(libSizes_per_stress, function(x) 
    unlist(x))

# put library sizes epr stress into a df
libSizes_per_stress_df<-
  as.data.frame(do.call("rbind", libSizes_per_stress))%>%rownames_to_column()

names(libSizes_per_stress_df)<-c("taxonomic_group", 
                                  "Library_size_Control", 
                                  "Library_size_MeJA", 
                                  "Library_size_SA")

#join with rest of 528 permanova df
perm_528_phylosig_pct_df<-
  left_join(perm_528_phylosig_pct_df, libSizes_per_stress_df)

# add column to define taxonomic rank
perm_528_phylosig_pct_df$taxonomic_rank<-
  factor(
        c(
           gsub(x = perm_528_phylosig_pct_df$taxonomic_group, pattern = "k__.*$",replacement = "Kingdom")%>%
           gsub(pattern = "p__.*$",replacement = "Phylum")%>%
           gsub(pattern = "o__.*$",replacement = "Order")%>%
           gsub(pattern = "c__.*$",replacement = "Class")%>%
           gsub(pattern = "f__.*$",replacement = "Family")%>%
           gsub(pattern = "g__.*$",replacement = "Genus")),
         levels = c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus"))

# define unexplained permanova variance
perm_528_phylosig_pct_df$residual_permanova<-
  1 - (perm_528_phylosig_pct_df$R2_Block +
          perm_528_phylosig_pct_df$R2_Sp_full_name +
          perm_528_phylosig_pct_df$R2_Sp_x_Stress + 
          perm_528_phylosig_pct_df$R2_Stress)

# define PCoA_1/ PCoA2 varaince aexplained ratio
perm_528_phylosig_pct_df$var_explined_ratio<-
  (perm_528_phylosig_pct_df$PCoA_1/perm_528_phylosig_pct_df$PCoA_2)



```



## define distance to centroids for the 3 experimental factors (Block, plant_sp, Stress)
```{r}

#define function to wrangle beta dispersion data
get_beta_disper_comparison<-function(Taxonomic_group_beta_disper){
  
dist_to_centroid_df<-
  as.data.frame(Taxonomic_group_beta_disper$beta_disper_block$distances)

names(dist_to_centroid_df)<-"distance_to_centroid"
dist_to_centroid_df<-rownames_to_column(dist_to_centroid_df, var = "sample_name")
dist_to_centroid_df$centroid_group<-"Block"
dist_to_centroid_df$group_ID<-Taxonomic_group_beta_disper$beta_disper_block$group


dist_to_centroid_df2<-
  as.data.frame(Taxonomic_group_beta_disper$beta_disper_sp$distances)

names(dist_to_centroid_df2)<-"distance_to_centroid"
dist_to_centroid_df2<-rownames_to_column(dist_to_centroid_df2, var = "sample_name")
dist_to_centroid_df2$centroid_group<-"plant_Sp"
dist_to_centroid_df2$group_ID<-Taxonomic_group_beta_disper$beta_disper_sp$group


dist_to_centroid_df3<-
  as.data.frame(Taxonomic_group_beta_disper$beta_disper_stress$distances)

names(dist_to_centroid_df3)<-"distance_to_centroid"
dist_to_centroid_df3<-rownames_to_column(dist_to_centroid_df3, var = "sample_name")
dist_to_centroid_df3$centroid_group<-"Stress"
dist_to_centroid_df3$group_ID<-Taxonomic_group_beta_disper$beta_disper_stress$group


dist_to_centroid_df<-
  rbind(dist_to_centroid_df,dist_to_centroid_df2,dist_to_centroid_df3)

#print(ggplot(dist_to_centroid_df, aes(x = centroid_group, y = distance_to_centroid))+
 # geom_boxplot())

return(dist_to_centroid_df)
}

#execute custom function above, generating a list of dfs
beta_disper_comparions_df_l<-lapply(permanovas_spXstress_528$beta_disper_l, function(z)
  get_beta_disper_comparison(Taxonomic_group_beta_disper = z))

#join list
beta_disper_comparions_df<-
  do.call("rbind", beta_disper_comparions_df_l)%>%rownames_to_column(var = "taxonomic_group")

#adjsut names
beta_disper_comparions_df$taxonomic_group<-
  gsub(pattern = "\\..*",replacement = "", x = beta_disper_comparions_df$taxonomic_group)

# add column to define taxonomic rank
beta_disper_comparions_df$taxonomic_rank<-
  factor(
        c(
           gsub(x = beta_disper_comparions_df$taxonomic_group, pattern = "k__.*$",replacement = "Kingdom")%>%
           gsub(pattern = "p__.*$",replacement = "Phylum")%>%
           gsub(pattern = "o__.*$",replacement = "Order")%>%
           gsub(pattern = "c__.*$",replacement = "Class")%>%
           gsub(pattern = "f__.*$",replacement = "Family")%>%
           gsub(pattern = "g__.*$",replacement = "Genus")),
         levels = c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus"))


#check summary
summarized_df_dist_to_centroid<-
group_by(.data = beta_disper_comparions_df, centroid_group, taxonomic_group, taxonomic_rank)%>%
  dplyr::summarize(mean_dist = mean(distance_to_centroid),
            median_dist = median(distance_to_centroid),
            sd_dist = sd(distance_to_centroid))


```



## define if beta distribution residuals are bimodal 
to be kept in the analysis, a taoxnomic group must have a unimodal distribution in distance to centroids for stress, plant species AND block

```{r}

#split beta diversity comparisons into a nested list for lapply
beta_disper_comparions_df_l<-
  split(x = beta_disper_comparions_df, f = beta_disper_comparions_df$taxonomic_group)

#run multimode test (takes several minutes to compute)
multimode_test_l_output<-
lapply(beta_disper_comparions_df_l, function(x){
  
  set.seed(12345)
  Block_modtest<-modetest(dplyr::filter(x, centroid_group =="Block")$distance_to_centroid)
    
  plantSp_modtest<-modetest(dplyr::filter(x, centroid_group =="plant_Sp")$distance_to_centroid)
  
  Stress_modtest<-modetest(dplyr::filter(x, centroid_group =="Stress")$distance_to_centroid)
  
  output<-list("Block_modtest" = Block_modtest,
               "plantSp_modtest" = plantSp_modtest,
               "Stress_modtest" = Stress_modtest)
  
  return(output)
  
})

# fix two odd family names that were compromising during data wrangling
names(multimode_test_l_output)[names(multimode_test_l_output) == "f__env"]<-"f__env.OPS_17"
names(multimode_test_l_output)[names(multimode_test_l_output) == "g__env"]<-"g__env.OPS_17"

#save externally
save(multimode_test_l_output, file = "./14_Phylogenetic_distance/r_objects/multimode_test_l_output.RData")
load(file = "./14_Phylogenetic_distance/r_objects/multimode_test_l_output.RData")


```

##  put multimode test results into a df, join with 528 permanova df
```{r}
#extract p values from the multimode test, put in lsit
multimode_p_l<-
lapply(multimode_test_l_output, function(x){
  block_p_multimode<-x$Block_modtest$p.value
  plantSp_p_multimode<-x$plantSp_modtest$p.value
  Stress_p_multimode<-x$Stress_modtest$p.value
  
  return(c("block_p_multimode" = block_p_multimode,
           "plantSp_p_multimode" = plantSp_p_multimode,
           "Stress_p_multimode" = Stress_p_multimode))
  
})

#turn list into df
multimode_p_df<-
  as.data.frame(do.call("rbind",multimode_p_l ))

#define if residuals are unimodal in the experimental factors ( number of experiemntal factros where residues are unimodal)
multimode_p_df$unimodal_residues<-rowCounts(multimode_p_df[,1:3]>0.05)
multimode_p_df$unimodal_residues001<-rowCounts(multimode_p_df[,1:3]>0.01)



#put the output of the beta dispersion unimodalisty testing in the permanova df
multimode_p_df<-rownames_to_column(multimode_p_df, var = "taxonomic_group")
perm_528_phylosig_pct_df<-left_join(perm_528_phylosig_pct_df,multimode_p_df)








```



## define if residual distributions are similar to control
```{r}


# run Kolmogorov-Smirnov to compare distribution of residuals with k__Bacteria
ks_l_output<-
lapply(beta_disper_comparions_df_l, function(x){
  # run Kolmogorov-Smirnov to define if two samples (k__bacteria and focal taxonomic group) derive from teh same distribution on a varaibles (scaled distances to centroids)
  
  ks_Block<-
    ks.test(x = scale(dplyr::filter(beta_disper_comparions_df_l$k__Bacteria, 
                         centroid_group == "Block" )$distance_to_centroid),
           y = scale(dplyr::filter(x, 
                             centroid_group == "Block" )$distance_to_centroid))
    
  ks_Stress<-
    ks.test(x = scale(dplyr::filter(beta_disper_comparions_df_l$k__Bacteria, 
                         centroid_group == "Stress" )$distance_to_centroid),
           y = scale(dplyr::filter(x, 
                             centroid_group == "Stress" )$distance_to_centroid))
  
   
   ks_plantSp<-
    ks.test(x = scale(dplyr::filter(beta_disper_comparions_df_l$k__Bacteria, 
                         centroid_group == "plant_Sp" )$distance_to_centroid),
           y = scale(dplyr::filter(x, 
                             centroid_group == "plant_Sp" )$distance_to_centroid))
  
  
  output<-list("ks_Block" = ks_Block,
               "ks_Stress" = ks_Stress,
               "ks_plantSp" = ks_plantSp)
  
  return(output)
  
})

#extract p values from the test, put in lsit
ks_l_pval<-
  lapply(ks_l_output, function(x){
    
  block_p_ks<-x$ks_Block$p.value
  plantSp_p_ks<-x$ks_plantSp$p.value
  Stress_p_ks<-x$ks_Stress$p.value
  
  return(c("block_p_ks" = block_p_ks,
           "plantSp_p_ks" = plantSp_p_ks,
           "Stress_p_ks" = Stress_p_ks))
  
})



   
   
   

#turn list into df
ks_pval_df<-
  as.data.frame(do.call("rbind",ks_l_pval ))

#define if residuals are unimodal in all experimental factors (as in number of experiemntal factros)
ks_pval_df$residue_dist_simialr_to_K__bacteria<-rowCounts(ks_pval_df[,1:3]>0.05)
ks_pval_df$residue_dist_simialr_to_K__bacteria001<-rowCounts(ks_pval_df[,1:3]>0.01)

#put the output of the beta dispersion unimodalisty testing in the permanova df
ks_pval_df<-rownames_to_column(ks_pval_df, var = "taxonomic_group")
perm_528_phylosig_pct_df<-left_join(perm_528_phylosig_pct_df,ks_pval_df)



```



## reshape mantel pipeline output to fit permanova df
```{r}

# remove Jaccard distances, as they provide the same conclusions as bray-curtis
bac_alltaxa_mantel_pipeline_df_reshaped<-
  dplyr::filter(bac_alltaxa_mantel_pipeline_df, distance =="Bray")

#remove column that identifies if  the test ws significant
bac_alltaxa_mantel_pipeline_df_reshaped<-
  dplyr::select(bac_alltaxa_mantel_pipeline_df_reshaped,  -significant)

# rename some variable to avoid confusion wiht abouheif tests
names(bac_alltaxa_mantel_pipeline_df_reshaped)[c(3,5)]<-c("distance_Mantel", "p_value_Mantel")

#widen df to have one row per taxonomic group
bac_alltaxa_mantel_pipeline_df_reshaped<-
  pivot_wider(bac_alltaxa_mantel_pipeline_df_reshaped, 
              names_from = "Treatment", 
              values_from = c("r2_Mantel", "p_value_Mantel"))



```



## add mantel pipeline outut to permanova df
```{r}

perm_528_phylosig_pct_df<-
  left_join(perm_528_phylosig_pct_df,bac_alltaxa_mantel_pipeline_df_reshaped)

```

## define PCoA variance explained by the PC with highest abouheif
```{r}
#re-arrange which is the hgiehr PC
perm_528_phylosig_pct_df$higher_PC_Control<-
  gsub(x = perm_528_phylosig_pct_df$trait_Control_Higher_PC, pattern = "_[a-z]_.*", replacement = "")

perm_528_phylosig_pct_df$higher_PC_MeJA<-
  gsub(x = perm_528_phylosig_pct_df$trait_MeJA_Higher_PC, pattern = "_[a-z]_.*", replacement = "")

perm_528_phylosig_pct_df$higher_PC_SA<-
  gsub(x = perm_528_phylosig_pct_df$trait_SA_Higher_PC, pattern = "_[a-z]_.*", replacement = "")

# define PCoA_1/ PCoA2 varaince aexplained ratio
perm_528_phylosig_pct_df$var_explined_ratio_Control<-
  (perm_528_phylosig_pct_df$PCoA_1_Control_only/perm_528_phylosig_pct_df$PCoA_2_Control_only)

perm_528_phylosig_pct_df$var_explined_ratio_MeJA<-
  (perm_528_phylosig_pct_df$PCoA_1_MeJA_only/perm_528_phylosig_pct_df$PCoA_2_MeJA_only)

perm_528_phylosig_pct_df$var_explined_ratio_SA<-
  (perm_528_phylosig_pct_df$PCoA_1_SA_only/perm_528_phylosig_pct_df$PCoA_2_SA_only)

# assing the abouheif value for each PCoA in each treatmet condition, (hard-coded)
perm_528_phylosig_pct_df$PCoA_1_Control_abouheif_value<-
  ifelse(test = perm_528_phylosig_pct_df$higher_PC_Control == "PCoA_1",
         yes = perm_528_phylosig_pct_df$value_Control_Higher_PC,
         no = perm_528_phylosig_pct_df$value_Control_Lower_PC)

perm_528_phylosig_pct_df$PCoA_2_Control_abouheif_value<-
  ifelse(test = perm_528_phylosig_pct_df$higher_PC_Control == "PCoA_2",
         yes = perm_528_phylosig_pct_df$value_Control_Higher_PC,
         no = perm_528_phylosig_pct_df$value_Control_Lower_PC)

perm_528_phylosig_pct_df$PCoA_1_MeJA_abouheif_value<-
  ifelse(test = perm_528_phylosig_pct_df$higher_PC_MeJA == "PCoA_1",
         yes = perm_528_phylosig_pct_df$value_MeJA_Higher_PC,
         no = perm_528_phylosig_pct_df$value_MeJA_Lower_PC)

perm_528_phylosig_pct_df$PCoA_2_MeJA_abouheif_value<-
  ifelse(test = perm_528_phylosig_pct_df$higher_PC_MeJA == "PCoA_2",
         yes = perm_528_phylosig_pct_df$value_MeJA_Higher_PC,
         no = perm_528_phylosig_pct_df$value_MeJA_Lower_PC)

perm_528_phylosig_pct_df$PCoA_1_SA_abouheif_value<-
  ifelse(test = perm_528_phylosig_pct_df$higher_PC_SA == "PCoA_1",
         yes = perm_528_phylosig_pct_df$value_SA_Higher_PC,
         no = perm_528_phylosig_pct_df$value_SA_Lower_PC)

perm_528_phylosig_pct_df$PCoA_2_SA_abouheif_value<-
  ifelse(test = perm_528_phylosig_pct_df$higher_PC_SA == "PCoA_2",
         yes = perm_528_phylosig_pct_df$value_SA_Higher_PC,
         no = perm_528_phylosig_pct_df$value_SA_Lower_PC)

# assing the % PCoA explained in the PCoA wth higer abouheif
perm_528_phylosig_pct_df$higher_PC_Control_pct<-
  ifelse(test = perm_528_phylosig_pct_df$higher_PC_Control == "PCoA_1",
         yes = perm_528_phylosig_pct_df$PCoA_1_Control_only,
         no = perm_528_phylosig_pct_df$PCoA_2_Control_only)

perm_528_phylosig_pct_df$lower_PC_Control_pct<-
  ifelse(test = perm_528_phylosig_pct_df$higher_PC_Control == "PCoA_2",
         yes = perm_528_phylosig_pct_df$PCoA_1_Control_only,
         no = perm_528_phylosig_pct_df$PCoA_2_Control_only)

perm_528_phylosig_pct_df$higher_PC_MeJA_pct<-
  ifelse(test = perm_528_phylosig_pct_df$higher_PC_MeJA == "PCoA_1",
         yes = perm_528_phylosig_pct_df$PCoA_1_MeJA_only,
         no = perm_528_phylosig_pct_df$PCoA_2_MeJA_only)

perm_528_phylosig_pct_df$lower_PC_MeJA_pct<-
  ifelse(test = perm_528_phylosig_pct_df$higher_PC_MeJA == "PCoA_2",
         yes = perm_528_phylosig_pct_df$PCoA_1_MeJA_only,
         no = perm_528_phylosig_pct_df$PCoA_2_MeJA_only)
          
perm_528_phylosig_pct_df$higher_PC_SA_pct<-
  ifelse(test = perm_528_phylosig_pct_df$higher_PC_SA == "PCoA_1",
         yes = perm_528_phylosig_pct_df$PCoA_1_SA_only,
         no = perm_528_phylosig_pct_df$PCoA_2_SA_only)

perm_528_phylosig_pct_df$lower_PC_SA_pct<-
  ifelse(test = perm_528_phylosig_pct_df$higher_PC_SA == "PCoA_2",
         yes = perm_528_phylosig_pct_df$PCoA_1_SA_only,
         no = perm_528_phylosig_pct_df$PCoA_2_SA_only)

```



### save processed 528 PERMANOVAS df

```{r}
# save df of 528 permanovas, % variance, Abouheif values, beta dispersion results
save(perm_528_phylosig_pct_df, file = "./14_Phylogenetic_distance/r_objects/perm_528_phylosig_pct_df.RData")
load(file = "./14_Phylogenetic_distance/r_objects/perm_528_phylosig_pct_df.RData")
view(perm_528_phylosig_pct_df)
```

# - 
# data structure check: scree plots, histograms, beta dispersion, PCoAs

## define selected taxonomic groups for a few examples
```{r}
#set a few groups as an example for beta dispersion comparsions (shapes of distributions)
taxonomic_groups_for_dispersion_example<-c("k__Bacteria",
                                            "p__Proteobacteria",
                                            "p__Deinococcota",
                                            "c__Vampirivibrionia",
                                            "c__Alphaproteobacteria",
                                            "o__Rhizobiales",
                                            "o__Fibrobacterales",
                                            "f__Burkholderiaceae",
                                            "f__Chitinophagaceae",
                                            "g__Massilia",
                                            "f__Spirosomaceae",
                                            "g__Bacteriovorax")

```


## unexplained residuals of the permanovas
```{r}

#check sumamry of PERMANOVA  unexplained varaicne, all taxonomic groups together
permanova_r2_residual_quartiles<-
  perm_528_phylosig_pct_df%>%
    dplyr::summarise(mean = mean(residual_permanova),
                     median = median(residual_permanova),
                     sd = sd(residual_permanova),
                     quartile_1st = quantile(residual_permanova)[[2]],
                     quartile_3rd = quantile(residual_permanova)[[4]],
                     max = max(residual_permanova))

#histogram of unexplauined permanova varince
residual_permanova_histogram<-
ggplot(perm_528_phylosig_pct_df, aes(x = residual_permanova))+
  geom_histogram()+
  geom_vline(xintercept = 0.85, color = "red")+
  geom_label(label= paste("3rd quartile of distribution:", 
                          round(permanova_r2_residual_quartiles[[1,5]],3), 
                          sep = " "),
             x = 0.65,
             y = 60)+
  xlab("R2 of residuals in PERMANOVAs")+
  ylab("Number of taxonomic groups")


permanova_r2_residual_quartiles_perrank<-
perm_528_phylosig_pct_df%>%
group_by(taxonomic_rank)%>%
  dplyr::summarise(mean = mean(residual_permanova),
                   median = median(residual_permanova),
                   sd = sd(residual_permanova),
                   quartile_1st = quantile(residual_permanova)[[2]],
                   quartile_3rd = quantile(residual_permanova)[[4]],
                   max = max(residual_permanova))


#histogram of unexplauined permanova varince
residual_permanova_histogram_perrank<-
ggplot(perm_528_phylosig_pct_df, aes(x = residual_permanova))+
  geom_histogram()+
  facet_wrap(~taxonomic_rank)+
  geom_vline(xintercept = 0.85, color = "red")+
  xlab("R2 of residuals in PERMANOVAs")+
  ylab("Number of taxonomic groups")

#save plot externally
ggsave(plot = ggarrange(residual_permanova_histogram,
                        residual_permanova_histogram_perrank,
                        labels = "AUTO"),
       filename = "./14_Phylogenetic_distance/results/Results_April2023/histogram_permanova_residuals.svg",
       width = 250,
       height = 90,
       units = "mm")

```


## screeplot of explained variance per PC, with broken stick
```{r}

# preapre to save pdf file
pdf("./14_Phylogenetic_distance/results/Results_April2023/Eigenvalues_f__Chitinophagaceae_top_f__Burkholderiaceae_bottom.pdf",
    width = 20,
    height = 7) 

# screeplot of explained variance per PC, with broken stick
evplot(PCoA_528_3x_plots_axis_pct$PCoA_l_Control$f__Chitinophagaceae$values$Eigenvalues)
evplot(PCoA_528_3x_plots_axis_pct$PCoA_l_Control$f__Burkholderiaceae$values$Eigenvalues)
dev.off() 


```


## selected PCoA plots
```{r}


# put the same taxons as PCoA plots
mode_type_PCoA_pannel_l<-
  PCoA_528_plots_axis_pct$ploted_ordination[taxonomic_groups_for_dispersion_example]

mode_type_PCoA_pannel<-
ggarrange(plotlist = mode_type_PCoA_pannel_l, common.legend = TRUE)

#put histograms next to PCoA plots
ggsave(plot = mode_type_PCoA_pannel,
       filename = "./14_Phylogenetic_distance/results/Results_April2023/PCoAs_selected_examples.pdf",
       width = 300,
       height = 250,
       units = "mm")

```

##  ratio of PCoA explained variance (PCoA1/PCoA2)
```{r}
#check sumamry of PCOA  varaicne 9PC_1 /PC_2)
var_explain_ratio_df<-
perm_528_phylosig_pct_df%>%
  dplyr::summarise(mean = mean(var_explined_ratio),
                   median = median(var_explined_ratio),
                   sd = sd(var_explined_ratio),
                   quartile_1st = quantile(var_explined_ratio)[[2]],
                   quartile_3rd = quantile(var_explined_ratio)[[4]],
                   max = max(var_explined_ratio))


#histogram of PCoA variance explained, all 528 taxonomic groups
hist_var_explained_ratio<-
ggplot(perm_528_phylosig_pct_df, aes(x = var_explined_ratio))+
  geom_histogram()+
  geom_vline(xintercept = 1.25, color = "red")+
  geom_label(label= paste("1st quartile of distribution:", 
                          round(var_explain_ratio_df[[1,4]],3), 
                          sep = " "),
             x = 4,
             y = 100)+
  xlab("% of variance explained in PCoA_1 / \n % of variance explained in PCoA_2")+
  ylab("Number of taxonomic groups")




# check quartiles per taxonomic rank
perm_528_phylosig_pct_df%>%
group_by(taxonomic_rank)%>%
  dplyr::summarise(mean = mean(var_explined_ratio),
                   median = median(var_explined_ratio),
                   sd = sd(var_explined_ratio),
                   quartile_1st = quantile(var_explined_ratio)[[2]],
                   quartile_3rd = quantile(var_explined_ratio)[[4]],
                   max = max(var_explined_ratio))


#histogram of PCoA variance explained, split per taxonomic rank
hist_var_explained_ratio_perrank<-
ggplot(perm_528_phylosig_pct_df, aes(x = var_explined_ratio))+
  geom_histogram()+
  facet_wrap(~taxonomic_rank)+
  geom_vline(xintercept = 1.25, color = "red")+
   xlab("% of variance explained in PCoA_1 / \n % of variance explained in PCoA_2")+
  ylab("Number of taxonomic groups")


#save plot externally
ggsave(plot = ggarrange(hist_var_explained_ratio,
                        hist_var_explained_ratio_perrank,
                        labels = "AUTO"),
       filename = "./14_Phylogenetic_distance/results/Results_April2023/histogram_var_explained_ratio.svg",
       width = 250,
       height = 90,
       units = "mm")


```




##  plots of beta dispersion
```{r}
# plots of beta dispersion
pdf("./14_Phylogenetic_distance/results/Results_April2023/betadisper_k__Bacteria_c__Vampirivibrionia.pdf",
    width = 7,
    height = 7) 


plot(permanovas_spXstress_528$beta_disper_l$k__Bacteria$beta_disper_block)
plot(permanovas_spXstress_528$beta_disper_l$k__Bacteria$beta_disper_sp)
plot(permanovas_spXstress_528$beta_disper_l$k__Bacteria$beta_disper_stress)

plot(permanovas_spXstress_528$beta_disper_l$c__Vampirivibrionia$beta_disper_block)
plot(permanovas_spXstress_528$beta_disper_l$c__Vampirivibrionia$beta_disper_sp)
plot(permanovas_spXstress_528$beta_disper_l$c__Vampirivibrionia$beta_disper_stress)
dev.off() 


                          


plot(permanovas_spXstress_528$beta_disper_l$f__Spirosomaceae$beta_disper_block)




```



##  beta dispersion histgram next to PCoA of a few taxonomies

```{r}
#subset df, adjust order of factors to reflect artefacts
df_for_betadisp_histogram<-
  dplyr::filter(beta_disper_comparions_df, taxonomic_group  %in% taxonomic_groups_for_dispersion_example)

df_for_betadisp_histogram$taxonomic_group<-factor(df_for_betadisp_histogram$taxonomic_group, 
                                                  levels = taxonomic_groups_for_dispersion_example)



# check p values for multimodal testing and similarity to distribution in control
multimode_p_df_filt<-
  multimode_p_df[multimode_p_df$taxonomic_group %in% taxonomic_groups_for_dispersion_example,1:4]

ks_pval_df_filt<-
  ks_pval_df[ks_pval_df$taxonomic_group %in% taxonomic_groups_for_dispersion_example,1:4]

ks_pval_df_filt[,2:4]<-round(ks_pval_df_filt[,2:4], 3) # round up numbers, reduce decimals






# join dfs
multimode_ks_df_filt<-
  left_join(multimode_p_df_filt, ks_pval_df_filt)

multimode_ks_df_filt<-
  pivot_longer(data = multimode_ks_df_filt, 
              cols = !taxonomic_group,
              names_to ="Experimental_factor",
              values_to = "p_values")

multimode_ks_df_filt<-
  tidyr::separate(data = multimode_ks_df_filt,
                  col = Experimental_factor,
                  into = c("centroid_group", "test"),
                  sep = "_p_")

# export table with results, so they can be pasted, by hand, on the histogram on inskcape. sorry!
write.table(multimode_ks_df_filt,
            file = "./14_Phylogenetic_distance/results/Results_April2023/betadisper_histograms_tests_pval.txt")

#check a few histograms of distances to centroids
Histogram_pannel<-
df_for_betadisp_histogram%>%
  ggplot(aes(x = distance_to_centroid, fill = centroid_group))+
           geom_histogram()+
  scale_fill_discrete(name = "Experimental factor", 
                      labels = c("Block", "Plant species", "Experimental treatment"))+
  theme(legend.position="top")+
  ylab("Number of datapoints (samples)")+
  xlab("Distance to respective centroid")+
  facet_wrap(~taxonomic_group) 





#save histograms of ebta dispersion
ggsave(plot =Histogram_pannel,
       filename = "./14_Phylogenetic_distance/results/Results_April2023/betadisper_histograms.pdf",
       width = 250,
       height = 200,
       units = "mm")






```

## mantel correlation r2 and p for chitinophageaceae and burkholderiacea

```{r}
#subset df
mantel_f__Chitinophagaceae_f__Burkholderiaceae_df<-
  dplyr::filter(bac_alltaxa_mantel_pipeline_df, 
                taxonomic_group %in% c("f__Chitinophagaceae", "f__Burkholderiaceae"),
                distance == "Bray")

#rename first column t fit old function
names(mantel_f__Chitinophagaceae_f__Burkholderiaceae_df)[1]<-"Community"

# make individual mantel plot
mantel_example_f__Chitinophagaceae_f__Burkholderiaceae<-
  plot_mantel_correlations(mantel_f__Chitinophagaceae_f__Burkholderiaceae_df, quoted_plot_title = "")

#export example mantel plot with r2 and p values
ggsave(mantel_example_f__Chitinophagaceae_f__Burkholderiaceae,
       width = 180,
       height = 120,
       units = "mm",
       filename = "./14_Phylogenetic_distance/results/Results_April2023/mantel_r2_p_f__Chitinophagaceae_f__Burkholderiaceae.pdf",)



#export example mantel plot pariiwise comparisons among samples
ggsave(ggarrange(bac_alltaxa_mantel_pipeline_bray_correlation_plot$f__Chitinophagaceae,
                 bac_alltaxa_mantel_pipeline_bray_correlation_plot$f__Burkholderiaceae,
                 ncol = 2),
       width = 450,
       height = 350,
       units = "mm",
       filename = "./14_Phylogenetic_distance/results/Results_April2023/mantel_pairwise_f__Chitinophagaceae_f__Burkholderiaceae.pdf",)

```



## PCoA plot of Chitinophagaceae for figure S2 
this can be useful for the figure S2 pannel
```{r}

# add extended lineage as metadata
  ps_example_taxa_l<- lapply(ps_example_taxa_l, function(x){
    
  x@sam_data$Sp_Lineage_Walden_extended<-
    if_else(condition = x@sam_data$Sp_abb_name %in% c("Ia", "Lm", "Co", "Bi"), 
            true ="lineage_II_extended",
            false =  x@sam_data$Sp_Lineage_Walden)
  
    return(x)
  
  })



# split datset for plotting with PCoA_calculate_extract_add_to_ps function
Chitinophagaceae_stress_l_l<-lapply(ps_example_taxa_l[7], function(x)
  phyloseq_sep_variable(physeq = x, variable = "Stress"))



# calcualte PCoA and add cordiantes t ps object
Chitinophagaceae_stress_l_l_coordinates<-
PCoA_calculate_extract_add_to_ps(input_ps_l_l = Chitinophagaceae_stress_l_l,
                                 original_ps_l = ps_example_taxa_l[7])


#extract metadata/PCoA coordinates of Chitinophagaceae SA as df
df_Chitinophagaceae<-as(sample_data(subset_samples(Chitinophagaceae_stress_l_l_coordinates$f__Chitinophagaceae, Stress =="SA")),
                 "data.frame")

#calcualte centroid
cent <- aggregate(cbind(PCoA_1, PCoA_2) ~ Sp_abb_name, data = df_Chitinophagaceae, FUN = mean)

#plot PCoA with enphais on Co
Chitinophagaceae_SA_enphasys_Co<-
ggplot(df_Chitinophagaceae, aes(x = PCoA_1, y = PCoA_2, color = Sp_Lineage_Walden_extended))+
  geom_point()+
  geom_point(size = ifelse(df_Chitinophagaceae$Sp_abb_name == "Co",4,1))+
    geom_label(data=cent, 
             aes(label=Sp_abb_name, 
                 color = NULL))+
  theme(plot.title = element_text(size = 10, face = "bold")) +
  theme(legend.position="top")

#save plto for further editing in inkscape
ggsave(filename = "./14_Phylogenetic_distance/results/Results_April2023/Chitinophagaceae_SA_PCoA_centroids.svg",
       plot = Chitinophagaceae_SA_enphasys_Co,
       width = 110, 
       height = 110, 
       units = "mm")




```


## abouheif histogram of Chitinophagaceae as part of figure 2 pannel
in this chunk we: 1) prepare to save the pdf file, 2) rerun pipeline with show_abouheif_histogram = TRUE, return_plot = FALSE; 3) close plot saving device, creating the pdf file
```{r}

#save Niastella  PCoA Abouheif histogram (last plot) as pdf
 pdf(file = "./14_Phylogenetic_distance/results/Results_April2023/Chitinophagaceae_Contol_MeJA_SA_PCoA_Abouheif_histogram.pdf",     
     width = 5, 
     height = 5) 
 

 phylogenetic_signal_pipeline(ps_l = ps_example_taxa_l[7],
                              phylogenetic_test = "Abouheif's Cmean",
                              ordination_method = "PCoA",
                              significant_signal_only = TRUE,
                              return_plot = FALSE, 
                              higher_PC_only = FALSE,
                              return_lambda = FALSE, 
                              show_abouheif_histogram = TRUE)


dev.off()

```


# - 
#  PCA visualization of permanova output: high abouheif PCoA VS low abouheif PCoA


this is more convinient, and adequate, than comparing PCoA_1 v PCoA_2 directlty

## PCA Of 528X3 PCOAS, PCoA with higest abouheif vs PCoA with lowest abouheif

```{r}
##### High_low PC instead of PCoA 1 and PCOA2

# subset, removing unecessary columns
abouhf_528x3_phylosig_df_subset_highlow<-
dplyr::select(perm_528_phylosig_pct_df, 
              pvalue_Sp_x_Stress, 
              pvalue_Sp_full_name,
              value_Control_Higher_PC,
              value_Control_Lower_PC,
              value_MeJA_Higher_PC, 
              value_MeJA_Lower_PC,
              value_SA_Higher_PC,
              value_SA_Lower_PC,
              higher_PC_Control_pct,
              higher_PC_MeJA_pct,
              higher_PC_SA_pct,
              lower_PC_Control_pct,
              lower_PC_MeJA_pct,
              lower_PC_SA_pct,
              var_explined_ratio_Control,
              var_explined_ratio_MeJA,
              var_explined_ratio_SA,
              Library_size_Control,
              Library_size_MeJA,
              Library_size_SA)


#tunr library size to log scale
abouhf_528x3_phylosig_df_subset_highlow$Library_size_Control<-
  log(abouhf_528x3_phylosig_df_subset_highlow$Library_size_Control)
abouhf_528x3_phylosig_df_subset_highlow$Library_size_MeJA<-
  log(abouhf_528x3_phylosig_df_subset_highlow$Library_size_MeJA)
abouhf_528x3_phylosig_df_subset_highlow$Library_size_SA<-
  log(abouhf_528x3_phylosig_df_subset_highlow$Library_size_SA)



# only significant  interaction
pca_res_528x3_abouheifs_highlow<-
  prcomp(abouhf_528x3_phylosig_df_subset_highlow,
         scale. = TRUE)


pca_plot_528x3_abouheifs_highlow<-
autoplot(pca_res_528x3_abouheifs_highlow, 
         data =  perm_528_phylosig_pct_df,
         colour = "taxonomic_rank",
         loadings = TRUE, 
         loadings.colour = 'blue', 
          size = 3,
         loadings.label = TRUE, 
         loadings.label.repel=T,
         loadings.label.size = 3)+
         theme(legend.position="bottom")




```





## Correlation plots related to the  528x3 PCoAs, PCoA with higest abouheif vs PCoA with lowest abouhei
```{r}
### plots of higher and lower PCs
corplot_control_abuheif_high_PC<-
ggplot(perm_528_phylosig_pct_df,
       aes(x = higher_PC_Control_pct,y = value_Control_Higher_PC, color =taxonomic_rank, alpha = 0.8))+
     geom_point( aes(size = log(Library_size_Control)))+
     #facet_wrap(~Stress)+
  stat_cor(method = "pearson",
           p.accuracy = 0.001, 
           r.accuracy = 0.01,
           aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~")),
           color = "Black")+
  geom_smooth(method = "lm",
              formula = y ~ x,
              colour = "Black" )  +
  xlab("% var. explained in PCoA axis with \n higher abouheif value, Control samples")+
  ylab("Higher of two Abouheif values, Control Samples")

corplot_control_abuheif_low_pc<-
ggplot(perm_528_phylosig_pct_df, 
       aes(x = lower_PC_Control_pct,y = value_Control_Lower_PC, color =taxonomic_rank, alpha = 0.8))+
     geom_point( aes(size = log(Library_size_Control)))+
     #facet_wrap(~Stress)+
  stat_cor(method = "pearson",
           p.accuracy = 0.001, 
           r.accuracy = 0.01,
           aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~")),
           color = "Black")+
  geom_smooth(method = "lm",
              formula = y ~ x,
              colour = "Black" )  +
  xlab("% var. explained in PCoA axis with \n lower abouheif value, Control samples")+
  ylab("Lower of two Abouheif values, Control Samples")




corplot_MeJA_abuheif_high_PC<-
ggplot(perm_528_phylosig_pct_df,
       aes(x = higher_PC_MeJA_pct,y = value_MeJA_Higher_PC, color =taxonomic_rank, alpha = 0.8))+
     geom_point( aes(size = log(Library_size_MeJA)))+
     #facet_wrap(~Stress)+
  stat_cor(method = "pearson",
           p.accuracy = 0.001, 
           r.accuracy = 0.01,
           aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~")),
           color = "Black")+
  geom_smooth(method = "lm",
              formula = y ~ x,
              colour = "Black" )  +
  xlab("% var. explained in PCoA axis with \n higher abouheif value, MeJA samples")+
  ylab("Higher of two Abouheif values, MeJA Samples")

corplot_MeJA_abuheif_low_pc<-
ggplot(perm_528_phylosig_pct_df, 
       aes(x = lower_PC_MeJA_pct,y = value_MeJA_Lower_PC, color =taxonomic_rank, alpha = 0.8))+
     geom_point( aes(size = log(Library_size_MeJA)))+
     #facet_wrap(~Stress)+
  stat_cor(method = "pearson",
           p.accuracy = 0.001, 
           r.accuracy = 0.01,
           aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~")),
           color = "Black")+
  geom_smooth(method = "lm",
              formula = y ~ x,
              colour = "Black" )  +
  xlab("% var. explained in PCoA axis with \n lower abouheif value, MeJA samples")+
  ylab("Lower of two Abouheif values, MeJA Samples")




corplot_SA_abuheif_high_PC<-
ggplot(perm_528_phylosig_pct_df,
       aes(x = higher_PC_SA_pct,y = value_SA_Higher_PC, color =taxonomic_rank, alpha = 0.8))+
     geom_point( aes(size = log(Library_size_SA)))+
     #facet_wrap(~Stress)+
  stat_cor(method = "pearson",
           p.accuracy = 0.001, 
           r.accuracy = 0.01,
           aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~")),
           color = "Black")+
  geom_smooth(method = "lm",
              formula = y ~ x,
              colour = "Black" )  +
  xlab("% var. explained in PCoA axis with \n higher abouheif value, SA samples")+
  ylab("Higher of two Abouheif values, SA Samples")

corplot_SA_abuheif_low_pc<-
ggplot(perm_528_phylosig_pct_df, 
       aes(x = lower_PC_SA_pct,y = value_SA_Lower_PC, color =taxonomic_rank, alpha = 0.8))+
     geom_point( aes(size = log(Library_size_SA)))+
     #facet_wrap(~Stress)+
  stat_cor(method = "pearson",
           p.accuracy = 0.001, 
           r.accuracy = 0.01,
           aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~")),
           color = "Black")+
  geom_smooth(method = "lm",
              formula = y ~ x,
              colour = "Black" )+
  xlab("% var. explained in PCoA axis with \n lower abouheif value, SA samples")+
  ylab("Lower of two Abouheif values, SA Samples")



#panels of high_low abouheif PCoA splits
panel_high_low_PCoA_Abouheif<-
ggarrange(corplot_control_abuheif_high_PC + facet_wrap(~taxonomic_rank, nrow = 1),
          corplot_MeJA_abuheif_high_PC + facet_wrap(~taxonomic_rank, nrow = 1),
          corplot_SA_abuheif_high_PC + facet_wrap(~taxonomic_rank, nrow = 1),
          corplot_control_abuheif_low_pc + facet_wrap(~taxonomic_rank, nrow = 1),
          corplot_MeJA_abuheif_low_pc + facet_wrap(~taxonomic_rank, nrow = 1),
          corplot_SA_abuheif_low_pc + facet_wrap(~taxonomic_rank, nrow = 1),
          common.legend = TRUE, 
          legend = "bottom",
            ncol = 1, 
            nrow = 6)




### plots of higher and lower PCs library size VS aouheif sinal
corplot_control_LibSize_abuheif_high_PC<-
ggplot(perm_528_phylosig_pct_df,
       aes(x = log(Library_size_Control),y = value_Control_Higher_PC, color =taxonomic_rank, alpha = 0.8))+
     geom_point( aes(size = log(Library_size_Control)))+
     #facet_wrap(~Stress)+
  stat_cor(method = "pearson",
           p.accuracy = 0.001, 
           r.accuracy = 0.01,
           aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~")),
           color = "Black")+
  geom_smooth(method = "lm",
              formula = y ~ x,
              colour = "Black" )  +
  xlab("log of Library size, Control samples")+
  ylab("Higher of two Abouheif values, Control Samples")

corplot_control_LibSize_abuheif_low_pc<-
ggplot(perm_528_phylosig_pct_df, 
       aes(x = log(Library_size_Control),y = value_Control_Lower_PC, color =taxonomic_rank, alpha = 0.8))+
     geom_point( aes(size = log(Library_size_Control)))+
     #facet_wrap(~Stress)+
  stat_cor(method = "pearson",
           p.accuracy = 0.001, 
           r.accuracy = 0.01,
           aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~")),
           color = "Black")+
  geom_smooth(method = "lm",
              formula = y ~ x,
              colour = "Black" )  +
  xlab("log of Library size, Control samples")+
  ylab("Lower of two Abouheif values, Control Samples")




corplot_MeJA_LibSize_abuheif_high_PC<-
ggplot(perm_528_phylosig_pct_df,
       aes(x = log(Library_size_MeJA),y = value_MeJA_Higher_PC, color =taxonomic_rank, alpha = 0.8))+
     geom_point( aes(size = log(Library_size_MeJA)))+
     #facet_wrap(~Stress)+
  stat_cor(method = "pearson",
           p.accuracy = 0.001, 
           r.accuracy = 0.01,
           aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~")),
           color = "Black")+
  geom_smooth(method = "lm",
              formula = y ~ x,
              colour = "Black" )  +
  xlab("log of Library size, MeJA samples")+
  ylab("Higher of two Abouheif values, MeJA Samples")

corplot_MeJA_LibSize_abuheif_low_pc<-
ggplot(perm_528_phylosig_pct_df, 
       aes(x = log(Library_size_MeJA),y = value_MeJA_Lower_PC, color =taxonomic_rank, alpha = 0.8))+
     geom_point( aes(size = log(Library_size_MeJA)))+
     #facet_wrap(~Stress)+
  stat_cor(method = "pearson",
           p.accuracy = 0.001, 
           r.accuracy = 0.01,
           aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~")),
           color = "Black")+
  geom_smooth(method = "lm",
              formula = y ~ x,
              colour = "Black" )  +
  xlab("log of Library size, MeJA samples")+
  ylab("Lower of two Abouheif values, MeJA Samples")




corplot_SA_LibSize_abuheif_high_PC<-
ggplot(perm_528_phylosig_pct_df,
       aes(x = log(Library_size_SA),y = value_SA_Higher_PC, color =taxonomic_rank, alpha = 0.8))+
     geom_point( aes(size = log(Library_size_SA)))+
     #facet_wrap(~Stress)+
  stat_cor(method = "pearson",
           p.accuracy = 0.001, 
           r.accuracy = 0.01,
           aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~")),
           color = "Black")+
  geom_smooth(method = "lm",
              formula = y ~ x,
              colour = "Black" )  +
  xlab("log of Library size, SA samples")+
  ylab("Higher of two Abouheif values, SA Samples")

corplot_SA_LibSize_abuheif_low_pc<-
ggplot(perm_528_phylosig_pct_df, 
       aes(x = log(Library_size_SA),y = value_SA_Lower_PC, color =taxonomic_rank, alpha = 0.8))+
     geom_point( aes(size = log(Library_size_SA)))+
     #facet_wrap(~Stress)+
  stat_cor(method = "pearson",
           p.accuracy = 0.001, 
           r.accuracy = 0.01,
           aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~")),
           color = "Black")+
  geom_smooth(method = "lm",
              formula = y ~ x,
              colour = "Black" )+
  xlab("log of Library size, SA samples")+
  ylab("Lower of two Abouheif values, SA Samples")
```

## make Panel with PCoA + correlations, 528x3 PCoA  with higest abouheif vs PCoA with lowest abouheif
```{r}
#panels of high_low abouheif PCoA splits
panel_high_low_PCoA_LibSize_Abouheif<-
ggarrange(corplot_control_LibSize_abuheif_high_PC + facet_wrap(~taxonomic_rank, nrow = 1),
          corplot_MeJA_LibSize_abuheif_high_PC + facet_wrap(~taxonomic_rank, nrow = 1),
          corplot_SA_LibSize_abuheif_high_PC + facet_wrap(~taxonomic_rank, nrow = 1),
          corplot_control_LibSize_abuheif_low_pc + facet_wrap(~taxonomic_rank, nrow = 1),
          corplot_MeJA_LibSize_abuheif_low_pc + facet_wrap(~taxonomic_rank, nrow = 1),
          corplot_SA_LibSize_abuheif_low_pc + facet_wrap(~taxonomic_rank, nrow = 1),
          common.legend = TRUE, 
          legend = "bottom",
            ncol = 1, 
            nrow = 6)



# join with respective PCA plot of varaibles
panel_high_low_Abouheif_528_PCA<-
  ggarrange(pca_plot_528x3_abouheifs_highlow,
            panel_high_low_PCoA_LibSize_Abouheif,
            panel_high_low_PCoA_Abouheif,
            ncol = 3, 
            widths = c(1,2,2))

ggsave(panel_high_low_Abouheif_528_PCA,
       filename = "./14_Phylogenetic_distance/results/Results_April2023/PCA_panel_highlow_528_Abouheif.svg",
       width = 700,
       height = 525,
       units = "mm")





```































#-
# filter the 528 permanovas

## define taxa that pass and fail the filtering

```{r}




########## filtering to permanovas of interest
filtered_permavonas<-
  dplyr::filter(perm_528_phylosig_pct_df, 
                pvalue_Sp_x_Stress<0.05 , # keep significant interaction 
                residual_permanova<0.85, # remove very ghie residuals (top qurtile of unexplained PERMANOVA varaince)
                var_explined_ratio>1.25,# remove taxas whose first PC dosen't explain at least 1.25 times PC2 (first quartile of PCOAs)
                unimodal_residues001 >=2,
               residue_dist_simialr_to_K__bacteria001 >=2)


#remove f__uncultured and g__uncultured.. they put different unidentified microbes together
filtered_permavonas<-filtered_permavonas[filtered_permavonas$taxonomic_group != "f__uncultured",]
filtered_permavonas<-filtered_permavonas[filtered_permavonas$taxonomic_group != "g__uncultured",]


#save externally
save(filtered_permavonas, file = "./14_Phylogenetic_distance/r_objects/filtered_permavonas.RData")


#define taxaonomic groups that passed all filters
filtered_taxonomic_groups<-filtered_permavonas$taxonomic_group

#make subset of plots,
filtered_PCOA_plots<-
  PCoA_528_plots_axis_pct$ploted_ordination[filtered_taxonomic_groups]

#arrange in a single plot
filtered_PCOA_plots<-
ggarrange(plotlist = filtered_PCOA_plots,
          common.legend = TRUE)

#save
ggsave(filtered_PCOA_plots, 
       filename = "./14_Phylogenetic_distance/results/Results_April2023/Filtered_PCOA_plots_3filt_p001min2.pdf",
     width = 1900, 
     height = 1900, 
     limitsize  = FALSE,
     units = "mm")



#define taxaonomic groups that FAILED all filters
discarded_taxonomic_groups<- dplyr::filter(perm_528_phylosig_pct_df,  
                                           !(taxonomic_group  %in% filtered_permavonas$taxonomic_group))$taxonomic_group

#make subset of plots,
discarded_PCOA_plots<-
  PCoA_528_plots_axis_pct$ploted_ordination[discarded_taxonomic_groups]

#arrange in a single plot
discarded_taxonomic_groups<-
ggarrange(plotlist = discarded_taxonomic_groups,
          common.legend = TRUE)


#save
ggsave(discarded_taxonomic_groups, 
       filename = "./14_Phylogenetic_distance/results/Results_April2023/Discarded_PCOA_plots_residual085_varratio125_KS2_unimodal2.pdf",
     width = 1900, 
     height = 1900, 
     limitsize  = FALSE,
     units = "mm")




         
```



## PCA of FILTERED PERMANVOAS

```{r}
##### High_low PC instead of PCoA 1 and PCOA2

# subset, removing unecessary columns
abouhf_528x3_phylosig_df_subset_highlow<-
dplyr::select(filtered_permavonas, 
              pvalue_Sp_x_Stress, 
              pvalue_Sp_full_name,
              value_Control_Higher_PC,
              value_Control_Lower_PC,
              value_MeJA_Higher_PC, 
              value_MeJA_Lower_PC,
              value_SA_Higher_PC,
              value_SA_Lower_PC,
              higher_PC_Control_pct,
              higher_PC_MeJA_pct,
              higher_PC_SA_pct,
              lower_PC_Control_pct,
              lower_PC_MeJA_pct,
              lower_PC_SA_pct,
              var_explined_ratio_Control,
              var_explined_ratio_MeJA,
              var_explined_ratio_SA,
              Library_size_Control,
              Library_size_MeJA,
              Library_size_SA)


#tunr library size to log scale
abouhf_528x3_phylosig_df_subset_highlow$Library_size_Control<-
  log(abouhf_528x3_phylosig_df_subset_highlow$Library_size_Control)
abouhf_528x3_phylosig_df_subset_highlow$Library_size_MeJA<-
  log(abouhf_528x3_phylosig_df_subset_highlow$Library_size_MeJA)
abouhf_528x3_phylosig_df_subset_highlow$Library_size_SA<-
  log(abouhf_528x3_phylosig_df_subset_highlow$Library_size_SA)



# only significant  interaction
pca_res_528x3_abouheifs_sig_int_highlow<-
  prcomp(dplyr::filter(abouhf_528x3_phylosig_df_subset_highlow, pvalue_Sp_x_Stress<0.05),
         scale. = TRUE)


pca_plot_528x3_abouheifs_sig_int_highlow<-
autoplot(pca_res_528x3_abouheifs_sig_int_highlow, 
         data =  dplyr::filter(perm_528_phylosig_pct_df, pvalue_Sp_x_Stress<0.05),
         colour = "taxonomic_rank",
         loadings = TRUE, 
         loadings.colour = 'blue', 
          size = 3,
         loadings.label = TRUE, 
         loadings.label.repel=T,
         loadings.label.size = 3)+
         theme(legend.position="bottom")




```





## Correlation plots related to the  528x3 PCoAs, PCoA with higest abouheif vs PCoA with lowest abouhei
```{r}
### plots of higher and lower PCs
corplot_control_abuheif_high_PC<-
ggplot(perm_528_phylosig_pct_df_sigInt,
       aes(x = higher_PC_Control_pct,y = value_Control_Higher_PC, color =taxonomic_rank, alpha = 0.8))+
     geom_point( aes(size = log(Library_size_Control)))+
     #facet_wrap(~Stress)+
  stat_cor(method = "pearson",
           p.accuracy = 0.001, 
           r.accuracy = 0.01,
           aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~")),
           color = "Black")+
  geom_smooth(method = "lm",
              formula = y ~ x,
              colour = "Black" )  +
  xlab("% var. explained in PCoA axis with \n higher abouheif value, Control samples")+
  ylab("Higher of two Abouheif values, Control Samples")

corplot_control_abuheif_low_pc<-
ggplot(perm_528_phylosig_pct_df_sigInt, 
       aes(x = lower_PC_Control_pct,y = value_Control_Lower_PC, color =taxonomic_rank, alpha = 0.8))+
     geom_point( aes(size = log(Library_size_Control)))+
     #facet_wrap(~Stress)+
  stat_cor(method = "pearson",
           p.accuracy = 0.001, 
           r.accuracy = 0.01,
           aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~")),
           color = "Black")+
  geom_smooth(method = "lm",
              formula = y ~ x,
              colour = "Black" )  +
  xlab("% var. explained in PCoA axis with \n lower abouheif value, Control samples")+
  ylab("Lower of two Abouheif values, Control Samples")




corplot_MeJA_abuheif_high_PC<-
ggplot(perm_528_phylosig_pct_df_sigInt,
       aes(x = higher_PC_MeJA_pct,y = value_MeJA_Higher_PC, color =taxonomic_rank, alpha = 0.8))+
     geom_point( aes(size = log(Library_size_MeJA)))+
     #facet_wrap(~Stress)+
  stat_cor(method = "pearson",
           p.accuracy = 0.001, 
           r.accuracy = 0.01,
           aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~")),
           color = "Black")+
  geom_smooth(method = "lm",
              formula = y ~ x,
              colour = "Black" )  +
  xlab("% var. explained in PCoA axis with \n higher abouheif value, MeJA samples")+
  ylab("Higher of two Abouheif values, MeJA Samples")

corplot_MeJA_abuheif_low_pc<-
ggplot(perm_528_phylosig_pct_df_sigInt, 
       aes(x = lower_PC_MeJA_pct,y = value_MeJA_Lower_PC, color =taxonomic_rank, alpha = 0.8))+
     geom_point( aes(size = log(Library_size_MeJA)))+
     #facet_wrap(~Stress)+
  stat_cor(method = "pearson",
           p.accuracy = 0.001, 
           r.accuracy = 0.01,
           aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~")),
           color = "Black")+
  geom_smooth(method = "lm",
              formula = y ~ x,
              colour = "Black" )  +
  xlab("% var. explained in PCoA axis with \n lower abouheif value, MeJA samples")+
  ylab("Lower of two Abouheif values, MeJA Samples")




corplot_SA_abuheif_high_PC<-
ggplot(perm_528_phylosig_pct_df_sigInt,
       aes(x = higher_PC_SA_pct,y = value_SA_Higher_PC, color =taxonomic_rank, alpha = 0.8))+
     geom_point( aes(size = log(Library_size_SA)))+
     #facet_wrap(~Stress)+
  stat_cor(method = "pearson",
           p.accuracy = 0.001, 
           r.accuracy = 0.01,
           aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~")),
           color = "Black")+
  geom_smooth(method = "lm",
              formula = y ~ x,
              colour = "Black" )  +
  xlab("% var. explained in PCoA axis with \n higher abouheif value, SA samples")+
  ylab("Higher of two Abouheif values, SA Samples")

corplot_SA_abuheif_low_pc<-
ggplot(perm_528_phylosig_pct_df_sigInt, 
       aes(x = lower_PC_SA_pct,y = value_SA_Lower_PC, color =taxonomic_rank, alpha = 0.8))+
     geom_point( aes(size = log(Library_size_SA)))+
     #facet_wrap(~Stress)+
  stat_cor(method = "pearson",
           p.accuracy = 0.001, 
           r.accuracy = 0.01,
           aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~")),
           color = "Black")+
  geom_smooth(method = "lm",
              formula = y ~ x,
              colour = "Black" )+
  xlab("% var. explained in PCoA axis with \n lower abouheif value, SA samples")+
  ylab("Lower of two Abouheif values, SA Samples")



#panels of high_low abouheif PCoA splits
panel_high_low_PCoA_Abouheif<-
ggarrange(corplot_control_abuheif_high_PC + facet_wrap(~taxonomic_rank, nrow = 1),
          corplot_MeJA_abuheif_high_PC + facet_wrap(~taxonomic_rank, nrow = 1),
          corplot_SA_abuheif_high_PC + facet_wrap(~taxonomic_rank, nrow = 1),
          corplot_control_abuheif_low_pc + facet_wrap(~taxonomic_rank, nrow = 1),
          corplot_MeJA_abuheif_low_pc + facet_wrap(~taxonomic_rank, nrow = 1),
          corplot_SA_abuheif_low_pc + facet_wrap(~taxonomic_rank, nrow = 1),
          common.legend = TRUE, 
          legend = "bottom",
            ncol = 1, 
            nrow = 6)




### plots of higher and lower PCs library size VS aouheif sinal
corplot_control_LibSize_abuheif_high_PC<-
ggplot(perm_528_phylosig_pct_df_sigInt,
       aes(x = log(Library_size_Control),y = value_Control_Higher_PC, color =taxonomic_rank, alpha = 0.8))+
     geom_point( aes(size = log(Library_size_Control)))+
     #facet_wrap(~Stress)+
  stat_cor(method = "pearson",
           p.accuracy = 0.001, 
           r.accuracy = 0.01,
           aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~")),
           color = "Black")+
  geom_smooth(method = "lm",
              formula = y ~ x,
              colour = "Black" )  +
  xlab("log of Library size, Control samples")+
  ylab("Higher of two Abouheif values, Control Samples")

corplot_control_LibSize_abuheif_low_pc<-
ggplot(perm_528_phylosig_pct_df_sigInt, 
       aes(x = log(Library_size_Control),y = value_Control_Lower_PC, color =taxonomic_rank, alpha = 0.8))+
     geom_point( aes(size = log(Library_size_Control)))+
     #facet_wrap(~Stress)+
  stat_cor(method = "pearson",
           p.accuracy = 0.001, 
           r.accuracy = 0.01,
           aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~")),
           color = "Black")+
  geom_smooth(method = "lm",
              formula = y ~ x,
              colour = "Black" )  +
  xlab("log of Library size, Control samples")+
  ylab("Lower of two Abouheif values, Control Samples")




corplot_MeJA_LibSize_abuheif_high_PC<-
ggplot(perm_528_phylosig_pct_df_sigInt,
       aes(x = log(Library_size_MeJA),y = value_MeJA_Higher_PC, color =taxonomic_rank, alpha = 0.8))+
     geom_point( aes(size = log(Library_size_MeJA)))+
     #facet_wrap(~Stress)+
  stat_cor(method = "pearson",
           p.accuracy = 0.001, 
           r.accuracy = 0.01,
           aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~")),
           color = "Black")+
  geom_smooth(method = "lm",
              formula = y ~ x,
              colour = "Black" )  +
  xlab("log of Library size, MeJA samples")+
  ylab("Higher of two Abouheif values, MeJA Samples")

corplot_MeJA_LibSize_abuheif_low_pc<-
ggplot(perm_528_phylosig_pct_df_sigInt, 
       aes(x = log(Library_size_MeJA),y = value_MeJA_Lower_PC, color =taxonomic_rank, alpha = 0.8))+
     geom_point( aes(size = log(Library_size_MeJA)))+
     #facet_wrap(~Stress)+
  stat_cor(method = "pearson",
           p.accuracy = 0.001, 
           r.accuracy = 0.01,
           aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~")),
           color = "Black")+
  geom_smooth(method = "lm",
              formula = y ~ x,
              colour = "Black" )  +
  xlab("log of Library size, MeJA samples")+
  ylab("Lower of two Abouheif values, MeJA Samples")




corplot_SA_LibSize_abuheif_high_PC<-
ggplot(perm_528_phylosig_pct_df_sigInt,
       aes(x = log(Library_size_SA),y = value_SA_Higher_PC, color =taxonomic_rank, alpha = 0.8))+
     geom_point( aes(size = log(Library_size_SA)))+
     #facet_wrap(~Stress)+
  stat_cor(method = "pearson",
           p.accuracy = 0.001, 
           r.accuracy = 0.01,
           aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~")),
           color = "Black")+
  geom_smooth(method = "lm",
              formula = y ~ x,
              colour = "Black" )  +
  xlab("log of Library size, SA samples")+
  ylab("Higher of two Abouheif values, SA Samples")

corplot_SA_LibSize_abuheif_low_pc<-
ggplot(perm_528_phylosig_pct_df_sigInt, 
       aes(x = log(Library_size_SA),y = value_SA_Lower_PC, color =taxonomic_rank, alpha = 0.8))+
     geom_point( aes(size = log(Library_size_SA)))+
     #facet_wrap(~Stress)+
  stat_cor(method = "pearson",
           p.accuracy = 0.001, 
           r.accuracy = 0.01,
           aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~")),
           color = "Black")+
  geom_smooth(method = "lm",
              formula = y ~ x,
              colour = "Black" )+
  xlab("log of Library size, SA samples")+
  ylab("Lower of two Abouheif values, SA Samples")
```

## make Panel with PCoA + correlations, 528x3 PCoA  with higest abouheif vs PCoA with lowest abouheif
```{r}
#panels of high_low abouheif PCoA splits
panel_high_low_PCoA_LibSize_Abouheif<-
ggarrange(corplot_control_LibSize_abuheif_high_PC + facet_wrap(~taxonomic_rank, nrow = 1),
          corplot_MeJA_LibSize_abuheif_high_PC + facet_wrap(~taxonomic_rank, nrow = 1),
          corplot_SA_LibSize_abuheif_high_PC + facet_wrap(~taxonomic_rank, nrow = 1),
          corplot_control_LibSize_abuheif_low_pc + facet_wrap(~taxonomic_rank, nrow = 1),
          corplot_MeJA_LibSize_abuheif_low_pc + facet_wrap(~taxonomic_rank, nrow = 1),
          corplot_SA_LibSize_abuheif_low_pc + facet_wrap(~taxonomic_rank, nrow = 1),
          common.legend = TRUE, 
          legend = "bottom",
            ncol = 1, 
            nrow = 6)



# join with respective PCA plot of varaibles
panel_PCoA1_PCoA2_Abouheif_sig_int_PCA<-
  ggarrange(pca_plot_528x3_abouheifs_sig_int_highlow,
            panel_high_low_PCoA_LibSize_Abouheif,
            panel_high_low_PCoA_Abouheif,
            ncol = 3, 
            widths = c(1,2,2))

ggsave(panel_PCoA1_PCoA2_Abouheif_sig_int_PCA,
       filename = "./14_Phylogenetic_distance/results/Results_April2023/panel_highlow_Abouheif_sig_int_PCA.svg",
       width = 700,
       height = 525,
       units = "mm")





```































* includes permanovas that were no significant
* check r, p, lib size and PCoA%
```{r}
# subset, removing unecessary columns
perm_filtered_phylosig_df_subset<-#perm_34_phylosig_df_wide
dplyr::select(filtered_permavonas, 
              R2_Sp_full_name, 
              R2_Stress, 
              R2_Sp_x_Stress, 
              R2_Block, 
        #      pvalue_Sp_full_name,
              pvalue_Stress,
              pvalue_Sp_x_Stress, 
              pvalue_Block,
          #    value_Control_Higher_PC,  
          #    value_MeJA_Higher_PC, 
          #    value_SA_Higher_PC, 
              PCoA_1, 
              PCoA_2, 
              Library_size)

#plot a PCA with loadings, large shapes for sinificant sp*stress interaction
library(ggfortify)
pca_res_filtered_Permanovas<-
  prcomp(perm_filtered_phylosig_df_subset, scale. = TRUE)

pca_plot_filtered_Permanovas<-
autoplot(pca_res_filtered_Permanovas, 
         data = filtered_permavonas,
         colour = "taxonomic_rank",
         loadings = TRUE, 
         loadings.colour = 'blue', 
         size = ifelse(test = filtered_permavonas$pvalue_Sp_x_Stress<0.05,
                       yes = 3 ,
                       no = 1),
         loadings.label = TRUE, 
         loadings.label.repel=T,
         loadings.label.size = 3)+
         theme(legend.position="bottom")





```


## Correlation plots related to the of 34 permanovas
```{r}

#p values VS library sizes
pinteract_x_LibSize_plot<-
ggplot(filtered_permavonas, 
       aes(x = log(pvalue_Sp_x_Stress) ,y = log(Library_size), color =R2_Sp_x_Stress, alpha = 0.8))+
     geom_point()+
     #facet_wrap(~Stress)+
  stat_cor(method = "pearson",
           p.accuracy = 0.001, 
           r.accuracy = 0.01,
           aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~")),
           color = "Black")+
  geom_smooth(method = "lm",
              formula = y ~ x,
              colour = "Black" )  +
  xlab("log pvalue Plant_Sp*Treatment")+
  scale_color_viridis()+
  ylab("log Library size (unnormalized counts)")+
  facet_wrap(~taxonomic_rank, nrow =1)


#PCoA1 VS library sizes
PCoA1_x_LibSize_plot<-
ggplot(filtered_permavonas, 
       aes(x = PCoA_1,y = log(Library_size), color =R2_Sp_x_Stress, alpha = 0.8))+
     geom_point()+
     #facet_wrap(~Stress)+
  stat_cor(method = "pearson",
           p.accuracy = 0.001, 
           r.accuracy = 0.01,
           aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~")),
           color = "Black")+
  geom_smooth(method = "lm",
              formula = y ~ x,
              colour = "Black" )  +
  xlab("% var. explained in PCoA_1")+
  ylab("log Library size (unnormalized counts)")+
  scale_color_viridis()+
  facet_wrap(~taxonomic_rank, nrow =1)

#PCoA2 VS library sizes
PCoA2_x_LibSize_plot<-
ggplot(filtered_permavonas, 
       aes(x = PCoA_2,y = log(Library_size), color =R2_Sp_x_Stress, alpha = 0.8))+
     geom_point()+
     #facet_wrap(~Stress)+
  stat_cor(method = "pearson",
           p.accuracy = 0.001, 
           r.accuracy = 0.01,
           aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~")),
           color = "Black")+
  geom_smooth(method = "lm",
              formula = y ~ x,
              colour = "Black" )  +
  xlab("% var. explained in PCoA_2")+
  ylab("log Library size (unnormalized counts)")+
  scale_color_viridis()+
  facet_wrap(~taxonomic_rank, nrow =1)


R2_sp_x_R2_block_plot<-
ggplot(filtered_permavonas, 
       aes(x = R2_Sp_full_name,y = R2_Block, color =R2_Sp_x_Stress, alpha = 0.8))+
     geom_point()+
     #facet_wrap(~Stress)+
  stat_cor(method = "pearson",
           p.accuracy = 0.001, 
           r.accuracy = 0.01,
           aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~")),
           color = "Black")+
  geom_smooth(method = "lm",
              formula = y ~ x,
              colour = "Black" )  +
  xlab("R2 Plant Sp.")+
  ylab("R2 Block")+
  scale_color_viridis()+
  facet_wrap(~taxonomic_rank, nrow =1)



R2_spXstress_x_R2_block_plot<-
ggplot(filtered_permavonas, 
       aes(x = R2_Sp_x_Stress ,y = R2_Block, color =R2_Sp_x_Stress, alpha = 0.8))+
     geom_point()+
     #facet_wrap(~Stress)+
  stat_cor(method = "pearson",
           p.accuracy = 0.001, 
           r.accuracy = 0.01,
           aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~")),
           color = "Black")+
  geom_smooth(method = "lm",
              formula = y ~ x,
              colour = "Black" )+
  xlab("R2 Sp*Stress")+
  ylab("R2 Block")+
  scale_color_viridis()+
  facet_wrap(~taxonomic_rank, nrow =1)







```

## Panel of 34 permanvoa plots: PCA + R2
```{r}
#arrange plot
panel_34_permanovas<-
ggarrange(pca_plot_34_Permanovas,
          ggarrange(pinteract_x_LibSize_plot,
                    PCoA1_x_LibSize_plot,
                    PCoA2_x_LibSize_plot,
                    R2_sp_x_R2_block_plot,
                    R2_spXstress_x_R2_block_plot,
                    ncol = 1, 
                    common.legend = TRUE, 
                    legend = "bottom"),
          ncol = 2, 
          widths = c(1,2))

#save plot
ggsave(filename = "./14_Phylogenetic_distance/results/Results_April2023/panel_34_permanovas.svg",
       plot = panel_34_permanovas,
       width = 400, 
       height = 350, 
       units = "mm")


```





# ------
# --- Fungi --- 
# ------


# load processed R objects, FUN
```{r}


# load output of pipeline, Fun 
load(file ="./14_Phylogenetic_distance/r_objects/fun_alltaxa_phylosig_PCoA_nonsig.RData")

#load hundred of permanovas calculated on the HPC (output of permanova_per_taxonomic_group)
load(file = "./14_Phylogenetic_distance/r_objects/permanovas_spXstress_70.RData") 

# load hundreds of PCoA plots
load(file = "./14_Phylogenetic_distance/r_objects/PCoA_Fun_70_plots_axis_pct.RData") # one per tax group
load(file = "./14_Phylogenetic_distance/r_objects/PCoA_Fun_70_3x_plots_axis_pct.RData") # one per tax group and per stress

# load ps objects with overwrriten library size per taxonomic group (from 11h_library_size_on_phylosig.Rmd)
load(file = "./14_Phylogenetic_distance/r_objects/Fun_ps_taxa_l_overwrittenLibSize_ps_l.RData")

#load output of mantel pipeline 
load(file = "./14_Phylogenetic_distance/r_objects/fun_alltaxa_mantel_pipeline_df.RData") #df
load(file = "./14_Phylogenetic_distance/r_objects/fun_alltaxa_mantel_pipeline_bray_correlation_plot.RData") #plots



```


# - 
# Prepare a df of results from 70 Permanovas + PCoAs


## extract p/r valeus from 70 permanovas, set base df
```{r}


#extract r/p values
perm_70_df<-
permanova_output_to_plot_input2(permanova_output_l = permanovas_spXstress_70$permanova_l)


# put % explained variaance per PC in a df
axis_pct_tax_group<-
  PCoA_plot_l_to_pct_variance_df(PCoA_Fun_70_plots_axis_pct)

axis_pct_tax_group_Control<-
  PCoA_plot_l_to_pct_variance_df(PCoA_Fun_70_3x_plots_axis_pct[1:3])
names(axis_pct_tax_group_Control)[2:3]<-paste0(names(axis_pct_tax_group_Control)[2:3], "_Control_only")

axis_pct_tax_group_MeJA<-
  PCoA_plot_l_to_pct_variance_df(PCoA_Fun_70_3x_plots_axis_pct[4:6])
names(axis_pct_tax_group_MeJA)[2:3]<-paste0(names(axis_pct_tax_group_MeJA)[2:3], "_MeJA_only")

axis_pct_tax_group_SA<-
  PCoA_plot_l_to_pct_variance_df(PCoA_Fun_70_3x_plots_axis_pct[7:9])
names(axis_pct_tax_group_SA)[2:3]<-paste0(names(axis_pct_tax_group_SA)[2:3], "_SA_only")

#put 70 PCoAs from the 3 stresses into 1 df
axis_pct_tax_group_per_stress<-
  cbind(axis_pct_tax_group_Control, axis_pct_tax_group_MeJA[,2:3], axis_pct_tax_group_SA[,2:3])


# widen the df
perm_70_df_wide<-
pivot_wider(data = perm_70_df,
            names_from = factor,
            values_from = c("pvalue", "R2"))

# widen phylosig df
df_phylosig_wide<-
pivot_wider(data = fun_alltaxa_phylosig_PCoA_nonsig$df_phylosig, 
            names_from = c("stress", "higher_lower_PC"),
            values_from = c("trait", "value", "pvalue", "pvalue_adjust"))

# remove unecessary info (phylosig test used, phylosig test adjsuted p)
df_phylosig_wide<-
  df_phylosig_wide[,-c(2, 21:26)]


#join 70 permanovas to widened df-phylosig
perm_70_phylosig_df_wide<-
  left_join(perm_70_df_wide, df_phylosig_wide, by = c("taxonomic_group" = "column_label"))

#add % explained varaince
perm_70_phylosig_pct_df<-
  left_join(perm_70_phylosig_df_wide, axis_pct_tax_group)

perm_70_phylosig_pct_df<-
  left_join(perm_70_phylosig_pct_df, axis_pct_tax_group_per_stress)



```

## Add library sizes to DF of 70 PERMANOVAS + PCoA
these updated library sizes objects were generated in script 11h_library_size_on_phylosig.Rmd
```{r}

#add library size (all 3 stresses together; that is the PERMANOVA input) to into a df 
libSizes<-
  lapply(ps_taxa_l_overwrittenLibSize_fun, function(x) sum(x@sam_data$library_size))

libSizes_df<-
  as.data.frame(do.call("rbind", libSizes))%>%rownames_to_column()

names(libSizes_df)<-c("taxonomic_group", "Library_size")

#join with rest of 70 permanova df
perm_70_phylosig_pct_df<-
  left_join(perm_70_phylosig_pct_df, libSizes_df)

#add library size under each stress condition(split by taxonomic gorupto df 
libSizes_per_stress<-
  lapply(ps_taxa_l_overwrittenLibSize_fun, function(x) {
    ps_per_stress<-phyloseq_sep_variable(x, var = "Stress")
      output<-lapply(ps_per_stress, function(z)
       sum(z@sam_data$library_size)) 
    return(output)
    })

libSizes_per_stress<-
  lapply(libSizes_per_stress, function(x) 
    unlist(x))

# put library sizes epr stress into a df
libSizes_per_stress_df<-
  as.data.frame(do.call("rbind", libSizes_per_stress))%>%rownames_to_column()

names(libSizes_per_stress_df)<-c("taxonomic_group", 
                                  "Library_size_Control", 
                                  "Library_size_MeJA", 
                                  "Library_size_SA")

#join with rest of 70 permanova df
perm_70_phylosig_pct_df<-
  left_join(perm_70_phylosig_pct_df, libSizes_per_stress_df)

# add column to define taxonomic rank
perm_70_phylosig_pct_df$taxonomic_rank<-
  factor(
        c(
           gsub(x = perm_70_phylosig_pct_df$taxonomic_group, pattern = "k__.*$",replacement = "Kingdom")%>%
           gsub(pattern = "p__.*$",replacement = "Phylum")%>%
           gsub(pattern = "o__.*$",replacement = "Order")%>%
           gsub(pattern = "c__.*$",replacement = "Class")%>%
           gsub(pattern = "f__.*$",replacement = "Family")%>%
           gsub(pattern = "g__.*$",replacement = "Genus")),
         levels = c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus"))

# define unexplained permanova variance
perm_70_phylosig_pct_df$residual_permanova<-
  1 - (perm_70_phylosig_pct_df$R2_Block +
          perm_70_phylosig_pct_df$R2_Sp_full_name +
          perm_70_phylosig_pct_df$R2_Sp_x_Stress + 
          perm_70_phylosig_pct_df$R2_Stress)

# define PCoA_1/ PCoA2 varaince aexplained ratio
perm_70_phylosig_pct_df$var_explined_ratio<-
  (perm_70_phylosig_pct_df$PCoA_1/perm_70_phylosig_pct_df$PCoA_2)



```



## define distance to centroids for the 3 experimental factors (Block, plant_sp, Stress)
```{r}

#define function to wrangle beta dispersion data
get_beta_disper_comparison<-function(Taxonomic_group_beta_disper){
  
dist_to_centroid_df<-
  as.data.frame(Taxonomic_group_beta_disper$beta_disper_block$distances)

names(dist_to_centroid_df)<-"distance_to_centroid"
dist_to_centroid_df<-rownames_to_column(dist_to_centroid_df, var = "sample_name")
dist_to_centroid_df$centroid_group<-"Block"
dist_to_centroid_df$group_ID<-Taxonomic_group_beta_disper$beta_disper_block$group


dist_to_centroid_df2<-
  as.data.frame(Taxonomic_group_beta_disper$beta_disper_sp$distances)

names(dist_to_centroid_df2)<-"distance_to_centroid"
dist_to_centroid_df2<-rownames_to_column(dist_to_centroid_df2, var = "sample_name")
dist_to_centroid_df2$centroid_group<-"plant_Sp"
dist_to_centroid_df2$group_ID<-Taxonomic_group_beta_disper$beta_disper_sp$group


dist_to_centroid_df3<-
  as.data.frame(Taxonomic_group_beta_disper$beta_disper_stress$distances)

names(dist_to_centroid_df3)<-"distance_to_centroid"
dist_to_centroid_df3<-rownames_to_column(dist_to_centroid_df3, var = "sample_name")
dist_to_centroid_df3$centroid_group<-"Stress"
dist_to_centroid_df3$group_ID<-Taxonomic_group_beta_disper$beta_disper_stress$group


dist_to_centroid_df<-
  rbind(dist_to_centroid_df,dist_to_centroid_df2,dist_to_centroid_df3)

#print(ggplot(dist_to_centroid_df, aes(x = centroid_group, y = distance_to_centroid))+
 # geom_boxplot())

return(dist_to_centroid_df)
}

#execute custom function above, generating a list of dfs
beta_disper_comparions_df_l<-lapply(permanovas_spXstress_70$beta_disper_l, function(z)
  get_beta_disper_comparison(Taxonomic_group_beta_disper = z))

#join list
beta_disper_comparions_df<-
  do.call("rbind", beta_disper_comparions_df_l)%>%rownames_to_column(var = "taxonomic_group")

#adjsut names
beta_disper_comparions_df$taxonomic_group<-
  gsub(pattern = "\\..*",replacement = "", x = beta_disper_comparions_df$taxonomic_group)

# add column to define taxonomic rank
beta_disper_comparions_df$taxonomic_rank<-
  factor(
        c(
           gsub(x = beta_disper_comparions_df$taxonomic_group, pattern = "k__.*$",replacement = "Kingdom")%>%
           gsub(pattern = "p__.*$",replacement = "Phylum")%>%
           gsub(pattern = "o__.*$",replacement = "Order")%>%
           gsub(pattern = "c__.*$",replacement = "Class")%>%
           gsub(pattern = "f__.*$",replacement = "Family")%>%
           gsub(pattern = "g__.*$",replacement = "Genus")),
         levels = c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus"))


#check summary
summarized_df_dist_to_centroid<-
group_by(.data = beta_disper_comparions_df, centroid_group, taxonomic_group, taxonomic_rank)%>%
  dplyr::summarize(mean_dist = mean(distance_to_centroid),
            median_dist = median(distance_to_centroid),
            sd_dist = sd(distance_to_centroid))


```



## define if beta distribution residuals are bimodal 
to be kept in the analysis, a taoxnomic group must have a unimodal distribution in distance to centroids for stress, plant species AND block

```{r}

#split beta diversity comparisons into a nested list for lapply
beta_disper_comparions_df_l<-
  split(x = beta_disper_comparions_df, f = beta_disper_comparions_df$taxonomic_group)

#run multimode test (takes several minutes to compute)
multimode_test_l_output<-
lapply(beta_disper_comparions_df_l, function(x){
  
  set.seed(12345)
  Block_modtest<-modetest(dplyr::filter(x, centroid_group =="Block")$distance_to_centroid)
    
  plantSp_modtest<-modetest(dplyr::filter(x, centroid_group =="plant_Sp")$distance_to_centroid)
  
  Stress_modtest<-modetest(dplyr::filter(x, centroid_group =="Stress")$distance_to_centroid)
  
  output<-list("Block_modtest" = Block_modtest,
               "plantSp_modtest" = plantSp_modtest,
               "Stress_modtest" = Stress_modtest)
  
  return(output)
  
})



#save externally
save(multimode_test_l_output, file = "./14_Phylogenetic_distance/r_objects/fun_multimode_test_l_output.RData")
load(file = "./14_Phylogenetic_distance/r_objects/fun_multimode_test_l_output.RData")


```

##  put multimode test results into a df, join with 70 permanova df
```{r}
#extract p values from the multimode test, put in lsit
multimode_p_l<-
lapply(multimode_test_l_output, function(x){
  block_p_multimode<-x$Block_modtest$p.value
  plantSp_p_multimode<-x$plantSp_modtest$p.value
  Stress_p_multimode<-x$Stress_modtest$p.value
  
  return(c("block_p_multimode" = block_p_multimode,
           "plantSp_p_multimode" = plantSp_p_multimode,
           "Stress_p_multimode" = Stress_p_multimode))
  
})

#turn list into df
multimode_p_df<-
  as.data.frame(do.call("rbind",multimode_p_l ))

#define if residuals are unimodal in the experimental factors ( number of experiemntal factros where residues are unimodal)
multimode_p_df$unimodal_residues<-rowCounts(multimode_p_df[,1:3]>0.05)
multimode_p_df$unimodal_residues001<-rowCounts(multimode_p_df[,1:3]>0.01)



#put the output of the beta dispersion unimodalisty testing in the permanova df
multimode_p_df<-rownames_to_column(multimode_p_df, var = "taxonomic_group")
perm_70_phylosig_pct_df<-left_join(perm_70_phylosig_pct_df,multimode_p_df)








```



## define if residual distributions are similar to control
```{r}


# run Kolmogorov-Smirnov to compare distribution of residuals with k__Fungi
ks_l_output<-
lapply(beta_disper_comparions_df_l, function(x){
  # run Kolmogorov-Smirnov to define if two samples (k__Fungi and focal taxonomic group) derive from teh same distribution on a varaibles (scaled distances to centroids)
  
  ks_Block<-
    ks.test(x = scale(dplyr::filter(beta_disper_comparions_df_l$k__Fungi, 
                         centroid_group == "Block" )$distance_to_centroid),
           y = scale(dplyr::filter(x, 
                             centroid_group == "Block" )$distance_to_centroid))
    
  ks_Stress<-
    ks.test(x = scale(dplyr::filter(beta_disper_comparions_df_l$k__Fungi, 
                         centroid_group == "Stress" )$distance_to_centroid),
           y = scale(dplyr::filter(x, 
                             centroid_group == "Stress" )$distance_to_centroid))
  
   
   ks_plantSp<-
    ks.test(x = scale(dplyr::filter(beta_disper_comparions_df_l$k__Fungi, 
                         centroid_group == "plant_Sp" )$distance_to_centroid),
           y = scale(dplyr::filter(x, 
                             centroid_group == "plant_Sp" )$distance_to_centroid))
  
  
  output<-list("ks_Block" = ks_Block,
               "ks_Stress" = ks_Stress,
               "ks_plantSp" = ks_plantSp)
  
  return(output)
  
})

#extract p values from the test, put in lsit
ks_l_pval<-
  lapply(ks_l_output, function(x){
    
  block_p_ks<-x$ks_Block$p.value
  plantSp_p_ks<-x$ks_plantSp$p.value
  Stress_p_ks<-x$ks_Stress$p.value
  
  return(c("block_p_ks" = block_p_ks,
           "plantSp_p_ks" = plantSp_p_ks,
           "Stress_p_ks" = Stress_p_ks))
  
})



   
   
   

#turn list into df
ks_pval_df<-
  as.data.frame(do.call("rbind",ks_l_pval ))

#define if residuals are unimodal in all experimental factors (as in number of experiemntal factros)
ks_pval_df$residue_dist_simialr_to_K__Fungi<-rowCounts(ks_pval_df[,1:3]>0.05)
ks_pval_df$residue_dist_simialr_to_K__Fungi001<-rowCounts(ks_pval_df[,1:3]>0.01)

#put the output of the beta dispersion unimodalisty testing in the permanova df
ks_pval_df<-rownames_to_column(ks_pval_df, var = "taxonomic_group")
perm_70_phylosig_pct_df<-left_join(perm_70_phylosig_pct_df,ks_pval_df)



```



## reshape mantel pipeline output to fit permanova df
```{r}

# remove Jaccard distances, as they provide the same conclusions as bray-curtis
fun_alltaxa_mantel_pipeline_df_reshaped<-
  dplyr::filter(fun_alltaxa_mantel_pipeline_df, distance =="Bray")

#remove column that identifies if  the test ws significant
fun_alltaxa_mantel_pipeline_df_reshaped<-
  dplyr::select(fun_alltaxa_mantel_pipeline_df_reshaped,  -significant)

# rename some variable to avoid confusion wiht abouheif tests
names(fun_alltaxa_mantel_pipeline_df_reshaped)[c(3,5)]<-c("distance_Mantel", "p_value_Mantel")

#widen df to have one row per taxonomic group
fun_alltaxa_mantel_pipeline_df_reshaped<-
  pivot_wider(fun_alltaxa_mantel_pipeline_df_reshaped, 
              names_from = "Treatment", 
              values_from = c("r2_Mantel", "p_value_Mantel"))



```



## add mantel pipeline outut to permanova df
```{r}

perm_70_phylosig_pct_df<-
  left_join(perm_70_phylosig_pct_df,fun_alltaxa_mantel_pipeline_df_reshaped)

```

## define PCoA variance explained by the PC with highest abouheif
```{r}
#re-arrange which is the hgiehr PC
perm_70_phylosig_pct_df$higher_PC_Control<-
  gsub(x = perm_70_phylosig_pct_df$trait_Control_Higher_PC, pattern = "_[a-z]_.*", replacement = "")

perm_70_phylosig_pct_df$higher_PC_MeJA<-
  gsub(x = perm_70_phylosig_pct_df$trait_MeJA_Higher_PC, pattern = "_[a-z]_.*", replacement = "")

perm_70_phylosig_pct_df$higher_PC_SA<-
  gsub(x = perm_70_phylosig_pct_df$trait_SA_Higher_PC, pattern = "_[a-z]_.*", replacement = "")

# define PCoA_1/ PCoA2 varaince aexplained ratio
perm_70_phylosig_pct_df$var_explined_ratio_Control<-
  (perm_70_phylosig_pct_df$PCoA_1_Control_only/perm_70_phylosig_pct_df$PCoA_2_Control_only)

perm_70_phylosig_pct_df$var_explined_ratio_MeJA<-
  (perm_70_phylosig_pct_df$PCoA_1_MeJA_only/perm_70_phylosig_pct_df$PCoA_2_MeJA_only)

perm_70_phylosig_pct_df$var_explined_ratio_SA<-
  (perm_70_phylosig_pct_df$PCoA_1_SA_only/perm_70_phylosig_pct_df$PCoA_2_SA_only)

# assing the abouheif value for each PCoA in each treatmet condition, (hard-coded)
perm_70_phylosig_pct_df$PCoA_1_Control_abouheif_value<-
  ifelse(test = perm_70_phylosig_pct_df$higher_PC_Control == "PCoA_1",
         yes = perm_70_phylosig_pct_df$value_Control_Higher_PC,
         no = perm_70_phylosig_pct_df$value_Control_Lower_PC)

perm_70_phylosig_pct_df$PCoA_2_Control_abouheif_value<-
  ifelse(test = perm_70_phylosig_pct_df$higher_PC_Control == "PCoA_2",
         yes = perm_70_phylosig_pct_df$value_Control_Higher_PC,
         no = perm_70_phylosig_pct_df$value_Control_Lower_PC)

perm_70_phylosig_pct_df$PCoA_1_MeJA_abouheif_value<-
  ifelse(test = perm_70_phylosig_pct_df$higher_PC_MeJA == "PCoA_1",
         yes = perm_70_phylosig_pct_df$value_MeJA_Higher_PC,
         no = perm_70_phylosig_pct_df$value_MeJA_Lower_PC)

perm_70_phylosig_pct_df$PCoA_2_MeJA_abouheif_value<-
  ifelse(test = perm_70_phylosig_pct_df$higher_PC_MeJA == "PCoA_2",
         yes = perm_70_phylosig_pct_df$value_MeJA_Higher_PC,
         no = perm_70_phylosig_pct_df$value_MeJA_Lower_PC)

perm_70_phylosig_pct_df$PCoA_1_SA_abouheif_value<-
  ifelse(test = perm_70_phylosig_pct_df$higher_PC_SA == "PCoA_1",
         yes = perm_70_phylosig_pct_df$value_SA_Higher_PC,
         no = perm_70_phylosig_pct_df$value_SA_Lower_PC)

perm_70_phylosig_pct_df$PCoA_2_SA_abouheif_value<-
  ifelse(test = perm_70_phylosig_pct_df$higher_PC_SA == "PCoA_2",
         yes = perm_70_phylosig_pct_df$value_SA_Higher_PC,
         no = perm_70_phylosig_pct_df$value_SA_Lower_PC)

# assing the % PCoA explained in the PCoA wth higer abouheif
perm_70_phylosig_pct_df$higher_PC_Control_pct<-
  ifelse(test = perm_70_phylosig_pct_df$higher_PC_Control == "PCoA_1",
         yes = perm_70_phylosig_pct_df$PCoA_1_Control_only,
         no = perm_70_phylosig_pct_df$PCoA_2_Control_only)

perm_70_phylosig_pct_df$lower_PC_Control_pct<-
  ifelse(test = perm_70_phylosig_pct_df$higher_PC_Control == "PCoA_2",
         yes = perm_70_phylosig_pct_df$PCoA_1_Control_only,
         no = perm_70_phylosig_pct_df$PCoA_2_Control_only)

perm_70_phylosig_pct_df$higher_PC_MeJA_pct<-
  ifelse(test = perm_70_phylosig_pct_df$higher_PC_MeJA == "PCoA_1",
         yes = perm_70_phylosig_pct_df$PCoA_1_MeJA_only,
         no = perm_70_phylosig_pct_df$PCoA_2_MeJA_only)

perm_70_phylosig_pct_df$lower_PC_MeJA_pct<-
  ifelse(test = perm_70_phylosig_pct_df$higher_PC_MeJA == "PCoA_2",
         yes = perm_70_phylosig_pct_df$PCoA_1_MeJA_only,
         no = perm_70_phylosig_pct_df$PCoA_2_MeJA_only)
          
perm_70_phylosig_pct_df$higher_PC_SA_pct<-
  ifelse(test = perm_70_phylosig_pct_df$higher_PC_SA == "PCoA_1",
         yes = perm_70_phylosig_pct_df$PCoA_1_SA_only,
         no = perm_70_phylosig_pct_df$PCoA_2_SA_only)

perm_70_phylosig_pct_df$lower_PC_SA_pct<-
  ifelse(test = perm_70_phylosig_pct_df$higher_PC_SA == "PCoA_2",
         yes = perm_70_phylosig_pct_df$PCoA_1_SA_only,
         no = perm_70_phylosig_pct_df$PCoA_2_SA_only)

```



### save processed 70 PERMANOVAS df

```{r}
# save df of 70 permanovas, % variance, Abouheif values, beta dispersion results
save(perm_70_phylosig_pct_df, file = "./14_Phylogenetic_distance/r_objects/perm_70_phylosig_pct_df.RData")
load(file = "./14_Phylogenetic_distance/r_objects/perm_70_phylosig_pct_df.RData")
view(perm_70_phylosig_pct_df)
```


# check mantel pairwise correlations of fungi
```{r}



#export example mantel plot pariiwise comparisons among samples
ggsave(ggarrange(fun_alltaxa_mantel_pipeline_bray_correlation_plot$k__Fungi,
                 fun_alltaxa_mantel_pipeline_bray_correlation_plot$c__Sordariomycetes,
                 fun_alltaxa_mantel_pipeline_bray_correlation_plot$o__Hypocreales,
                 ncol = 3),
       width = 450,
       height = 350,
       units = "mm",
       filename = "./14_Phylogenetic_distance/results/Results_April2023/mantel_pairwise_Fungi.pdf",)

```


```

