---
title: "14c_phylosig_plots_heattrees"
author: "Pedro_Beschoren"
date: "20/04/2023"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: consolehtml_document
---


# 14.1 - set up working environemnt


```{r}
# this will load all key libraries for the whole project
source("./renv/0_load_libraries.R")
library(plyr)
library("umap")
library(RColorBrewer)
library(cowplot) 
library(multiblock)
library(data.table)
library(ggExtra)

# IMPORTANT: load the custom functions necessary to run this script
source("./14_Phylogenetic_distance/code/custom_functions_PB.R")


library(picante)
library(ape)
library(adephylo)
library(ade4)
library(phylobase)
library(geiger)
library(phytools)
library(vegan)
library(phyloseq)
library(dplyr)


# increases memory limit used by R
memory.limit(size = 350000)
```

# 14.2 - load ps, remove species outside tree 

```{R}

#load the output of the phylogenetic signal pipeline - all taxa in 1 list
load(file = "./14_Phylogenetic_distance/r_objects/bac_alltaxa_phylosig_output_PCoA_nonsig.RData") 

#load the output of the phylogenetic signal pipeline - one list per taxonomic rank
load(file = "./14_Phylogenetic_distance/r_objects/per_level_phylosig_PPCA.RData") 

# a df with various metrics of the 528 permanovas
load(file = "./14_Phylogenetic_distance/r_objects/perm_528_phylosig_pct_df.RData")

# a df with of permanovas filtered absed on various metrics
load(file = "./14_Phylogenetic_distance/r_objects/filtered_permavonas.RData")

#load output of mantel pipeline
load(file = "./14_Phylogenetic_distance/r_objects/bac_alltaxa_mantel_pipeline_df.RData")

# load hundreds of PCoA plots
load(file = "./14_Phylogenetic_distance/r_objects/PCoA_528_plots_axis_pct.RData") # one per tax group
load(file = "./14_Phylogenetic_distance/r_objects/PCoA_528_3x_plots_axis_pct.RData") # one per tax group and per stress

# load PS data
load("./Data/phyloseq_objects/CSS_BacFun_rhizoplane_HighFilt_ps_l.RData")

# remove all species outside the MY trees
CSS_BacFun_rhizoplane_HighFilt_ps_l <- 
  lapply(CSS_BacFun_rhizoplane_HighFilt_ps_l, function(x) 
    subset_samples(x, MY_tree_presence == "inside_MY_tree"))



```

#### ####################
## ## Trait_on_tree plots ##
#### ####################

# Trait_on_tree plot: heatmap of PCoA coodinaties per plant species

This is the basis of the main figures for the paper.
All plots only contains the taxonomic groups that passed permanova/PCOA filtering

## plot brassicaceae phylogeny tree, export as pdf


```{r}
#load tree
Bra_tree_for_plot<-load_prepare_MY_tree()

#save tree plot
 pdf(file = "./14_Phylogenetic_distance/results/Results_April2023/bra_tree_hendriks.pdf",
     width = 4, 
     height = 6) 
 
 #plot
 plot(Bra_tree_for_plot)
  dev.off()

  
  
  

```


## Trait_on_tree: prepare df for heatmap

```{r}

# create df with PCoA coordinates for a heatmap plot
re_created_traitontree<-
lapply(bac_alltaxa_phylosig_output_PCoA_nonsig$ordination_coordinates_l, function(x){
  
  
  #extract emtadata
  metadata<-as(x@sam_data, "data.frame")
  
  #add extended lineage  
  metadata$Sp_Lineage_Walden_extended<-  
        if_else(condition = metadata$Sp_abb_name %in% c("Ia", "Lm", "Co", "Bi"), 
              true ="lineage_II_extended",
              false =  metadata$Sp_Lineage_Walden)
        
        
        
  
  #split by stress
  metadata_split<-split(metadata, f = metadata$Stress)
  

  #calcualte centroid
centroid_l<-
  lapply(metadata_split, function(z){
    
    #calculate centroid
    centroid <- aggregate(cbind(PCoA_1, PCoA_2) ~ Sp_abb_name, data = z, FUN = mean)

    #if Sar is positive on the PC1, rotate the axis by multiplying it by -1
    if(filter(centroid,Sp_abb_name == "Sar")$PCoA_1>0) {
       centroid$PCoA_1<-centroid$PCoA_1*-1 
    }
    
    #if Sar is positive on the PC2, rotate the axis by multiplying it by -1
    if(filter(centroid,Sp_abb_name == "Sar")$PCoA_2>0) {
       centroid$PCoA_2<-centroid$PCoA_2*-1 
    }


    #scale and centr PCoA1 and 2
    centroid$PCoA_1_centroid_scaled<-scale(centroid$PCoA_1,center = TRUE, scale = TRUE)
    centroid$PCoA_2_centroid_scaled<-scale(centroid$PCoA_2,center = TRUE, scale = TRUE)
    
    return(centroid)
  })

# extract per-species metadata
unique_meta_l<-lapply(metadata_split, function(y)
  dplyr::select(.data = y,
                        Sp_abb_name, 
                        Stress)%>%unique)

#join centroids with per-species emtadata
centroid_with_metadata_l<-
  mapply(function (a,b)
    left_join(a,b),
    a = centroid_l,
    b = unique_meta_l,
    SIMPLIFY = FALSE)

centroid_with_metadata<-
  do.call("rbind", centroid_with_metadata_l)

rownames(centroid_with_metadata)<-NULL

return(centroid_with_metadata)

})

# prepare df for heatmap plot 
re_created_traitontree_df<-
  do.call("rbind", re_created_traitontree)%>%rownames_to_column(var = "taxonomic_group")

re_created_traitontree_df$taxonomic_group<-
  gsub(x = re_created_traitontree_df$taxonomic_group, pattern = "\\.\\d+", replacement = "")


#make table longer so all scaled PCs are in one column
re_created_traitontree_df<-
  pivot_longer(data = re_created_traitontree_df, cols = 5:6, names_to = "scaled_PCoA_axis", values_to = "scaled_PCoA")

  #re-join pcoa axis with taxonomic group
trait_naming_output<-
  mapply(function (z,y)
         gsub(x = z,
              pattern = "centroid_scaled",
              replacement = y),
         z = re_created_traitontree_df$scaled_PCoA_axis,
         y = re_created_traitontree_df$taxonomic_group,
         SIMPLIFY = TRUE )
re_created_traitontree_df$trait<-as.vector(trait_naming_output)

#join hi/low pc information
heatmap_input<-
  left_join(re_created_traitontree_df,
            dplyr::select(bac_alltaxa_phylosig_output_PCoA_nonsig$df_phylosig, higher_lower_PC, trait,stress))


#prepare order of facotrs
taxonomic_group_ordering<-
  factor(bac_alltaxa_phylosig_output_PCoA_nonsig$df_phylosig$column_label%>%unique,
         levels = bac_alltaxa_phylosig_output_PCoA_nonsig$df_phylosig$column_label%>%unique)

#adjust order of factors
heatmap_input$taxonomic_group<-
  factor(heatmap_input$taxonomic_group, levels = levels(taxonomic_group_ordering))

perm_528_phylosig_pct_df$taxonomic_group<-
  factor(perm_528_phylosig_pct_df$taxonomic_group, levels = levels(taxonomic_group_ordering))

# change order of the species to reflect newst phylogeny
heatmap_input$Sp_abb_name<-factor(heatmap_input$Sp_abb_name, levels = Bra_tree_for_plot$tip.label)%>%
  fct_rev() # rever order to keet At as the first
  



```

## Trait_on_tree: filter df (significant interaction, abouheif) and plot heatmap
this plot only contains the taxonomic groups that passed permanova/PCOA filtering
these are further filtered to contain only taxa with significat sp*stress interactions and significat abouheif signals

plots heatmaps + barplots with abouheif

*ATTENTION* this is the chunk that defines teh filtering of taxonomic groups that will be showin in figure 1!
```{r}



#####  ------------------- #####
#####  different taxonomic groups per stress (panel of 3 plots) #####
#####  ------------------- #####



# only keep taxonomic groups with...
  # significant interaction 
  # significant abouheif
  # at least 15% variance explained by permanova (residual smaller than 0.8, simialr to K__Bacteria)
  # PCoA_1 explains at least 1.25 time the variance of PCOA_2 (sufficient variance explained in 2 dimensions, simialr to K__Bacteria)
  # at least 2 of the 3 factors (block, plant species, stress) presents unimodal distanes to centroids (p<0.001)
  # at least 2 of the 3 factors (block, plant species, stress) presents distanes to centroids distribution similar to k__Bacteria (p<0.001)


taxonomic_groups_sig_interaction_Control<-
  dplyr::filter(perm_528_phylosig_pct_df, 
                pvalue_Sp_x_Stress<0.05,
                residual_permanova<0.85,
                var_explined_ratio>1.25,
                unimodal_residues001 >=2,
                residue_dist_simialr_to_K__bacteria001 >=2,
                pvalue_Control_Higher_PC<0.05,
                p_value_Mantel_Control<0.05) 

taxonomic_groups_sig_interaction_MeJA<-
  dplyr::filter(perm_528_phylosig_pct_df, 
                pvalue_Sp_x_Stress<0.05, 
                residual_permanova<0.85,
                var_explined_ratio>1.25,
                unimodal_residues001 >=2,
                residue_dist_simialr_to_K__bacteria001 >=2,
                pvalue_MeJA_Higher_PC<0.05,
                p_value_Mantel_MeJA<0.05) 

taxonomic_groups_sig_interaction_SA<-
  dplyr::filter(perm_528_phylosig_pct_df, 
                pvalue_Sp_x_Stress<0.05, 
                residual_permanova<0.85,
                var_explined_ratio>1.25,
                unimodal_residues001 >=2,
                residue_dist_simialr_to_K__bacteria001 >=2,
                pvalue_SA_Higher_PC<0.05,
                p_value_Mantel_SA<0.05) 

#filter df to the taxonomies and samples above
heatmap_input_filtered_interaction_Control<-
  dplyr::filter(heatmap_input, 
                heatmap_input$taxonomic_group %in% taxonomic_groups_sig_interaction_Control$taxonomic_group,
                higher_lower_PC == "Higher_PC", 
                Stress =="Control") 

heatmap_input_filtered_interaction_MeJA<-
  dplyr::filter(heatmap_input, 
                heatmap_input$taxonomic_group %in% taxonomic_groups_sig_interaction_MeJA$taxonomic_group,
                higher_lower_PC == "Higher_PC", 
                Stress =="MeJA") 
  

heatmap_input_filtered_interaction_SA<-
  dplyr::filter(heatmap_input, 
                heatmap_input$taxonomic_group %in% taxonomic_groups_sig_interaction_SA$taxonomic_group,
                higher_lower_PC == "Higher_PC", 
                Stress =="SA") 
  
  
#heatmap plots
heatmap_PCoA_coord_005_Control<-
ggplot(data = heatmap_input_filtered_interaction_Control, aes(y = Sp_abb_name, x = taxonomic_group, fill = scaled_PCoA))+
  geom_tile()+
  scale_y_discrete(limits=rev)+
    scale_fill_viridis(option = "B")+
    scale_color_viridis(option = "B")+
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1))+
  facet_wrap(~Stress)+
  ylab(NULL)

heatmap_PCoA_coord_005_MeJA<-
ggplot(data = heatmap_input_filtered_interaction_MeJA, aes(y = Sp_abb_name, x = taxonomic_group, fill = scaled_PCoA))+
  geom_tile()+
  scale_y_discrete(limits=rev)+
    scale_fill_viridis(option = "B")+
    scale_color_viridis(option = "B")+
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1))+
  facet_wrap(~Stress)+
  ylab(NULL)

heatmap_PCoA_coord_005_SA<-
ggplot(data = heatmap_input_filtered_interaction_SA, aes(y = Sp_abb_name, x = taxonomic_group, fill = scaled_PCoA))+
  geom_tile()+
  scale_y_discrete(limits=rev)+
    scale_fill_viridis(option = "B")+
    scale_color_viridis(option = "B")+
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1))+
  facet_wrap(~Stress)+
  ylab(NULL)

#arrange plots proportionally
heamap_diff_taxonomic_group_per_stress<-
ggarrange(heatmap_PCoA_coord_005_Control,
          heatmap_PCoA_coord_005_MeJA,
          heatmap_PCoA_coord_005_SA,
          nrow = 1, 
          common.legend = TRUE, legend = "bottom",
          align = "h",  
          widths = c(dim(heatmap_input_filtered_interaction_Control)[1],
                     dim(heatmap_input_filtered_interaction_MeJA)[1],
                     dim(heatmap_input_filtered_interaction_SA)[1]))



#prepare barpltos to show on top of heatmaps

barplot_Control_abouheif<-
ggplot(data = taxonomic_groups_sig_interaction_Control, aes(x = taxonomic_group, y = value_Control_Higher_PC))+
  geom_bar(stat = "identity", 
           fill = colour_v[5],
           alpha = if_else(condition = taxonomic_groups_sig_interaction_Control$pvalue_Control_Higher_PC<0.05,
                     true = 1 ,
                     false = 0.3))+
    geom_point(aes(x = taxonomic_group, 
                 y = r2_Mantel_Control), 
             alpha = if_else(condition = taxonomic_groups_sig_interaction_Control$p_value_Mantel_Control<0.05,
                     true = 1 ,
                     false = 0.3),
             size = 2)+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        legend.position = "none")+
  ylim(c(0,0.61))+
#  ylim(c(0,1))+
  ylab("Control - Bars: Abouheif's C Mean / Dots: Mantel R2")

barplot_MeJA_abouheif<-
ggplot(data = taxonomic_groups_sig_interaction_MeJA, aes(x = taxonomic_group, y = value_MeJA_Higher_PC))+
  geom_bar(stat = "identity",
           fill = colour_v[3],
           alpha = if_else(condition = taxonomic_groups_sig_interaction_MeJA$pvalue_MeJA_Higher_PC<0.05,
                   true = 1 ,
                   false = 0.3))+
    geom_point(aes(x = taxonomic_group, 
                 y = r2_Mantel_MeJA), 
             alpha = if_else(condition = taxonomic_groups_sig_interaction_MeJA$p_value_Mantel_MeJA<0.05,
                     true = 1 ,
                     false = 0.3 ),
             size = 2)+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        legend.position = "none")+
  ylim(c(0,0.61))+
  #  ylim(c(0,1))+
    ylab("MeJA - Bars: Abouheif's C Mean / Dots: Mantel R2")


barplot_SA_abouheif<-
ggplot(data = taxonomic_groups_sig_interaction_SA, aes(x = taxonomic_group, y = value_SA_Higher_PC))+
  geom_bar(stat = "identity", 
           fill = colour_v[4],
           alpha = if_else(condition = taxonomic_groups_sig_interaction_SA$pvalue_SA_Higher_PC<0.05,
                  true = 1 ,
                  false = 0.3))+
  geom_point(aes(x = taxonomic_group, 
                 y = r2_Mantel_SA), 
             alpha = if_else(condition = taxonomic_groups_sig_interaction_SA$p_value_Mantel_SA<0.05,
                     true = 1 ,
                     false = 0.3 ),
             size = 2)+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
         legend.position = "none")+
  ylim(c(0,0.61))+
   # ylim(c(0,1))+
    ylab("SA - Bars: Abouheif's C Mean / Dots: Mantel R2")




#arrange plots proportionally
barplot_abouheif_taxonomic_group_per_stress<-
ggarrange(barplot_Control_abouheif,
          barplot_MeJA_abouheif,
          barplot_SA_abouheif,
          nrow = 1, 
          common.legend = TRUE, 
          align = "h",  
          widths = c(dim(heatmap_input_filtered_interaction_Control)[1],
                     dim(heatmap_input_filtered_interaction_MeJA)[1],
                     dim(heatmap_input_filtered_interaction_SA)[1]))


#put barplots on top of heatmaps, proportionally
barplots_Abouheif_heatmaps_PCoA<-
ggarrange(barplot_abouheif_taxonomic_group_per_stress,
          heamap_diff_taxonomic_group_per_stress,
          nrow = 2,
          heights = c(1,3),
          align = "hv")




#save heatmap, significant interactions and abouheif at each stress
ggsave(barplots_Abouheif_heatmaps_PCoA, 
       filename = "./14_Phylogenetic_distance/results/Results_April2023/trait_on_tree_heatmap_barplot_005Interaction_005Abouheif__PCratio125_residual085_unimodal2_SK2_mantelR2dot_overlap.pdf",
     width = 525, 
     height = 375, 
     units = "mm")



```



## Trait_on_tree:  plot heatmap of taxonomies that failed the interaction and abouheif significane selection

here we still only look at the taxonomies that passed the ermanova filters
coordinates of taxonomic groups with NO significatn sp*stress interactions AND no significant abouheif

heatmaps + barplots with abouheif

```{r}



#####  ------------------- #####
#####  different taxonomic groups per stress (panel of 3 plots) #####
#####  ------------------- #####



# only keep taxonomic groups with...
# not present in taxonomic_groups_sig_interaction_Control, taxonomic_groups_sig_interaction_MeJA, taxonomic_groups_sig_interaction_SA
  # at least 15% variance explained by permanova (residual smaller than 0.8, simialr to K__Bacteria)
  # PCoA_1 explains at least 1.25 time the variance of PCOA_2 (sufficient variance explained in 2 dimensions, simialr to K__Bacteria)
  # at least 2 of the 3 factors (block, plant species, stress) presents unimodal distanes to centroids (p<0.001)
  # at least 2 of the 3 factors (block, plant species, stress) presents distanes to centroids distribution similar to k__Bacteria (p<0.001)

taxonomic_groups_nonsig_Control<-
  dplyr::filter(perm_528_phylosig_pct_df, 
                !taxonomic_group %in% taxonomic_groups_sig_interaction_Control$taxonomic_group,
                residual_permanova<0.85,
                var_explined_ratio>1.25,
                unimodal_residues001 >=2,
                residue_dist_simialr_to_K__bacteria001 >=2) 

taxonomic_groups_nonsig_MeJA<-
  dplyr::filter(perm_528_phylosig_pct_df,
                !taxonomic_group %in% taxonomic_groups_sig_interaction_MeJA$taxonomic_group,
                residual_permanova<0.85,
                var_explined_ratio>1.25,
                unimodal_residues001 >=2,
                residue_dist_simialr_to_K__bacteria001 >=2) 

taxonomic_groups_nonsig_SA<-
  dplyr::filter(perm_528_phylosig_pct_df, 
                !taxonomic_group %in% taxonomic_groups_sig_interaction_SA$taxonomic_group,
                residual_permanova<0.85,
                var_explined_ratio>1.25,
                unimodal_residues001 >=2,
                residue_dist_simialr_to_K__bacteria001 >=2)

#filter df to the taxonomies and samples above
heatmap_input_filtered_nonsig_Control<-
  dplyr::filter(heatmap_input, 
                heatmap_input$taxonomic_group %in% taxonomic_groups_nonsig_Control$taxonomic_group,
                higher_lower_PC == "Higher_PC", 
                Stress =="Control") 

heatmap_input_filtered_nonsig_MeJA<-
  dplyr::filter(heatmap_input, 
                heatmap_input$taxonomic_group %in% taxonomic_groups_nonsig_MeJA$taxonomic_group,
                higher_lower_PC == "Higher_PC", 
                Stress =="MeJA") 
  

heatmap_input_filtered_nonsig_SA<-
  dplyr::filter(heatmap_input, 
                heatmap_input$taxonomic_group %in% taxonomic_groups_nonsig_SA$taxonomic_group,
                higher_lower_PC == "Higher_PC", 
                Stress =="SA") 
  
  
#heatmap plots
heatmap_PCoA_coord_nonsig_Control<-
ggplot(data = heatmap_input_filtered_nonsig_Control, aes(y = Sp_abb_name, x = taxonomic_group, fill = scaled_PCoA))+
  geom_tile()+
  scale_y_discrete(limits=rev)+
    scale_fill_viridis(option = "B")+
    scale_color_viridis(option = "B")+
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1))+
  facet_wrap(~Stress)+
  ylab(NULL)

heatmap_PCoA_coord_nonsig_MeJA<-
ggplot(data = heatmap_input_filtered_nonsig_MeJA, aes(y = Sp_abb_name, x = taxonomic_group, fill = scaled_PCoA))+
  geom_tile()+
  scale_y_discrete(limits=rev)+
    scale_fill_viridis(option = "B")+
    scale_color_viridis(option = "B")+
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1))+
  facet_wrap(~Stress)+
  ylab(NULL)

heatmap_PCoA_coord_nonsig_SA<-
ggplot(data = heatmap_input_filtered_nonsig_SA, aes(y = Sp_abb_name, x = taxonomic_group, fill = scaled_PCoA))+
  geom_tile()+
  scale_y_discrete(limits=rev)+
    scale_fill_viridis(option = "B")+
    scale_color_viridis(option = "B")+
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1))+
  facet_wrap(~Stress)+
  ylab(NULL)

#arrange plots proportionally
heamap_diff_taxonomic_group_per_stress_nonsig<-
ggarrange(heatmap_PCoA_coord_nonsig_Control,
          heatmap_PCoA_coord_nonsig_MeJA,
          heatmap_PCoA_coord_nonsig_SA,
          nrow = 1, 
          common.legend = TRUE, legend = "bottom",
          align = "h",  
          widths = c(dim(heatmap_input_filtered_nonsig_Control)[1],
                     dim(heatmap_input_filtered_nonsig_MeJA)[1],
                     dim(heatmap_input_filtered_nonsig_SA)[1]))



#prepare barpltos to show on top of heatmaps
barplot_Control_nonsig_abouheif<-
ggplot(data = taxonomic_groups_nonsig_Control, aes(x = taxonomic_group, y = value_Control_Higher_PC))+
  geom_bar(stat = "identity", 
           fill = colour_v[5],
           alpha = if_else(condition = taxonomic_groups_nonsig_Control$pvalue_Control_Higher_PC<0.05,
                     true = 1 ,
                     false = 0.3))+
    geom_point(aes(x = taxonomic_group, 
                 y = r2_Mantel_Control), 
             alpha = if_else(condition = taxonomic_groups_nonsig_Control$p_value_Mantel_Control<0.05,
                     true = 1 ,
                     false = 0.3),
             size = 2)+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        legend.position = "none")+  
    ylim(c(-0.2,0.61))+
    ylab("Control - Bars: Abouheif's C Mean / Dots: Mantel R2")

barplot_MeJA_nonsig_abouheif<-
ggplot(data = taxonomic_groups_nonsig_MeJA, aes(x = taxonomic_group, y = value_MeJA_Higher_PC))+
 geom_bar(stat = "identity",
           fill = colour_v[3],
           alpha = if_else(condition = taxonomic_groups_nonsig_MeJA$pvalue_MeJA_Higher_PC<0.05,
                   true = 1 ,
                   false = 0.3))+
    geom_point(aes(x = taxonomic_group, 
                 y = r2_Mantel_MeJA), 
             alpha = if_else(condition = taxonomic_groups_nonsig_MeJA$p_value_Mantel_MeJA<0.05,
                     true = 1 ,
                     false = 0.3 ),
             size = 2)+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        legend.position = "none")+
  ylim(c(-0.2,0.61))+
  #  ylim(c(0,1))+
    ylab("MeJA - Bars: Abouheif's C Mean / Dots: Mantel R2")


barplot_SA_nonsig_abouheif<-
ggplot(data = taxonomic_groups_nonsig_SA, aes(x = taxonomic_group, y = value_SA_Higher_PC))+
 geom_bar(stat = "identity", 
           fill = colour_v[4],
           alpha = if_else(condition = taxonomic_groups_nonsig_SA$pvalue_SA_Higher_PC<0.05,
                  true = 1 ,
                  false = 0.3))+
  geom_point(aes(x = taxonomic_group, 
                 y = r2_Mantel_SA), 
             alpha = if_else(condition = taxonomic_groups_nonsig_SA$p_value_Mantel_SA<0.05,
                     true = 1 ,
                     false = 0.3 ),
             size = 2)+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
         legend.position = "none")+
  ylim(c(-0.2,0.61))+
   # ylim(c(0,1))+
    ylab("SA - Bars: Abouheif's C Mean / Dots: Mantel R2")



#arrange plots proportionally
barplot_abouheif_taxonomic_group_per_stress_nonsig<-
ggarrange(barplot_Control_nonsig_abouheif,
          barplot_MeJA_nonsig_abouheif,
          barplot_SA_nonsig_abouheif,
          nrow = 1, 
          common.legend = TRUE, 
          align = "h",  
          widths = c(dim(heatmap_input_filtered_nonsig_Control)[1],
                     dim(heatmap_input_filtered_nonsig_MeJA)[1],
                     dim(heatmap_input_filtered_nonsig_SA)[1]))


#put barplots on top of heatmaps, proportionally
barplots_Abouheif_heatmaps_PCoA_nonsig<-
ggarrange(barplot_abouheif_taxonomic_group_per_stress_nonsig,
          heamap_diff_taxonomic_group_per_stress_nonsig,
          nrow = 2,
          heights = c(0.8,3),
          align = "hv")



#save heatmap, significant interactions and abouheif at each stress
ggsave(barplots_Abouheif_heatmaps_PCoA_nonsig, 
       filename = "./14_Phylogenetic_distance/results/Results_April2023/trait_on_tree_heatmap_barplot_nonsig__PCratio125_residual085_unimodal2_SK2_overlap.pdf",
     width = 700, 
     height = 300, 
     units = "mm")



```


#### ####################
### ## heat_tree plots ##
#### ####################


# heat_tree: prepare phylosig pipeline output heat tree visualization

we only keep teh higehr of 2 PCoAs for simplification. this means that microbial  taxa with a significat abouheif value in both PCs is treated jsut like taxa with significant abouheif value in a single PC. the comparision is between the higeest abouheif value we could find
```{r}

# filter to higher PC only; remove f/g__uncultured that is meaningless
all_taxa_phylosig<-dplyr::filter(bac_alltaxa_phylosig_output_PCoA_nonsig$df_phylosig, 
                                 higher_lower_PC == "Higher_PC", 
                                 column_label!="NAs")%>%
  filter(!grepl("uncultured", column_label))


```



# heat_tree: create taxmap object for phylosig heat tree
```{r}

# fitler phyloseq to remove taxa abset from df
to_parse<-CSS_BacFun_rhizoplane_HighFilt_ps_l$Bac

# turn NA in taxonomy (such as c__alphaproteobacteria; NA;NA;NA) into "" to avoid removal from tree
to_parse@tax_table@.Data[is.na(to_parse@tax_table)]<-""

#remove sp and taxonomy confidence
to_parse@tax_table<-to_parse@tax_table[,1:6]

#we won't need reference sequences, lets remove as they were giving arnings
to_parse@refseq<-NULL

#parse phylsoeq to metacoder
metacoder_bac<-parse_phyloseq(to_parse)

```


## heat_tree: filter and prepare taxmap object
```{r}

#define the taxonomies inside our taxmap object
library(purrr)
char_taxmap<-lapply(metacoder_bac$taxa, function (x)
    x$get_name())%>%
    purrr::map(1)

# subset taxonomies, keeping only those present in all_taxa_phylosig 
metacoder_bac<-metacoder::filter_taxa(metacoder_bac, 
                                      char_taxmap%>%as.character() %in% unique(filtered_permavonas$taxonomic_group),
                                      reassign_obs = FALSE)


#rearrange phylosig df to fit metacoder
phylosig_to_taxmap<-
pivot_wider(data = dplyr::select(all_taxa_phylosig, column_label, stress, value, pvalue),
            names_from = stress, 
            values_from = c(value, pvalue))

#########################################
#### IMPORTANT, PLOT-AFFECTING STEP #### 
#filter df remove taxonomic groups that were dsicarded due to permanova/PCOA quality filtering
phylosig_to_taxmap<-
  phylosig_to_taxmap[phylosig_to_taxmap$column_label %in% filtered_permavonas$taxonomic_group,]
#########################################

#check if the taxmap and phylisg_df tazons are identical (in teh same order)
identical(lapply(metacoder_bac$taxa, function (x)
           x$get_name())%>%
           purrr::map(1)%>%as.character(),  
          unique(phylosig_to_taxmap$column_label))

#add tax map codes (aab, aac...) to the phylosig df
metacoder_taxon_names<-data.frame(taxon_id = as.character(taxon_names(metacoder_bac)%>%names), 
                                  column_label = taxon_names(metacoder_bac)%>%as.character)

phylosig_to_taxmap<-left_join(phylosig_to_taxmap, metacoder_taxon_names)

# add phylosig df to taxmap                                  
metacoder_bac$data$phylosig_to_taxmap<-phylosig_to_taxmap

# check again if taxonomy orders match
identical(taxon_names(metacoder_bac)%>%names, metacoder_bac$data$phylosig_to_taxmap$taxon_id)

#count number of unique with significant phylosig (per taxon level)
metacoder_bac$data$tax_data$Kingdom%>%unique%>%length()
metacoder_bac$data$tax_data$Phylum%>%unique%>%length()
metacoder_bac$data$tax_data$Class%>%unique%>%length()
metacoder_bac$data$tax_data$Order%>%unique%>%length()
metacoder_bac$data$tax_data$Family%>%unique%>%length()
metacoder_bac$data$tax_data$Genus%>%unique%>%length()

```

## heat_tree: add mantel test results to the phylosig_to_taxmap
```{r}
#subset 528_permanova_df to only contain mantel test values
mantel_to_taxmap_df<-
  dplyr:: select(perm_528_phylosig_pct_df, 
                 taxonomic_group,
                 r2_Mantel_Control,
                 r2_Mantel_MeJA,
                 r2_Mantel_SA,
                 p_value_Mantel_Control,
                 p_value_Mantel_MeJA,
                 p_value_Mantel_SA)
               
# add mantel test results to the phylosig_to_taxmap object
phylosig_to_taxmap<-left_join(phylosig_to_taxmap, mantel_to_taxmap_df, by = c("column_label" ="taxonomic_group"))
```



### heat_tree: Compare and plot control and MeJA with metacoder
```{r}

# remove SA
metacoder_Control_MeJA<-
filter_obs(obj = metacoder_bac, 
           data = c("sample_data"),
           Stress %in% c("Control","MeJA"), 
           drop_obs = TRUE)




#abundance per taxon 
metacoder_Control_MeJA$data$tax_abund<-calc_taxon_abund(obj = metacoder_Control_MeJA, 
                                            data = "otu_table", 
                                            cols = metacoder_Control_MeJA$data$sample_data$sample_id) 

#get occurrence of ASVs per treatment
metacoder_Control_MeJA$data$tax_occ_3treatments <- calc_n_samples(obj = metacoder_Control_MeJA, 
                                                      data = "tax_abund", 
                                                      cols = metacoder_Control_MeJA$data$sample_data$sample_id,
                                                      groups = metacoder_Control_MeJA$data$sample_data$Stress) 

# calculate log2 median ratios 
metacoder_Control_MeJA$data$diff_table_3treatments <- compare_groups(obj = metacoder_Control_MeJA,
                                                            data = "tax_abund",
                                                            cols = metacoder_Control_MeJA$data$sample_data$sample_id, 
                                                            groups = metacoder_Control_MeJA$data$sample_data$Stress) 

# define column with the difference on phylosig value of treatment 1 vc treatment 2 in a format metacoder understands
metacoder_Control_MeJA$data$diff_table_3treatments$t1_t2_phylosig<- 
  metacoder_Control_MeJA$data$phylosig_to_taxmap$value_Control - metacoder_Control_MeJA$data$phylosig_to_taxmap$value_MeJA


# save vector of taxonomies in the taxmap object for the label colors operations below
taxmap_label<-metacoder_Control_MeJA$data$phylosig_to_taxmap$column_label


# make heat tree of phylogenetic signal
heat_tree(.input = metacoder_Control_MeJA,
                 node_label = taxon_names,
                 node_color = metacoder_Control_MeJA$data$diff_table_3treatments$t1_t2_phylosig, 
                 node_color_range =  c(colour_v[3], "#FAF9F6", colour_v[5]),
                 node_color_trans = "linear", 
                 node_color_interval = c(-0.5, 0.5), 
                 edge_color_interval = c(-0.5, 0.5), 
                 node_size_axis_label = "Size: large for significant Abouheif & Mantel",
                 node_size = ifelse(phylosig_to_taxmap$pvalue_MeJA<0.05 &
                                      phylosig_to_taxmap$p_value_Mantel_MeJA<0.05 | 
                                      phylosig_to_taxmap$pvalue_Control <0.05 &
                                      phylosig_to_taxmap$p_value_Mantel_Control<0.05, 
                                    yes = 2.5,
                                    no = 0.25),
                 node_color_axis_label = "Color: difference in Abouheif", #; Yellow MeJA Grey Control
                 initial_layout = "reingold-tilford",
                 output_file = "./14_Phylogenetic_distance/results/Results_April2023/phylosig_tree_PCoA_Control_MeJA_PERMVANOVA_filt.pdf")   


```


### heat_tree: Compare and plot control and SA with metacoder

```{r}

# remove MeJA
metacoder_Control_SA<-
filter_obs(obj = metacoder_bac, 
           data = c("sample_data"),
           Stress %in% c("Control","SA"), 
           drop_obs = TRUE)




#abundance per taxon 
metacoder_Control_SA$data$tax_abund<-calc_taxon_abund(obj = metacoder_Control_SA, 
                                            data = "otu_table", 
                                            cols = metacoder_Control_SA$data$sample_data$sample_id) 

#get occurrence of ASVs per treatment
metacoder_Control_SA$data$tax_occ_3treatments <- calc_n_samples(obj = metacoder_Control_SA, 
                                                      data = "tax_abund", 
                                                      cols = metacoder_Control_SA$data$sample_data$sample_id,
                                                      groups = metacoder_Control_SA$data$sample_data$Stress) 

# calculate log2 median ratios 
metacoder_Control_SA$data$diff_table_3treatments <- compare_groups(obj = metacoder_Control_SA,
                                                            data = "tax_abund",
                                                            cols = metacoder_Control_SA$data$sample_data$sample_id, 
                                                            groups = metacoder_Control_SA$data$sample_data$Stress) 

# define column with the difference on phylosig value of treatment 1 vc treatment 2 in a format emtacoder understands
metacoder_Control_SA$data$diff_table_3treatments$t1_t2_phylosig<- 
  metacoder_Control_SA$data$phylosig_to_taxmap$value_Control - metacoder_Control_SA$data$phylosig_to_taxmap$value_SA


# save vector of taxonomies in the taxmap object for the label colors operations below
taxmap_label<-metacoder_Control_SA$data$phylosig_to_taxmap$column_label


# in red, golden taxa that overlap between methods (most on SA)
set.seed(1)
heat_tree(.input = metacoder_Control_SA,
                 node_label = taxon_names,
                 node_color = metacoder_Control_SA$data$diff_table_3treatments$t1_t2_phylosig, 
                 node_color_range =  c(colour_v[4], "#FAF9F6", colour_v[5]),
                 node_color_trans = "linear", 
                 node_color_interval = c(-0.5, 0.5), 
                 edge_color_interval = c(-0.5, 0.5), 
                 node_size_axis_label = "Size: large for significat Abouheif",
                 node_size = ifelse(phylosig_to_taxmap$pvalue_SA<0.05 &
                                       phylosig_to_taxmap$p_value_Mantel_SA<0.05 | 
                                      phylosig_to_taxmap$pvalue_Control <0.05 &
                                      phylosig_to_taxmap$p_value_Mantel_Control<0.05, 
                                    yes = 2.5,
                                    no = 0.25),
                 node_color_axis_label = "Color: difference in Abouheif", #; Green SA Grey Control
                 layout = "davidson-harel", # The primary layout algorithm
                 initial_layout = "reingold-tilford",
                 output_file = "./14_Phylogenetic_distance/results/Results_April2023/phylosig_tree_PCoA_Control_SA_PERMVANOVA_filt.pdf")   




```






# -----------------
# FUNGAL phylogenetic signal
# -----------------

# 14.2 - load ps, remove species outside tree 

```{R}

#load the output of the phylogenetic signal pipeline - all taxa in 1 list
load(file = "./14_Phylogenetic_distance/r_objects/fun_alltaxa_phylosig_PCoA_nonsig.RData") 

#load the output of the phylogenetic signal pipeline - one list per taxonomic rank
load(file = "./14_Phylogenetic_distance/r_objects/per_level_fun_phylosig_PPCA.RData") 

# a df with various metrics of the 70 permanovas
load(file = "./14_Phylogenetic_distance/r_objects/perm_70_phylosig_pct_df.RData")

# a df with of permanovas filtered absed on various metrics
# load(file = "./14_Phylogenetic_distance/r_objects/filtered_permavonas.RData")

#load output of mantel pipeline
load(file = "./14_Phylogenetic_distance/r_objects/fun_alltaxa_mantel_pipeline_df.RData")

# load hundreds of PCoA plots
load(file = "./14_Phylogenetic_distance/r_objects/PCoA_Fun_70_plots_axis_pct.RData") # one per tax group
load(file = "./14_Phylogenetic_distance/r_objects/PCoA_Fun_70_3x_plots_axis_pct.RData") # one per tax group and per stress

# load PS data
load("./Data/phyloseq_objects/CSS_BacFun_rhizoplane_HighFilt_ps_l.RData")

# remove all species outside the MY trees
CSS_BacFun_rhizoplane_HighFilt_ps_l <- 
  lapply(CSS_BacFun_rhizoplane_HighFilt_ps_l, function(x) 
    subset_samples(x, MY_tree_presence == "inside_MY_tree"))



```


#### ####################
## ## Trait_on_tree plots ##
#### ####################

# Trait_on_tree plot: heatmap of PCoA coodinaties per plant species

This is the basis of the main figures for the paper.
All plots only contains the taxonomic groups that passed permanova/PCOA filtering

## plot brassicaceae phylogeny tree, export as pdf


```{r}
#load tree
Bra_tree_for_plot<-load_prepare_MY_tree()

#save tree plot
 pdf(file = "./14_Phylogenetic_distance/results/Results_April2023/bra_tree_hendriks.pdf",
     width = 4, 
     height = 6) 
 
 #plot
 plot(Bra_tree_for_plot)
  dev.off()

  
  
  

```


## Trait_on_tree: prepare df for heatmap

```{r}

# create df with PCoA coordinates for a heatmap plot
re_created_traitontree<-
lapply(fun_alltaxa_phylosig_PCoA_nonsig$ordination_coordinates_l, function(x){
  
  
  #extract emtadata
  metadata<-as(x@sam_data, "data.frame")
  
  #add extended lineage  
  metadata$Sp_Lineage_Walden_extended<-  
        if_else(condition = metadata$Sp_abb_name %in% c("Ia", "Lm", "Co", "Bi"), 
              true ="lineage_II_extended",
              false =  metadata$Sp_Lineage_Walden)
        
        
        
  
  #split by stress
  metadata_split<-split(metadata, f = metadata$Stress)
  

  #calcualte centroid
centroid_l<-
  lapply(metadata_split, function(z){
    
    #calculate centroid
    centroid <- aggregate(cbind(PCoA_1, PCoA_2) ~ Sp_abb_name, data = z, FUN = mean)

    #if Sar is positive on the PC1, rotate the axis by multiplying it by -1
    if(filter(centroid,Sp_abb_name == "Sar")$PCoA_1>0) {
       centroid$PCoA_1<-centroid$PCoA_1*-1 
    }
    
    #if Sar is positive on the PC2, rotate the axis by multiplying it by -1
    if(filter(centroid,Sp_abb_name == "Sar")$PCoA_2>0) {
       centroid$PCoA_2<-centroid$PCoA_2*-1 
    }


    #scale and centr PCoA1 and 2
    centroid$PCoA_1_centroid_scaled<-scale(centroid$PCoA_1,center = TRUE, scale = TRUE)
    centroid$PCoA_2_centroid_scaled<-scale(centroid$PCoA_2,center = TRUE, scale = TRUE)
    
    return(centroid)
  })

# extract per-species metadata
unique_meta_l<-lapply(metadata_split, function(y)
  dplyr::select(.data = y,
                        Sp_abb_name, 
                        Stress)%>%unique)

#join centroids with per-species emtadata
centroid_with_metadata_l<-
  mapply(function (a,b)
    left_join(a,b),
    a = centroid_l,
    b = unique_meta_l,
    SIMPLIFY = FALSE)

centroid_with_metadata<-
  do.call("rbind", centroid_with_metadata_l)

rownames(centroid_with_metadata)<-NULL

return(centroid_with_metadata)

})

# prepare df for heatmap plot 
re_created_traitontree_df<-
  do.call("rbind", re_created_traitontree)%>%rownames_to_column(var = "taxonomic_group")

re_created_traitontree_df$taxonomic_group<-
  gsub(x = re_created_traitontree_df$taxonomic_group, pattern = "\\.\\d+", replacement = "")


#make table longer so all scaled PCs are in one column
re_created_traitontree_df<-
  pivot_longer(data = re_created_traitontree_df, cols = 5:6, names_to = "scaled_PCoA_axis", values_to = "scaled_PCoA")

  #re-join pcoa axis with taxonomic group
trait_naming_output<-
  mapply(function (z,y)
         gsub(x = z,
              pattern = "centroid_scaled",
              replacement = y),
         z = re_created_traitontree_df$scaled_PCoA_axis,
         y = re_created_traitontree_df$taxonomic_group,
         SIMPLIFY = TRUE )
re_created_traitontree_df$trait<-as.vector(trait_naming_output)

#join hi/low pc information
heatmap_input<-
  left_join(re_created_traitontree_df,
            dplyr::select(fun_alltaxa_phylosig_PCoA_nonsig$df_phylosig, higher_lower_PC, trait,stress))


#prepare order of facotrs
taxonomic_group_ordering<-
  factor(fun_alltaxa_phylosig_PCoA_nonsig$df_phylosig$column_label%>%unique,
         levels = fun_alltaxa_phylosig_PCoA_nonsig$df_phylosig$column_label%>%unique)

#adjust order of factors
heatmap_input$taxonomic_group<-
  factor(heatmap_input$taxonomic_group, levels = levels(taxonomic_group_ordering))

perm_70_phylosig_pct_df$taxonomic_group<-
  factor(perm_70_phylosig_pct_df$taxonomic_group, levels = levels(taxonomic_group_ordering))

# change order of the species to reflect newst phylogeny
heatmap_input$Sp_abb_name<-factor(heatmap_input$Sp_abb_name, levels = Bra_tree_for_plot$tip.label)%>%
  fct_rev() # rever order to keet At as the first
  



```

## Trait_on_tree: filter df (significant interaction, abouheif) and plot heatmap
this plot only contains the taxonomic groups that passed permanova/PCOA filtering
these are further filtered to contain only taxa with significat sp*stress interactions and significat abouheif signals

plots heatmaps + barplots with abouheif

*ATTENTION* this is the chunk that defines teh filtering of taxonomic groups that will be showin in figure 1!
```{r}



#####  ------------------- #####
#####  different taxonomic groups per stress (panel of 3 plots) #####
#####  ------------------- #####



# only keep taxonomic groups with...
  # significant interaction 
  # at least 15% variance explained by permanova (residual smaller than 0.85, simialr to K__fungi)
  # PCoA_1 explains at least 1.15 time the variance of PCOA_2 (sufficient variance explained in 2 dimensions, simialr to K__fungi)
  # the 3 factors (block, plant species, stress) presents unimodal distanes to centroids (p<0.05)
  #  the 3 factors (block, plant species, stress) presents distanes to centroids distribution similar to k__Bacteria (p<0.05)


taxonomic_groups_sig_interaction_Control<-
  dplyr::filter(perm_70_phylosig_pct_df, 
                pvalue_Sp_x_Stress<0.05,
                residual_permanova<0.85,
                var_explined_ratio>1.15,
                unimodal_residues >=3,
                residue_dist_simialr_to_K__Fungi >=3)


taxonomic_groups_sig_interaction_MeJA<-
  dplyr::filter(perm_70_phylosig_pct_df, 
                pvalue_Sp_x_Stress<0.05, 
                residual_permanova<0.85,
                var_explined_ratio>1.15,
                unimodal_residues >=3,
                residue_dist_simialr_to_K__Fungi >=3)


taxonomic_groups_sig_interaction_SA<-
  dplyr::filter(perm_70_phylosig_pct_df, 
                pvalue_Sp_x_Stress<0.05, 
                residual_permanova<0.85,
                var_explined_ratio>1.15,
                unimodal_residues >=3,
                residue_dist_simialr_to_K__Fungi >=3)


#filter df to the taxonomies and samples above
heatmap_input_filtered_interaction_Control<-
  dplyr::filter(heatmap_input, 
                heatmap_input$taxonomic_group %in% taxonomic_groups_sig_interaction_Control$taxonomic_group,
                higher_lower_PC == "Higher_PC", 
                Stress =="Control") 

heatmap_input_filtered_interaction_MeJA<-
  dplyr::filter(heatmap_input, 
                heatmap_input$taxonomic_group %in% taxonomic_groups_sig_interaction_MeJA$taxonomic_group,
                higher_lower_PC == "Higher_PC", 
                Stress =="MeJA") 
  

heatmap_input_filtered_interaction_SA<-
  dplyr::filter(heatmap_input, 
                heatmap_input$taxonomic_group %in% taxonomic_groups_sig_interaction_SA$taxonomic_group,
                higher_lower_PC == "Higher_PC", 
                Stress =="SA") 
  
  
#heatmap plots
heatmap_PCoA_coord_005_Control<-
ggplot(data = heatmap_input_filtered_interaction_Control, aes(y = Sp_abb_name, x = taxonomic_group, fill = scaled_PCoA))+
  geom_tile()+
  scale_y_discrete(limits=rev)+
    scale_fill_viridis(option = "B")+
    scale_color_viridis(option = "B")+
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1))+
  facet_wrap(~Stress)+
  ylab(NULL)

heatmap_PCoA_coord_005_MeJA<-
ggplot(data = heatmap_input_filtered_interaction_MeJA, aes(y = Sp_abb_name, x = taxonomic_group, fill = scaled_PCoA))+
  geom_tile()+
  scale_y_discrete(limits=rev)+
    scale_fill_viridis(option = "B")+
    scale_color_viridis(option = "B")+
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1))+
  facet_wrap(~Stress)+
  ylab(NULL)

heatmap_PCoA_coord_005_SA<-
ggplot(data = heatmap_input_filtered_interaction_SA, aes(y = Sp_abb_name, x = taxonomic_group, fill = scaled_PCoA))+
  geom_tile()+
  scale_y_discrete(limits=rev)+
    scale_fill_viridis(option = "B")+
    scale_color_viridis(option = "B")+
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1))+
  facet_wrap(~Stress)+
  ylab(NULL)

#arrange plots proportionally
heamap_diff_taxonomic_group_per_stress<-
ggarrange(heatmap_PCoA_coord_005_Control,
          heatmap_PCoA_coord_005_MeJA,
          heatmap_PCoA_coord_005_SA,
          nrow = 1, 
          common.legend = TRUE, legend = "bottom",
          align = "h",  
          widths = c(dim(heatmap_input_filtered_interaction_Control)[1],
                     dim(heatmap_input_filtered_interaction_MeJA)[1],
                     dim(heatmap_input_filtered_interaction_SA)[1]))



#prepare barpltos to show on top of heatmaps

barplot_Control_abouheif<-
ggplot(data = taxonomic_groups_sig_interaction_Control, aes(x = taxonomic_group, y = value_Control_Higher_PC))+
  geom_bar(stat = "identity", 
           fill = colour_v[5],
           alpha = if_else(condition = taxonomic_groups_sig_interaction_Control$pvalue_Control_Higher_PC<0.05,
                     true = 1 ,
                     false = 0.3))+
    geom_point(aes(x = taxonomic_group, 
                 y = r2_Mantel_Control), 
             alpha = if_else(condition = taxonomic_groups_sig_interaction_Control$p_value_Mantel_Control<0.05,
                     true = 1 ,
                     false = 0.3),
             size = 2)+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        legend.position = "none")+
  ylim(c(0,0.61))+
#  ylim(c(0,1))+
  ylab("Control - Bars: Abouheif's C Mean / Dots: Mantel R2")

barplot_MeJA_abouheif<-
ggplot(data = taxonomic_groups_sig_interaction_MeJA, aes(x = taxonomic_group, y = value_MeJA_Higher_PC))+
  geom_bar(stat = "identity",
           fill = colour_v[3],
           alpha = if_else(condition = taxonomic_groups_sig_interaction_MeJA$pvalue_MeJA_Higher_PC<0.05,
                   true = 1 ,
                   false = 0.3))+
    geom_point(aes(x = taxonomic_group, 
                 y = r2_Mantel_MeJA), 
             alpha = if_else(condition = taxonomic_groups_sig_interaction_MeJA$p_value_Mantel_MeJA<0.05,
                     true = 1 ,
                     false = 0.3 ),
             size = 2)+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        legend.position = "none")+
  ylim(c(0,0.61))+
  #  ylim(c(0,1))+
    ylab("MeJA - Bars: Abouheif's C Mean / Dots: Mantel R2")


barplot_SA_abouheif<-
ggplot(data = taxonomic_groups_sig_interaction_SA, aes(x = taxonomic_group, y = value_SA_Higher_PC))+
  geom_bar(stat = "identity", 
           fill = colour_v[4],
           alpha = if_else(condition = taxonomic_groups_sig_interaction_SA$pvalue_SA_Higher_PC<0.05,
                  true = 1 ,
                  false = 0.3))+
  geom_point(aes(x = taxonomic_group, 
                 y = r2_Mantel_SA), 
             alpha = if_else(condition = taxonomic_groups_sig_interaction_SA$p_value_Mantel_SA<0.05,
                     true = 1 ,
                     false = 0.3 ),
             size = 2)+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
         legend.position = "none")+
  ylim(c(0,0.61))+
   # ylim(c(0,1))+
    ylab("SA - Bars: Abouheif's C Mean / Dots: Mantel R2")




#arrange plots proportionally
barplot_abouheif_taxonomic_group_per_stress<-
ggarrange(barplot_Control_abouheif,
          barplot_MeJA_abouheif,
          barplot_SA_abouheif,
          nrow = 1, 
          common.legend = TRUE, 
          align = "h",  
          widths = c(dim(heatmap_input_filtered_interaction_Control)[1],
                     dim(heatmap_input_filtered_interaction_MeJA)[1],
                     dim(heatmap_input_filtered_interaction_SA)[1]))


#put barplots on top of heatmaps, proportionally
barplots_Abouheif_heatmaps_PCoA<-
ggarrange(barplot_abouheif_taxonomic_group_per_stress,
          heamap_diff_taxonomic_group_per_stress,
          nrow = 2,
          heights = c(1,3),
          align = "hv")




#save heatmap, significant interactions and abouheif at each stress
ggsave(barplots_Abouheif_heatmaps_PCoA, 
       filename = "./14_Phylogenetic_distance/results/Results_April2023/trait_on_tree_FUN_heatmap_barplot_005Interaction_005Abouheif__PCratio125_residual085_unimodal2_SK2_mantelR2dot_overlap.pdf",
     width = 275, 
     height = 375, 
     units = "mm")



```


