---
title: "1_loading_and_pre_processing"
author: "Pedro Beschoren da Costa"
date: "August 24, 2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

Code writen by Pedro and modified by Zulema to have unfiltered phyloseq object input for alpha diversity testing. 


# 1.0 load all libraries, custom functions, setup environemnt
```{r results="hide"}


# loads all libraries 
source("./renv/0_load_libraries.R")

install.packages("remotes")
remotes::install_github("benjjneb/decontam")
library("decontam")

# custom functions related to data loading and decontamination
source("./01_Preprocessing/code/functions_loading_and_decontamination.R")

# increases memory limit used by R
memory.limit(size = 350000)

```


# 1.0 load microbiome data from biom file
our data (processed in qiime2) was saved in BIOM format, making importing easier since there are less files to handle



## 1.1.1 - Loading 16S data 
our 16S data takes 7.3Gb in memory if not filtered before saving in the R environment. after filtering (removing host DNA and ASVs occuring less than 7 times) the object takes 2.1Gb, which is much more manageable
```{r results="hide"}
##### Loading 16S  data ##########

# loads the OTU table, immediatetly saving it as a phyloseq OTU table (lighter than a df)
raw_bac_otutab<-otu_table(object = read.table(file = "./data/Amplicon_sequencing/16S_seqtab_final.txt", 
                                              header = TRUE,
                                              row.names = 1, # first column has row names (ASV names)
                                              check.names = FALSE), # prevents "X" to be added to column names, such as X49_16S,
                           taxa_are_rows = TRUE)


#updated tax
sklearn_bac_taxtab<-read.table(file = "./data/Amplicon_sequencing/16S_taxonomy_sklearn.tsv", 
                               header = TRUE,
                               sep = "\t",
                               row.names = 1, # first column has row names (ASV names)
                               check.names = FALSE) # prevents "X" to be added to column names, such as X49_16S,

# adjusts number and name of columns
sklearn_bac_taxtab<- separate(data = sklearn_bac_taxtab,
                               col = Taxon,
                               into =c("Kingdom", 
                                       "Phylum", 
                                       "Class", 
                                       "Order", 
                                       "Family",
                                       "Genus", 
                                       "Species"),
                               sep = ";")

# change from d__Bacteria to k__bacteria, mathcing fungi dataset
sklearn_bac_taxtab$Kingdom<-gsub("d__", "k__", sklearn_bac_taxtab$Kingdom)

# saves taxa as phyloseq object
sklearn_bac_taxtab<-tax_table(object = as.matrix(sklearn_bac_taxtab))

#change name and remove old object
raw_bac_taxtab<-sklearn_bac_taxtab
rm(sklearn_bac_taxtab)

# loads the representative sequences table, immediatetly saving it as a phyloseq OTU table (lighter than a df)
raw_bac_refseq<-refseq(physeq = Biostrings::readDNAStringSet(filepath = "./data/Amplicon_sequencing/16S_repset.fasta", use.names = TRUE)) 
taxa_names(raw_bac_refseq)<-gsub(" .*", "", taxa_names(raw_bac_refseq)) # drops taxonomy from ASV names

# loads the mapping tile, immediatetly saving it as a phyloseq OTU table (lighter than a df)
raw_bac_metadata<-sample_data(object = read.table(file = "./data/Amplicon_sequencing/Mapping_file_16S_Family_experiment.txt", 
                                                  header = TRUE,
                                                  sep = "\t",
                                                  row.names = 1, # first column has row names (ASV names)
                                                  check.names = FALSE)) # prevents "X" to be added to column names, such as X49_16S,

# build the main 16S phyloseq object by putting all these phyloseq-class objects (otu table, tax table, ref seq, metadta) into a single ps
raw_bac_ps<-merge_phyloseq(raw_bac_otutab,
                           raw_bac_taxtab,
                           raw_bac_refseq,
                           raw_bac_metadata)

# remove old objects to reduce memory use
rm(raw_bac_otutab,raw_bac_taxtab,raw_bac_refseq,raw_bac_metadata)

# change names from ASV to fASV, so they can be distinguished from bacterial ASVs
taxa_names(raw_bac_ps)<-paste("b", taxa_names(raw_bac_ps), sep = "")
tax_table(raw_bac_ps)<-gsub(" ", "", tax_table(raw_bac_ps)) # drops a space charather from taxa names. I don't know how that charather ended up in there

# let's check the imported objects. Often errors will arise from typos when filling up the data sheets 
otu_table(raw_bac_ps)[1:10,1:10]
sample_data(raw_bac_ps)[1:10,1:10]
tax_table(raw_bac_ps)[1:10,1:6]
refseq(raw_bac_ps)



# run garbage collection after creating large objects
gc()

```




## 1.1.2 - Loading ITS data 
This will load all the essential data (OTU frequencies,  metadata, taxonomy) for a single phyloseq object
```{r results="hide"}

# loads the OTU table, inmediatetly saving it as a phyloseq OTU table (lighter than a df)
raw_fun_otutab<-otu_table(object = read.table(file = "./data/Amplicon_sequencing/ITS_seqtab_final.txt", 
                                              header = TRUE,
                                              row.names = 1, # first column has row names (ASV names)
                                              check.names = FALSE), # prevents "X" to be added to column names, such as X49_ITS,
                           taxa_are_rows = TRUE)



#updated tax
sklearn_fun_taxtab<-read.table(file = "./data/Amplicon_sequencing/ITS_taxonomy_sklearn.tsv", 
                               header = TRUE,
                               sep = "\t",
                               row.names = 1, # first column has row names (ASV names)
                               check.names = FALSE) # prevents "X" to be added to column names, such as X49_16S,

# adjusts number and name of columns
sklearn_fun_taxtab<- separate(data = sklearn_fun_taxtab,
                               col = Taxon,
                               into =c("Kingdom", 
                                       "Phylum", 
                                       "Class", 
                                       "Order", 
                                       "Family",
                                       "Genus", 
                                       "Species"),
                               sep = ";")

# saves taxa as phyloseq object
sklearn_fun_taxtab<-tax_table(object = as.matrix(sklearn_fun_taxtab))

# renames object, remve old version7
raw_fun_taxtab<-sklearn_fun_taxtab
rm(sklearn_fun_taxtab)



# loads the representative sequences table, immediatetly saving it as a phyloseq OTU table (lighter than a df)
raw_fun_refseq<-refseq(physeq = Biostrings::readDNAStringSet(filepath = "./data/Amplicon_sequencing/ITS_repset.fasta", use.names = TRUE)) 
taxa_names(raw_fun_refseq)<-gsub(" .*", "", taxa_names(raw_fun_refseq)) # drops taxonomy from ASV names

# loads the mapping tile, immediatetly saving it as a phyloseq OTU table (lighter than a df)
raw_fun_metadata<-sample_data(object = read.table(file = "./data/Amplicon_sequencing/Mapping_file_ITS_Family_experiment.txt", 
                                                  header = TRUE,
                                                  sep = "\t",
                                                  row.names = 1, # first column has row names (ASV names)
                                                  check.names = FALSE)) # prevents "X" to be added to column names, such as X49_ITS,

# build the main ITS phyloseq object by putting all these phyloseq-class objects (otu table, tax table, ref seq, metadta) into a single ps
raw_fun_ps<-merge_phyloseq(raw_fun_otutab,
                           raw_fun_taxtab,
                           raw_fun_refseq,
                           raw_fun_metadata)

# remove old objects to reduce memory use
rm(raw_fun_otutab,raw_fun_taxtab,raw_fun_refseq,raw_fun_metadata)

# change names from ASV to fASV, so they can be distinguished from bacterial ASVs
taxa_names(raw_fun_ps)<-paste("f", taxa_names(raw_fun_ps), sep = "")

# let's check the imported objects. Often errors will arise from typos when filling up the data sheets 
otu_table(raw_fun_ps)[1:10,1:10]
sample_data(raw_fun_ps)[1:10,1:10]
tax_table(raw_fun_ps)[1:10,1:7]
refseq(raw_fun_ps)



# run garbage collection after creating large objects
gc()


```




# 1.2 - remove bad ASVs

here we will remove sequences associated witht he host, background prokatyotes, low abundance ASVs, poorly identified sequences, and other bad ASVs

## 1.2.1 - Remove non-bacterial sequences & filter ASVs
note that the input data (380k taxa) has been pre-filtered to remove any ASVs that occur less than 3 times in the dataset. this was necessary bcause the dada2-associated taxonomy assingment tools could not assing taxonomies to the full dataset (700k taxa) with 1 GB ram on the HPC.

details of host DNA contamination can be found in "./src/Amplicon_sequencing/functions/host_DNA_contamination.r"
  * 35% of 16S reads are plastids
  * 8% of 15S reads are mitochondria


```{r}
# here we see that sequences of only 50bp were still retained by the updated dada2 pipeline
hist(as.data.frame(refseq(raw_bac_ps)@ranges)$width, breaks = 300)
summary(as.data.frame(refseq(raw_bac_ps)@ranges)$width)

# here we get rid of sequences shorter than 380bp. this reduced the total number of taxa by ~4k (from 384089 to 380244 taxa )
raw_bac_ps<-prune_taxa(taxa = as.data.frame(refseq(raw_bac_ps)@ranges)$width>380,
                       x = raw_bac_ps)


# keeps only ASVs identified as bacterial
raw_bac_ps<-subset_taxa(raw_bac_ps, Kingdom=="k__Bacteria") # drops from 380244 to 380019 taxa, barely any difference 

# define library sizes as metadata before filtering
raw_bac_ps@sam_data$library_sizes_prefiltering<-sample_sums(raw_bac_ps)

# removes taxa occuring less than 8 times in the dataser, as VSEARCH standard
#otu_table(raw_bac_ps) <- otu_table(raw_bac_ps)[which (rowSums(otu_table(raw_bac_ps)) > 7),]  #drops from 380K to 168K taxa, from 222M to 221M sequences (843k seqeunces removed)


# filter ASV according samples (min 0 reads, present in 3 samples)
#filter <- phyloseq::genefilter_sample(raw_bac_ps, filterfun_sample(function(x) x > 0), A = 3) 
#raw_bac_ps <- prune_taxa(filter, raw_bac_ps) #drops from 168k to 72k taxa, drops from 221M  to 219M reads











#define plastid, mitochondria and host plant contamination ps objects
Mitochondria_ps<-subset_taxa(raw_bac_ps, Family == "f__Mitochondria" | Family == "Mitochondria")
Plastid_ps<-subset_taxa(raw_bac_ps, Order == "o__Chloroplast" | Order == "Chloroplast") 
host_plant_ps<-merge_phyloseq(Mitochondria_ps,Plastid_ps)

#quick histogram showing plant DNA contamination
hist(sample_sums(host_plant_ps)/sample_sums(raw_bac_ps)*100, breaks = 100)

# define host plant 16S contamination as metadata
raw_bac_ps@sam_data$Mitochondria_reads<-sample_sums(Mitochondria_ps)
raw_bac_ps@sam_data$Plastid_reads<-sample_sums(Plastid_ps)
raw_bac_ps@sam_data$Host_DNA_n_reads<-sample_sums(host_plant_ps)
raw_bac_ps@sam_data$Host_DNA_contamination_pct<-sample_sums(host_plant_ps)/sample_sums(raw_bac_ps)*100

# remove plant host sequences (plastid and mitochndrial DNA) 
raw_bac_ps<-remove_Chloroplast_Mitochondria(raw_bac_ps) # drops from 72k to 69.9k taxa, from 219M to 121M reads (almost 45% of our 16S sequencing effort! that's about 12k Euro!)






# add library sizes as part of metadata
sample_data(raw_bac_ps)$library_size<-sample_sums(raw_bac_ps)

#check librabzy size distribution
hist(sample_data(raw_bac_ps)$library_size)
summary(sample_data(raw_bac_ps)$library_size)
sort(sample_data(raw_bac_ps)$library_size) #one sample has 0 reads

# remove 1 sample with library size = 0  (non-informative; failed PCR/sequencing)
raw_bac_ps<-subset_samples(raw_bac_ps, library_size >0)

# run garbage collection after creating large objects
gc()

```




## 1.2.2 - Remove non-fungal sequences & filter ASVs
details of host DNA contamination can be found in "./src/Amplicon_sequencing/functions/host_DNA_contamination.r"
  * ~0.15% of total reads are plant ITS
  * ~5.77% of total reads are protozoa ITS
  * ~9.92% of ITS reads are unidentified

```{r}


# define library sizes as metadata before filtering
raw_fun_ps@sam_data$library_sizes_prefiltering<-sample_sums(raw_fun_ps)

# removes taxa occuring less than 8 times in the dataser, as VSEARCH standard
#otu_table(raw_fun_ps) <- otu_table(raw_fun_ps)[which (rowSums(otu_table(raw_fun_ps)) > 7),]  #drops from 15.5K to 14.1K taxa, only 6K reads removed

# filter  samples (min 0 reads, present in 3 samples)
#filter <- phyloseq::genefilter_sample(raw_fun_ps, filterfun_sample(function(x) x > 0), A = 3) 

#raw_fun_ps <- prune_taxa(filter, raw_fun_ps) #drops from 14.1k to 3.1k taxa, drops from 206M  to 195M reads
#rm(filter) # we no longer need this string

# removes g__Malassezia, a known human pathogen (saunders et al 2012, https://journals.plos.org/plospathogens/article?id=10.1371/journal.ppat.1002701)
raw_fun_ps<-prune_taxa(!(taxa_names(raw_fun_ps) %in% taxa_names(subset_taxa(raw_fun_ps, Genus=="g__Malassezia"))), raw_fun_ps) 







# take  all plant ITS sequences here to calculate host contamination
raw_PlantITS_ps<-subset_taxa(raw_fun_ps, Kingdom=="k__Viridiplantae") 

# keeps only ASVs identified as fungal
raw_fun_ps<-subset_taxa(raw_fun_ps, Kingdom=="k__Fungi") # drops from 18.5K to 15.5K taxa, from 206M reads to 201M reads

# define ITS host contamination as sample emtadata
raw_fun_ps@sam_data$Host_DNA_contamination_pct<-sample_sums(raw_PlantITS_ps)/sample_sums(raw_fun_ps)*100
raw_fun_ps@sam_data$Host_DNA_n_reads<-sample_sums(raw_PlantITS_ps)

#check final sequence lenght. ITS sequences smaller than 100bp should be discarded
summary(raw_fun_ps@refseq@ranges@width)
sort(raw_fun_ps@refseq@ranges@width) #only one sample has below 100bp, we keep it

# add library sizes as part of metadata
sample_data(raw_fun_ps)$library_size<-sample_sums(raw_fun_ps)
sort(raw_fun_ps@refseq@ranges@width)

#remove unecessary objects and collect garbage
rm(raw_PlantITS_ps)
gc()


```


#1.3  - Decontaminate phyloseq objects
the decontam package will use blank DNA samples to remove possible contaminants. here we use a single custom function to run decontamination, generate plots, and return a clean phyloseq object

We will also check the number of reads in a blank sample (average +SD) so we can compare those blank samples with libraries with low number of reads. samples that cannot be distinguished from a blank (in terms of library size) 


##1.3.1. Decontaminate 16S

```{r}
#Pedro's function to decontaminate didn't work with the 16S data, to big dataset for my laptop. I did it then step by step

# set negatives as TRUE in a  new column 
  sample_data(raw_bac_ps)$is.neg <- sample_data(raw_bac_ps)$Stress  == "Blank" 
  
  # add library size to metadata
  sample_data(raw_bac_ps)$library_size<-sample_sums(raw_bac_ps)
  
  # decontaminate! 
  decontam_output_bac <- isContaminant(raw_bac_ps,
                                   neg="is.neg",
                                   batch = "Speed",
                                   conc = "Amplicon_concentation_ngul",
                                   method = "combined",
                                   threshold= 0.05)  
  
  #save output 
  save(decontam_output_bac, file = "./01_Preprocessing/r_objects/decontam_output_bac.RData")
  
  
  table(decontam_output$contaminant) #260 ASVs contaminants     
  which(decontam_output$contaminant)
  

  ## Check which taxa are contaminants:
  ps_contam <- prune_taxa(decontam_output$contaminant, raw_bac_ps)
  tax_table(ps_contam) #lots of most abundant ASVs
  
  plot_bar(ps_contam, "Phylum", fill="Class")
  plot_bar(ps_contam, "Phylum", fill="Class", facet_grid=~Speed) #slow samples have more contaminants than medium and fast
  #Proteobacteria and actinobacteria are the main contaminants
  
  #plot frequency of contaminant ASVs vs DNA concentration
  plot_frequency(raw_bac_ps, taxa_names(raw_bac_ps)[sample(which(decontam_output_bac$contaminant),20)],
                 conc="Amplicon_concentation_ngul")+
    xlab("DNA concentration (ngul)")
  
  #Which samples have the highest abundance of contaminants?
  p_cont <- plot_bar(ps_contam, fill="Phylum")
  
  #What proportion of our count data were removed as contaminants?
  ps_noncontam <- prune_taxa(!decontam_output_bac$contaminant, raw_bac_ps)
  
  pre <- sum(sample_sums(raw_bac_ps))
  post <- sum(sample_sums(ps_noncontam))
  
  (pre-post) / pre #only 0.0005%

  #clean physeq object by removing blank samples
  unfiltered_unnormalized_bac_ps <- subset_samples(ps_noncontam, Stress != "Blank") #37 blanks removed
  
  
 # add library sizes as part of metadata
sample_data(unfiltered_unnormalized_bac_ps)$library_size_decont<-sample_sums(unfiltered_unnormalized_bac_ps)

# let's look into how many reads our blank had (average+sd) so we can compare those to libraries with very low number of reads
blank_reads_bac<-sort(sample_sums(subset_samples(raw_bac_ps, Stress == "Blank")))
plot(blank_reads_bac) #almost all below 5000 reads


# now let's check the lower end of our library sizes

sort(sample_sums(unfiltered_unnormalized_bac_ps))
plot(sort(sample_sums(unfiltered_unnormalized_bac_ps)))
summary(sample_sums(unfiltered_unnormalized_bac_ps))
# 5 are below 5000 reads, just like the blanks
# 12 samples are above 200.000, we also remove them

unfiltered_unnormalized_bac_ps <- subset_samples(unfiltered_unnormalized_bac_ps, library_size_decont>5000)

unfiltered_unnormalized_bac_ps <- subset_samples(unfiltered_unnormalized_bac_ps, library_size_decont<200000)

plot(sort(sample_sums(unfiltered_unnormalized_bac_ps)))
summary(sample_sums(unfiltered_unnormalized_bac_ps))

#remove those ASVs with now 0 reads 
unfiltered_unnormalized_bac_ps<-prune_taxa(taxa_sums(unfiltered_unnormalized_bac_ps) >0, unfiltered_unnormalized_bac_ps)


# set sp_speed as factor, and then order it properly

sample_data(unfiltered_unnormalized_bac_ps)$Sp_speed <- factor(sample_data(unfiltered_unnormalized_bac_ps)$Sp_speed, 
                                 levels = c("F1","F2","F3","F4","F5","F6","F7","F8","F9","F10","F11","F12",
                                            "M1","M2","M3","M4","M5","M6","M7","M8","M9","M10","M11","M12",
                                            "S1","S2","S3","S4","S5","S6","S7","S8","S9","S10","S11","S12",
                                            "pre_planting"))

# set Sp_full_name as factor, and then order it according phylogeny

sample_data(unfiltered_unnormalized_bac_ps)$Sp_full_name<- factor(sample_data(unfiltered_unnormalized_bac_ps)$Sp_full_name, 
                                 levels = c("Arabidopsis_thaliana",
                                            "Arabidopsis_lyrata",
                                            "Camelina_sativa",
                                            "Turritis_glabra",
                                            "Erysium_cheirantoides",
                                            "Malcomia_maritima",
                                            "Barbarea_vulgaris",
                                            "Lepidium_ruderale",
                                            "Descurainia_sophiades",
                                            "Brassica_nigra",
                                            "Sinapis_arvensis",
                                            "Sinapis_alba",
                                            "Hirschfeldia_incana",
                                            "Brassica_rapa",
                                            "Brassica_oleraceae_Rivera_Fast",
                                            "Brassica_oleraceae_Rivera_Medium",
                                            "Brassica_oleraceae_Rivera_Slow",
                                            "Brassica_oleraceae_Kimmeridge",
                                            "Diplotaxis_tenuifolia",
                                            "Sysimbrium_irio",
                                            "Isatis_tinctoria",
                                            "Cochlearia_officinalis",
                                            "Lobularia_maritima",
                                            "Iberis_amara",
                                            "Berteroa_incana",
                                            "Euclydium_syriacum",
                                            "Solanum_lycopersicum",
                                            "Nicotiana_benthamiana",
                                            "Pisum_sativum",
                                            "Medicago_truncatula",
                                            "Zea_mays",
                                            "Brachypodium_distachyon",
                                            "Cucumis_sativus",
                                            "Soil_Fast",
                                            "Soil_Medium",
                                            "Soil_Slow",
                                            "Soil_pre_planting"))
                                           


# set Sp_ab_name as factor, and then order it according phylogeny

sample_data(unfiltered_unnormalized_bac_ps)$Sp_abb_name<- factor(sample_data(unfiltered_unnormalized_bac_ps)$Sp_abb_name, 
                                 levels = c("At",
                                            "Al",
                                            "Cas",
                                            "Tg",
                                            "Ec",
                                            "Mm",
                                            "Bv",
                                            "Lr",
                                            "Ds",
                                            "Bn",
                                            "Sar",
                                            "Sal",
                                            "Hi",
                                            "Br",
                                            "Bo_F",
                                            "Bo_M",
                                            "Bo_S",
                                            "BoK",
                                            "Dt",
                                            "Si",
                                            "It",
                                            "Co",
                                            "Lm",
                                            "Ia",
                                            "Bi",
                                            "Es",
                                            "Sl",
                                            "Nb",
                                            "Ps",
                                            "Mt",
                                            "Zm",
                                            "Bd",
                                            "Cus",
                                            "S_F",
                                            "S_M",
                                            "S_S",
                                            "S_pre"))
                                       
#let's save these unnormalized objects as RData so they can be loaded in other scripts
save(unfiltered_unnormalized_bac_ps, file = "./data/phyloseq_objects/unfiltered_unnormalized_bac_ps.RData")
 
```


##1.3.2. Decontaminate ITS
```{r}

#decontaminate!
#we the following function we remove contaminants and after we also remove blank samples
unfiltered_unnormalized_bac_l<-decontaminate_and_plot(raw_bac_ps)
unfiltered_unnormalized_fun_l<-decontaminate_and_plot(raw_fun_ps)


# let's look into how many reads our blank had (average+sd) so we can compare those to libraries with very low number of reads
blank_reads_bac<-sort(sample_sums(subset_samples(raw_bac_ps, Stress == "Blank")))
plot(blank_reads_bac)


blank_reads_fun<-sort(sample_sums(subset_samples(raw_fun_ps, Stress == "Blank")))
plot(blank_reads_fun)





#clean up enviroment to free memory
#rm(raw_bac_ps, raw_fun_ps)

# after checking the decontamination plots and reports, we can remove them and focus on the phyloseq oject
unnormalized_bac_ps<-unlist(unnormalized_bac_ps[[1]])

unfiltered_unnormalized_fun_ps<-unlist(unfiltered_unnormalized_fun_l[[1]])


# add library sizes as part of metadata
sample_data(unfiltered_unnormalized_fun_ps)$library_size_decont<-sample_sums(unfiltered_unnormalized_fun_ps)


# now let's save them as list, add names to this list, and remove unlisted ps objects to release memory
#unnormalized_BacFun_ps_l<-list(unnormalized_bac_ps,unnormalized_fun_ps)
#names(unnormalized_BacFun_ps_l)<-c("Bac","Fun")
#rm(unnormalized_bac_ps,unnormalized_fun_ps)


# now let's check the lower end of our library sizes
sort(sample_sums(unnormalized_BacFun_ps_l$Bac))
plot(sort(sample_sums(unnormalized_BacFun_ps_l$Bac)))
# 5 samples from 440-1362 reads, and 16S blank averages + 3xSD = 3752 reads. the next sample has 6200 reads. let's remove this 5 samples even from CSS and unnormalized as they essentially failed amplification

sort(sample_sums(unfiltered_unnormalized_fun_ps))
plot(sort(sample_sums(unfiltered_unnormalized_fun_ps)))
summary(sample_sums(unfiltered_unnormalized_fun_ps))
# 52 samples were from 504-3227 reads, we remove them (all below 5000)
# 1 sample is above 230.000, we also remove it

unfiltered_unnormalized_fun_ps <- subset_samples(unfiltered_unnormalized_fun_ps, library_size_decont>5000)

unfiltered_unnormalized_fun_ps <- subset_samples(unfiltered_unnormalized_fun_ps, library_size_decont<230000)

plot(sort(sample_sums(unfiltered_unnormalized_fun_ps)))
summary(sample_sums(unfiltered_unnormalized_fun_ps))

#remove those ASVs with now 0 reads 
prune_taxa(taxa_sums(unfiltered_unnormalized_fun_ps) >0, unfiltered_unnormalized_fun_ps)

#and ITS blank averages + 3xSD = 1598 reads, so we will remove them. there are other 4 samples between 1678 and 2151 reads before a jump to 5751 reads, so we will also remove them (they would likely be outliers in any test)

sample_data(unfiltered_unnormalized_fun_ps)




# set sp_speed as factor, and then order it properly
#unnormalized_BacFun_ps_l<- lapply(unnormalized_BacFun_ps_l, function (x){
#sample_data(x)$Sp_speed<- factor(sample_data(x)$Sp_speed, 
 #                                levels = c("F1","F2","F3","F4","F5","F6","F7","F8","F9","F10","F11","F12",
                                            #"M1","M2","M3","M4","M5","M6","M7","M8","M9","M10","M11","M12",
                                            #"S1","S2","S3","S4","S5","S6","S7","S8","S9","S10","S11","S12",
 #                                           "pre_planting"))
  #                                          return(x)
   #                                                                     })

# set sp_speed as factor, and then order it properly

sample_data(unfiltered_unnormalized_fun_ps)$Sp_speed <- factor(sample_data(unfiltered_unnormalized_fun_ps)$Sp_speed, 
                                 levels = c("F1","F2","F3","F4","F5","F6","F7","F8","F9","F10","F11","F12",
                                            "M1","M2","M3","M4","M5","M6","M7","M8","M9","M10","M11","M12",
                                            "S1","S2","S3","S4","S5","S6","S7","S8","S9","S10","S11","S12",
                                            "pre_planting"))

# set Sp_full_name as factor, and then order it according phylogeny

sample_data(unfiltered_unnormalized_fun_ps)$Sp_full_name<- factor(sample_data(unfiltered_unnormalized_fun_ps)$Sp_full_name, 
                                 levels = c("Arabidopsis_thaliana",
                                            "Arabidopsis_lyrata",
                                            "Camelina_sativa",
                                            "Turritis_glabra",
                                            "Erysium_cheirantoides",
                                            "Malcomia_maritima",
                                            "Barbarea_vulgaris",
                                            "Lepidium_ruderale",
                                            "Descurainia_sophiades",
                                            "Brassica_nigra",
                                            "Sinapis_arvensis",
                                            "Sinapis_alba",
                                            "Hirschfeldia_incana",
                                            "Brassica_rapa",
                                            "Brassica_oleraceae_Rivera_Fast",
                                            "Brassica_oleraceae_Rivera_Medium",
                                            "Brassica_oleraceae_Rivera_Slow",
                                            "Brassica_oleraceae_Kimmeridge",
                                            "Diplotaxis_tenuifolia",
                                            "Sysimbrium_irio",
                                            "Isatis_tinctoria",
                                            "Cochlearia_officinalis",
                                            "Lobularia_maritima",
                                            "Iberis_amara",
                                            "Berteroa_incana",
                                            "Euclydium_syriacum",
                                            "Solanum_lycopersicum",
                                            "Nicotiana_benthamiana",
                                            "Pisum_sativum",
                                            "Medicago_truncatula",
                                            "Zea_mays",
                                            "Brachypodium_distachyon",
                                            "Cucumis_sativus",
                                            "Soil_Fast",
                                            "Soil_Medium",
                                            "Soil_Slow",
                                            "Soil_pre_planting"))
                                           


# set Sp_ab_name as factor, and then order it according phylogeny

sample_data(unfiltered_unnormalized_fun_ps)$Sp_abb_name<- factor(sample_data(unfiltered_unnormalized_fun_ps)$Sp_abb_name, 
                                 levels = c("At",
                                            "Al",
                                            "Cas",
                                            "Tg",
                                            "Ec",
                                            "Mm",
                                            "Bv",
                                            "Lr",
                                            "Ds",
                                            "Bn",
                                            "Sar",
                                            "Sal",
                                            "Hi",
                                            "Br",
                                            "Bo_F",
                                            "Bo_M",
                                            "Bo_S",
                                            "BoK",
                                            "Dt",
                                            "Si",
                                            "It",
                                            "Co",
                                            "Lm",
                                            "Ia",
                                            "Bi",
                                            "Es",
                                            "Sl",
                                            "Nb",
                                            "Ps",
                                            "Mt",
                                            "Zm",
                                            "Bd",
                                            "Cus",
                                            "S_F",
                                            "S_M",
                                            "S_S",
                                            "S_pre"))
                                       
#let's save these unnormalized objects as RData so they can be loaded in other scripts
save(unfiltered_unnormalized_fun_ps, file = "./data/phyloseq_objects/unfiltered_unnormalized_fun_ps.RData")

load("./data/phyloseq_objects/unfiltered_unnormalized_fun_ps.RData")




```






# 1.4-  Plot library sizes per treatment

```{r}

# let's check the number of reads in a couple histograms
hist(sample_sums(unfiltered_unnormalized_bac_ps), breaks = 200)
summary(sample_sums(unfiltered_unnormalized_bac_ps))

hist(sample_sums(unfiltered_unnormalized_fun_ps), breaks = 200)
summary(sample_sums(unfiltered_unnormalized_fun_ps))


#check plot for some minimum library sizes on rhizoplane samples

  
ggplot(data = sample_data(subset_samples(unfiltered_unnormalized_bac_ps, sample_type=="rhizoplane")), mapping = aes(x=library_size, y=Sp_abb_name, color = Stress, fill = Stress))+
    geom_jitter()+
    geom_boxplot()+
    scale_y_discrete(limits=rev)+
    geom_hline(yintercept=30000, color="green")+
    geom_hline(yintercept=25000, color="yellow")+
    geom_hline(yintercept=20000, color="blue")+
    geom_hline(yintercept=15000, color="red")+
    labs(x = "Library size",
         y = "Plant species and speed")+
theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))

ggplot(data = sample_data(subset_samples(unfiltered_unnormalized_fun_ps, sample_type=="rhizoplane")), mapping = aes(x=library_size, y=Sp_abb_name, color = Stress, fill = Stress))+
    geom_jitter()+
    geom_boxplot()+
    scale_y_discrete(limits=rev)+
    geom_hline(yintercept=30000, color="green")+
    geom_hline(yintercept=25000, color="yellow")+
    geom_hline(yintercept=20000, color="blue")+
    geom_hline(yintercept=15000, color="red")+
    labs(x = "Library size",
         y = "Plant species and speed")+
theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))


#check plot for some minimum library sizes on top soil samples

  ggplot(data = sample_data(subset_samples(unfiltered_unnormalized_bac_ps, sample_type!="rhizoplane")), mapping = aes(x=Sp_speed, y=library_size, color = Stress, fill = Stress))+
    geom_jitter()+
    geom_boxplot()+
    geom_hline(yintercept=30000, color="green")+
    geom_hline(yintercept=25000, color="yellow")+
    geom_hline(yintercept=20000, color="blue")+
    geom_hline(yintercept=15000, color="red")+
theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))

  ggplot(data = sample_data(subset_samples(unfiltered_unnormalized_fun_ps, sample_type!="rhizoplane")), mapping = aes(x=Sp_speed, y=library_size, color = Stress, fill = Stress))+
    geom_jitter()+
    geom_boxplot()+
    geom_hline(yintercept=30000, color="green")+
    geom_hline(yintercept=25000, color="yellow")+
    geom_hline(yintercept=20000, color="blue")+
    geom_hline(yintercept=15000, color="red")+
theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))  

#check bacterial 16S plant host contamination

ggplot(data = sample_data(subset_samples(unfiltered_unnormalized_bac_ps, sample_type=="rhizoplane")), mapping = aes(x=Host_DNA_contamination_pct, y=Sp_abb_name, color = Stress, fill = Stress))+
    geom_jitter()+
    geom_boxplot()+
    scale_y_discrete(limits=rev)+
    geom_hline(yintercept=30000, color="green")+
    geom_hline(yintercept=25000, color="yellow")+
    geom_hline(yintercept=20000, color="blue")+
    geom_hline(yintercept=15000, color="red")+
    labs(x = "Plant host DNA contamination",
         y = "Plant species")+
    theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))

ggplot(data = sample_data(subset_samples(unfiltered_unnormalized_fun_ps, sample_type=="rhizoplane")), mapping = aes(x=Host_DNA_contamination_pct, y=Sp_abb_name, color = Stress, fill = Stress))+
    geom_jitter()+
    geom_boxplot()+
    scale_y_discrete(limits=rev)+
    geom_hline(yintercept=30000, color="green")+
    geom_hline(yintercept=25000, color="yellow")+
    geom_hline(yintercept=20000, color="blue")+
    geom_hline(yintercept=15000, color="red")+
    labs(x = "Plant host DNA contamination",
         y = "Plant species")+
    theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))



#check percentage of NA in taxonomy
lapply(unnormalized_BacFun_ps_l, function (x) check_n_taxa_NA_percentage(x, 100))
lapply(unnormalized_BacFun_ps_l, function (x) check_n_taxa_NA_percentage(x, 1000))
lapply(unnormalized_BacFun_ps_l, function (x) check_n_taxa_NA_percentage(x, 10000))

lapply(unnormalized_BacFun_ps_l, function (x) subset_taxa(x, Class == "c__uncultured" | Class == "c__unidentified"))
lapply(unnormalized_BacFun_ps_l, function (x) subset_taxa(x, Family == "f__uncultured" | Family == "f__unidentified"))
lapply(unnormalized_BacFun_ps_l, function (x) subset_taxa(x, Genus == "g__uncultured" | Genus == "g__unidentified"))


check_n_taxa_NA_percentage(unfiltered_unnormalized_bac_ps, 100) #23
check_n_taxa_NA_percentage(unfiltered_unnormalized_bac_ps, 1000) #13.3
check_n_taxa_NA_percentage(unfiltered_unnormalized_bac_ps, 10000) #9.42

subset_taxa(unfiltered_unnormalized_bac_ps, Class == "c__uncultured" | Class == "c__unidentified") #1872
subset_taxa(unfiltered_unnormalized_bac_ps, Family == "f__uncultured" | Family == "f__unidentified") #2930
subset_taxa(unfiltered_unnormalized_bac_ps, Genus == "g__uncultured" | Genus == "g__unidentified") #69074


check_n_taxa_NA_percentage(unfiltered_unnormalized_fun_ps, 100) #15
check_n_taxa_NA_percentage(unfiltered_unnormalized_fun_ps, 1000) #22
check_n_taxa_NA_percentage(unfiltered_unnormalized_fun_ps, 10000) #21.77

subset_taxa(unfiltered_unnormalized_fun_ps, Class == "c__uncultured" | Class == "c__unidentified") #286
subset_taxa(unfiltered_unnormalized_fun_ps, Family == "f__uncultured" | Family == "f__unidentified") #754
subset_taxa(unfiltered_unnormalized_fun_ps, Genus == "g__uncultured" | Genus == "g__unidentified") #1161

```



# 1.5 export filtered unnormalized ps object
```{r}

#let's prepare 2 different  versions: one without only with rhizoplane and other only with top/bulksoils 



unfiltered_unnormalized_bac_rhizo_ps <- subset_samples(unfiltered_unnormalized_bac_ps, 
                  sample_type == "rhizoplane" & 
                  Stress != "Soil" )

unfiltered_unnormalized_fun_rhizo_ps <- subset_samples(unfiltered_unnormalized_fun_ps, 
                  sample_type == "rhizoplane" & 
                  Stress != "Soil" )

unfiltered_unnormalized_bulksoil_bac_ps <- subset_samples(unfiltered_unnormalized_bac_ps, 
                  sample_type != "rhizoplane" & 
                  Stress == "Soil" )

unfiltered_unnormalized_bulksoil_fun_ps <- subset_samples(unfiltered_unnormalized_fun_ps, 
                  sample_type != "rhizoplane" & 
                  Stress == "Soil" )



# remove non-informative rows

unfiltered_unnormalized_bac_rhizo_ps <- prune_taxa(taxa_sums(otu_table(unfiltered_unnormalized_bac_rhizo_ps)) > 0, unfiltered_unnormalized_bac_rhizo_ps)

unfiltered_unnormalized_bulksoil_bac_ps <- prune_taxa(taxa_sums(otu_table(unfiltered_unnormalized_bulksoil_bac_ps)) > 0, unfiltered_unnormalized_bulksoil_bac_ps)

unfiltered_unnormalized_fun_rhizo_ps <- prune_taxa(taxa_sums(otu_table(unfiltered_unnormalized_fun_rhizo_ps)) > 0, unfiltered_unnormalized_fun_rhizo_ps)

unfiltered_unnormalized_bulksoil_fun_ps <- prune_taxa(taxa_sums(otu_table(unfiltered_unnormalized_bulksoil_fun_ps)) > 0, unfiltered_unnormalized_bulksoil_fun_ps)
  


# now we save the different versions
save(unfiltered_unnormalized_bac_rhizo_ps, 
     file = "./data/phyloseq_objects/unfiltered_unnormalized_bac_rhizo_ps.RData")
save(unfiltered_unnormalized_bulksoil_bac_ps, 
     file = "./data/phyloseq_objects/unfiltered_unnormalized_bulksoil_bac_ps.RData")
save(unfiltered_unnormalized_fun_rhizo_ps, 
     file = "./data/phyloseq_objects/unfiltered_unnormalized_fun_rhizo_ps.RData")
save(unfiltered_unnormalized_bulksoil_fun_ps, 
     file = "./data/phyloseq_objects/unfiltered_unnormalized_bulksoil_fun_ps.RData")





```

