---
title: "4_neutral_models_AllSpecies"
author: "Marcela Aragon"
date: "3/29/2022"
output: 
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

# 3.1 - load libraries

```{r}
# this will load all key libraries for the whole project
source("./renv/0_load_libraries.R")

# increases memory limit used by R
memory.limit(size = 350000)

```

# 3.2 - load data

```{R}
# load data 
load("./Data/phyloseq_objects/rarefied_BacFun_rhizoplane_HighFilt_ps.RData")

# remove all species outside the MY trees
rarefied_BacFun_rhizoplane_HighFilt_ps<-lapply(rarefied_BacFun_rhizoplane_HighFilt_ps, function(x)  subset_samples(x, MY_tree_presence == "inside_MY_tree"))

```

# 3.3 - split ps objects into suitable lists for the NM pipeline

```{R}
#create empty list
neutral_pairwise_ps<-list()

# make separate objects wint onyl Meja-control or SA-control
neutral_pairwise_ps[1:2]<-lapply(rarefied_BacFun_rhizoplane_HighFilt_ps, function(x)  subset_samples(x, Stress != "SA"))
neutral_pairwise_ps[3:4]<-lapply(rarefied_BacFun_rhizoplane_HighFilt_ps, function(x)  subset_samples(x, Stress != "MeJA"))

#adjsut list anmes
names(neutral_pairwise_ps)<-c("Bac_MeJA_Control", "Fun_MeJA_Control", "Bac_SA_Control", "Fun_SA_Control")


# split by 22 species
BacFun_22sp_l_l <- lapply(neutral_pairwise_ps, function(x) metagMisc::phyloseq_sep_variable(x, variable = "Sp_abb_name")) 

```


# 3.4 - create 66 models(sp*stress), using all samples from a species as the pool
```{r}


# define a function to split a species in 2 stresses, retuning 4 models per species
split_by_stress_and_fit_sncm<- function(phyloseq){
  
   input_for_model_split_by_stress<- metagMisc::phyloseq_sep_variable(phyloseq, variable = "Stress")
   
   neutral_model_output<-lapply(input_for_model_split_by_stress, function(x)
                          tyRa::fit_sncm(spp = t(otu_table(x)), 
                                         pool = t(otu_table(phyloseq)), 
                                         taxon = data.frame(tax_table(x))))
               
   return(neutral_model_output)

}



#run the custom function for all stresses of all species in both Bac and Fun
neutral_Bacfun_22sp_2pairwise_l_l_l<-lapply(BacFun_22sp_l_l, function(x)
  lapply(x, function(y)
    split_by_stress_and_fit_sncm(y)))



# plot a few neutral models to check
plot_sncm_fit(neutral_Bacfun_22sp_2pairwise_l_l_l$Bac_MeJA_Control$At$MeJA)
plot_sncm_fit(neutral_Bacfun_22sp_2pairwise_l_l_l$Bac_MeJA_Control$At$Control)



```



## 3.4.1 - extract R2 fit and m parameters 


```{r}




# fit stats of all neutral models, keeps only M and R square of that list
fitstats_l_l<-lapply(neutral_Bacfun_22sp_2pairwise_l_l_l, function (y)
  lapply(y, function(x){
  
Control_fitstats <- data.frame(x[[1]]$fitstats[c(1,4, 11:12)])
Stress_fitstats <-   data.frame(x[[2]]$fitstats[c(1,4, 11:12)])

output<-rbind(Control_fitstats,
              Stress_fitstats)

output$Stress<-c("Control", names(x)[2])

return(output)  
  
}))

#put all species in a list
fitstats_df_l<-lapply(fitstats_l_l, function(x)
  as.data.frame(do.call(rbind, x)))

#adjust rownames as species names
fitstats_df_l<-lapply(fitstats_df_l, function(y){
df_out<-rownames_to_column(.data = y,var = "Sp_abb_name")
df_out$Sp_abb_name<-gsub(pattern = "\\..", replacement = "", x= df_out$Sp_abb_name)
return(df_out)
})

# put bac and fun in the same DF
fitstats_df_l$Bac_MeJA_Control$Kingdom<-"Bac"
fitstats_df_l$Bac_SA_Control$Kingdom<-"Bac"
fitstats_df_l$Fun_MeJA_Control$Kingdom<-"Fun"
fitstats_df_l$Fun_SA_Control$Kingdom<-"Fun"
fitstats_df<-rbind(fitstats_df_l[[1]],
                   fitstats_df_l[[2]],
                   fitstats_df_l[[3]],
                   fitstats_df_l[[4]])

# make the Sp_abb_name a factor that fallows the phylogeny level
fitstats_df$Sp_abb_name<-factor(fitstats_df$Sp_abb_name, levels = levels(rarefied_BacFun_rhizoplane_HighFilt_ps$Bac@sam_data$Sp_abb_name))



#this output df has key neutral model metrics
fitstats_df




```

## 3.4.2 -  plot the R and m values

```{r}

#plot R2 and m, more details to follow (or to be skipped)
plot_r_neutral<-ggplot(fitstats_df, aes(x= Rsqr, y= Sp_abb_name, color = Kingdom) ) + 
  geom_point(size=2)+
  facet_wrap(~Stress)+
  scale_y_discrete(limits=rev)+
  theme_bw()+
  theme(text= element_text(size=10))

#plot R2 and m, more details to follow (or to be skipped)
plot_m_neutral<-ggplot(fitstats_df, aes(x= m, y= Sp_abb_name, color = Kingdom) ) + 
  geom_point(size=2)+
  facet_wrap(~Stress)+
  scale_y_discrete(limits=rev)+
  theme_bw()+
  theme(text= element_text(size=10))

#save and as a single plot
plot_r_m_neutral_2pairwise<-ggarrange(plot_r_neutral, plot_m_neutral, 
                                      labels = c("NM_r2_2pairwise", "NM_m_2pairwise"), 
                                      common.legend = TRUE,
                                      vjust = 0)

#export plot
ggsave(plot_r_m_neutral_2pairwise, 
       filename = "./04_Neutral_models/plots_PB/plot_r_m_neutral_2pairwise.pdf",
       height = 20,
       width = 20,
       units = "cm")


```

## 3.4.3 -  extract ASVs accuring above and below neutrality
```{r}

# Extracts above-neutral bASVs ad fASVs
Above_neutral_2pairwise_bASVs_fASVs<-lapply(neutral_Bacfun_22sp_2pairwise_l_l_l, function (z)
  lapply(z, function(y)
   lapply(y, function(x)
    filter(x$predictions, x$predictions$fit_class == "Above prediction")%>%rownames()
   )
  )
)



# Extracts below-neutral bASVs ad fASVs
Below_neutral_2pairwise_bASVs_fASVs<-lapply(neutral_Bacfun_22sp_2pairwise_l_l_l, function (z)
  lapply(z, function(y)
   lapply(y, function(x)
    filter(x$predictions, x$predictions$fit_class == "Below prediction")%>%rownames()
   )
  )
)


# check bASVs and fASVs dsitributions and structure
unlist(Above_neutral_2pairwise_bASVs_fASVs$Bac_MeJA_Control)%>%table%>%hist(breaks = 50)
unlist(Above_neutral_2pairwise_bASVs_fASVs$Bac_MeJA_Control)%>%table%>%sort(decreasing = TRUE)

unlist(Below_neutral_2pairwise_bASVs_fASVs$Bac_MeJA_Control)%>%table%>%hist(breaks = 50)
unlist(Below_neutral_2pairwise_bASVs_fASVs$Bac_MeJA_Control)%>%table%>%sort(decreasing = TRUE)



```
## 3.4.4 - export lists of above-below per species and stress for bacteria and fungi

```{r}
#save vecor of above-neutral ASVs externally
save(Above_neutral_2pairwise_bASVs_fASVs, file = "./04_Neutral_models/r_objects/Above_neutral_2pairwise_bASVs_fASVs.RData")

#save vecor of Below-neutral ASVs externally
save(Below_neutral_2pairwise_bASVs_fASVs, file = "./04_Neutral_models/r_objects/Below_neutral_2pairwise_bASVs_fASVs.RData")

```




# 3.5 -  extract ASVs accuring above neutrality in Stress but not above enutrality in control; and those occuring below neutrality in stress but not below neutrality in control
This better reflects the differences between control and stress, and also better reflects teh RF and DA methods we are using

 AGREED WITH MELISSA THAT THIS VERISON IS BETTER

```{r}

# remove ASVs that are above neutral in both control and stress (they do not differentiate both conditions)
differing_ASVs_above_l_l<-
lapply(Above_neutral_2pairwise_bASVs_fASVs, function(x)
    lapply(x, function(z){
      
    #join control and stress
    full_list<-union(z[[1]], z[[2]]) 
    
    #define control adn stress overlap
    to_remove<-intersect(z[[1]], z[[2]])
    
    # remove overlap from control and stress
    to_keep<-full_list[!(full_list %in% to_remove)]
    
    return(to_keep)
      
}))

# remove ASVs that are below neutral in both control and stress (they do not differentiate both conditions)
differing_ASVs_below_l_l<-
lapply(Below_neutral_2pairwise_bASVs_fASVs, function(x)
    lapply(x, function(z){
      
    #join control and stress
    full_list<-union(z[[1]], z[[2]]) 
    
    #define control adn stress overlap
    to_remove<-intersect(z[[1]], z[[2]])
    
    # remove overlap from control and stress
    to_keep<-full_list[!(full_list %in% to_remove)]
    
    return(to_keep)
      
}))

# join these into a flat list of 22 sp (containng above and below)
differing_ASVs_above_l_22sp<-lapply(differing_ASVs_above_l_l, function(x) unlist(x)%>%unique)
differing_ASVs_below_l_22sp<-lapply(differing_ASVs_below_l_l, function(x) unlist(x)%>%unique)

differing_ASVs_above_below_l_22sp<- mapply(function (x,y)
  c(x,y)%>%unique,
x = differing_ASVs_above_l_22sp,
y = differing_ASVs_below_l_22sp,
SIMPLIFY = FALSE)

# adjust order to better reflect next steps in the golden ASV joining scripts
differing_ASVs_above_below_l_22sp<-differing_ASVs_above_below_l_22sp[c(1,3,2,4)]

# save externally as .RData
save(differing_ASVs_above_below_l_22sp, file = "./04_Neutral_models/r_objects/differing_ASVs_above_below_l_22sp.RData")



```


# 3.6 - cluster the species according microbiome types
```{r}
# define species present in each microbiome type
type_LI <- c("At", "Al", "Cas", "Tg", "Ec", "Mm", "Bv",
              "Lr", "Ds", "Bi",   "Co", "Lm", "Ia", "Es")

type_LII <- c("Br", "Bo_M","Bn", "Sar", "Hi", "Si", "It","Dt")





###################################
##### warning, bad code start #####
###################################

# looping/lapply failed after trying 10 minutes of trying, I will do this with hard code

# microbiome type_LI, above
differing_bASVs_above_MeJA_Control_14sp<-
  differing_ASVs_above_l_l[[1]][names(differing_ASVs_above_l_l[[1]]) %in% type_LI ]

differing_fASVs_above_MeJA_Control_14sp<-
  differing_ASVs_above_l_l[[2]][names(differing_ASVs_above_l_l[[2]]) %in% type_LI ]

differing_bASVs_above_SA_Control_14sp<-
  differing_ASVs_above_l_l[[3]][names(differing_ASVs_above_l_l[[3]]) %in% type_LI ]

differing_fASVs_above_SA_Control_14sp<-
  differing_ASVs_above_l_l[[4]][names(differing_ASVs_above_l_l[[4]]) %in% type_LI ]


# microbiome type_LI, below
differing_bASVs_below_MeJA_Control_14sp<-
  differing_ASVs_below_l_l[[1]][names(differing_ASVs_below_l_l[[1]]) %in% type_LI ]

differing_fASVs_below_MeJA_Control_14sp<-
  differing_ASVs_below_l_l[[2]][names(differing_ASVs_below_l_l[[2]]) %in% type_LI ]

differing_bASVs_below_SA_Control_14sp<-
  differing_ASVs_below_l_l[[3]][names(differing_ASVs_below_l_l[[3]]) %in% type_LI ]

differing_fASVs_below_SA_Control_14sp<-
  differing_ASVs_below_l_l[[4]][names(differing_ASVs_below_l_l[[4]]) %in% type_LI ]




# microbiome type_LII, above
differing_bASVs_above_MeJA_Control_8sp<-
  differing_ASVs_above_l_l[[1]][names(differing_ASVs_above_l_l[[1]]) %in% type_LII ]

differing_fASVs_above_MeJA_Control_8sp<-
  differing_ASVs_above_l_l[[2]][names(differing_ASVs_above_l_l[[2]]) %in% type_LII ]

differing_bASVs_above_SA_Control_8sp<-
  differing_ASVs_above_l_l[[3]][names(differing_ASVs_above_l_l[[3]]) %in% type_LII ]

differing_fASVs_above_SA_Control_8sp<-
  differing_ASVs_above_l_l[[4]][names(differing_ASVs_above_l_l[[4]]) %in% type_LII ]


# microbiome type_LII, below
differing_bASVs_below_MeJA_Control_8sp<-
  differing_ASVs_below_l_l[[1]][names(differing_ASVs_below_l_l[[1]]) %in% type_LII ]

differing_fASVs_below_MeJA_Control_8sp<-
  differing_ASVs_below_l_l[[2]][names(differing_ASVs_below_l_l[[2]]) %in% type_LII ]

differing_bASVs_below_SA_Control_8sp<-
  differing_ASVs_below_l_l[[3]][names(differing_ASVs_below_l_l[[3]]) %in% type_LII ]

differing_fASVs_below_SA_Control_8sp<-
  differing_ASVs_below_l_l[[4]][names(differing_ASVs_below_l_l[[4]]) %in% type_LII ]


# join together above and blow ASVs, uninting all species of the same microbiome type
#type_LI
NM_bASVs_MeJA_Control_14sp<-
union(as.character(unlist(differing_bASVs_above_MeJA_Control_14sp)), 
      as.character(unlist(differing_bASVs_below_MeJA_Control_14sp)))%>%unique()

NM_fASVs_MeJA_Control_14sp<-
union(as.character(unlist(differing_fASVs_above_MeJA_Control_14sp)), 
      as.character(unlist(differing_fASVs_below_MeJA_Control_14sp)))%>%unique()

NM_bASVs_SA_Control_14sp<-
union(as.character(unlist(differing_bASVs_above_SA_Control_14sp)), 
      as.character(unlist(differing_bASVs_below_SA_Control_14sp)))%>%unique()

NM_fASVs_SA_Control_14sp<-
union(as.character(unlist(differing_fASVs_above_SA_Control_14sp)), 
      as.character(unlist(differing_fASVs_below_SA_Control_14sp)))%>%unique()



#type_LII
NM_bASVs_MeJA_Control_8sp<-
union(as.character(unlist(differing_bASVs_above_MeJA_Control_8sp)), 
      as.character(unlist(differing_bASVs_below_MeJA_Control_8sp)))%>%unique()

NM_fASVs_MeJA_Control_8sp<-
union(as.character(unlist(differing_fASVs_above_MeJA_Control_8sp)), 
      as.character(unlist(differing_fASVs_below_MeJA_Control_8sp)))%>%unique()

NM_bASVs_SA_Control_8sp<-
union(as.character(unlist(differing_bASVs_above_SA_Control_8sp)), 
      as.character(unlist(differing_bASVs_below_SA_Control_8sp)))%>%unique()

NM_fASVs_SA_Control_8sp<-
union(as.character(unlist(differing_fASVs_above_SA_Control_8sp)), 
      as.character(unlist(differing_fASVs_below_SA_Control_8sp)))%>%unique()



# join output into a single list
# here we change teh order of bac and fun; L_I and L_II
NM_BacFun_MeJASA_differentiating_typeLI<-list(
"NM_bASVs_MeJA_Control_14sp" = NM_bASVs_MeJA_Control_14sp,
"NM_bASVs_SA_Control_14sp" = NM_bASVs_SA_Control_14sp,
"NM_fASVs_MeJA_Control_14sp" = NM_fASVs_MeJA_Control_14sp,
"NM_fASVs_SA_Control_14sp" = NM_fASVs_SA_Control_14sp)




NM_BacFun_MeJASA_differentiating_typeLII<-list(
"NM_bASVs_MeJA_Control_8sp" = NM_bASVs_MeJA_Control_8sp,
"NM_bASVs_MeJA_Control_8sp" = NM_bASVs_SA_Control_8sp,
"NM_fASVs_MeJA_Control_8sp" = NM_fASVs_MeJA_Control_8sp,
"NM_fASVs_SA_Control_8sp" = NM_fASVs_SA_Control_8sp)

###################################
##### bad code ends - sorry!! #####
###################################



```

#3.7 - export NM list of 14sp in Li and 8 spi in LII
```{r}
#this is our output: Non-neutral ASVs, excluding ASVs that are above selected in both stress and treatment, and clsutered by microbiom type, and separaed by fungi and bacteria
NM_BacFun_MeJASA_differentiating_typeLI
NM_BacFun_MeJASA_differentiating_typeLII


#save them externally as Li and LII NM ASVs
save(NM_BacFun_MeJASA_differentiating_typeLI, file = "./04_Neutral_models/r_objects/NM_BacFun_MeJASA_differentiating_typeLI.RData")

save(NM_BacFun_MeJASA_differentiating_typeLII, file = "./04_Neutral_models/r_objects/NM_BacFun_MeJASA_differentiating_typeLII.RData")


```

