---
title: "3_Alpha_Diversity_CLEAN"
author: "Zulema Carracedo Lorenzo"
date: "2023-08-02"
output: html_document
editor_options: 
  chunk_output_type: console
---

---
This script has as a goal to analyse the alpha diversity of the Brassica species included in the million year tree.
```{r results="hide"}

#Load libraries

library(phyloseq)
library(microbiome)
library(vegan)
library(ggplot2)
library(tidyr)
library(dplyr)
library(ggpubr)
library(tibble)
library(car)
library(purrr)
library(ggbreak) #for break barplot


# increases memory limit used by R
memory.limit(size = 350000)

# this will load the unfiltered unnormalized phyloseq objects for 16S and ITS 

load("./Data/phyloseq_objects/unfiltered_unnormalized_bac_rhizo_ps.RData") 
load("./Data/phyloseq_objects/unfiltered_unnormalized_fun_rhizo_ps.RData")

#load needed functions for this script

source("./renv/functions/lapply_lapply.R")


# we will select only the Brassica samples from the million year tree
#define species present in the MY tree
MY_tree <- c("At", "Al", "Cas", "Tg", "Ec", "Mm", "Bv",
                "Lr", "Ds", "Bi", "Bn", "Sar", "Hi", "Br", "Bo_M", 
                "Dt", "Si", "It", "Co", "Lm", "Ia", "Es")

#define function to update ps objects
add_MY_metadata<-function(ps_obj){
ps_obj@sam_data$MY_tree_presence<-ifelse(test = ps_obj@sam_data$Sp_abb_name %in% MY_tree,
                                       yes = "inside_MY_tree",
                                       no = "outside_MY_tree")
return(ps_obj)
}

unfiltered_unnormalized_bac_rhizo_ps <- add_MY_metadata(unfiltered_unnormalized_bac_rhizo_ps) #now tree data is added

unfiltered_unnormalized_fun_rhizo_ps <- add_MY_metadata(unfiltered_unnormalized_fun_rhizo_ps) #now tree data is added

#Now make a new phyloseq object with only plant species present in the phylogenetic tree

keep_tree_species <- function(x) subset_samples(x, MY_tree_presence == "inside_MY_tree")

tree_bac_rhizo_ps <- keep_tree_species(unfiltered_unnormalized_bac_rhizo_ps)

tree_bac_rhizo_ps <- prune_taxa(taxa_sums(tree_bac_rhizo_ps) >0,tree_bac_rhizo_ps) #we need to prune the ASV that are not present in the subset

tree_fun_rhizo_ps <- keep_tree_species(unfiltered_unnormalized_fun_rhizo_ps)

tree_fun_rhizo_ps <- prune_taxa(taxa_sums(tree_fun_rhizo_ps) >0,tree_fun_rhizo_ps) #we need to prune the ASV that are not present in the subset

tree_BacFun_rhizo_ps <- list(tree_bac_rhizo_ps, tree_fun_rhizo_ps) #make a list with the two phyloseq objects

rm(unfiltered_unnormalized_bac_rhizo_ps) #we don't need the full phyloseq object
rm(unfiltered_unnormalized_fun_rhizo_ps) #we don't need the full phyloseq object


#Color blind friendly palette for plots
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

phylogeny_vector <- c("Arabidopsis_lyrata", "Arabidopsis_thaliana", "Camelina_sativa", "Turritis_glabra","Erysium_cheirantoides","Malcomia_maritima","Descurainia_sophiades","Lepidium_ruderale", "Barbarea_vulgaris",  "Berteroa_incana", "Euclydium_syriacum", "Brassica_rapa",  "Brassica_oleraceae_Rivera_Medium","Hirschfeldia_incana","Sinapis_arvensis","Diplotaxis_tenuifolia","Brassica_nigra", "Isatis_tinctoria", "Sysimbrium_irio","Cochlearia_officinalis", "Lobularia_maritima", "Iberis_amara") #needed for the plots
```

# 3.0 Basic alpha diversity analysis


Alpha diversity will show you the quantifiable diversity within a sample - this way you compare which treatment is more or less diverse or species-rich than the other.

Alpha diversity requires counts from samples with equal samples sizes. This means we cannot use the CSS transformation from the metagenomeseq package, we will use unfiltered data. 



#3.1. Alpha diversity testing

##3.1.1. Overall ANOVA (Supplementary Table S3)
```{r}

#define function to calculate alpha diversity and then add sample metadata
add_diversity_to_physeq_object<-function (phyloseq_object){

#Calculate richness  of a single phyloseq object

total_diversity<-estimate_richness(phyloseq_object)%>%
  rownames_to_column(var = "Sample")

total_diversity$Sample<-sub("X", "", total_diversity$Sample)# this removes the X that appears at the beginning of the first column

total_diversity<-column_to_rownames(total_diversity, var= "Sample")


#### add diversity metrics to mapping file of phyloseq objects
# we do this so we can perform anovas, access metadata, make nicer plots, etc
merg_to_ps<-sample_data(total_diversity) # makes the diversity calculations  sample_data for phyloseq object
sample_data(phyloseq_object)<-merge_phyloseq(phyloseq_object,merg_to_ps) # merge the new phyloseq object with the old phyloseq object
diversity<-as(sample_data(phyloseq_object),"data.frame") # forces sample data of updated phyloseq object into a dataframe

return (diversity)
}

# Calculate diversity for all samples
diversity_all <- lapply(tree_BacFun_rhizo_ps, add_diversity_to_physeq_object)
#For bacteria

save(diversity_all, file = "./03_Alpha_Diversity/r_objects/Diversity_all.RData")

diversity_all[[1]]$Shannon #here we have the Shannon indexes for 16S samples

aov_bac <- aov(diversity_all[[1]]$Shannon ~ Block + Sp_full_name + Stress, data = diversity_all[[1]])

overall_anova_bac <- summary(aov_bac) #Supplementary Table S3-A



#For fungi

diversity_all[[2]]$Shannon #here we have the Shannon indexes for 16S samples

aov_fun <- aov(diversity_all[[2]]$Shannon ~ Block + Sp_full_name + Stress, data = diversity_all[[2]])

overall_anova_fun <- summary(aov_fun) #Supplementary Table S3-B


```


##3.1.2. ANOVA per plant species
```{r}

# Calculate diversity per species
tree_species_ps_l <- lapply(tree_BacFun_rhizo_ps, function (x) metagMisc::phyloseq_sep_variable(x, variable = "Sp_full_name")) 

diversity <- lapply_lapply(tree_species_ps_l, add_diversity_to_physeq_object)
names(diversity) <- c('Bac','Fun')

save(diversity_all, file = "./03_Alpha_Diversity/r_objects/Diversity_per_species.RData")

# Calculate ANOVA

#first make a function that builds the model. We want to test the effect of block and stress 
aov_list <- function (diversity_output){aov(lm((Shannon) ~ Stress + Block, data = diversity_output))}

aov_all_l <- lapply_lapply(diversity, aov_list) 
save(aov_all_l, file = "./03_Alpha_Diversity/r_objects/aov_all_l.RData")

anova_function <- function (aov_l){anova(aov_l)}
anova <- lapply_lapply(aov_all_l, anova_function) #calculates anova
anova

#Function for extracting p-values and R2 of stress and block effects 
anova_output_to_plot_input<-function(anova_output_l){

p_values <-map(anova_output_l, 5)
list_lenght<-length(p_values)

#make vectors with data of interest
vector_of_p_stress <- as.vector(unlist(map(p_values,1)))
vector_of_p_block <- as.vector(unlist(map(p_values,2)))


# make a vector with all p values
vector_p <- c(vector_of_p_stress,
              vector_of_p_block)


#defines factors
factor_text<-c(rep("Stress", list_lenght),
          rep("Block", list_lenght))


species_name<-names(c(p_values, p_values))


#join vectors in a dataframe
p_df<- data.frame(species_name, # names of species in list
                  as.factor(factor_text), # Stress and block as factors
                      vector_p) # p values for the stress and then for the block


#change anmes of columns in dataframe
names(p_df)<-c("species","factor", "pvalue")

#output
return(p_df)

}

#Lets first check the anova results for 16S

anova_results_Bac_df <- anova_output_to_plot_input(anova$Bac)

anova_results_Bac_df$fdr_pvalue <- stats::p.adjust(p = anova_results_Bac_df$pvalue, method = "fdr") #Add FDR pvalue correction



save(anova_results_Bac_df, 
    file = "./03_Alpha_Diversity/r_objects/anova_results_Bac_df.RData")

#load(file = "./03_Alpha_Diversity/r_objects/Final_paper/anova_results_Bac_df.RData")



anova_results_Bac_df <- anova_results_Bac_df %>%
  arrange(match(factor, c("Stress", "Block")))%>%
    arrange(match(species, phylogeny_vector)) 
     #arrange order according to phylogeny

anova_results_Bac_df$species <- factor(anova_results_Bac_df$species, levels = phylogeny_vector) 
anova_results_Bac_df$factor <- factor(anova_results_Bac_df$factor, levels = c("Stress", "Block")) #arrange order for ggplot

write.csv(anova_results_Bac_df, "./03_Alpha_diversity/results/pval_Bac_Shannon_brassica_df.csv", row.names=FALSE) 



#Now lets look to ITS samples
anova_results_Fun_df <- anova_output_to_plot_input(anova$Fun)
anova_results_Fun_df$fdr_pvalue <- stats::p.adjust(p = anova_results_Fun_df$pvalue, method = "fdr") #Add FDR pvalue correction

save(anova_results_Fun_df, 
     file = "./03_Alpha_Diversity/r_objects/anova_results_Fun_df.RData")
#load(file = "./03_Alpha_Diversity/r_objects/anova_results_Fun_df.RData")


anova_results_Fun_df <- anova_results_Fun_df %>%
  arrange(match(factor, c("Stress",  "Block")))%>%
    arrange(match(species, phylogeny_vector)) 
     #arrange order according to phylogeny

anova_results_Fun_df$species <- factor(anova_results_Fun_df$species, levels = phylogeny_vector) 
anova_results_Fun_df$factor <- factor(anova_results_Fun_df$factor, levels = c("Stress", "Block")) #arrange order for ggplot

write.csv(anova_results_Fun_df, "./03_Alpha_diversity/results/pval_Fun_Shannon_brassica_df.csv", row.names=FALSE) 


```

##3.1.2. Tukey (Statistics Suppl. Figure S5)

```{r}
#post-hoc -> check tukey

tukey_function <- function(aov_list){TukeyHSD(aov_list, "Stress")}

tukey_output_l <- lapply_lapply(aov_all_l, tukey_function)
tukey_output_l

tukey_output_to_plot_input <- function(tukey_output_l){
  
#select the p-values for each pair of stress factor  
  
p_values_MeJA_Cntrl_l <- map(tukey_output_l, 10) #MeJA Cntrl
p_values_SA_Cntrl_l <- map(tukey_output_l, 11) #SA Cntrl
p_values_SA_MeJA_l <- map(tukey_output_l, 12) #SA MeJa
list_lenght<-length(p_values_MeJA_Cntrl_l)

vector_MeJA_Cntrl <- as.vector(unlist(p_values_MeJA_Cntrl_l, use.names = FALSE))
vector_SA_Cntrl <- as.vector(unlist(p_values_SA_Cntrl_l, use.names = FALSE))
vector_SA_MeJA <- as.vector(unlist(p_values_SA_MeJA_l, use.names = FALSE))

p_values <- c(vector_MeJA_Cntrl,
            vector_SA_Cntrl,
            vector_SA_MeJA)

list_lenght<-length(p_values_MeJA_Cntrl_l)

stress <-c(rep("MeJA-Cntrl", list_lenght),
               rep("SA-Cntrl", list_lenght),
               rep("SA-MeJA", list_lenght))


#join vectors in a dataframe
p_df<- data.frame(p_values, # p values for the stress and then for the block
                     # species_name, # names of species in list
                     stress) # Stress and block as factors

}

output <- lapply_lapply(tukey_output_l, tukey_output_to_plot_input)




#Results Tukey 16S
tukey_bac <- output$Bac

tukey_bac_df <- do.call(rbind, tukey_bac[phylogeny_vector]) #convert the Tukey output list into a dataframe

tukey_bac_df <- rownames_to_column(tukey_bac_df)

colnames(tukey_bac_df) <- c("species", "p_values", "stress")

tukey_bac_df$species <- gsub("\\.[1-3]$", "", as.character(tukey_bac_df$species)) #fixes the species names

#We should perfom Tukey test only on the species were there was a significant effect of stress in the anova. so let's remove the species that were not having a significant stress effect
sig_species_Bac <- anova_results_Bac_df %>% subset(factor == "Stress") %>% filter(pvalue < 0.05) %>% select(species)

tukey_bac_df <- merge(tukey_bac_df, sig_species_Bac, by="species")

#make a new column with -log10pvalue
tukey_bac_df <- tukey_bac_df %>%arrange(match(species, phylogeny_vector)) #arrange order according to phylogeny

tukey_bac_df <- tukey_bac_df%>% mutate(Significance = 
                                                         case_when(p_values <= 0.001 ~ "***",
                                                        p_values <= 0.01 ~ "**",
                                                        p_values <= 0.05 ~ "*",
                                                        p_values > 0.05 ~ "NS"))
     
write.csv(tukey_bac_df, "./03_Alpha_diversity/results/tukey_bac_df.csv", row.names=FALSE) #Statistics Supplementary Figure S5 Bacteria


#Results Tukey ITS
tukey_fun <- output$Fun

tukey_fun_df <- do.call(rbind, tukey_fun[phylogeny_vector]) #convert the Tukey output list into a dataframe

tukey_fun_df <- rownames_to_column(tukey_fun_df)

colnames(tukey_fun_df) <- c("species", "p_values", "stress")

tukey_fun_df$species <- gsub("\\.[1-3]$", "", as.character(tukey_fun_df$species)) #fixes the species names

#We should perfom Tukey test only on the species were there was a significant effect of stress in the anova. so let's remove the species that were not having a significant stress effect
sig_species_Fun <- anova_results_Fun_df %>% subset(factor == "Stress") %>% filter(pvalue < 0.05) %>% select(species)

tukey_fun_df <- merge(tukey_fun_df, sig_species_Fun, by = "species")

#make a new column with -log10pvalue

tukey_fun_df <- tukey_fun_df %>%arrange(match(species, phylogeny_vector)) #arrange order according to phylogeny

tukey_fun_df <- tukey_fun_df%>% mutate(Significance = 
                                                         case_when(p_values <= 0.001 ~ "***",
                                                        p_values <= 0.01 ~ "**",
                                                        p_values <= 0.05 ~ "*",
                                                        p_values > 0.05 ~ "NS"))
     
write.csv(tukey_fun_df, "./03_Alpha_diversity/results/tukey_fun_df.csv", row.names=FALSE) #Statistics Supplementary Figure S5 Fungi




```




#3.2. Supplementry figure S5

```{r}
diversity_all
names(diversity_all) <- c("Bac","Fun")


##Fix dataframe for plots

diversity_all$Bac$Sp_full_name <- gsub("_", " ", diversity_all$Bac$Sp_full_name) #fixes the species names
diversity_all$Fun$Sp_full_name <- gsub("_", " ", diversity_all$Fun$Sp_full_name) #fixes the species names

 diversity_all$Bac <- diversity_all$Bac %>%  mutate(Sp_full_name2 =
                                                     case_when(Sp_full_name == "Brassica oleraceae Rivera Medium" ~ "Brassica oleracea",
                                                     TRUE ~ Sp_full_name)) #Renaming B. oleracea

diversity_all$Fun <- diversity_all$Fun %>%  mutate(Sp_full_name2 =
                                                     case_when(Sp_full_name == "Brassica oleraceae Rivera Medium" ~ "Brassica oleracea",
                                                     TRUE ~ Sp_full_name)) #Renaming B. oleracea


phylogeny_vector <- c("Arabidopsis lyrata", "Arabidopsis thaliana", "Camelina sativa", "Turritis glabra","Erysium cheirantoides","Malcomia maritima","Descurainia sophiades","Lepidium ruderale", "Barbarea vulgaris",  "Berteroa incana", "Euclydium syriacum", "Brassica rapa",  "Brassica oleracea","Hirschfeldia incana","Sinapis arvensis","Diplotaxis tenuifolia","Brassica nigra", "Isatis tinctoria", "Sysimbrium irio","Cochlearia officinalis", "Lobularia maritima", "Iberis amara") #needed for the plots

diversity_all$Bac$Sp_full_name2 <- factor(diversity_all$Bac$Sp_full_name2, levels = phylogeny_vector)#arrange order for ggplot
diversity_all$Fun$Sp_full_name2 <- factor(diversity_all$Fun$Sp_full_name2, levels = phylogeny_vector)#arrange order for ggplot

names(diversity_all$Bac)[names(diversity_all$Bac) == "Stress"] <- "Treatment" #change stress for treatment
names(diversity_all$Fun)[names(diversity_all$Fun) == "Stress"] <- "Treatment"  #change stress for treatment


##Put bacteria and fungi in the same dataframe and re-do the plot

#Bac
names(diversity_all$Bac) #I need columns 5, 52 and 56

diversity_Bac <- diversity_all$Bac %>% select(5, 52, 56)

diversity_Bac <- tibble::rownames_to_column(diversity_Bac, "Sample")

diversity_Bac <- diversity_Bac %>%
  add_column(Kingdom = "Bacteria")

#Fun
names(diversity_all$Fun) #I need columns 5, 52 and 56

diversity_Fun <- diversity_all$Fun %>% select(5, 50, 54)

diversity_Fun <- tibble::rownames_to_column(diversity_Fun, "Sample")

diversity_Fun <- diversity_Fun %>%
  add_column(Kingdom = "Fungi")

#Merge
diversity_BacFun <- rbind(diversity_Bac, diversity_Fun)

diversity_BacFun$Sp_full_name2 <- factor(diversity_BacFun$Sp_full_name2, levels = phylogeny_vector)#arrange order for ggplot
diversity_BacFun$Treatment <- factor(diversity_BacFun$Treatment, levels=c("SA", "MeJA", "Control"))

#plot

shannon_H_BacFun_boxplot2 <- ggplot(diversity_BacFun, aes(x = Shannon, y = Sp_full_name2, fill = Treatment, colour = Treatment)) +
  geom_boxplot(width = 0.9) +
  scale_colour_manual(values=c("Control" = "#999999", 
                                      "MeJA" = "#E69F00",
                                      "SA" = "#009E73")) +
  scale_fill_manual(values=c("Control" = "#999999", 
                                      "MeJA" = "#E69F00",
                                      "SA" = "#009E73")) +
  xlab("Shannon's index H") +
  ylab("Plant species") +
  scale_y_discrete(limits=rev) + #to have phylogeny order
  guides(fill = guide_legend(reverse=TRUE), colour = guide_legend(reverse=TRUE)) +
  theme_light() +
  facet_wrap(~Kingdom, scales = "free_x") +
  theme(text = element_text(size = 30), axis.text.y=element_text(hjust=1, face=c(rep("italic", 22), "plain"))) #for italicizing plant species names


ggsave("Supp_Fig_S5.png", path = "./03_Alpha_diversity/results",
 width = 40, height = 20, units = "in", dpi = 700,limitsize = FALSE )

```

