---
title: "5_differential_abundance"
author: "Melissa Uribe"
date: "March 25, 2022"
output: Rmd file
editor_options: 
  chunk_output_type: console
---
## 5.01 load libraries, improve memory use and subset objects for the desired treatment comparisons
```{r results="hide"}
# this will load all key libraries for the whole project
source("./renv/0_load_libraries.R")

# increases memory limit used by R
memory.limit(size = 350000)    

#Load object with bac and fun and separate to each group
load("./Data/phyloseq_objects/unnormalized_BacFun_rhizoplane_HighFilt_ps_l.RData")
unnorm_Bac_ps <- unnormalized_BacFun_rhizoplane_HighFilt_ps_l[["Bac"]]
unnorm_Fun_ps <- unnormalized_BacFun_rhizoplane_HighFilt_ps_l[["Fun"]]

#BACTERIA
##Make vectors for Meja and SA 
MeJA <-c("MeJA")
SA <-c("SA")

MeJA_unnorm_Bac_ps <- subset_samples(unnorm_Bac_ps, !Stress %in% SA)
SA_unnorm_Bac_ps <- subset_samples(unnorm_Bac_ps, !Stress %in% MeJA)

#separate to each plant species
MeJA_unnorm_Bac_ps_l <- metagMisc::phyloseq_sep_variable(MeJA_unnorm_Bac_ps, variable = "Sp_abb_name")
SA_unnorm_Bac_ps_l <- metagMisc::phyloseq_sep_variable(SA_unnorm_Bac_ps, variable = "Sp_abb_name")

#Save the unfiltered (with zeroes) object for Standard DESeq
MeJA_unnorm_Bac_ps_l_unfilt <- MeJA_unnorm_Bac_ps_l
SA_unnorm_Bac_ps_l_unfilt <- SA_unnorm_Bac_ps_l

save(MeJA_unnorm_Bac_ps_l_unfilt,  
     file = "./05_Differential_abundance/Objects/MeJA_unnorm_Bac_ps_l_unfilt.RData")
save(SA_unnorm_Bac_ps_l_unfilt,  
     file = "./05_Differential_abundance/Objects/SA_unnorm_Bac_ps_l_unfilt.RData")

##For ZINBWAve, remove taxa with a total sum of 0:
MeJA_unnorm_Bac_ps_l <- lapply(MeJA_unnorm_Bac_ps_l, function(x)
         prune_taxa(taxa_sums(x) >0,x))
SA_unnorm_Bac_ps_l <- lapply(SA_unnorm_Bac_ps_l, function(x)
         prune_taxa(taxa_sums(x) >0,x))

save(MeJA_unnorm_Bac_ps_l,  
     file = "./05_Differential_abundance/Objects/MeJA_unnorm_Bac_ps_l.RData")
save(SA_unnorm_Bac_ps_l,  
     file = "./05_Differential_abundance/Objects/SA_unnorm_Bac_ps_l.RData")


#FUNGI
MeJA_unnorm_Fun_ps <- subset_samples(unnorm_Fun_ps, !Stress %in% SA)
SA_unnorm_Fun_ps <- subset_samples(unnorm_Fun_ps, !Stress %in% MeJA)

#separate to each plant species
MeJA_unnorm_Fun_ps_l <- metagMisc::phyloseq_sep_variable(MeJA_unnorm_Fun_ps, variable = "Sp_abb_name")
SA_unnorm_Fun_ps_l <- metagMisc::phyloseq_sep_variable(SA_unnorm_Fun_ps, variable = "Sp_abb_name")

#Save the unfiltered (with zeroes) object for Standard DESeq
MeJA_unnorm_Fun_ps_l_unfilt <- MeJA_unnorm_Fun_ps_l
SA_unnorm_Fun_ps_l_unfilt <- SA_unnorm_Fun_ps_l

save(MeJA_unnorm_Fun_ps_l_unfilt,  
     file = "./05_Differential_abundance/Objects/MeJA_unnorm_Fun_ps_l_unfilt.RData")
save(SA_unnorm_Fun_ps_l_unfilt,  
     file = "./05_Differential_abundance/Objects/SA_unnorm_Fun_ps_l_unfilt.RData")

##For ZINBWAve, remove taxa with a total sum of 0:
MeJA_unnorm_Fun_ps_l <- lapply(MeJA_unnorm_Fun_ps_l, function(x)
         prune_taxa(taxa_sums(x) >0,x))
SA_unnorm_Fun_ps_l <- lapply(SA_unnorm_Fun_ps_l, function(x)
         prune_taxa(taxa_sums(x) >0,x))

save(MeJA_unnorm_Fun_ps_l,  
     file = "./05_Differential_abundance/Objects/MeJA_unnorm_Fun_ps_l.RData")
save(SA_unnorm_Fun_ps_l,  
     file = "./05_Differential_abundance/Objects/SA_unnorm_Fun_ps_l.RData")
```

#5.1 DeSeq and ZINBDeSeq
##5.1.1 Convert phyloseq object to DeSeq object with lists
```{r}
#Convert unfiltered object to DESeq object for standard DESeq
dds_MeJA_Bac_ps_unfilt <- lapply(MeJA_unnorm_Bac_ps_l_unfilt, function(x) phyloseq_to_deseq2(x, ~ Stress))
save(dds_MeJA_Bac_ps_unfilt, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/dds_MeJA_Bac_ps_unfilt.RData")

dds_SA_Bac_ps_unfilt <- lapply(SA_unnorm_Bac_ps_l_unfilt, function(x) phyloseq_to_deseq2(x, ~ Stress))
save(dds_SA_Bac_ps_unfilt, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/dds_SA_Bac_ps_unfilt.RData")

dds_MeJA_Fun_ps_unfilt <- lapply(MeJA_unnorm_Fun_ps_l_unfilt, function(x) phyloseq_to_deseq2(x, ~ Stress))
save(dds_MeJA_Fun_ps_unfilt, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/dds_MeJA_Fun_ps_unfilt.RData")

dds_SA_Fun_ps_unfilt <- lapply(SA_unnorm_Fun_ps_l_unfilt, function(x) phyloseq_to_deseq2(x, ~ Stress))
save(dds_SA_Fun_ps_unfilt, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/dds_SA_Fun_ps_unfilt.RData")


#Convert filtered object to DeSeq object for ZINBWave
dds_MeJA_Bac_ps <- lapply(MeJA_unnorm_Bac_ps_l, function(x) phyloseq_to_deseq2(x, ~ Stress))
save(dds_MeJA_Bac_ps, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/dds_MeJA_Bac_ps.RData")

dds_SA_Bac_ps <- lapply(SA_unnorm_Bac_ps_l, function(x) phyloseq_to_deseq2(x, ~ Stress))
save(dds_SA_Bac_ps, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/dds_SA_Bac_ps.RData")

dds_MeJA_Fun_ps <- lapply(MeJA_unnorm_Fun_ps_l, function(x) phyloseq_to_deseq2(x, ~ Stress))
save(dds_MeJA_Fun_ps, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/dds_MeJA_Fun_ps.RData")

dds_SA_Fun_ps <- lapply(SA_unnorm_Fun_ps_l, function(x) phyloseq_to_deseq2(x, ~ Stress))
save(dds_SA_Fun_ps, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/dds_SA_Fun_ps.RData")

```

##5.1.2 DESeq
```{r}
#MeJA_Bac
#Standard DESeq
stddds2_Wald_MeJA_Bac <- lapply(dds_MeJA_Bac_ps_unfilt, function(x) 
                  DESeq(x, 
                  test="Wald",  
                  minReplicatesForReplace=Inf,
                  fitType = "local")) 

save(stddds2_Wald_MeJA_Bac, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/stddds2_Wald_MeJA_Bac.RData")

#Dataframe generation
lfcres_stddds2_Wald_MeJA_Bac <- lapply(stddds2_Wald_MeJA_Bac, function(x) lfcShrink(x,  coef="Stress_MeJA_vs_Control", type="normal"))

save(lfcres_stddds2_Wald_MeJA_Bac, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/lfcres_stddds2_Wald_MeJA_Bac.RData")

#Filter to only p adjusted under alpha
alpha = 0.05
sigtab_stddds2_Wald_MeJA_Bac <- lapply(lfcres_stddds2_Wald_MeJA_Bac, function(x) x[which(x$padj < alpha), ])

save(sigtab_stddds2_Wald_MeJA_Bac, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/sigtab_stddds2_Wald_MeJA_Bac.RData")

#SA_Bac
#DeSeq with Wald test
stddds2_Wald_SA_Bac <- lapply(dds_SA_Bac_ps_unfilt, function(x) 
                  DESeq(x, 
                  test="Wald",  
                  minReplicatesForReplace=Inf,
                  fitType = "local")) 

save(stddds2_Wald_SA_Bac, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/stddds2_Wald_SA_Bac.RData")

#Dataframe generation
lfcres_stddds2_Wald_SA_Bac <- lapply(stddds2_Wald_SA_Bac, function(x) lfcShrink(x,  coef="Stress_SA_vs_Control", type="normal"))

save(lfcres_stddds2_Wald_SA_Bac, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/lfcres_stddds2_Wald_SA_Bac.RData")

#Filter to only p adjusted under alpha
sigtab_stddds2_Wald_SA_Bac <- lapply(lfcres_stddds2_Wald_SA_Bac, function(x) x[which(x$padj < alpha), ])

save(sigtab_stddds2_Wald_SA_Bac, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/sigtab_stddds2_Wald_SA_Bac.RData")


#MeJA_Fun
#Standard DESeq
stddds2_Wald_MeJA_Fun <- lapply(dds_MeJA_Fun_ps_unfilt, function(x) 
                  DESeq(x, 
                  test="Wald",  
                  minReplicatesForReplace=Inf,
                  fitType = "local")) 

save(stddds2_Wald_MeJA_Fun, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/stddds2_Wald_MeJA_Fun.RData")

alpha = 0.05

#Dataframe generation
lfcres_stddds2_Wald_MeJA_Fun <- lapply(stddds2_Wald_MeJA_Fun, function(x) lfcShrink(x,  coef="Stress_MeJA_vs_Control", type="normal"))

save(lfcres_stddds2_Wald_MeJA_Fun, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/lfcres_stddds2_Wald_MeJA_Fun.RData")

#Filter to only p adjusted under alpha
alpha = 0.05
sigtab_stddds2_Wald_MeJA_Fun <- lapply(lfcres_stddds2_Wald_MeJA_Fun, function(x) x[which(x$padj < alpha), ])

save(sigtab_stddds2_Wald_MeJA_Fun, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/sigtab_stddds2_Wald_MeJA_Fun.RData")



#SA_Fun
#Standard DESeq
stddds2_Wald_SA_Fun <- lapply(dds_SA_Fun_ps_unfilt, function(x) 
                  DESeq(x, 
                  test="Wald",  
                  minReplicatesForReplace=Inf,
                  fitType = "local")) 

save(stddds2_Wald_SA_Fun, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/stddds2_Wald_SA_Fun.RData")

alpha = 0.05

#Dataframe generation
lfcres_stddds2_Wald_SA_Fun <- lapply(stddds2_Wald_SA_Fun, function(x) lfcShrink(x,  coef="Stress_SA_vs_Control", type="normal"))

save(lfcres_stddds2_Wald_SA_Fun, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/lfcres_stddds2_Wald_SA_Fun.RData")

#Filter to only p adjusted under alpha
alpha = 0.05
sigtab_stddds2_Wald_SA_Fun <- lapply(lfcres_stddds2_Wald_SA_Fun, function(x) x[which(x$padj < alpha), ])

save(sigtab_stddds2_Wald_SA_Fun, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/sigtab_stddds2_Wald_SA_Fun.RData")

```

##5.1.3 ZINBDESeq
####5.1.3.1 Calculate weights
```{r}
#library("scran") DESeq can compute the size factors already from the weights output of zinbwave
library(zinbwave)

#Calculate observational weights with ZINBwave

#MeJA_Bac
zw_MeJA_Bac <- lapply(dds_MeJA_Bac_ps, function(x)
                                zinbwave(x,
                               epsilon = 1e10,
                               verbose = TRUE,
                               K = 0,
                               observationalWeights = TRUE,
                               BPPARAM = BiocParallel::SerialParam()))

save(zw_MeJA_Bac, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/zw_MeJA_Bac.RData")

#Check that weights have been calculated for each assay in each smaple
assay(zw_MeJA_Bac$At, "weights")[1:20, 1:20]

#SA_Bac
zw_SA_Bac <- lapply(dds_SA_Bac_ps, function(x)
                                zinbwave(x,
                               epsilon = 1e10,
                               verbose = TRUE,
                               K = 0,
                               observationalWeights = TRUE,
                               BPPARAM = BiocParallel::SerialParam()))
save(zw_SA_Bac, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/zw_SA_Bac.RData")

 
#MeJA_Fun
zw_MeJA_Fun <- lapply(dds_MeJA_Fun_ps, function(x)
                                zinbwave(x,
                               epsilon = 1e10,
                               verbose = TRUE,
                               K = 0,
                               observationalWeights = TRUE,
                               BPPARAM = BiocParallel::SerialParam()))

save(zw_MeJA_Fun, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/zw_MeJA_Fun.RData")



#SA_Fun
zw_SA_Fun <- lapply(dds_SA_Fun_ps, function(x)
                                zinbwave(x,
                               epsilon = 1e10,
                               verbose = TRUE,
                               K = 0,
                               observationalWeights = TRUE,
                               BPPARAM = BiocParallel::SerialParam()))

save(zw_SA_Fun, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/zw_SA_Fun.RData")

```
 

####5.1.3.2 Estimate size factors and run DeSeq
```{r}
load("./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/zw_MeJA_Bac.RData")
load("./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/zw_SA_Bac.RData")
load("./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/zw_MeJA_Fun.RData")
load("./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/zw_SA_Fun.RData")

#MeJA_Bac
#Compute size factors with DeSeq
dds2_zw_MeJA_Bac <- lapply(zw_MeJA_Bac, 
                          function(x) DESeqDataSet(x, design = ~ Stress))
dds2_zw_MeJA_Bac <- lapply(dds2_zw_MeJA_Bac, 
                            function(x) estimateSizeFactors(x, type="poscounts"))
save(dds2_zw_MeJA_Bac, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/dds2_zw_MeJA_Bac.RData")

#DeSeq with LRT test
dds2_LRT_zinb_MeJA_Bac <- lapply(dds2_zw_MeJA_Bac, function(x) 
                  DESeq(x, 
                  test="LRT", 
                  reduced=~1,
                  sfType="poscounts", 
                  minmu=1e-6, 
                  minReplicatesForReplace=Inf,
                  fitType = "local"))

save(dds2_LRT_zinb_MeJA_Bac, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/dds2_LRT_zinb_MeJA_Bac.RData")

#DeSeq with Wald test test
dds2_Wald_zinb_MeJA_Bac <- lapply(dds2_zw_MeJA_Bac, function(x) 
                  DESeq(x, 
                  test="Wald",
                  sfType="poscounts",
                  useT=TRUE,
                  minmu=1e-6, 
                  minReplicatesForReplace=Inf,
                  fitType = "local")) 

save(dds2_Wald_zinb_MeJA_Bac, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/dds2_Wald_zinb_MeJA_Bac.RData")


#SA_Bac
#Compute size factors with DeSeq
dds2_zw_SA_Bac <- lapply(zw_SA_Bac, 
                          function(x) DESeqDataSet(x, design = ~ Stress))
dds2_zw_SA_Bac <- lapply(dds2_zw_SA_Bac, 
                            function(x) estimateSizeFactors(x, type="poscounts"))

save(dds2_zw_SA_Bac, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/dds2_zw_SA_Bac.RData")

#DeSeq with LRT test
dds2_LRT_zinb_SA_Bac <- lapply(dds2_zw_SA_Bac, function(x) 
                  DESeq(x, 
                  test="LRT", 
                  reduced=~1,
                  sfType="poscounts", 
                  minmu=1e-6, 
                  minReplicatesForReplace=Inf,
                  fitType = "local")) 

save(dds2_LRT_zinb_SA_Bac, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/dds2_LRT_zinb_SA_Bac.RData")

#DeSeq with Wald test
dds2_Wald_zinb_SA_Bac <- lapply(dds2_zw_SA_Bac, function(x) 
                  DESeq(x, 
                  test="Wald",
                  sfType="poscounts",
                  useT=TRUE,
                  minmu=1e-6, 
                  minReplicatesForReplace=Inf,
                  fitType = "local")) 

save(dds2_Wald_zinb_SA_Bac, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/dds2_Wald_zinb_MeJA_Bac.RData")

#MeJA_Fun
#Compute size factors with DeSeq
dds2_zw_MeJA_Fun <- lapply(zw_MeJA_Fun, 
                          function(x) DESeqDataSet(x, design = ~ Stress))
dds2_zw_MeJA_Fun <- lapply(dds2_zw_MeJA_Fun, 
                            function(x) estimateSizeFactors(x, type="poscounts"))
save(dds2_zw_MeJA_Fun, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/dds2_zw_MeJA_Fun.RData")

#DeSeq with LRT test
dds2_LRT_zinb_MeJA_Fun <- lapply(dds2_zw_MeJA_Fun, function(x) 
                  DESeq(x, 
                  test="LRT", 
                  reduced=~1,
                  sfType="poscounts", 
                  minmu=1e-6, 
                  minReplicatesForReplace=Inf,
                  fitType = "local"))

save(dds2_LRT_zinb_MeJA_Fun, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/dds2_LRT_zinb_MeJA_Fun.RData")

#DeSeq with Wald test test
dds2_Wald_zinb_MeJA_Fun <- lapply(dds2_zw_MeJA_Fun, function(x) 
                  DESeq(x, 
                  test="Wald",
                  sfType="poscounts",
                  useT=TRUE,
                  minmu=1e-6, 
                  minReplicatesForReplace=Inf,
                  fitType = "local")) 

save(dds2_Wald_zinb_MeJA_Fun, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/dds2_Wald_zinb_MeJA_Fun.RData")

#SA_Fun
#Compute size factors with DeSeq
dds2_zw_SA_Fun <- lapply(zw_SA_Fun, 
                          function(x) DESeqDataSet(x, design = ~ Stress))
dds2_zw_SA_Fun <- lapply(dds2_zw_SA_Fun, 
                            function(x) estimateSizeFactors(x, type="poscounts"))
save(dds2_zw_SA_Fun, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/dds2_zw_SA_Fun.RData")

#DeSeq with LRT test
dds2_LRT_zinb_SA_Fun <- lapply(dds2_zw_SA_Fun, function(x) 
                  DESeq(x, 
                  test="LRT", 
                  reduced=~1,
                  sfType="poscounts", 
                  minmu=1e-6, 
                  minReplicatesForReplace=Inf,
                  fitType = "local"))

save(dds2_LRT_zinb_SA_Fun, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/dds2_LRT_zinb_SA_Fun.RData")

#DeSeq with Wald test test
dds2_Wald_zinb_SA_Fun <- lapply(dds2_zw_SA_Fun, function(x) 
                  DESeq(x, 
                  test="Wald",
                  sfType="poscounts",
                  useT=TRUE,
                  minmu=1e-6, 
                  minReplicatesForReplace=Inf,
                  fitType = "local")) 

save(dds2_Wald_zinb_SA_Fun, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/dds2_Wald_zinb_SA_Fun.RData")

```

####5.1.3.3 Dataframe generation
```{r}
alpha = 0.05
#MeJA_Bac
##LRT
res_zinbLRT_MeJA_Bac <- lapply(dds2_LRT_zinb_MeJA_Bac, function(x) results(x, contrast = c("Stress", "MeJA", "Control")))
save(res_zinbLRT_MeJA_Bac, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/res_zinbLRT_MeJA_Bac.RData")
#Filter to only p adjusted under alpha
sigtab_zinbLRT_MeJA_Bac <- lapply(res_zinbLRT_MeJA_Bac, function(x) x[which(x$padj < alpha), ])
save(sigtab_zinbLRT_MeJA_Bac, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/sigtab_zinbLRT_MeJA_Bac.RData")

##Wald
res_zinbWald_MeJA_Bac <- lapply(dds2_Wald_zinb_MeJA_Bac, function(x) results(x, contrast = c("Stress", "MeJA", "Control")))
save(res_zinbWald_MeJA_Bac, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/res_zinbWald_MeJA_Bac.RData")
#Filter to only p adjusted under alpha
sigtab_zinbWald_MeJA_Bac <- lapply(res_zinbWald_MeJA_Bac, function(x) x[which(x$padj < alpha), ])
save(sigtab_zinbWald_MeJA_Bac, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/sigtab_zinbWald_MeJA_Bac.RData")

#SA_Bac
##LRT
res_zinbLRT_SA_Bac <- lapply(dds2_LRT_zinb_SA_Bac, function(x) results(x, contrast = c("Stress", "SA", "Control")))
save(res_zinbLRT_SA_Bac, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/res_zinbLRT_SA_Bac.RData")
#Filter to only p adjusted under alpha
sigtab_zinbLRT_SA_Bac <- lapply(res_zinbLRT_SA_Bac, function(x) x[which(x$padj < alpha), ])
save(sigtab_zinbLRT_SA_Bac, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/sigtab_zinbLRT_SA_Bac.RData")

##Wald
res_zinbWald_SA_Bac <- lapply(dds2_Wald_zinb_SA_Bac, function(x) results(x, contrast = c("Stress", "SA", "Control")))
save(res_zinbWald_SA_Bac, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/res_zinbWald_SA_Bac.RData")
#Filter to only p adjusted under alpha
sigtab_zinbWald_SA_Bac <- lapply(res_zinbWald_SA_Bac, function(x) x[which(x$padj < alpha), ])
save(sigtab_zinbWald_SA_Bac, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/sigtab_zinbWald_SA_Bac.RData")

#MeJA_Fun
##LRT
res_zinbLRT_MeJA_Fun <- lapply(dds2_LRT_zinb_MeJA_Fun, function(x) results(x, contrast = c("Stress", "MeJA", "Control")))
save(res_zinbLRT_MeJA_Fun, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/res_zinbLRT_MeJA_Fun.RData")
#Filter to only p adjusted under alpha
sigtab_zinbLRT_MeJA_Fun <- lapply(res_zinbLRT_MeJA_Fun, function(x) x[which(x$padj < alpha), ])
save(sigtab_zinbLRT_MeJA_Fun, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/sigtab_zinbLRT_MeJA_Fun.RData")

##Wald
res_zinbWald_MeJA_Fun <- lapply(dds2_Wald_zinb_MeJA_Fun, function(x) results(x, contrast = c("Stress", "MeJA", "Control")))
save(res_zinbWald_MeJA_Fun, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/res_zinbWald_MeJA_Fun.RData")
#Filter to only p adjusted under alpha
sigtab_zinbWald_MeJA_Fun <- lapply(res_zinbWald_MeJA_Fun, function(x) x[which(x$padj < alpha), ])
save(sigtab_zinbWald_MeJA_Fun, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/sigtab_zinbWald_MeJA_Fun.RData")

#SA_Fun
##LRT
res_zinbLRT_SA_Fun <- lapply(dds2_LRT_zinb_SA_Fun, function(x) results(x, contrast = c("Stress", "SA", "Control")))
save(res_zinbLRT_SA_Fun, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/res_zinbLRT_SA_Fun.RData")
#Filter to only p adjusted under alpha
sigtab_zinbLRT_SA_Fun <- lapply(res_zinbLRT_SA_Fun, function(x) x[which(x$padj < alpha), ])
save(sigtab_zinbLRT_SA_Fun, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/sigtab_zinbLRT_SA_Fun.RData")

##Wald
res_zinbWald_SA_Fun <- lapply(dds2_Wald_zinb_SA_Fun, function(x) results(x, contrast = c("Stress", "SA", "Control")))
save(res_zinbWald_SA_Fun, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/res_zinbWald_SA_Fun.RData")
#Filter to only p adjusted under alpha
sigtab_zinbWald_SA_Fun <- lapply(res_zinbWald_SA_Fun, function(x) x[which(x$padj < alpha), ])
save(sigtab_zinbWald_SA_Fun, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/sigtab_zinbWald_SA_Fun.RData")

```


#5.2 ANCOM-BC
##5.2.1 ANCOM including structural zeroes detection
```{r}
library(ANCOMBC)

#Bac_MeJA
ancomWZ_Bac_MeJA <- lapply(MeJA_unnorm_Bac_ps_l, function (x)
                            ancombc(phyloseq = x, 
                             formula = "Stress", 
                             p_adj_method = "fdr", 
                             zero_cut = 0.90, 
                             lib_cut = 0, 
                             group = "Stress", 
                             struc_zero = TRUE, 
                             neg_lb = TRUE, 
                             tol = 1e-5,               
                             max_iter = 100, 
                             conserve = TRUE, 
                             alpha = 0.05, 
                             global = FALSE))


save(ancomWZ_Bac_MeJA, 
     file = "./05_Differential_abundance/Objects/5.2ANCOM/ancomWZ_Bac_MeJA.RData")

#Bac_SA
ancomWZ_Bac_SA<- lapply(SA_unnorm_Bac_ps_l, function (x)
                            ancombc(phyloseq = x, 
                             formula = "Stress", 
                             p_adj_method = "fdr", 
                             zero_cut = 0.90, 
                             lib_cut = 0, 
                             group = "Stress", 
                             struc_zero = TRUE, 
                             neg_lb = TRUE, 
                             tol = 1e-5,               
                             max_iter = 100, 
                             conserve = TRUE, 
                             alpha = 0.05, 
                             global = FALSE))


save(ancomWZ_Bac_SA, 
     file = "./05_Differential_abundance/Objects/5.2ANCOM/ancomWZ_Bac_SA.RData")



#Fun_MeJA
load("./05_Differential_abundance/Objects/MeJA_unnorm_Fun_ps_l.RData")
ancomWZ_Fun_MeJA <- lapply(MeJA_unnorm_Fun_ps_l, function (x)
                            ancombc(phyloseq = x, 
                             formula = "Stress", 
                             p_adj_method = "fdr", 
                             zero_cut = 0.90, 
                             lib_cut = 0, 
                             group = "Stress", 
                             struc_zero = TRUE, 
                             neg_lb = TRUE, 
                             tol = 1e-5,               
                             max_iter = 100, 
                             conserve = TRUE, 
                             alpha = 0.05, 
                             global = FALSE))


save(ancomWZ_Fun_MeJA, 
     file = "./05_Differential_abundance/Objects/5.2ANCOM/ancomWZ_Fun_MeJA.RData")

#Fun_SA
load("./05_Differential_abundance/Objects/SA_unnorm_Fun_ps_l.RData")
ancomWZ_Fun_SA<- lapply(SA_unnorm_Fun_ps_l, function (x)
                            ancombc(phyloseq = x, 
                             formula = "Stress", 
                             p_adj_method = "fdr", 
                             zero_cut = 0.90, 
                             lib_cut = 0, 
                             group = "Stress", 
                             struc_zero = TRUE, 
                             neg_lb = TRUE, 
                             tol = 1e-5,               
                             max_iter = 100, 
                             conserve = TRUE, 
                             alpha = 0.05, 
                             global = FALSE))


save(ancomWZ_Fun_SA, 
     file = "./05_Differential_abundance/Objects/5.2ANCOM/ancomWZ_Fun_SA.RData")

#Organize and filter results
##MeJA_Bac
ancomWZ_res_df_MeJA_Bac <- lapply(ancomWZ_Bac_MeJA, function (x)
  data.frame(
  Species = row.names(x$res$beta),
  beta = unlist(x$res$beta),
  se = unlist(x$res$se),
  W = unlist(x$res$W),
  p_val = unlist(x$res$p_val),
  q_val = unlist(x$res$q_val),
  diff_abn = unlist(x$res$diff_abn)))

fdr_ancom_MeJA_Bac <- lapply(ancomWZ_res_df_MeJA_Bac, function(x)
  (x %>%
  dplyr::filter(q_val < 0.05)))

head(fdr_ancom_MeJA_Bac)

##SA_Bac
ancomWZ_res_df_SA_Bac <- lapply(ancomWZ_Bac_SA, function (x)
  data.frame(
  Species = row.names(x$res$beta),
  beta = unlist(x$res$beta),
  se = unlist(x$res$se),
  W = unlist(x$res$W),
  p_val = unlist(x$res$p_val),
  q_val = unlist(x$res$q_val),
  diff_abn = unlist(x$res$diff_abn)))

fdr_ancom_SA_Bac <- lapply(ancomWZ_res_df_SA_Bac, function(x)
  (x %>%
  dplyr::filter(q_val < 0.05)))



save(fdr_ancom_MeJA_Bac, 
     file = "./05_Differential_abundance/Objects/5.2ANCOM/fdr_ancomWZ_MeJA_Bac.RData")
save(fdr_ancom_SA_Bac, 
     file = "./05_Differential_abundance/Objects/5.2ANCOM/fdr_ancomWZ_SA_Bac.RData")

#Organize and filter results
##MeJA_Fun
ancomWZ_res_df_MeJA_Fun <- lapply(ancomWZ_Fun_MeJA, function (x)
  data.frame(
  Species = row.names(x$res$beta),
  beta = unlist(x$res$beta),
  se = unlist(x$res$se),
  W = unlist(x$res$W),
  p_val = unlist(x$res$p_val),
  q_val = unlist(x$res$q_val),
  diff_abn = unlist(x$res$diff_abn)))

fdr_ancom_MeJA_Fun <- lapply(ancomWZ_res_df_MeJA_Fun, function(x)
  (x %>%
  dplyr::filter(q_val < 0.05)))

##SA_Fun
ancomWZ_res_df_SA_Fun <- lapply(ancomWZ_Fun_SA, function (x)
  data.frame(
  Species = row.names(x$res$beta),
  beta = unlist(x$res$beta),
  se = unlist(x$res$se),
  W = unlist(x$res$W),
  p_val = unlist(x$res$p_val),
  q_val = unlist(x$res$q_val),
  diff_abn = unlist(x$res$diff_abn)))

fdr_ancom_SA_Fun <- lapply(ancomWZ_res_df_SA_Fun, function(x)
  (x %>%
  dplyr::filter(q_val < 0.05)))



save(fdr_ancom_MeJA_Fun, 
     file = "./05_Differential_abundance/Objects/5.2ANCOM/fdr_ancomWZ_MeJA_Fun.RData")
save(fdr_ancom_SA_Fun, 
     file = "./05_Differential_abundance/Objects/5.2ANCOM/fdr_ancomWZ_SA_Fun.RData")


```

##5.2.1 ANCOM WITHOUT structural zeroes detection
```{r}
load("./05_Differential_abundance/Objects/MeJA_unnorm_Bac_ps_l.RData")
load("./05_Differential_abundance/Objects/SA_unnorm_Bac_ps_l.RData")

#Bac_MeJA
ancomNoZ_Bac_MeJA <- lapply(MeJA_unnorm_Bac_ps_l, function (x)
                            ancombc(phyloseq = x, 
                             formula = "Stress", 
                             p_adj_method = "fdr", 
                             zero_cut = 0.90, 
                             lib_cut = 0, 
                             group = "Stress", 
                             struc_zero = FALSE, 
                             neg_lb = TRUE, 
                             tol = 1e-5,               
                             max_iter = 100, 
                             conserve = TRUE, 
                             alpha = 0.05, 
                             global = FALSE))
save(ancomNoZ_Bac_MeJA, 
     file = "./05_Differential_abundance/Objects/5.2ANCOM/ancomNoZ_Bac_MeJA.RData")


#Bac_SA
ancomNoZ_Bac_SA <- lapply(SA_unnorm_Bac_ps_l, function (x)
                            ancombc(phyloseq = x, 
                             formula = "Stress", 
                             p_adj_method = "fdr", 
                             zero_cut = 0.90, 
                             lib_cut = 0, 
                             group = "Stress", 
                             struc_zero = FALSE, 
                             neg_lb = TRUE, 
                             tol = 1e-5,               
                             max_iter = 100, 
                             conserve = TRUE, 
                             alpha = 0.05, 
                             global = FALSE))


save(ancomNoZ_Bac_SA, 
     file = "./05_Differential_abundance/Objects/5.2ANCOM/ancomNoZ_Bac_SA.RData")


#Fun_MeJA
ancomNoZ_Fun_MeJA <- lapply(MeJA_unnorm_Fun_ps_l, function (x)
                            ancombc(phyloseq = x, 
                             formula = "Stress", 
                             p_adj_method = "fdr", 
                             zero_cut = 0.90, 
                             lib_cut = 0, 
                             group = "Stress", 
                             struc_zero = FALSE, 
                             neg_lb = TRUE, 
                             tol = 1e-5,               
                             max_iter = 100, 
                             conserve = TRUE, 
                             alpha = 0.05, 
                             global = FALSE))
save(ancomNoZ_Fun_MeJA, 
     file = "./05_Differential_abundance/Objects/5.2ANCOM/ancomNoZ_Fun_MeJA.RData")


#Fun_SA
ancomNoZ_Fun_SA <- lapply(SA_unnorm_Fun_ps_l, function (x)
                            ancombc(phyloseq = x, 
                             formula = "Stress", 
                             p_adj_method = "fdr", 
                             zero_cut = 0.90, 
                             lib_cut = 0, 
                             group = "Stress", 
                             struc_zero = FALSE, 
                             neg_lb = TRUE, 
                             tol = 1e-5,               
                             max_iter = 100, 
                             conserve = TRUE, 
                             alpha = 0.05, 
                             global = FALSE))


save(ancomNoZ_Fun_SA, 
     file = "./05_Differential_abundance/Objects/5.2ANCOM/ancomNoZ_Fun_SA.RData")



#Organize and filter results
##MeJA_Bac
ancomNoZ_res_df_MeJA_Bac <- lapply(ancomNoZ_Bac_MeJA, function (x)
  data.frame(
  Species = row.names(x$res$beta),
  beta = unlist(x$res$beta),
  se = unlist(x$res$se),
  W = unlist(x$res$W),
  p_val = unlist(x$res$p_val),
  q_val = unlist(x$res$q_val),
  diff_abn = unlist(x$res$diff_abn)))

fdr_ancomNoZ_MeJA_Bac <- lapply(ancomNoZ_res_df_MeJA_Bac, function(x)
  (x %>%
  dplyr::filter(q_val < 0.05)))

head(fdr_ancomNoZ_MeJA_Bac)

##SA_Bac
ancomNoZ_res_df_SA_Bac <- lapply(ancomNoZ_Bac_SA, function (x)
  data.frame(
  Species = row.names(x$res$beta),
  beta = unlist(x$res$beta),
  se = unlist(x$res$se),
  W = unlist(x$res$W),
  p_val = unlist(x$res$p_val),
  q_val = unlist(x$res$q_val),
  diff_abn = unlist(x$res$diff_abn)))

fdr_ancomNoZ_SA_Bac <- lapply(ancomNoZ_res_df_SA_Bac, function(x)
  (x %>%
  dplyr::filter(q_val < 0.05)))


save(fdr_ancomNoZ_MeJA_Bac, 
     file = "./05_Differential_abundance/Objects/5.2ANCOM/fdr_ancomNoZ_MeJA_Bac.RData")
save(fdr_ancomNoZ_SA_Bac, 
     file = "./05_Differential_abundance/Objects/5.2ANCOM/fdr_ancomNoZ_SA_Bac.RData")

#Organize and filter results
##MeJA_Fun
ancomNoZ_res_df_MeJA_Fun <- lapply(ancomNoZ_Fun_MeJA, function (x)
  data.frame(
  Species = row.names(x$res$beta),
  beta = unlist(x$res$beta),
  se = unlist(x$res$se),
  W = unlist(x$res$W),
  p_val = unlist(x$res$p_val),
  q_val = unlist(x$res$q_val),
  diff_abn = unlist(x$res$diff_abn)))

fdr_ancomNoZ_MeJA_Fun <- lapply(ancomNoZ_res_df_MeJA_Fun, function(x)
  (x %>%
  dplyr::filter(q_val < 0.05)))

head(fdr_ancomNoZ_MeJA_Fun)

##SA_Fun
ancomNoZ_res_df_SA_Fun <- lapply(ancomNoZ_Fun_SA, function (x)
  data.frame(
  Species = row.names(x$res$beta),
  beta = unlist(x$res$beta),
  se = unlist(x$res$se),
  W = unlist(x$res$W),
  p_val = unlist(x$res$p_val),
  q_val = unlist(x$res$q_val),
  diff_abn = unlist(x$res$diff_abn)))

fdr_ancomNoZ_SA_Fun <- lapply(ancomNoZ_res_df_SA_Fun, function(x)
  (x %>%
  dplyr::filter(q_val < 0.05)))


save(fdr_ancomNoZ_MeJA_Fun, 
     file = "./05_Differential_abundance/Objects/5.2ANCOM/fdr_ancomNoZ_MeJA_Fun.RData")
save(fdr_ancomNoZ_SA_Fun, 
     file = "./05_Differential_abundance/Objects/5.2ANCOM/fdr_ancomNoZ_SA_Fun.RData")

```

#5.3 Summarizing results
##5.3.1 Create lists of ASVs detected per treatment/dataset
```{r}
#MeJA_Bac
##Load outputs from each test
load("./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/sigtab_stddds2_Wald_MeJA_Bac.RData")
load("./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/sigtab_zinbLRT_MeJA_Bac.RData")
load("./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/sigtab_zinbWald_MeJA_Bac.RData")
load("./05_Differential_abundance/Objects/5.2ANCOM/fdr_ancomWZ_MeJA_Bac.RData")
load("./05_Differential_abundance/Objects/5.2ANCOM/fdr_ancomNoZ_MeJA_Bac.RData")

##From each dataframe, extract the ASVIDs caught by each test
MeJA_Bac_DESeqWald_ASVs <- lapply(sigtab_stddds2_Wald_MeJA_Bac, function(x) rownames(x))
MeJA_Bac_zinbLRT_ASVs <- lapply(sigtab_zinbLRT_MeJA_Bac, function(x) rownames(x))
MeJA_Bac_zinbWald_ASVs <- lapply(sigtab_zinbWald_MeJA_Bac, function(x) rownames(x))
MeJA_Bac_ancomWZ_ASVs <- lapply(fdr_ancom_MeJA_Bac, function(x) as.vector(x[,'Species']))
MeJA_Bac_ancomNoZ_ASVs <- lapply(fdr_ancomNoZ_MeJA_Bac, function(x) as.vector(x[,'Species']))

#Within each plant concatenate the identified asvs
#MeJA_Bac
MeJA_Bac_conca <- mapply(function(t,w,x,y,z){
  input<-c(t,w,x,y,z)
  },
  t = MeJA_Bac_DESeqWald_ASVs,
  w = MeJA_Bac_zinbLRT_ASVs,
  x = MeJA_Bac_zinbWald_ASVs,
  y = MeJA_Bac_ancomWZ_ASVs,
  z = MeJA_Bac_ancomNoZ_ASVs,
  SIMPLIFY = FALSE)

save(MeJA_Bac_conca, 
     file = "./05_Differential_abundance/Objects/MeJA_Bac_conca.RData")

#Remove duplicated asvs names from that list
MeJA_Bac_conca_nodup <- lapply(MeJA_Bac_conca, function(x)
  unique(x))

save(MeJA_Bac_conca_nodup, 
     file = "./05_Differential_abundance/Objects/MeJA_Bac_conca_nodup.RData")

#Filter to ASVs detected by at least 2 tests
MeJA_Bac_conca_two <- lapply(MeJA_Bac_conca, function(x) 
  x[duplicated(x)])
MeJA_Bac_conca_two_f <- lapply(MeJA_Bac_conca_two, function(x) 
  x[!duplicated(x)])

save(MeJA_Bac_conca_two_f, 
     file = "./05_Differential_abundance/Objects/MeJA_Bac_conca_two_f.RData")



#SA_Bac
##Load outputs from each test
load("./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/sigtab_stddds2_Wald_SA_Bac.RData")
load("./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/sigtab_zinbLRT_SA_Bac.RData")
load("./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/sigtab_zinbWald_SA_Bac.RData")
load("./05_Differential_abundance/Objects/5.2ANCOM/fdr_ancomWZ_SA_Bac.RData")
load("./05_Differential_abundance/Objects/5.2ANCOM/fdr_ancomNoZ_SA_Bac.RData")

##From each dataframe, extract the ASVIDs caught by each test
SA_Bac_DESeqWald_ASVs <- lapply(sigtab_stddds2_Wald_SA_Bac, function(x) rownames(x))
SA_Bac_zinbLRT_ASVs <- lapply(sigtab_zinbLRT_SA_Bac, function(x) rownames(x))
SA_Bac_zinbWald_ASVs <- lapply(sigtab_zinbWald_SA_Bac, function(x) rownames(x))
SA_Bac_ancomWZ_ASVs <- lapply(fdr_ancom_SA_Bac, function(x) as.vector(x[,'Species']))
SA_Bac_ancomNoZ_ASVs <- lapply(fdr_ancomNoZ_SA_Bac, function(x) as.vector(x[,'Species']))

#Within each plant concatenate the identified asvs
#SA_Bac
SA_Bac_conca <- mapply(function(t,w,x,y,z){
  input<-c(t,w,x,y,z)
  },
  t = SA_Bac_DESeqWald_ASVs,
  w = SA_Bac_zinbLRT_ASVs,
  x = SA_Bac_zinbWald_ASVs,
  y = SA_Bac_ancomWZ_ASVs,
  z = SA_Bac_ancomNoZ_ASVs,
  SIMPLIFY = FALSE)

save(SA_Bac_conca, 
     file = "./05_Differential_abundance/Objects/SA_Bac_conca.RData")

#Remove duplicated asvs names from that list
SA_Bac_conca_nodup <- lapply(SA_Bac_conca, function(x)
  unique(x))

save(SA_Bac_conca_nodup, 
     file = "./05_Differential_abundance/Objects/SA_Bac_conca_nodup.RData")

#Filter to ASVs detected by at least 2 tests
SA_Bac_conca_two <- lapply(SA_Bac_conca, function(x) 
  x[duplicated(x)])
SA_Bac_conca_two_f <- lapply(SA_Bac_conca_two, function(x) 
  x[!duplicated(x)])

save(SA_Bac_conca_two_f, 
     file = "./05_Differential_abundance/Objects/SA_Bac_conca_two_f.RData")


#MeJA_Fun
##Load outputs from each test
load("./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/sigtab_stddds2_Wald_MeJA_Fun.RData")
load("./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/sigtab_zinbLRT_MeJA_Fun.RData")
load("./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/sigtab_zinbWald_MeJA_Fun.RData")
load("./05_Differential_abundance/Objects/5.2ANCOM/fdr_ancomWZ_MeJA_Fun.RData")
load("./05_Differential_abundance/Objects/5.2ANCOM/fdr_ancomNoZ_MeJA_Fun.RData")

##From each dataframe, extract the ASVIDs caught by each test
MeJA_Fun_DESeqWald_ASVs <- lapply(sigtab_stddds2_Wald_MeJA_Fun, function(x) rownames(x))
MeJA_Fun_zinbLRT_ASVs <- lapply(sigtab_zinbLRT_MeJA_Fun, function(x) rownames(x))
MeJA_Fun_zinbWald_ASVs <- lapply(sigtab_zinbWald_MeJA_Fun, function(x) rownames(x))
MeJA_Fun_ancomWZ_ASVs <- lapply(fdr_ancom_MeJA_Fun, function(x) as.vector(x[,'Species']))
MeJA_Fun_ancomNoZ_ASVs <- lapply(fdr_ancomNoZ_MeJA_Fun, function(x) as.vector(x[,'Species']))

#Within each plant concatenate the identified asvs
#MeJA_Fun
MeJA_Fun_conca <- mapply(function(t,w,x,y,z){
  input<-c(t,w,x,y,z)
  },
  t = MeJA_Fun_DESeqWald_ASVs,
  w = MeJA_Fun_zinbLRT_ASVs,
  x = MeJA_Fun_zinbWald_ASVs,
  y = MeJA_Fun_ancomWZ_ASVs,
  z = MeJA_Fun_ancomNoZ_ASVs,
  SIMPLIFY = FALSE)

save(MeJA_Fun_conca, 
     file = "./05_Differential_abundance/Objects/MeJA_Fun_conca.RData")

#Remove duplicated asvs names from that list
MeJA_Fun_conca_nodup <- lapply(MeJA_Fun_conca, function(x)
  unique(x))

save(MeJA_Fun_conca_nodup, 
     file = "./05_Differential_abundance/Objects/MeJA_Fun_conca_nodup.RData")

#Filter to ASVs detected by at least 2 tests
MeJA_Fun_conca_two <- lapply(MeJA_Fun_conca, function(x) 
  x[duplicated(x)])
MeJA_Fun_conca_two_f <- lapply(MeJA_Fun_conca_two, function(x) 
  x[!duplicated(x)])

save(MeJA_Fun_conca_two_f, 
     file = "./05_Differential_abundance/Objects/MeJA_Fun_conca_two_f.RData")

#SA_Fun
##Load outputs from each test
load("./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/sigtab_stddds2_Wald_SA_Fun.RData")
load("./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/sigtab_zinbLRT_SA_Fun.RData")
load("./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/sigtab_zinbWald_SA_Fun.RData")
load("./05_Differential_abundance/Objects/5.2ANCOM/fdr_ancomWZ_SA_Fun.RData")
load("./05_Differential_abundance/Objects/5.2ANCOM/fdr_ancomNoZ_SA_Fun.RData")

##From each dataframe, extract the ASVIDs caught by each test
SA_Fun_DESeqWald_ASVs <- lapply(sigtab_stddds2_Wald_SA_Fun, function(x) rownames(x))
SA_Fun_zinbLRT_ASVs <- lapply(sigtab_zinbLRT_SA_Fun, function(x) rownames(x))
SA_Fun_zinbWald_ASVs <- lapply(sigtab_zinbWald_SA_Fun, function(x) rownames(x))
SA_Fun_ancomWZ_ASVs <- lapply(fdr_ancom_SA_Fun, function(x) as.vector(x[,'Species']))
SA_Fun_ancomNoZ_ASVs <- lapply(fdr_ancomNoZ_SA_Fun, function(x) as.vector(x[,'Species']))

#Within each plant concatenate the identified asvs
#SA_Fun
SA_Fun_conca <- mapply(function(t,w,x,y,z){
  input<-c(t,w,x,y,z)
  },
  t = SA_Fun_DESeqWald_ASVs,
  w = SA_Fun_zinbLRT_ASVs,
  x = SA_Fun_zinbWald_ASVs,
  y = SA_Fun_ancomWZ_ASVs,
  z = SA_Fun_ancomNoZ_ASVs,
  SIMPLIFY = FALSE)

save(SA_Fun_conca, 
     file = "./05_Differential_abundance/Objects/SA_Fun_conca.RData")

#Remove duplicated asvs names from that list
SA_Fun_conca_nodup <- lapply(SA_Fun_conca, function(x)
  unique(x))

save(SA_Fun_conca_nodup, 
     file = "./05_Differential_abundance/Objects/SA_Fun_conca_nodup.RData")

#Filter to ASVs detected by at least 2 tests
SA_Fun_conca_two <- lapply(SA_Fun_conca, function(x) 
  x[duplicated(x)])
SA_Fun_conca_two_f <- lapply(SA_Fun_conca_two, function(x) 
  x[!duplicated(x)])

save(SA_Fun_conca_two_f, 
     file = "./05_Differential_abundance/Objects/SA_Fun_conca_two_f.RData")

```

##5.3.2 UpsetR plots to summarize results for all species
###5.3.2_1 For Upset plots per plant to work, identify which plants have 0 ASVs detected in 4 out of the 5 tests as this will not be plotted by upset causing the function to crash
```{r}
##Load outputs from each test
load("./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/sigtab_stddds2_Wald_MeJA_Fun.RData")
load("./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/sigtab_zinbLRT_MeJA_Fun.RData")
load("./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/sigtab_zinbWald_MeJA_Fun.RData")
load("./05_Differential_abundance/Objects/5.2ANCOM/fdr_ancomWZ_MeJA_Fun.RData")
load("./05_Differential_abundance/Objects/5.2ANCOM/fdr_ancomNoZ_MeJA_Fun.RData")

len_MeJA_Fun_DESeqWald_ASVs <- lapply(sigtab_stddds2_Wald_MeJA_Fun, function(x) length(rownames(x)))
len_MeJA_Fun_zinbLRT_ASVs <- lapply(sigtab_zinbLRT_MeJA_Fun, function(x) length(rownames(x)))
len_MeJA_Fun_zinbWald_ASVs <- lapply(sigtab_zinbWald_MeJA_Fun, function(x) length(rownames(x)))
len_MeJA_Fun_ancomWZ_ASVs <- lapply(fdr_ancom_MeJA_Fun, function(x) length(as.vector(x[,'Species'])))
len_MeJA_Fun_ancomNoZ_ASVs <- lapply(fdr_ancomNoZ_MeJA_Fun, function(x) length(as.vector(x[,'Species'])))

len_MeJA_Fun_conca_noANCOMWZ <- list(deseq=len_MeJA_Fun_DESeqWald_ASVs, zinblrt=len_MeJA_Fun_zinbLRT_ASVs, zinbwald=len_MeJA_Fun_zinbWald_ASVs, ancomNoZ=len_MeJA_Fun_ancomNoZ_ASVs)

lst <- lapply(len_MeJA_Fun_conca_noANCOMWZ, table)

out <- do.call(rbind, lapply(seq_along(lst), function(x) 
  cbind(id = x, data.frame(lst[[x]]))))

#  id At Al Cas Tg Ec Mm Bv Lr Ds Bn Sar Sal Hi Br Bo_F Bo_M Bo_S BoK Dt Si It Co Lm Ia Bi Es Sl Nb Ps Mt Zm Bd Cus Freq
#1  1  0  0   0  0  0  0  0  6  0  1   1   0  2  0    0    3    6   0  1  0  1  1  0  0 10  0  3  2  2  3 10  1   3    1
#2  2  0  0   1  0  0  0  0  0  0  0   2   0  4  1    0    2    3   1  0  0  3  1  0  0  9  0  5  8  2  0  5  0  10    1
#3  3  1  1  10  0  0  0  2  5  0  0   8  10  8  0    0    2    4   1  1  0  8  4  2  0  9  0  0 14  4  8  0  0  17    1
#4  4  0  1   0  0  0  1  0  0  0  0   2   0  1  0    2    1    6   0  0  1  0  1  1  0  0  1  0  0  0  1  0  0   0    1

#Would need to remove: Tg, Ec, Ds and Ia 

##Load outputs from each test
load("./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/sigtab_stddds2_Wald_SA_Fun.RData")
load("./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/sigtab_zinbLRT_SA_Fun.RData")
load("./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/sigtab_zinbWald_SA_Fun.RData")
load("./05_Differential_abundance/Objects/5.2ANCOM/fdr_ancomWZ_SA_Fun.RData")
load("./05_Differential_abundance/Objects/5.2ANCOM/fdr_ancomNoZ_SA_Fun.RData")

len_SA_Fun_DESeqWald_ASVs <- lapply(sigtab_stddds2_Wald_SA_Fun, function(x) length(rownames(x)))
len_SA_Fun_zinbLRT_ASVs <- lapply(sigtab_zinbLRT_SA_Fun, function(x) length(rownames(x)))
len_SA_Fun_zinbWald_ASVs <- lapply(sigtab_zinbWald_SA_Fun, function(x) length(rownames(x)))
len_SA_Fun_ancomWZ_ASVs <- lapply(fdr_ancom_SA_Fun, function(x) length(as.vector(x[,'Species'])))
len_SA_Fun_ancomNoZ_ASVs <- lapply(fdr_ancomNoZ_SA_Fun, function(x) length(as.vector(x[,'Species'])))

len_SA_Fun_conca_noANCOMWZ <- list(deseq=len_SA_Fun_DESeqWald_ASVs, zinblrt=len_SA_Fun_zinbLRT_ASVs, zinbwald=len_SA_Fun_zinbWald_ASVs, ancomNoZ=len_SA_Fun_ancomNoZ_ASVs)

lst <- lapply(len_SA_Fun_conca_noANCOMWZ, table)

out <- do.call(rbind, lapply(seq_along(lst), function(x) 
  cbind(id = x, data.frame(lst[[x]]))))

#> out
#  id At Al Cas Tg Ec Mm Bv Lr Ds Bn Sar Sal Hi Br Bo_F Bo_M Bo_S BoK Dt Si It Co Lm Ia Bi Es Sl Nb Ps Mt Zm Bd Cus Freq
#1  1  0  2   0  0  4  0  0  0  0  1   1   0  2  0    0    1   23   0  1  0  6  1  0  1  0  0  1  2  1  0  6  0   1    1
#2  2  2  1   0  0  1  1  0  4  0  0   0   0  2  0    2    3    7   1  0  1  4  2  0  1  0  0  2  3  3  2  2  0   7    1
#3  3  4  1   1  0  2  0  3 14  0  0   6   1 11  0   10    7    0   6  0  1  7 19  6  3  0  0  0  9  5  2  0  0   8    1
#4  4  0  1   0  1  1  1  0  0  0  0   1   0  2  2    1    0    0   3  1  1  5  0  0  1  0  2  0  0  1  1  0  1   1    1

#Would need to remove: Ds and Bi

```


###5.3.2_2 Upset ASVs detected by each test per species
```{r}
library(UpSetR)
#Make sure the ASVs from each test are loaded (chunk 5.3.1)

#MeJA_Bac
##use upset function. make sure to have names of the objects inside the lists!
MeJA_Bac_upset_5DAtests <- mapply(function(t,w,x,y,z){
  input<-list(t,w,x,y,z)
  names(input)<- c("Std_DESeq", # adds names to the list of tests
                   "zinb_DESeq_LRT",
                   "zinb_DESeq_Wald",
                   "ancomWZ", 
                   "ancomNoZ" )
output<-upset(fromList(input), order.by = "freq", nsets = 12, nintersects=100)
return(output)
  },
  t = MeJA_Bac_DESeqWald_ASVs,
  w = MeJA_Bac_zinbLRT_ASVs,
  x = MeJA_Bac_zinbWald_ASVs,
  y = MeJA_Bac_ancomWZ_ASVs,
  z = MeJA_Bac_ancomNoZ_ASVs, 
  SIMPLIFY = FALSE)

#Save files
save(MeJA_Bac_upset_5DAtests, 
     file = "./05_Differential_abundance/Results/UpSetRPlots/MeJA_Bac_upset_5DAtests.RData")

ggexport(MeJA_Bac_upset_5DAtests, filename = "./05_Differential_abundance/Results/UpSetRPlots/MeJA_Bac_upset_5DAtests.pdf",
 ncol = 1, nrow = 1, height=10, width = 10)

load("./05_Differential_abundance/Results/UpSetRPlots/MeJA_Bac_upset_5DAtests.RData")

#Adjusting the plot to fit bigger slides
MeJA_Bac_upset_5DAtests_2 <- mapply(function(t,w,x,y,z){
  input<-list(t,w,x,y,z)
  names(input)<- c("Std_DESeq", # adds names to the list of tests
                   "zinb_DESeq_LRT",
                   "zinb_DESeq_Wald",
                   "ancomWZ", 
                   "ancomNoZ" )
output<-upset(fromList(input), order.by = "freq", nsets = 12, nintersects=100, text.scale = 3, set_size.angles = 0,
  set_size.show = FALSE, set_size.numbers_size = NULL)
return(output)
  },
  t = MeJA_Bac_DESeqWald_ASVs,
  w = MeJA_Bac_zinbLRT_ASVs,
  x = MeJA_Bac_zinbWald_ASVs,
  y = MeJA_Bac_ancomWZ_ASVs,
  z = MeJA_Bac_ancomNoZ_ASVs, 
  SIMPLIFY = FALSE)

MeJA_Bac_upset_5DAtests_2$At
MeJA_Bac_upset_5DAtests_2$Bo_M

ggexport(MeJA_Bac_upset_5DAtests_2, filename = "./05_Differential_abundance/Results/UpSetRPlots/MeJA_Bac_upset_5DAtests_2.pdf",
 ncol = 1, nrow = 1, height=10, width = 10)

#SA_Bac
##use upset function. make sure to have names of the objects inside the lists!
SA_Bac_upset_5DAtests <- mapply(function(t,w,x,y,z){
  input<-list(t,w,x,y,z)
  names(input)<- c("Std_DESeq", # adds names to the list of tests
                   "zinb_DESeq_LRT",
                   "zinb_DESeq_Wald",
                   "ancomWZ", 
                   "ancomNoZ" )
output<-upset(fromList(input), order.by = "freq", nsets = 12, nintersects=100)
return(output)
  },
  t = SA_Bac_DESeqWald_ASVs,
  w = SA_Bac_zinbLRT_ASVs,
  x = SA_Bac_zinbWald_ASVs,
  y = SA_Bac_ancomWZ_ASVs,
  z = SA_Bac_ancomNoZ_ASVs, 
  SIMPLIFY = FALSE)

#Save files
save(SA_Bac_upset_5DAtests, 
     file = "./05_Differential_abundance/Results/UpSetRPlots/SA_Bac_upset_5DAtests.RData")

ggexport(SA_Bac_upset_5DAtests, filename = "./05_Differential_abundance/Results/UpSetRPlots/SA_Bac_upset_5DAtests.pdf",
 ncol = 1, nrow = 1, height=10, width = 10)

#MeJA_Fun
#Remove plants with 0 detected ASVs in 4 out of 5 tests (identified in chunk 5.3.2_1)
plants29_MeJA_Fun_DESeqWald_ASVs = within(MeJA_Fun_DESeqWald_ASVs, rm("Ia", "Ec", "Tg", "Ds"))
plants29_MeJA_Fun_zinbLRT_ASVs = within(MeJA_Fun_zinbLRT_ASVs, rm("Ia", "Ec", "Tg", "Ds"))
plants29_MeJA_Fun_zinbWald_ASVs = within(MeJA_Fun_zinbWald_ASVs, rm("Ia", "Ec", "Tg", "Ds"))
plants29_MeJA_Fun_ancomWZ_ASVs = within(MeJA_Fun_ancomWZ_ASVs, rm("Ia", "Ec", "Tg", "Ds"))
plants29_MeJA_Fun_ancomNoZ_ASVs = within(MeJA_Fun_ancomNoZ_ASVs, rm("Ia", "Ec", "Tg", "Ds"))
  
##use upset function. make sure to have names of the objects inside the lists!  
MeJA_Fun_upset_5DAtests <- mapply(function(t,w,x,y,z){
  input<-list(t,w,x,y,z)
  names(input)<- c("Std_DESeq", # adds names to the list of tests
                   "zinb_DESeq_LRT",
                   "zinb_DESeq_Wald",
                   "ancomWZ", 
                   "ancomNoZ" )
output<-upset(fromList(input), order.by = "freq", nsets = 12, nintersects=100)
return(output)
  },
  t = plants29_MeJA_Fun_DESeqWald_ASVs,
  w = plants29_MeJA_Fun_zinbLRT_ASVs,
  x = plants29_MeJA_Fun_zinbWald_ASVs,
  y = plants29_MeJA_Fun_ancomWZ_ASVs,
  z = plants29_MeJA_Fun_ancomNoZ_ASVs, 
  SIMPLIFY = FALSE)

#Save files
save(MeJA_Fun_upset_5DAtests, 
     file = "./05_Differential_abundance/Results/UpSetRPlots/MeJA_Fun_upset_5DAtests.RData")

ggexport(MeJA_Fun_upset_5DAtests, filename = "./05_Differential_abundance/Results/UpSetRPlots/MeJA_Fun_upset_5DAtests.pdf",
 ncol = 1, nrow = 1, height=10, width = 10)



#SA_Fun

#Remove plants with 0 detected ASVs in 4 out of 5 tests (identified in chunk 5.3.2_1)
plants31_SA_Fun_DESeqWald_ASVs = within(SA_Fun_DESeqWald_ASVs, rm("Bi", "Ds"))
plants31_SA_Fun_zinbLRT_ASVs = within(SA_Fun_zinbLRT_ASVs, rm("Bi", "Ds"))
plants31_SA_Fun_zinbWald_ASVs = within(SA_Fun_zinbWald_ASVs, rm("Bi", "Ds"))
plants31_SA_Fun_ancomWZ_ASVs = within(SA_Fun_ancomWZ_ASVs, rm("Bi", "Ds"))
plants31_SA_Fun_ancomNoZ_ASVs = within(SA_Fun_ancomNoZ_ASVs, rm("Bi", "Ds"))

##use upset function. make sure to have names of the objects inside the lists!
SA_Fun_upset_5DAtests <- mapply(function(t,w,x,y,z){
  input<-list(t,w,x,y,z)
  names(input)<- c("Std_DESeq", # adds names to the list of tests
                   "zinb_DESeq_LRT",
                   "zinb_DESeq_Wald",
                   "ancomWZ", 
                   "ancomNoZ" )
output<-upset(fromList(input), order.by = "freq", nsets = 12, nintersects=100)
return(output)
  },
  t = plants31_SA_Fun_DESeqWald_ASVs,
  w = plants31_SA_Fun_zinbLRT_ASVs,
  x = plants31_SA_Fun_zinbWald_ASVs,
  y = plants31_SA_Fun_ancomWZ_ASVs,
  z = plants31_SA_Fun_ancomNoZ_ASVs, 
  SIMPLIFY = FALSE)

#Save files
save(SA_Fun_upset_5DAtests, 
     file = "./05_Differential_abundance/Results/UpSetRPlots/SA_Fun_upset_5DAtests.RData")

ggexport(SA_Fun_upset_5DAtests, filename = "./05_Differential_abundance/Results/UpSetRPlots/SA_Fun_upset_5DAtests.pdf",
 ncol = 1, nrow = 1, height=10, width = 10)


```

#5.3.3 Dfs comparing ASVs detected by x amount of tests
```{r}
library("tuple")

#df with the amount of ASVs detected by at least one test and those detected by 2, 3 or more tests

#MeJA_Bac
#First amount of ASVs detected by at least one test
load("./05_Differential_abundance/Objects/MeJA_Bac_conca_nodup.RData")
MeJA_Bac_conca_nodup_num <- lapply(MeJA_Bac_conca_nodup, function(x) length(x))
MeJA_Bac_conca_nodup_df <- as.data.frame(MeJA_Bac_conca_nodup_num)

#Detected by at least two tests
load("./05_Differential_abundance/Objects/MeJA_Bac_conca.RData")
MeJA_Bac_conca_two <- lapply(MeJA_Bac_conca, function(x) 
  x[duplicated(x)])
MeJA_Bac_conca_two_f <- lapply(MeJA_Bac_conca_two, function(x) 
  x[!duplicated(x)])
MeJA_Bac_conca_two_f_num <- lapply(MeJA_Bac_conca_two_f, function(x) length(x))
MeJA_Bac_conca_two_df <- as.data.frame(MeJA_Bac_conca_two_f_num)

#Three tests
#Keep only asvs that are at least three times 
MeJA_Bac_conca_three <- lapply(MeJA_Bac_conca, function(x)
  x[triplicated(x)])
#vectors that appear more than 3 times may be duplicated inthe new object, so make sure all vectors are unique
MeJA_Bac_conca_three_f <- lapply(MeJA_Bac_conca_three, function(x)
  x[!duplicated(x)])
MeJA_Bac_conca_three_f_num <- lapply(MeJA_Bac_conca_three_f, function(x) length(x))
MeJA_Bac_conca_three_df <- as.data.frame(MeJA_Bac_conca_three_f_num)

MeJA_Bac_detected_listed <- list(all = MeJA_Bac_conca_nodup_df, two_at_least = MeJA_Bac_conca_two_df, three_at_least = MeJA_Bac_conca_three_df)
MeJA_Bac_DA_detected <- merge((t(MeJA_Bac_conca_nodup_df)), (t(MeJA_Bac_conca_two_df)), by = 0, all = T)
rownames(MeJA_Bac_DA_detected) <- MeJA_Bac_DA_detected$Row.names; MeJA_Bac_DA_detected$Row.names <- NULL
MeJA_Bac_DA_detected_final <- merge(MeJA_Bac_DA_detected, (t(MeJA_Bac_conca_three_df)), by = "row.names", all = T)
rownames(MeJA_Bac_DA_detected_final) <- MeJA_Bac_DA_detected_final$Row.names; MeJA_Bac_DA_detected_final$Row.names <- NULL
colnames(MeJA_Bac_DA_detected_final) <- c("All_detected_MeJA_Bac", "two_tests_or_more", "three_tests_or_more")

#Save table
write.table(MeJA_Bac_DA_detected_final, file="./05_Differential_abundance/Results/MeJA_Bac_DA_detected_final.txt",
sep = "\t", quote = F, row.names = T)



#SA_Bac
#First amount of ASVs detected by at least one test
load("./05_Differential_abundance/Objects/SA_Bac_conca_nodup.RData")
SA_Bac_conca_nodup_num <- lapply(SA_Bac_conca_nodup, function(x) length(x))
SA_Bac_conca_nodup_df <- as.data.frame(SA_Bac_conca_nodup_num)

#Detected by at least two tests
load("./05_Differential_abundance/Objects/SA_Bac_conca.RData")
SA_Bac_conca_two <- lapply(SA_Bac_conca, function(x) 
  x[duplicated(x)])
SA_Bac_conca_two_f <- lapply(SA_Bac_conca_two, function(x) 
  x[!duplicated(x)])
SA_Bac_conca_two_f_num <- lapply(SA_Bac_conca_two_f, function(x) length(x))
SA_Bac_conca_two_df <- as.data.frame(SA_Bac_conca_two_f_num)

#Three tests
#Keep only asvs that are at least three times 
SA_Bac_conca_three <- lapply(SA_Bac_conca, function(x)
  x[triplicated(x)])
#vectors that appear more than 3 times may be duplicated inthe new object, so make sure all vectors are unique
SA_Bac_conca_three_f <- lapply(SA_Bac_conca_three, function(x)
  x[!duplicated(x)])
SA_Bac_conca_three_f_num <- lapply(SA_Bac_conca_three_f, function(x) length(x))
SA_Bac_conca_three_df <- as.data.frame(SA_Bac_conca_three_f_num)

SA_Bac_detected_listed <- list(all = SA_Bac_conca_nodup_df, two_at_least = SA_Bac_conca_two_df, three_at_least = SA_Bac_conca_three_df)
SA_Bac_DA_detected <- merge((t(SA_Bac_conca_nodup_df)), (t(SA_Bac_conca_two_df)), by = 0, all = T)
rownames(SA_Bac_DA_detected) <- SA_Bac_DA_detected$Row.names; SA_Bac_DA_detected$Row.names <- NULL
SA_Bac_DA_detected_final <- merge(SA_Bac_DA_detected, (t(SA_Bac_conca_three_df)), by = "row.names", all = T)
rownames(SA_Bac_DA_detected_final) <- SA_Bac_DA_detected_final$Row.names; SA_Bac_DA_detected_final$Row.names <- NULL
colnames(SA_Bac_DA_detected_final) <- c("All_detected_SA_Bac", "two_tests_or_more", "three_tests_or_more")

#Save table
write.table(SA_Bac_DA_detected_final, file="./05_Differential_abundance/Results/SA_Bac_DA_detected_final.txt",
sep = "\t", quote = F, row.names = T)


#MeJA_Fun
#First amount of ASVs detected by at least one test
load("./05_Differential_abundance/Objects/MeJA_Fun_conca_nodup.RData")
MeJA_Fun_conca_nodup_num <- lapply(MeJA_Fun_conca_nodup, function(x) length(x))
MeJA_Fun_conca_nodup_df <- as.data.frame(MeJA_Fun_conca_nodup_num)

#Detected by at least two tests
load("./05_Differential_abundance/Objects/MeJA_Fun_conca.RData")
MeJA_Fun_conca_two <- lapply(MeJA_Fun_conca, function(x) 
  x[duplicated(x)])
MeJA_Fun_conca_two_f <- lapply(MeJA_Fun_conca_two, function(x) 
  x[!duplicated(x)])
MeJA_Fun_conca_two_f_num <- lapply(MeJA_Fun_conca_two_f, function(x) length(x))
MeJA_Fun_conca_two_df <- as.data.frame(MeJA_Fun_conca_two_f_num)

#Three tests
#Keep only asvs that are at least three times 
MeJA_Fun_conca_three <- lapply(MeJA_Fun_conca, function(x)
  x[triplicated(x)])
#vectors that appear more than 3 times may be duplicated inthe new object, so make sure all vectors are unique
MeJA_Fun_conca_three_f <- lapply(MeJA_Fun_conca_three, function(x)
  x[!duplicated(x)])
MeJA_Fun_conca_three_f_num <- lapply(MeJA_Fun_conca_three_f, function(x) length(x))
MeJA_Fun_conca_three_df <- as.data.frame(MeJA_Fun_conca_three_f_num)

MeJA_Fun_detected_listed <- list(all = MeJA_Fun_conca_nodup_df, two_at_least = MeJA_Fun_conca_two_df, three_at_least = MeJA_Fun_conca_three_df)
MeJA_Fun_DA_detected <- merge((t(MeJA_Fun_conca_nodup_df)), (t(MeJA_Fun_conca_two_df)), by = 0, all = T)
rownames(MeJA_Fun_DA_detected) <- MeJA_Fun_DA_detected$Row.names; MeJA_Fun_DA_detected$Row.names <- NULL
MeJA_Fun_DA_detected_final <- merge(MeJA_Fun_DA_detected, (t(MeJA_Fun_conca_three_df)), by = "row.names", all = T)
rownames(MeJA_Fun_DA_detected_final) <- MeJA_Fun_DA_detected_final$Row.names; MeJA_Fun_DA_detected_final$Row.names <- NULL
colnames(MeJA_Fun_DA_detected_final) <- c("All_detected_MeJA_Fun", "two_tests_or_more", "three_tests_or_more")

#Save table
write.table(MeJA_Fun_DA_detected_final, file="./05_Differential_abundance/Results/MeJA_Fun_DA_detected_final.txt",
sep = "\t", quote = F, row.names = T)




#SA_Fun
#First amount of ASVs detected by at least one test
load("./05_Differential_abundance/Objects/SA_Fun_conca_nodup.RData")
SA_Fun_conca_nodup_num <- lapply(SA_Fun_conca_nodup, function(x) length(x))
SA_Fun_conca_nodup_df <- as.data.frame(SA_Fun_conca_nodup_num)

#Detected by at least two tests
load("./05_Differential_abundance/Objects/SA_Fun_conca.RData")
SA_Fun_conca_two <- lapply(SA_Fun_conca, function(x) 
  x[duplicated(x)])
SA_Fun_conca_two_f <- lapply(SA_Fun_conca_two, function(x) 
  x[!duplicated(x)])
SA_Fun_conca_two_f_num <- lapply(SA_Fun_conca_two_f, function(x) length(x))
SA_Fun_conca_two_df <- as.data.frame(SA_Fun_conca_two_f_num)

#Three tests
#Keep only asvs that are at least three times 
SA_Fun_conca_three <- lapply(SA_Fun_conca, function(x)
  x[triplicated(x)])
#vectors that appear more than 3 times may be duplicated inthe new object, so make sure all vectors are unique
SA_Fun_conca_three_f <- lapply(SA_Fun_conca_three, function(x)
  x[!duplicated(x)])
SA_Fun_conca_three_f_num <- lapply(SA_Fun_conca_three_f, function(x) length(x))
SA_Fun_conca_three_df <- as.data.frame(SA_Fun_conca_three_f_num)

SA_Fun_detected_listed <- list(all = SA_Fun_conca_nodup_df, two_at_least = SA_Fun_conca_two_df, three_at_least = SA_Fun_conca_three_df)
SA_Fun_DA_detected <- merge((t(SA_Fun_conca_nodup_df)), (t(SA_Fun_conca_two_df)), by = 0, all = T)
rownames(SA_Fun_DA_detected) <- SA_Fun_DA_detected$Row.names; SA_Fun_DA_detected$Row.names <- NULL
SA_Fun_DA_detected_final <- merge(SA_Fun_DA_detected, (t(SA_Fun_conca_three_df)), by = "row.names", all = T)
rownames(SA_Fun_DA_detected_final) <- SA_Fun_DA_detected_final$Row.names; SA_Fun_DA_detected_final$Row.names <- NULL
colnames(SA_Fun_DA_detected_final) <- c("All_detected_SA_Fun", "two_tests_or_more", "three_tests_or_more")

#Save table
write.table(SA_Fun_DA_detected_final, file="./05_Differential_abundance/Results/SA_Fun_DA_detected_final.txt",
sep = "\t", quote = F, row.names = T)


```



#Comparing accross plants

###Upset comparing ASVs across BRASSICACEAE
```{r}
library(UpSetR)
#MeJA_Bac
#upset plot
load("./05_Differential_abundance/Objects/MeJA_Bac_conca_two_f.RData")

#Detected by two tests UpSet plot
MeJA_Bac_allsp_twotest_all <- upset(fromList(MeJA_Bac_conca_two_f), nsets = 50, nintersects = NA, order.by = "freq")
save(MeJA_Bac_allsp_twotest_all, 
     file = "./05_Differential_abundance/Results/UpSetRPlots/MeJA_Bac_allsp_twotest_all.RData")

```

#Add taxonomy to the list of detected ASVs so you can make tests at different taxonomic levels
```{r}
#MeJA_Bac
#Filter phyloseq objects to contain only ASVs detected by at least one DA test
#load the list with these ASVs per plant species and the phyloseq object (already subsetted to MeJA_Bac and listed per plant species)
load("./05_Differential_abundance/Objects/MeJA_Bac_conca_two_f.RData")
load("./05_Differential_abundance/Objects/MeJA_unnorm_Bac_ps_l.RData")

#Check MeJA_Bac_conca_two_f and identify plants that have 0 ASVs detected, these need to be removed from both objects for the functions to run

#In this case these are: Al and Sar

#Make copies of the objects to then remove plants with 0 or 1 asvs detected by 2 tests
MeJA_Bac_conca_two_f_31plants <- MeJA_Bac_conca_two_f
MeJA_unnorm_Bac_ps_l_31plants <- MeJA_unnorm_Bac_ps_l

#Remove plants with 0 ASVs detected by two tests
#Remove Lm, Al and Sar because they have 0 or 1 ASVs detected 
MeJA_Bac_conca_two_f_31plants$Al <- NULL
MeJA_Bac_conca_two_f_31plants$Sar <- NULL
#Also from the phyloseq object
MeJA_unnorm_Bac_ps_l_31plants$Al <- NULL
MeJA_unnorm_Bac_ps_l_31plants$Sar <- NULL

#Subset to those lists
twotests_MeJA_Bac_ps_unnorm_31plants <-mapply (function (list_1,list_2) 
            prune_taxa(taxa = list_1, x = list_2), #here, x as the name of an argument of prune_taxa, and refers to a phyloseq object
          list_1 = MeJA_Bac_conca_two_f_31plants, # List of ASVs to keep on the phyloseq object
          list_2 = MeJA_unnorm_Bac_ps_l_31plants, #phyloseq object
          SIMPLIFY = FALSE)

save(twotests_MeJA_Bac_ps_unnorm_31plants, 
     file = "./05_Differential_abundance/Objects/twotests_MeJA_Bac_ps_unnorm_31plants.RData")
save(MeJA_unnorm_Bac_ps_l_31plants, 
     file = "./05_Differential_abundance/Objects/MeJA_unnorm_Bac_ps_l_31plants.RData")



#SA_Bac
#Filter phyloseq objects to contain only ASVs detected by at least one DA test
#load the list with these ASVs per plant species and the phyloseq object (already subsetted to SA_Bac and listed per plant species)
load("./05_Differential_abundance/Objects/SA_Bac_conca_two_f.RData")
load("./05_Differential_abundance/Objects/SA_unnorm_Bac_ps_l.RData")


#Check SA_Bac_conca_two_f and identify plants that have 0 ASVs detected, these need to be removed from both objects for the functions to run

#In this case this is only Al

#Make copies of the objects to then remove plants with 0 or 1 asvs detected by 2 tests
SA_Bac_conca_two_f_32plants <- SA_Bac_conca_two_f
SA_unnorm_Bac_ps_l_32plants <- SA_unnorm_Bac_ps_l

#Remove plants with 0 ASVs detected by two tests
SA_Bac_conca_two_f_32plants$Al <- NULL
#Also from the phyloseq object
SA_unnorm_Bac_ps_l_32plants$Al <- NULL

#Subset to those lists
twotests_SA_Bac_ps_unnorm_32plants <-mapply (function (list_1,list_2) 
            prune_taxa(taxa = list_1, x = list_2), #here, x as the name of an argument of prune_taxa, and refers to a phyloseq object
          list_1 = SA_Bac_conca_two_f_32plants, # List of ASVs to keep on the phyloseq object
          list_2 = SA_unnorm_Bac_ps_l_32plants, #phyloseq object
          SIMPLIFY = FALSE)

save(twotests_SA_Bac_ps_unnorm_32plants, 
     file = "./05_Differential_abundance/Objects/twotests_SA_Bac_ps_unnorm_32plants.RData")
save(SA_unnorm_Bac_ps_l_32plants, 
     file = "./05_Differential_abundance/Objects/SA_unnorm_Bac_ps_l_32plants.RData")

#MeJA_Fun
#Filter phyloseq objects to contain only ASVs detected by at least one DA test
#load the list with these ASVs per plant species and the phyloseq object (already subsetted to MeJA_Fun and listed per plant species)
load("./05_Differential_abundance/Objects/MeJA_Fun_conca_two_f.RData")
load("./05_Differential_abundance/Objects/MeJA_unnorm_Fun_ps_l.RData")

#Check MeJA_Fun_conca_two_f and identify plants that have 0 ASVs detected, these need to be removed from both objects for the functions to run

#In this case these are: Tg, Ec, Ds, Bn, Ia

#Make copies of the objects to then remove plants with 0 or 1 asvs detected by 2 tests
MeJA_Fun_conca_two_f_28plants <- MeJA_Fun_conca_two_f
MeJA_unnorm_Fun_ps_l_28plants <- MeJA_unnorm_Fun_ps_l

#Remove plants with 0 ASVs detected by two tests
MeJA_Fun_conca_two_f_28plants$Tg <- NULL
MeJA_Fun_conca_two_f_28plants$Ec <- NULL
MeJA_Fun_conca_two_f_28plants$Ds <- NULL
MeJA_Fun_conca_two_f_28plants$Bn <- NULL
MeJA_Fun_conca_two_f_28plants$Ia <- NULL
#Also from the phyloseq object
MeJA_unnorm_Fun_ps_l_28plants$Tg <- NULL
MeJA_unnorm_Fun_ps_l_28plants$Ec <- NULL
MeJA_unnorm_Fun_ps_l_28plants$Ds <- NULL
MeJA_unnorm_Fun_ps_l_28plants$Bn <- NULL
MeJA_unnorm_Fun_ps_l_28plants$Ia <- NULL

#Subset to those lists
twotests_MeJA_Fun_ps_unnorm_28plants <-mapply (function (list_1,list_2) 
            prune_taxa(taxa = list_1, x = list_2), #here, x as the name of an argument of prune_taxa, and refers to a phyloseq object
          list_1 = MeJA_Fun_conca_two_f_28plants, # List of ASVs to keep on the phyloseq object
          list_2 = MeJA_unnorm_Fun_ps_l_28plants, #phyloseq object
          SIMPLIFY = FALSE)

save(twotests_MeJA_Fun_ps_unnorm_28plants, 
     file = "./05_Differential_abundance/Objects/twotests_MeJA_Fun_ps_unnorm_28plants.RData")
save(MeJA_unnorm_Fun_ps_l_28plants, 
     file = "./05_Differential_abundance/Objects/MeJA_unnorm_Fun_ps_l_28plants.RData")



#SA_Fun
#Filter phyloseq objects to contain only ASVs detected by at least one DA test
#load the list with these ASVs per plant species and the phyloseq object (already subsetted to SA_Fun and listed per plant species)
load("./05_Differential_abundance/Objects/SA_Fun_conca_two_f.RData")
load("./05_Differential_abundance/Objects/SA_unnorm_Fun_ps_l.RData")


#Check SA_Fun_conca_two_f and identify plants that have 0 ASVs detected, these need to be removed from both objects for the functions to run

#In this case these are: Ds, Bn, Sal and Bi

#Make copies of the objects to then remove plants with 0 or 1 asvs detected by 2 tests
SA_Fun_conca_two_f_29plants <- SA_Fun_conca_two_f
SA_unnorm_Fun_ps_l_29plants <- SA_unnorm_Fun_ps_l

#Remove plants with 0 ASVs detected by two tests
SA_Fun_conca_two_f_29plants$Sal <- NULL
SA_Fun_conca_two_f_29plants$Bi <- NULL
SA_Fun_conca_two_f_29plants$Ds <- NULL
SA_Fun_conca_two_f_29plants$Bn <- NULL
#Also from the phyloseq object
SA_unnorm_Fun_ps_l_29plants$Sal <- NULL
SA_unnorm_Fun_ps_l_29plants$Bi <- NULL
SA_unnorm_Fun_ps_l_29plants$Ds <- NULL
SA_unnorm_Fun_ps_l_29plants$Bn <- NULL


#Subset to those lists
twotests_SA_Fun_ps_unnorm_29plants <-mapply (function (list_1,list_2) 
            prune_taxa(taxa = list_1, x = list_2), #here, x as the name of an argument of prune_taxa, and refers to a phyloseq object
          list_1 = SA_Fun_conca_two_f_29plants, # List of ASVs to keep on the phyloseq object
          list_2 = SA_unnorm_Fun_ps_l_29plants, #phyloseq object
          SIMPLIFY = FALSE)

save(twotests_SA_Fun_ps_unnorm_29plants, 
     file = "./05_Differential_abundance/Objects/twotests_SA_Fun_ps_unnorm_29plants.RData")
save(SA_unnorm_Fun_ps_l_29plants, 
     file = "./05_Differential_abundance/Objects/SA_unnorm_Fun_ps_l_29plants.RData")

```


###UpSet comparing shared accross plant species at family level
```{r}
#Agglomerate to Family level
twotests_MeJA_Bac_ps_unnorm_Family <- lapply(twotests_MeJA_Bac_ps_unnorm, function(x)
  tax_glom(x, taxrank = "Family"))

save(twotests_MeJA_Bac_ps_unnorm_Family, 
     file = "./05_Differential_abundance/Objects/twotests_MeJA_Bac_ps_unnorm_Family.RData")

#Turn families into vectors
##Make tax_table a df
twotests_MeJA_Bac_ps_unnorm_Family_df <- lapply(twotests_MeJA_Bac_ps_unnorm_Family, function(x) as.data.frame(tax_table(x)))
##Subset that df to only keep column family
twotests_MeJA_Bac_ps_unnorm_Family_vector <- lapply(twotests_MeJA_Bac_ps_unnorm_Family_df, function(x) as.vector(x$Family))
#Remove duplicated families inside each plant list
twotests_MeJA_Bac_ps_unnorm_Family_vector_nodup <- lapply(twotests_MeJA_Bac_ps_unnorm_Family_vector, function(x) 
  x[!duplicated(x)])

#Upset plot
upset_Family_twotests_MeJA_Bac_allsp <- upset(fromList(twotests_MeJA_Bac_ps_unnorm_Family_vector_nodup), nsets = 50, nintersects = NA, order.by = "freq")

pdf(file = "./05_Differential_abundance/Results/UpSetRPlots/upset_Family_twotests_MeJA_Bac_allsp.pdf",   # The directory you want to save the file in
    width = 17, # The width of the plot in inches
    height = 10)
upset_Family_twotests_MeJA_Bac_allsp
dev.off()

#Quantify Shared and unique
twotests_MeJA_Bac_ps_unnorm_Family_vector2_nodup <- stack(twotests_MeJA_Bac_ps_unnorm_Family_vector_nodup)$values 
Family_twotests_MeJA_Bac_shared2plants <- 
  twotests_MeJA_Bac_ps_unnorm_Family_vector2_nodup[duplicated(twotests_MeJA_Bac_ps_unnorm_Family_vector2_nodup)]
Family_twotests_MeJA_Bac_shared2plants_f <-
  Family_twotests_MeJA_Bac_shared2plants[!duplicated(Family_twotests_MeJA_Bac_shared2plants)]
#number of shared families accross plant species
numbershared <- length(Family_twotests_MeJA_Bac_shared2plants_f) #55

#Unique
Family_twotests_MeJA_Bac_unique <- unique(twotests_MeJA_Bac_ps_unnorm_Family_vector2_nodup) #99
#Remove the shared bacterial families by getting the intersection
Family_twotests_MeJA_Bac_unique_rmshared <- setdiff(Family_twotests_MeJA_Bac_unique, Family_twotests_MeJA_Bac_shared2plants_f) #number of unique families accross plant species
numberunique <- length(Family_twotests_MeJA_Bac_unique_rmshared) #44

```


#Example of how it works to Keep ASVs detected by two or three tests using the ASV names as vectors and the functions duplicated and triplicated
```{r}
library("tuple")

#2023-03-17
fake_asvs <- c("bASVlala", "bASV_14")
fake_asvs2 <- c("bASV_14", "bASVlala")
fake_asvs3 <- c("bASVlala")
fake_asvs4 <- c("bASVlala")
fake_asvs5 <- c("bASVnini", "bASVtete", "bASVlili")
fake_asvs6 <- c("bASVnini")

fake_conca <- c(fake_asvs, fake_asvs2, fake_asvs3, fake_asvs4, fake_asvs5, fake_asvs6)
dupli_keep <- fake_conca[duplicated(fake_conca)]
#[1] "bASV_14"  "bASVlala" "bASVlala" "bASVlala" "bASVnini"
dupli_keep_unique <- dupli_keep[!duplicated(dupli_keep)]
#[1] "bASV_14"  "bASVlala" "bASVnini"

#for 3 or more use https://rdrr.io/cran/tuple/man/triplicate.html
tripli_keep <- fake_conca[triplicated(fake_conca)]
#[1] "bASVlala" "bASVlala"
tripli_keep_unique <- tripli_keep[!duplicated(tripli_keep)]
#[1] "bASVlala"
```


#Metacoder with Pedro
```{r}
#Following Pedro's workshop https://github.com/PedroBeschoren/metacoder_workshop/blob/master/code/metacoder_workshop.Rmd

#MeJA_Bac
#Filter phyloseq objects to contain only ASVs detected by at least one DA test
#load the list with these ASVs per plant species and the phyloseq object (already subsetted to MeJA_Bac and listed per plant species)
load("./05_Differential_abundance/Objects/MeJA_Bac_conca_two_f.RData")
load("./05_Differential_abundance/Objects/MeJA_unnorm_Bac_ps_l_unfilt.RData")

# filter your list of phyloliseq objects to only contain the ASVs defined as "importat"
MeJA_Bac_conca_twotests_ps<-mapply (function (list_1,list_2) #mapply will let you manipulate 2 or more lists at once
            prune_taxa(taxa = list_1, x = list_2), #here, x as the name of an argument of prune_taxa, and refers to a phyloseq object
          list_1 = MeJA_Bac_conca_two_f, # here you define what R object is list_1
          list_2 = MeJA_unnorm_Bac_ps_l_unfilt, # here you define what R object is list_2
          SIMPLIFY = FALSE)

MeJA_Bac_conca_ps$At@tax_table


#Subset taxonomy to Family level
MeJA_Bac_conca_Family <- lapply(MeJA_Bac_conca_ps, function(x) {
  tax_table(x)<-tax_table(x)[,2:6]
  return(x)
  })

#Genus
MeJA_Bac_conca_genus <- lapply(MeJA_Bac_conca_ps, function(x) {
  tax_table(x)<-tax_table(x)[,2:7]
  return(x)
  })

# transform from phyloseq to taxmap object
#taxmap_MeJA_Bac_conca_Family <-lapply(MeJA_Bac_conca_Family[1:3], function(x)
#                                      parse_phyloseq(x))






```


###Metacoder with Pedro: define custom function to create heat trees from ps objects, cpopied from chunk 3.8.1 - Define a custom function of the metacoder workshop from pedro]
```{r}
# define funciton to quickly plot a heat tree from a phtloseq object
phyloseq_to_heat_tree_matrix<-function(ps_object, sample_group){
  
  # this function output is a heat trees comparing metadata. the input is based on a phyloseq object
  
  # ps object =  a phyloseq object, containing sample metadata, otu table, and taxonomy table
  # sample_group = the name of a column in your emtadata that you want to compare in the heat tree. it has to be quoted.
  
  # this function will create a matrix of heat trees if your metadata has more than 2 groups. it should fail if it has only 1 group
  

#remove unecessary taxonomic info (dada2id, "s__" and "ASV_id) by updating the tax table with a subset of the tax table
#tax_table(ps_object)<-tax_table(ps_object)[,1:6]


# let's remove the "r__"ranks from the taxonomy, they can be useful but will pollute our plot
tax_table(ps_object)[, colnames(tax_table(ps_object))] <- gsub(pattern = "[a-z]__", # regular expression pattern to search for
                                                                   x = tax_table(ps_object)[, colnames(tax_table(ps_object))], # "df"
                                                                   replacement = "") # replacement for pattern
# transform from phyloseq to taxmap object
taxmap_obj<-parse_phyloseq(ps_object)

#get abundance per taxon
taxmap_obj$data$tax_abund<-calc_taxon_abund(obj = taxmap_obj, 
                                      data = "otu_table",
                                      cols = taxmap_obj$data$sample_data$sample_id) 

#get occurrence of ASVs per treatment
# the sample groups needs some wrangling to fit within the soft code of the function
taxmap_obj$data$tax_occ<- calc_n_samples(obj = taxmap_obj, 
                                                      data = "tax_abund", 
                                                      cols = taxmap_obj$data$sample_data$sample_id,
                                                      groups = taxmap_obj$data$sample_data[colnames(taxmap_obj$data$sample_data)==sample_group][[1]]) 

# calculate log2 median ratios and p values for a wilcoxcon test within taxas in this stress treatment groups
# the sample groups needs some wrangling to fit within the soft code of the function
taxmap_obj$data$diff_table <- compare_groups(obj = taxmap_obj,
                                                        data = "tax_abund",
                                                        cols = taxmap_obj$data$sample_data$sample_id, 
                                                        groups = taxmap_obj$data$sample_data[colnames(taxmap_obj$data$sample_data)==sample_group][[1]]) 

# set differential log ratio to 0 based on adjusted p values
taxmap_obj$data$diff_table$log2_median_ratio[taxmap_obj$data$diff_table$wilcox_p_value > 0.05] <- 0

# define number of compared factors
factors_compared<-taxmap_obj$data$sample_data[colnames(taxmap_obj$data$sample_data)==sample_group][[1]] 

# draw the plot based on an if else statement: if there are 2 groups, plot a a heat tree comparing abundances between both groups, else plot a matrix of ehat trees. this function will fail if you only have 1 sample group! 

if (length(unique(factors_compared)) == 2) {

set.seed(1)
 output<- taxmap_obj %>%
heat_tree(
            node_label = taxon_names,
      #      data = "diff_table", 
            node_size = n_obs,
            node_color = log2_median_ratio,
            node_color_interval = c(-3, 3), # The range of `log2_median_ratio` to display
            node_color_range = c("cyan", "gray", "tan"), # The color palette used
            layout = "davidson-harel",
            initial_layout = "reingold-tilford")

 } else {

set.seed(1)
output<-heat_tree_matrix(taxmap_obj,
                         data = "diff_table", # this is the table with the data you want to plot
                         node_size = n_obs, # n_obs is a function that calculates, in this case, the number of OTUs per taxon
                         node_label = taxon_names,
                         node_color = log2_median_ratio, # A column from `taxmap_obj$data$diff_table_3treatments`
                         node_color_range = diverging_palette(), # The built-in palette for diverging data
                         node_color_interval = c(-3, 3), # The range of `log2_median_ratio` to display
                         edge_color_interval = c(-3, 3), # The range of `log2_median_ratio` to display
                         node_size_axis_label = "Number of OTUs",
                         node_color_axis_label = "Log2 ratio median proportions",
                         layout = "davidson-harel", # The primary layout algorithm
                         initial_layout = "reingold-tilford") # The layout algorithm that initializes node locations

   
   
 }


# clearly define the output object you will get from the function
return(output)   


}


#define function that calculates proportion tests between 2 phyloseq objects
fisher_all_taxa_groups<-function(ps_important_taxa, ps_all_taxa){

#**********************************************#
############# fisher_all_taxa_groups ################
#**********************************************#

# this function takes 2 phyloseq objects as 2 arguments:
# ps_important_taxa =  a subset of imporntant taxa, such as ASVs tagged as important by differential abundance, random forest, and netowrk analysis
# ps_all_taxa = the full phyloseq objects from where you obtained the imporntat subset (and likely your input for differential abundance, random forest, and netowork analysis)

# Then it runs a fisher test, comparing the proportions of every taxonomic level accuting more than twice in both datasets
# it's just a test to compare proportions - is 6 out of 17 a similar proportion to 193 out of 2243?

  
  
# first, get the taxonomic groups of the taxa defined as important
  
  #this will get us a list of (taxa_level) that appears in the important subset more than once
  imp_phylum_l<-as.character(tax_table(ps_important_taxa)[,"Phylum"]) # gets a char vector  of "families" shown as relevant
  imp_phylum_l<-imp_phylum_l[imp_phylum_l != "p__uncultured"] # removes any taxonomy set as "uncultured"
  imp_phylum_l<-names(which(table(imp_phylum_l)>2))%>% # get names of taxa that occur ate least 1 time NOTE: CHANGED FROM >2
    na.omit()%>% #remove NA from classifications
    unique()%>%  #dereplicates repetitive values (avoids "f__Chitinophagaceae" "f__Chitinophagaceae" "f__Oxalobacteraceae")
    as.list(c()) # save the dereplicated values as a list
  
  # now, perfom the same as above for every taxa level... it's hard-coded, but it works well enough. Pedro tried automating this and gave up after a few hours
  imp_class_l<-as.character(tax_table(ps_important_taxa)[,"Class"]) 
  imp_class_l<-imp_class_l[imp_class_l != "c__uncultured"] 
  imp_class_l<-names(which(table(imp_class_l)>2))%>% 
    na.omit()%>% 
    unique()%>%  
    as.list(c()) 
  
  imp_order_l<-as.character(tax_table(ps_important_taxa)[,"Order"])
  imp_order_l<-imp_order_l[imp_order_l != "o__uncultured"]
  imp_order_l<-names(which(table(imp_order_l)>2))%>%
    na.omit()%>% 
    unique()%>%
    as.list(c()) 
  
  imp_fam_l<-as.character(tax_table(ps_important_taxa)[,"Family"])
  imp_fam_l<-imp_fam_l[imp_fam_l != "f__uncultured"]
  imp_fam_l<-names(which(table(imp_fam_l)>2))%>%
    na.omit()%>% 
    unique()%>% 
    as.list(c()) 
  
  imp_genus_l<-as.character(tax_table(ps_important_taxa)[,"Genus"])
  imp_genus_l<-imp_genus_l[imp_genus_l != "g__uncultured"]
  imp_genus_l<-names(which(table(imp_genus_l)>2))%>%
    na.omit()%>% 
    unique()%>% 
    as.list(c()) 
  
  
  
  
  
  
  
  
  
  # second, get the number of taxa occuring in each taxonomic group, within the imporntat taxa subset
  # note: using phyloseq::subset_taxa with a for loop will cause issues as your taxa is intepreted as "i"! to avoid this, use phyloseq::prune_taxa instead
  
  
  
  #make one empty lists to store results
  target_in_important_n<-list()
  
  # obtain the number of reads of the target taxa in the important subset
  for(i in imp_phylum_l) { #for every Phylum "i" with representatives defined as important....
    target_in_important_n[i]<-phyloseq::ntaxa(prune_taxa(taxa = tax_table(ps_important_taxa)[,"Phylum"] %in% i, # get only the taxa that belong to that particular phylum "i", then define the number of taxa in there, then save it in the empty list we just made to store the results
                                                         x = ps_important_taxa))
  }
  
  # now, perfom the same as above for every taxa level... it's hard-coded, but it works well enough. Pedro tried automating this and gave up after a few hours
  for(i in imp_class_l) {
    target_in_important_n[i]<-phyloseq::ntaxa(prune_taxa(taxa = tax_table(ps_important_taxa)[,"Class"] %in% i,
                                                         x = ps_important_taxa))
  }
  
  for(i in imp_order_l) {
    target_in_important_n[i]<-phyloseq::ntaxa(prune_taxa(taxa = tax_table(ps_important_taxa)[,"Order"] %in% i,
                                                         x = ps_important_taxa))
  }
  
  
  for(i in imp_fam_l) {
    target_in_important_n[i]<-phyloseq::ntaxa(prune_taxa(taxa = tax_table(ps_important_taxa)[,"Family"] %in% i,
                                                         x = ps_important_taxa))
  }
  
  for(i in imp_genus_l) {
    target_in_important_n[i]<-phyloseq::ntaxa(prune_taxa(taxa = tax_table(ps_important_taxa)[,"Genus"] %in% i,
                                                         x = ps_important_taxa))
  }
  
  
  
  
  
  
  
  
  
  # third, get the number of taxa occuring in each taxonomic group, within the full dataset 
  
  #make one empty lists to store results
  target_in_all_n<-list()
  
  # obtain the number of reads of the target taxa in the important subset
  for(i in imp_phylum_l) { #for every Phylum "i" with representatives defined as important....
    target_in_all_n[i]<-phyloseq::ntaxa(prune_taxa(taxa = tax_table(ps_all_taxa)[,"Phylum"] %in% i, # get only the taxa that belong to that particular phylum "i", then define the number of taxa in there, then save it in the empty list we just made to store the results
                                                   x = ps_all_taxa))
  }
  
  # now, perfom the same as above for every taxa level... it's hard-coded, but it works well enough. Pedro tried automating this and gave up after a few hours
  for(i in imp_class_l) {
    target_in_all_n[i]<-phyloseq::ntaxa(prune_taxa(taxa = tax_table(ps_all_taxa)[,"Class"] %in% i,
                                                   x = ps_all_taxa))
  }
  
  for(i in imp_order_l) {
    target_in_all_n[i]<-phyloseq::ntaxa(prune_taxa(taxa = tax_table(ps_all_taxa)[,"Order"] %in% i,
                                                   x = ps_all_taxa))
  }
  
  
  for(i in imp_fam_l) {
    target_in_all_n[i]<-phyloseq::ntaxa(prune_taxa(taxa = tax_table(ps_all_taxa)[,"Family"] %in% i,
                                                   x = ps_all_taxa))
  }
  
  for(i in imp_genus_l) {
    target_in_all_n[i]<-phyloseq::ntaxa(prune_taxa(taxa = tax_table(ps_all_taxa)[,"Genus"] %in% i,
                                                   x = ps_all_taxa))
  }
  
  
  
  
  
  
  
  
  
  # now, get the total number of taxa in the imporntat subset and in the full dataset
  
  # all important taxa
  all_taxa_in_important_n<-phyloseq::ntaxa(ps_important_taxa)
  
  #all taxa 
  all_taxa_in_all_n<-phyloseq::ntaxa(ps_all_taxa)
  
  
  
  
  
  
  
  
  
  # now perform fisher tests over lists; check online tutorials for fisher.test() if need
  
  # this contigency table: (summed marginal totals is equal to the total number of taxa the the full object)
  fisher_result<-mapply(function (target_in_important_n,target_in_all_n)
    fisher.test(matrix(c(target_in_important_n, 
                         all_taxa_in_important_n - target_in_important_n, 
                         target_in_all_n - target_in_important_n, 
                         all_taxa_in_all_n - all_taxa_in_important_n - target_in_all_n),
                       ncol=2),alternative = "greater" ), 
    target_in_all_n = target_in_all_n,
    target_in_important_n = target_in_important_n,
    SIMPLIFY = FALSE)
  
  return(fisher_result)
  
}

#define function that puts fisher tests in metacoder heat trees
fisher_to_heatTree<-function(fisher_output, ps_important_taxa){
#**********************************************#
############# fisher_to_heatTree ################
#**********************************************#
  
# this function takes 1 phyloseq object and the putput of the fisher_all_taxa_groups() function as 2 arguments:
# fisher_output =  The output of the function defined above, fisher_all_taxa_groups
# ps_important_taxa = a phtloseq object with the taxa classified as "important". it is the same input as you used in the function fisher_all_taxa_groups  

# this function will:
  # 1)  create a taxmap object from the phyloseq object
  # 2)  extract the p values from the fisher test and apply an fdr correction
  # 3)  extract the taxa names form the heat tree, and match it to the p values of the fisher test
  # 4)  extract the odds ratio from the fisher test and add it to the extracted p values adn taxonomies
  # 5)  creates a heat map of the important taxa, with colour for adjusted p values and node size for odds ratio

  
  
#######################################
######## make metacoder object ######## 
#######################################
imp_heat_ps<-ps_important_taxa

#remove unecessary taxonomic indo (dada2id, "S__" and" above_selected)
tax_table(imp_heat_ps)<-tax_table(imp_heat_ps)[,2:7]

# let's remove the "r__"ranks from the taxonomy, they can be useful but will pollute our plot
#tax_table(imp_heat_ps)[, colnames(tax_table(imp_heat_ps))] <- gsub(pattern = "[a-z]__", # regular expression pattern to search for
 #                                                                  x = tax_table(imp_heat_ps)[, colnames(tax_table(imp_heat_ps))], # "df"
  #                                                                 replacement = "") # replacement for pattern

# transform from phyloseq to  taxmap object
imp_heat<-parse_phyloseq(imp_heat_ps)



#######################################
######## aplly fdr to fisher ######## 
#######################################

#apply FDR correction to the p values

  #single df with all p values
  pvector<-do.call(rbind.data.frame, map(fisher_output,1))
  
  # adjust the p value with fdr, then save as a list
  adjusted_p<-p.adjust(p = pvector[[1]],
                          method = "fdr")%>%
                      as.list()
  # adjust names
  names(adjusted_p)<-names(fisher_output)

  

#this is the name that mustch match our fisher p list
taxon_id_metacoder<-lapply(imp_heat$taxa, function (x)
                                                    x$get_name())%>%
                                                    map(1)


# now turn that list into a df
taxon_id_metacoder<-do.call(rbind.data.frame, map(taxon_id_metacoder,1))

# and change column name
colnames(taxon_id_metacoder)<-"taxa_id"

# truns the p values into a dataframe
adjusted_p_input<-as.data.frame(t(as.data.frame(adjusted_p)))%>%
  rownames_to_column()

#change column names
colnames(adjusted_p_input)<-c("taxa_id", "fdr_p")
  
# now join both df
fdr_for_heatTree<-left_join(taxon_id_metacoder,
                            adjusted_p_input,
                            by = "taxa_id")



#######################################################################
################### add fold changes! ########################################
#######################################################################


# get odds ratio for each taxa
odd_list<- map(fisher_output,3)%>%
  map(1)

# truns the p values into a dataframe
odd_list<-as.data.frame(t(as.data.frame(odd_list)))%>%
  rownames_to_column()

#change column names
colnames(odd_list)<-c("taxa_id", "odds_ratio")

# now join both df
fisher_for_heatTree<-left_join(fdr_for_heatTree,
                            odd_list,
                            by = "taxa_id")




#NAs will crash metadore. let's change them with 0.9999 for now
fisher_for_heatTree[is.na(fisher_for_heatTree)]<-"1"

#change structure to numeric
fisher_for_heatTree$fdr_p<-as.numeric(fisher_for_heatTree$fdr_p)
fisher_for_heatTree$odds_ratio<-as.numeric(fisher_for_heatTree$odds_ratio)


#now let's dim p values above 0.1, or else we can get counfounded
#fisher_for_heatTree$fdr_p[fisher_for_heatTree$fdr_p>0.1]<-1







#######################################################################
############### now plot the heat tree ############### 
#######################################################################



set.seed(1)
      output<- heat_tree(imp_heat,
                 node_size = fisher_for_heatTree$odds_ratio, # n_obs is a function that calculates, in this case, the number of OTUs per taxon
                 node_color = fisher_for_heatTree$fdr_p,
                 node_label = taxon_names,
                 node_size_axis_label = "Size: oods_ratio",
                 node_color_interval = c(0, 0.15),
                 edge_color_interval = c(0, 0.15),
                 node_color_range = c("darkgreen", "seagreen", "yellowgreen", "grey"),
                 edge_color_range = c("darkgreen", "seagreen", "yellowgreen", "grey"),
                 node_color_axis_label = "Color: FDR-adjusted p",
                 layout = "davidson-harel", # The primary layout algorithm
                 initial_layout = "reingold-tilford") # The layout algorithm that initializes node locations

      return(output)
}

# define function to extract significat p values from the fisher output into a df(filtered to p<0.1)
fisherOut_to_df_pfilter<-function(fisher_all_taxa_groups_output){
  
  #single df with all p values
  pvector<-do.call(rbind.data.frame, map(fisher_all_taxa_groups_output,1))
  # odsvector<-do.call(rbind.data.frame, map(fisher_all_taxa_groups_output,3)) # we might want to include odds ratio with this
  
  # adjust the p value with fdr, then save as a list
  adjusted_p<-p.adjust(p = pvector[[1]],
                          method = "fdr")%>%
                      as.list()
  # adjust names
  names(adjusted_p)<-names(fisher_all_taxa_groups_output)

  # adjust df format
  adjusted_p_df<- as.data.frame(unlist(adjusted_p))
  
  #move rownames
  adjusted_p_df<-rownames_to_column(adjusted_p_df,var = "taxa" )
  
  #change column names
  colnames(adjusted_p_df)<-c("taxa","pvalue")
  
  #filter based on p values
  adjusted_p_df<-adjusted_p_df[adjusted_p_df$pvalue <0.1,] # this column name can be a problem

  return(adjusted_p_df)


}




```


###Metacoder with Pedro: use fisher test on diff abund output
```{r}
load("./05_Differential_abundance/Objects/twotests_MeJA_Bac_ps_unnorm_31plants.RData")
load("./05_Differential_abundance/Objects/MeJA_unnorm_Bac_ps_l_31plants.RData")

#MeJA_Bac
#run custom function, making a fisher tests that compares proportions of each taxa in the diff_abund list to the rest of the dataset
MeJA_Bac_fisher_out_v2 <-mapply(function(x,y) 
  fisher_all_taxa_groups(ps_important_taxa =  x ,
                                  ps_all_taxa = y),
  x = twotests_MeJA_Bac_ps_unnorm_31plants,
  y = MeJA_unnorm_Bac_ps_l_31plants,
  SIMPLIFY = FALSE)

#Filter out plants with 0 taxa detected as "important"
f_MeJA_Bac_fisher_out_v2 <- MeJA_Bac_fisher_out_v2[lapply(MeJA_Bac_fisher_out_v2,length)>0]

#run custom function over lists of fisher test outputs
MeJA_Bac_df_pval_fisherComparison_l<-lapply(f_MeJA_Bac_fisher_out_v2, fisherOut_to_df_pfilter)

# make a table of p values <0.1
MeJA_Bac_df_pval_fisherComparison<-do.call(rbind.data.frame,MeJA_Bac_df_pval_fisherComparison_l)
table(MeJA_Bac_df_pval_fisherComparison_l$taxa)%>%sort

write.table(MeJA_Bac_df_pval_fisherComparison, file="./05_Differential_abundance/Results/FisherHeattree_onTwoTestsASVs/MeJA_Bac_PerPlant_df_pval_fisherComparison_l_DAASVs.txt",
sep = "\t", quote = F, row.names = T)

#SA_Bac
load("./05_Differential_abundance/Objects/twotests_SA_Bac_ps_unnorm_32plants.RData")
load("./05_Differential_abundance/Objects/SA_unnorm_Bac_ps_l_32plants.RData")

#SA_Bac
#run custom function, making a fisher tests that compares proportions of each taxa in the diff_abund list to the rest of the dataset
SA_Bac_fisher_out_v2 <-mapply(function(x,y) 
  fisher_all_taxa_groups(ps_important_taxa =  x ,
                                  ps_all_taxa = y),
  x = twotests_SA_Bac_ps_unnorm_32plants,
  y = SA_unnorm_Bac_ps_l_32plants,
  SIMPLIFY = FALSE)

#Filter out plants with 0 taxa detected as "important"
f_SA_Bac_fisher_out_v2 <- SA_Bac_fisher_out_v2[lapply(SA_Bac_fisher_out_v2,length)>0]

#run custom function over lists of fisher test outputs
SA_Bac_df_pval_fisherComparison_l<-lapply(f_SA_Bac_fisher_out_v2, fisherOut_to_df_pfilter)

# make a table of p values <0.1
SA_Bac_df_pval_fisherComparison<-do.call(rbind.data.frame,SA_Bac_df_pval_fisherComparison_l)

write.table(SA_Bac_df_pval_fisherComparison, file="./05_Differential_abundance/Results/FisherHeattree_onTwoTestsASVs/SA_Bac_PerPlant_df_pval_fisherComparison_l_DAASVs.txt",
sep = "\t", quote = F, row.names = T)


#MeJA_Fun
load("./05_Differential_abundance/Objects/twotests_MeJA_Fun_ps_unnorm_28plants.RData")
load("./05_Differential_abundance/Objects/MeJA_unnorm_Fun_ps_l_28plants.RData")

#MeJA_Fun
#run custom function, making a fisher tests that compares proportions of each taxa in the diff_abund list to the rest of the dataset
MeJA_Fun_fisher_out_v2 <-mapply(function(x,y) 
  fisher_all_taxa_groups(ps_important_taxa =  x ,
                                  ps_all_taxa = y),
  x = twotests_MeJA_Fun_ps_unnorm_28plants,
  y = MeJA_unnorm_Fun_ps_l_28plants,
  SIMPLIFY = FALSE)

#Filter out plants with 0 taxa detected as "important"
f_MeJA_Fun_fisher_out_v2 <- MeJA_Fun_fisher_out_v2[lapply(MeJA_Fun_fisher_out_v2,length)>0]

#run custom function over lists of fisher test outputs
MeJA_Fun_df_pval_fisherComparison_l<-lapply(f_MeJA_Fun_fisher_out_v2, fisherOut_to_df_pfilter)

# make a table of p values <0.1
MeJA_Fun_df_pval_fisherComparison<-do.call(rbind.data.frame,MeJA_Fun_df_pval_fisherComparison_l)

write.table(MeJA_Fun_df_pval_fisherComparison, file="./05_Differential_abundance/Results/FisherHeattree_onTwoTestsASVs/MeJA_Fun_PerPlant_df_pval_fisherComparison_l_DAASVs.txt",
sep = "\t", quote = F, row.names = T)



#SA_Fun
load("./05_Differential_abundance/Objects/twotests_SA_Fun_ps_unnorm_29plants.RData")
load("./05_Differential_abundance/Objects/SA_unnorm_Fun_ps_l_29plants.RData")


#SA_Fun
#run custom function, making a fisher tests that compares proportions of each taxa in the diff_abund list to the rest of the dataset
SA_Fun_fisher_out_v2 <-mapply(function(x,y) 
  fisher_all_taxa_groups(ps_important_taxa =  x ,
                                  ps_all_taxa = y),
  x = twotests_SA_Fun_ps_unnorm_29plants,
  y = SA_unnorm_Fun_ps_l_29plants,
  SIMPLIFY = FALSE)

#Filter out plants with 0 taxa detected as "important"
f_SA_Fun_fisher_out_v2 <- SA_Fun_fisher_out_v2[lapply(SA_Fun_fisher_out_v2,length)>0]

#run custom function over lists of fisher test outputs
SA_Fun_df_pval_fisherComparison_l<-lapply(f_SA_Fun_fisher_out_v2, fisherOut_to_df_pfilter)

# make a table of p values <0.1
SA_Fun_df_pval_fisherComparison<-do.call(rbind.data.frame,SA_Fun_df_pval_fisherComparison_l)

write.table(SA_Fun_df_pval_fisherComparison, file="./05_Differential_abundance/Results/FisherHeattree_onTwoTestsASVs/SA_Fun_PerPlant_df_pval_fisherComparison_l_DAASVs.txt",
sep = "\t", quote = F, row.names = T)

```


###Metacoder with Pedro: put fisher tests results ina  tree
```{r}
#plor one species opn heat tree (fisher tests on deseq output)
fisher_to_heatTree(fisher_output = fisher_out,
                   ps_important_taxa =MeJA_Bac_conca_ps$At )


```


###Test effect of not filtering absent ASVs from the plants on standard DeSeq output
```{r}
#MeJA_Bac
#Standard DESeq
stddds2withfilt_Wald_MeJA_Bac <- lapply(dds_MeJA_Bac_ps, function(x) 
                  DESeq(x, 
                  test="Wald",  
                  minReplicatesForReplace=Inf,
                  fitType = "local")) 

save(stddds2withfilt_Wald_MeJA_Bac, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/stddds2withfilt_Wald_MeJA_Bac.RData")

#Dataframe generation
res_stddds2withfilt_Wald_MeJA_Bac <- lapply(stddds2withfilt_Wald_MeJA_Bac, function(x) results(x, contrast = c("Stress", "MeJA", "Control")))

save(res_stddds2withfilt_Wald_MeJA_Bac, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/res_stddds2withfilt_Wald_MeJA_Bac.RData")

#Filter to only p adjusted under alpha
alpha = 0.05
sigtab_stddds2withfilt_Wald_MeJA_Bac <- lapply(res_stddds2withfilt_Wald_MeJA_Bac, function(x) x[which(x$padj < alpha), ])

save(sigtab_stddds2withfilt_Wald_MeJA_Bac, 
     file = "./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/sigtab_stddds2withfilt_Wald_MeJA_Bac.RData")


#Take significant asvs from each test in a vector
Load("./05_Differential_abundance/Objects/5.1deseqandZINBdeseq/sigtab_stddds2_Wald_MeJA_Bac.RData")
MeJA_Bac_DESeqWald_ASVs_unfilt <- lapply(sigtab_stddds2_Wald_MeJA_Bac, function(x) rownames(x))
MeJA_Bac_DESeqWald_ASVs_filt <- lapply(sigtab_stddds2withfilt_Wald_MeJA_Bac, function(x) rownames(x))

#Check if there's differences
diff_filt_unfilt_stddds <- setdiff(MeJA_Bac_DESeqWald_ASVs_filt, MeJA_Bac_DESeqWald_ASVs_unfilt)
#no differences
```

