---
title: "Untitled"
author: "Marcela Aragon"
date: "31/1/2024"
output: html_document
---

# 1.0 set up working environemnt

## load libraries

```{r}

renv::deactivate()

# set working directory to project directory 

#data
library("phyloseq") #microbial data
library("vegan") #microbial data
library("ggpubr") 
library("dplyr")
library("tidyr")
library("tibble")
library("purrr") #works with lists
library("readxl") 
library("plyr")
library("ggplotify") #change as.grob
library("gridExtra")

#plots/visualizations
library("ggplot2")
library("heatmaply")
library("pheatmap")
library("RColorBrewer")
library("viridis")
library("grid") #change as.grob
library("lattice")
library("ggplotify") #change as.grob
#library("ggpmisc") #plots R2 and regression lines

#stats
library("car") # for Levene's test
library("ade4") #multi-variate analysis
library("stats") #cophenetic distances 
 
#phylo
library("seqinr") #to read alignments
library("adephylo")
library("phylobase")
library("phylosignal")
library("ape") #handles phylogenetic objects
library("TreeTools")
library("tidytree")
library("phytools") 
library("phangorn") #UPGMA & NJ
library("distory") #to compare trees
library("geiger")

# increases memory limit used by R
memory.limit(size = 500000)

#load some relevant custom functionseveyrone uses, like lapply_lapply
source("./renv/functions/lapply_lapply.R")

```


## 1.02 Settings for plots

```{r}
# Set axis looks
theme_set(theme_bw())
axis_looks <- theme(axis.text.x = element_text(colour = "black", size = 10,
                                               face = "bold"))+
  theme(axis.text.y = element_text(colour = "black", size = 12, face = "bold"))+
  theme(axis.title=element_text(size=15, face = "bold"))
```


## 1.03 Load data & metadata

I'll load the .fas concatenated file first as well as the excel file with the species code from Kasper and our metadata to change later the species names.


**I need to re-order plant phylogeny in same folder, instead of in 14_Phylogenetic_distance**

```{r}

#load species code from .xsl file
sp_Kasper <- read_excel("./14_Phylogenetic_distance/Genetic_alignments_KasperH/2023-11-16_Brassicaceae_crop_genetic_data_for_MarcelaAG.xlsx")

#Load alignment data
alignment <- ape::read.dna(file = "./14_Phylogenetic_distance/Genetic_alignments_KasperH/2023-11-20_Concatenated_codon_aware_alignments_for_Brassica_crop_project_Marcela_Aragon-Gomez.fas", format = "fasta")

#Load new plant phylogenetic tree
Bra_tree <- read.tree("./17_Plant_phylogeny/Results/BrassiToL_supermatrix.tree")
plot(Bra_tree) #coded as sample name
summary(Bra_tree) #23 tips 

#Load phyloseq metadata
# this will load the CSS-normalized 16S and ITS data as a list of two objects
load("./Data/phyloseq_objects/CSS_BacFun_rhizoplane_HighFilt_ps_l.RData")

#View
CSS_BacFun_rhizoplane_HighFilt_ps_l

#Get only metadata df
sp_Family <- sample_data(CSS_BacFun_rhizoplane_HighFilt_ps_l$Bac)

#clean
rm(CSS_BacFun_rhizoplane_HighFilt_ps_l)

```

## 1.04 Arranging data

```{r}

#Clean Family metadata

#Get info from plant species kept for the analysis 
sp_Family2 <- sp_Family %>%
                   as_tibble() %>%
                   dplyr::filter(Sp_Family == "Brassicaceae")%>%
                   dplyr::select(Sp_full_name, Sp_abb_name, MY_tree_presence)

sp_Family3 <- sp_Family2%>%
                   distinct(Sp_full_name, Sp_abb_name, MY_tree_presence) #there are 26 different species originally

#Remove Kimmeridge and the pther two speeds (F and S) of B.ole as they were not included in the overall analysis 
sp_Family4 <- sp_Family3%>%
                   dplyr::filter(!Sp_abb_name %in% c("Bo_F", "Bo_S", "BoK"))%>%
                   droplevels() #23 species now 

#Arrange Kasper's data 
#Now, make a single df with Kasper's coding
sp_Kasper

#For this, I need to match our species names with those of Kasper
sp_Family4$Sp_full_name #23 species 
sp_Kasper$Species #23 species (all of the species without B_Kimmerdige)

#Let's add a "_" to Kasper's species names
sp_Kasper$Species <- sub(" ", "_", sp_Kasper$Species)
sp_Kasper$Species <- as.factor(sp_Kasper$Species)
sp_Kasper$Sample <- as.factor(sp_Kasper$Sample)
sp_Kasper$Tribe <- as.factor(sp_Kasper$Tribe)

#For now, let's change the name of Mutarda_nigra and Mutarda_arvensis for Brassica and Sinapis, respectively, as well as same name for Bole
sp_Kasper2 <- sp_Kasper
sp_Kasper2$Species <- revalue(sp_Kasper2$Species, c("Mutarda_arvensis"="Sinapis_arvensis", "Mutarda_nigra"="Brassica_nigra", "Brassica_oleracea"="Brassica_oleraceae_Rivera_Medium")) #works

#change colname of 'Species' to same as family
sp_Kasper3 <- sp_Kasper2
colnames(sp_Kasper3)[3] <- colnames(sp_Family4)[1]

#let's check
test <- full_join(sp_Family4, sp_Kasper3, by="Sp_full_name")
View(test)

# 5 names of our species are wrong (sorry!) for now I'll leave them misspelled but will change later in metadata
sp_Kasper4 <- sp_Kasper3
sp_Kasper4$Sp_full_name <- revalue(sp_Kasper4$Sp_full_name, 
                                 c("Descurainia_sophia"="Descurainia_sophiades",
                                   "Erysimum_cheiranthoides"="Erysium_cheirantoides",
                                   "Euclidium_syriacum"="Euclydium_syriacum",
                                   "Malcolmia_maritima"="Malcomia_maritima",
                                   "Sisymbrium_irio"= "Sysimbrium_irio")) 

#let's check again
df_species <- full_join(sp_Family4, sp_Kasper4, by="Sp_full_name") #now it works!
View(df_species)

#let's clean 
rm(sp_Family, sp_Family2, sp_Family3, sp_Family4, sp_Kasper, sp_Kasper2, sp_Kasper3, sp_Kasper4, test)

```

**Change species names for correct ones in our data**


## 1.05 Arranging plant phylogeny

```{r}

Bra_tree
plot(Bra_tree)

#get tip labels
tips <- Bra_tree$tip.label
#make if a df
tips2 <- tips%>%
         as_tibble()%>%
         dplyr::rename(., Sample=value)
#get only Sample and Sp code 
sel <- df_species%>%
       select(Sample, Sp_abb_name)
#make a new df 
tips3 <- left_join(tips2, sel)

#check if same order as tree tips
tips
tips3$Sample #yes

#make character vector with new sp names in order
new_tips <- as.character(tips3$Sp_abb_name)

#now replace with sp names
Bra_tree$tip.label <- new_tips
plot(Bra_tree)
axisPhylo()

#check the tree
summary.phylo(Bra_tree)
is.rooted(Bra_tree) #no
is.ultrametric(Bra_tree) #no

#save it
write.tree(Bra_tree, file = "./17_Plant_phylogeny/Results/Supermatrix_Kasper_tree_renamed.tre")

#The order is weird, so in order to rotate clades as we had them in old tree I will export the Supermatrix_Kasper_tree_renamed.tre file
#Into iTOL and rotate branches in there (as it is way easier than in R)
#then I will export that treee and import it again here 

#change order in iTOL, see if it works. Yes it does! :)
Bra_tre2 <- read.tree("./17_Plant_phylogeny/Results/Supermatrix_Kasper_tree_renamed_rotated.tre")
plot(Bra_tre2) #much better

#Still it is a bit weird that it is not rooted, so I'll do that again in iTOl and export it back
Bra_tre3 <- read.tree("./17_Plant_phylogeny/Results/Supermatrix_Kasper_tree_renamed_rotated_rooted.tre")
plot(Bra_tre3) #even better

#save it
write.tree(Bra_tre3, file = "./17_Plant_phylogeny/Results/Supermatrix_Kasper_tree_Final_28-08-24.tre")

#This is our final tree, now it can be used for figures and analysis

```

